### 2020-05-07 chapter1.Rmd
"initial" commit

```{r libraries/definitions, include=FALSE}
library(stringi) # %s+%
library(lubridate) # floor_date
library(xts) # zoo
library(fpp) # auto.arima
library(tidyverse) # map_df, read_csv, group_by, summarize
library(tsbox) # https://cran.r-project.org/web/packages/tsbox/vignettes/tsbox.html
print_dec <- function(x, decimals=2) trimws(format(round(x, decimals), nsmall=decimals))

setwd("/media/data/Documents/Uni/spatiotemp/villalobos/project/")

# UCCS
domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"
# MEMS
# domain <- c(-7200,7200); pattern <- "MEM.*.csv"; name <- "MEM"; description <- "Membach"
```
```{r import dataset, include=FALSE}
seis <-
  list.files(path="csv/", pattern=pattern) %>%
  map_df(~read_csv("csv/" %s+% .))
```

# Data preparation
```{r dataset description, echo=FALSE}
long_description <- description %s+% " (" %s+% name %s+% " station)"
long_description
"Domain values: " %s+% domain[1] %s+% " to " %s+% domain[2] %s+% " nm"
"Data from " %s+% format.POSIXct(min(seis$utc_second), "%Y-%m-%d %H:%M:%S") %s+% " to " %s+% format.POSIXct(max(seis$utc_second), "%Y-%m-%d %H:%M:%S")
```

```{r cleanup, echo=FALSE}
out_domain <- (seis$min < domain[1] | seis$max > domain[2])
invalid_values <- (seis$min == -1 & seis$max == 1) | (seis$min == 0 & seis$max == 0) | out_domain
"Removing " %s+% sum(out_domain) %s+% " values out of domain, " %s+% (sum(invalid_values) - sum(out_domain)) %s+% " explicit NA"
seis <-
  seis %>%
  mutate(min=replace(min, invalid_values, NA), max=replace(max, invalid_values, NA)) %>%
  drop_na()
# %>% mutate(belgian_second=with_tz(utc_second, "Europe/Brussels"))
remove(out_domain, invalid_values) # free ~350 Mb from RAM
seis.present <- length(seis$utc_second)
seis.total <- max(as.numeric(seis$utc_second)) - min(as.numeric(seis$utc_second))
seis.missing <- seis.total - seis.present
seis.present %s+% " present values, " %s+% seis.missing %s+% " missing points (" %s+% print_dec(seis.present / seis.total * 100, 2) %s+% "% / " %s+% print_dec(seis.missing / seis.total * 100, 2) %s+% "%)"

```

```{r build_ts function}
build_ts <- function(dset, periods=NULL, start=NULL, end=NULL, force_ms=T) {
  dset.ts <- ts_zoo(dset)
  if (!is.null(start) | !is.null(end)) {
    dset.ts <- ts_span(dset.ts, start=start, end=end)
  }
  dset.ts <- ts_regular(dset.ts)
  if (is.null(periods)) {
    return(dset.ts)
  }
  if (is.vector(periods) & length(periods) == 1) {
    periods <- periods[1]
  }
  if (is.integer(periods)){
    if (force_ms) {
      periods <- c(periods)
    } else {
      frequency(dset.ts) <- periods
      return(dset.ts)
    }
  }
  return(msts(dset.ts, periods))
}
```

```{r group by hour}
seis.h <- 
  seis %>%
  mutate(utc_hour=floor_date(with_tz(utc_second, "Europe/Brussels"), "1 hour")) %>%
  group_by(utc_hour=as_datetime(date(utc_hour)) + hour(utc_hour) * 3600) %>% # this forces the hour to represent bruxelles local time in utc. tz-ed time is normally evaluated in utc by ts-related functions
  summarize(mean_movement_s=mean(max-min)) 
# we'll use ts_regular instead of completing data here
# %>% complete(utc_hour=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = as.difftime("1:0:0")))
```

```{r seasonplots}

seis.h.start <- with_tz(as_datetime('2018-11-01T00:00:00'), "Europe/Brussels")
seis.h.end <- with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels")
seis.h.msts <- build_ts(seis.h, periods = c(24), start = seis.h.start, end = seis.h.end)
seasonplot(seis.h.msts, type="l", col=rainbow(7))
legend("topleft", legend=weekdays(seis.h.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
tsdisplay(seis.h.msts, lag.max = 168*2)
# tsdisplay(diff(seis.h.msts, 8), lag.max = 168*2)
tsdisplay(diff(seis.h.msts, 24), lag.max = 168*2)
tsdisplay(diff(seis.h.msts, 168), lag.max = 168*2)
tsdisplay(diff(diff(seis.h.msts, 24), 168), lag.max = 168*2)

seis.h.list <- as_datetime('2018-11-01T00:00:00') %m+% months(c(0:16))
for (i in 1:(length(seis.h.list)-1)) {
  seis.h.msts <- build_ts(seis.h, periods = c(24), start = seis.h.list[i], end = seis.h.list[i+1])
  seasonplot(seis.h.msts,
             type="l",
             xlab="Hour (" %s+% seis.h.list[i] %s+% " - " %s+% print_dec(length(seis.h.msts) / as.numeric(seis.h.list[i+1] - seis.h.list[i])) %s+% " points per day)",
             col=rainbow(7)
             )
  legend("topleft", legend=weekdays(seis.h.list[i] + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
}

```

```{r test of ts, eval=FALSE, include=FALSE}
seis.h.xts <- build_ts(seis.h, start = seis.h.start, end = seis.h.end)
# https://stats.stackexchange.com/questions/120806/frequency-value-for-seconds-minutes-intervals-data-in-r
# https://robjhyndman.com/hyndsight/seasonal-periods/
seis.h.msts <- build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end)
taylor.fit <- tbats(seis.h.msts, num.cores = NULL)
taylor.fit
time(seis.h.msts)
plot(taylor.fit)
plot(forecast(taylor.fit))
ts_summary(seis.h.msts)
ts_plot(ts_trend(seis.h.xts))
ts_ggplot(seis.h.xts, title = 'Hourly seismic movement', subtitle = long_description %s+% " from " %s+% seis.h.start %s+% " to " %s+% seis.h.end, ylab='mean movement per second (nm)')
str(seis.h.xts)
plot(seis.h.xts)
seis.h.msts.dec <- decompose(seis.h.msts)
#seis.h.msts.dec$
plot(seis.h.msts.dec)
length(seis.h.xts)
frequency(seis.h.xts)
deltat(seis.h.xts)
```

```{r parallel fourier, include=FALSE}

combinations <- function(v1, v2=v1) {
  A <- matrix(nrow=0, ncol=2)
  for (i in v1) {
    for (j in v2) {
      A <- rbind(A, c(i, j))
    }
  }
  return(A)
}
auto.arima.fourier <- function(msts.data, periods, ...) {
  xreg.d <- fourier(msts.data, periods)
  print("Starting auto.arima with # of fourier factors " %s+% periods)
  fitma <- auto.arima(msts.data, ..., max.P=0, D=0, max.Q=0, seasonal=F, xreg=xreg.d)
  print("Finished auto.arima with # of fourier factors " %s+% periods)
  return(fitma)
}

library(parallel)
if (!is.null(getDefaultCluster())) {
  setDefaultCluster(makeCluster(6))
}

seis.h.start <- with_tz(as_datetime('2018-03-01T00:00:00'), "Europe/Brussels")
seis.h.end <- with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels")
seis.h.msts <- build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end)
co <- combinations(7:12)
co <- co[order(rowSums(co)+co[,2], decreasing=T),]
clusterExport(getDefaultCluster(), c("seis.h.msts", "co", "auto.arima.fourier", "fourier", "auto.arima", "%s+%"))
fitma.list <- parLapply(getDefaultCluster(), 1:nrow(co), function(r) auto.arima.fourier(seis.h.msts, co[r,])) # TODO set arima parameters
stopCluster(getDefaultCluster())

aic_vals <- data.frame(Fourier24=integer(), Fourier168=integer(), AIC=integer())
for (fitma in fitma.list) {
  f24_l <- as.numeric(str_match(na.remove(str_extract(names(fitma$coef), "S\\d+-24")), "S(\\d+)-24")[,2])
  f24 <- max(f24_l)
  f168_l <- as.numeric(str_match(na.remove(str_extract(names(fitma$coef), "S\\d+-168")), "S(\\d+)-168")[,2])
  f168 <- max(f168_l)
  desc <- paste0("Seasonality component for ARMA(", fitma$arma[1], ",", fitma$arma[2], ")")
  subdesc <- paste0("Fourier terms [24h: ", length(f24_l), " - max ", f24, "] [168h: ", length(f168_l), " - max ", f168, "]")
  # plot(rowSums(fitma$xreg[,(c(1:(length(f24_l)*2)))][1:(168*2),]), type="l", col="red", ylim=c(-12,+24), main=desc)
  # lines(rowSums(fitma$xreg[,(c((length(f24_l)*2+1):(length(f168_l)*2+length(f24_l)*2)))][1:(168*2),]), type="l", col="blue")
  drift <- length(na.remove(str_extract(names(fitma$coef), "drift")))
  plot(rowSums(fitma$xreg[(1):(168*2),(drift+1):ncol(fitma$xreg)]), type="l", col="red", ylim=c(-12,+24), main=desc, sub=subdesc, xlab="Hours")
  aic_vals[nrow(aic_vals) + 1, ] <- c(length(f24_l), length(f168_l), fitma$aic)
}
p <- ggplot(aic_vals, aes(Fourier24, Fourier168)) + geom_tile(aes(fill = AIC), colour = "white") + scale_fill_gradient(low = "white", high = "steelblue")
plot(p)

min_aic <- min(aic_vals$AIC)
minvals <- aic_vals[which(aic_vals$AIC == min_aic),]
minvals

fitted <- fitma.list[order(aic_vals$AIC)][[2]]
tsdisplay(residuals(fitted))

```

```{r plot fourier}
str(fitted$xreg)

```

```{r fourier}
seis.h.start <- with_tz(as_datetime('2018-11-01T00:00:00'), "Europe/Brussels")
seis.h.end <- with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels")
seis.h.msts <- build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end)
fitted <- auto.arima(seis.h.msts, D=0, max.P=0, max.Q=0, xreg=fourier(seis.h.msts, c(3, 4)))
plot(fitted)
plot(residuals(fitted))
tsdisplay(residuals(fitted))
res.xts <- ts_zoo(residuals(fitted))
res.msts <- msts(res.xts, c(24, 168))
res.aic_vals_temp <- NULL
res.aic_vals <- NULL
for (i in 1:5) {
  for (j in 1:5) {
    #for (k in 1:5) {
      res.xreg.d <- fourier(res.msts, c(i, j)  )  # daily  seasonality
      res.fitma <- auto.arima(res.msts, D=0, max.P=0, max.Q=0, xreg=res.xreg.d)
      res.aic_vals_temp <- cbind(i, j, res.fitma$aic)  #,k
      res.aic_vals <- rbind(res.aic_vals, res.aic_vals_temp)
      print("(" %s+% i %s+% ", " %s+% j %s+% ") = " %s+% res.fitma$aic)
    #}
  }
}
remove(res.aic_vals_temp, res.xreg.d, res.fitma)
colnames(res.aic_vals) <- c("Fourier24", "Fourier168", "AIC")
res.aic_vals <- data.frame(res.aic_vals)
res.min_aic <- min(res.aic_vals$AIC)
res.minvals <- res.aic_vals[which(res.aic_vals$AIC == res.min_aic),]
res.p <- ggplot(res.aic_vals, aes(Fourier24, Fourier168)) + geom_tile(aes(fill = AIC), colour = "white") + scale_fill_gradient(low = "white", high = "steelblue")
plot(res.p)
res.minvals
res.fitted <- auto.arima(res.msts, D=0, max.P=0, max.Q=0, xreg=fourier(res.msts, c(res.minvals$Fourier24, res.minvals$Fourier168)))
plot(res.fitted)
tsdisplay(residuals(fitted))
tsdisplay(residuals(res.fitted))

```

```{r periodgram}
library(spec)
#pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2018-11-01T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2019-05-03T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)

topfreq <- pgram$freq[order(pgram$spec, decreasing = T)][1:100]
topseas <- 1/topfreq
topseas.h <- topseas/3600

top95pc <- pgram$spec[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*1)]
top95freq <- pgram$freq[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*1)][order(top95pc, decreasing = T)]
top95pc <- pgram$spec[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*1)][order(top95pc, decreasing = T)]
top95seas.s <- 1/top95freq
top95seas.h <- top95seas.s/3600
top95pc/sd(pgram$spec)
top95seas.h

ylims <- 10^par("usr")[3:4]
axis(side=3, at=top95freq, labels=as.integer(top95seas.h), padj = 1, las=2, cex.axis=0.8)
abline(h=mean(pgram$spec) + sd(pgram$spec)*c(.5,1:3), lty="dotted", col=c("orange", "darkgreen", "blue", "red"))
for (i in 1:length(top95freq)) {
  color <- ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*3), "red", ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*2), "blue", ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*1), "darkgreen", "orange")))
  lines(x=c(top95freq[i], top95freq[i]), y=c(top95pc[i],ylims[2]), lty="dotted", col=color)
  points(x=top95freq[i], y=top95pc[i], col=color)
}
axis(side=4, at=mean(pgram$spec) + sd(pgram$spec) * c(.5, 1:3), labels=c(".5", "1", "2", "3"), cex.axis=0.8)
mtext(side=4, "spectrum sd ratio", line=1)
```

### 2020-05-07 crawl.py

@@ -1,0 +1,120 @@
+import datetime
+import os
+import pytz
+
+from settings import START_DATE, CSV_FOLDER, LAST_DAY_FILE
+from requests_session import SESSION
+
+TZ_BELGIAN = pytz.timezone('Europe/Brussels')
+TZ_UTC = pytz.utc
+TODAY = datetime.date.today()
+
+
+def get_csv_url_for(station_code: str, day: datetime.date):
+    year = day.year
+    year_start = datetime.date(year=year, month=1, day=1)
+    day_delta = day - year_start
+    padded_day_of_year = f"{day_delta.days + 1:0>3}"
+    return f"http://seismologie.be/data/csv/SEIS-{year}-{padded_day_of_year}-{station_code}.csv"
+
+
+def get_csv_for(station_code: str, day: datetime.date):
+    csv_url = get_csv_url_for(station_code, day)
+    with SESSION.get(csv_url) as f:
+        if f.ok:
+            return f.content.decode().split()
+        return []
+
+
+def iter_data_for(station_code: str, day: datetime.date):
+    day_data = get_csv_for(station_code, day)
+    if len(day_data) > 0:
+        utc_second = TZ_UTC.localize(datetime.datetime(year=day.year, month=day.month, day=day.day))
+        # belgian_second = TZ_BELGIAN.normalize(utc_second)
+        second = datetime.timedelta(seconds=1)
+        for row in day_data[1:]:
+            data = [utc_second.isoformat()]  # , belgian_second.isoformat()]
+            data.extend(row.split(','))
+            yield data
+            utc_second += second
+            # belgian_second += second
+
+
+def get_events_url_on(day: datetime.date):
+    return f"http://seismologie.be/data/csv/events_{day.isoformat()}.json"
+
+
+def get_events_on(day: datetime.date):
+    json_url = get_events_url_on(day)
+    with SESSION.get(json_url) as f:
+        if f.ok:
+            return f.json()
+        return {}
+
+
+def save_month(station_code: str, month: datetime.date, force: bool = False):
+    day = month.replace(day=1)
+    filename = f"{station_code}-{day:%Y-%m}.csv"
+    path = os.path.join(CSV_FOLDER, filename)
+    if os.path.exists(path) and not force:
+        while day.month == month.month and day < TODAY:
+            day += datetime.timedelta(days=1)
+        return day
+    print(f"Loading {day:%Y-%m}", end='... ')
+    try:
[…]

### 2020-05-07 first_analysis.R

@@ -1,0 +1,224 @@
+
+# var domains = {
+# mems: [-7200,7200],
+# uccs: [-350000,350000],
+# ebn:  [-216000,216000],
+# dou:  [-22500,22500],
+# rch:  [-4000,4000],
+# zev:  [-2000000,2000000],
+# sti:  [-14400,14400],
+# skq:  [-60000,60000],
+# ostb: [-9400,9400]
+# };
+
+library(stringi) # %s+%
+library(lubridate) # floor_date
+library(xts) # zoo
+library(fpp) # auto.arima
+library(tidyverse) # map_df, read_csv, group_by, summarize
+
+setwd("/media/data/Documents/Uni/spatiotemp/villalobos/project/")
+# UCCS
+domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"
+# MEMS
+domain <- c(-7200,7200); pattern="MEM.*.csv"
+seis <-
+  list.files(path="csv/", pattern=pattern) %>%
+  map_df(~read_csv("csv/" %s+% .))
+
+
+out_domain <- (seis$min < domain[1] | seis$max > domain[2])
+invalid_values <- (seis$min == -1 & seis$max == 1) | (seis$min == 0 & seis$max == 0) | out_domain
+seis <-
+  seis %>%
+  mutate(min=replace(min, invalid_values, NA), max=replace(max, invalid_values, NA)) %>%
+  complete(utc_second=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = as.difftime("0:0:1")))
+# %>% mutate(belgian_second=with_tz(utc_second, "Europe/Brussels"))
+remove(out_domain, invalid_values) # free ~300 Mb from RAM
+
+
+
+summary(seis)
+head(seis, n=5)
+nrow(seis)
+nrow(na.omit(seis))
+nrow(seis) - nrow(na.omit(seis))
+
+
+
+seis.h <- 
+  na.exclude(seis) %>%
+  filter(utc_second < with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels") ) %>%
+  group_by(utc_hour=floor_date(with_tz(utc_second, "Europe/Brussels"), "1 hour")) %>%
+  # filter(between(utc_hour, as_datetime('2020-04-01T00:00:00'), as_datetime('2020-05-01T00:00:00') )) %>%
[…]

### 2020-05-07 fs_cache.py

@@ -1,0 +1,75 @@
+"""
+File-system cache for requests_cache.
+Source: https://github.com/codypiersall/vismap/blob/9b5051db86bdb4ab4527ffb2a0c21b54e6dc3395/vismap/fs_cache.py
+"""
+
+try:
+    import cPickle as pickle
+except ImportError:
+    import pickle
+from collections import MutableMapping
+import os
+
+from requests_cache.backends.base import BaseCache
+
+
+class FSDict(MutableMapping):
+    """File-system based dict.  No care is taken to ensure that the key
+    given to the dict can actually exist as a file on the filesystem.  For
+    example, a key with a \0 in it will raise an Exception.
+    """
+    def __init__(self, root='cache'):
+        self.root = os.path.abspath(root)
+        os.makedirs(self.root, exist_ok=True)
+
+    def _key_to_path(self, key):
+        path = os.path.join(self.root, key)
+        return path
+
+    def __getitem__(self, key):
+        path = self._key_to_path(key)
+        if not os.path.exists(path):
+            raise KeyError
+        with open(path, 'rb') as f:
+            return f.read()
+
+    def __delitem__(self, key):
+        path = self._key_to_path(key)
+        os.unlink(path)
+
+    def __setitem__(self, key, value):
+        path = self._key_to_path(key)
+        if isinstance(value, str):
+            value = value.encode()
+        with open(path, 'wb') as f:
+            f.write(value)
+
+    def __iter__(self):
+        files = os.listdir(self.root)
+        return (os.path.join(self.root, p) for p in files)
+
+    def __len__(self):
+        return len(os.listdir(self.root))
+
+
+class FSPickleDict(FSDict):
+    """Like FSDict but pickles values"""
+
+    def __getitem__(self, key):
+        sup = super(FSPickleDict, self)
+        return pickle.loads(sup.__getitem__(key))
+
+    def __setitem__(self, key, item):
+        sup = super(FSPickleDict, self)
+        sup.__setitem__(key, pickle.dumps(item))
+
+
+class FSCache(BaseCache):
+    """ Filesystem cache backend."""
+    def __init__(self, location='cache', **options):
+        """
+        :param location: root directory on filesystem cache
+        """
+        super(FSCache, self).__init__(**options)
+        self.responses = FSPickleDict(os.path.join(location, 'responses'))
+        self.keys_map = FSDict(os.path.join(location, 'urls'))

### 2020-05-07 requests_session.py

@@ -1,0 +1,32 @@
+import time
+
+import requests_cache
+
+from settings import CACHE_FOLDER
+from fs_cache import FSCache
+
+__all__ = [
+    'SESSION',
+]
+
+
+def make_throttle_hook(timeout=1.0):
+    """
+    Returns a response hook function which sleeps for `timeout` seconds if
+    response is not cached
+    """
+
+    def hook(response, *args, **kwargs):
+        if not getattr(response, 'from_cache', False):
+            # print(f"Got {response.url}")
+            time.sleep(timeout)
+        return response
+
+    return hook
+
+
+SESSION = requests_cache.CachedSession('.cache', backend=FSCache(location=CACHE_FOLDER))
+SESSION.hooks = {'response': make_throttle_hook(0.1)}
+SESSION.headers['User-Agent'] = 'Mozilla/5.0 (X11; Linux x86_64)' \
+                                ' AppleWebKit/537.36 (KHTML, like Gecko)' \
+                                ' Chrome/80.0.3987.163 Safari/537.36'

### 2020-05-07 settings.py

@@ -1,0 +1,8 @@
+CACHE_FOLDER = '.cache'
+CSV_FOLDER = 'csv'
+LAST_DAY_FILE = 'last_day'
+START_DATE = {
+  'year': 2018,
+  'month': 10,
+  'day': 25,
+}

### 2020-05-10 chapter1.Rmd
edited scratchpad chapter1 and made more organic project.Rmd

@@
 seis.present %s+% " present values, " %s+% seis.missing %s+% " missing points (" %s+% print_dec(seis.present / seis.total * 100, 2) %s+% "% / " %s+% print_dec(seis.missing / seis.total * 100, 2) %s+% "%)"
-
 ```
@@
 # %>% complete(utc_hour=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = as.difftime("1:0:0")))
+```
+
+```{r}
+head(seis.h)
+tides <- read.csv("dailymax.csv")
+days <-seq(from=as.Date("1983-1-01"), length.out = nrow(tides), by="day")
+ts.tides <- xts(tides$Level, order.by = days, frequency = 365)
+#xts objects have frequency set to daily by default, so we change it
+attr(ts.tides, 'frequency') <- 365
+plot(stl(as.ts(ts.tides), s.window = "periodic"))
 ```
@@
 
-fitted <- fitma.list[order(aic_vals$AIC)][[2]]
+fitted <- fitma.list[order(aic_vals$AIC)][[1]]
+fitted
 tsdisplay(residuals(fitted))
@@
 tsdisplay(residuals(res.fitted))
-
 ```
@@
 library(spec)
-#pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2018-11-01T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
-pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2019-05-03T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
+pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2018-11-01T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
+#pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2019-05-03T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
 
+library(spec)
+pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2018-11-01T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
 topfreq <- pgram$freq[order(pgram$spec, decreasing = T)][1:100]
@@
 
-top95pc <- pgram$spec[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*1)]
-top95freq <- pgram$freq[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*1)][order(top95pc, decreasing = T)]
-top95pc <- pgram$spec[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*1)][order(top95pc, decreasing = T)]
+top95pc <- pgram$spec[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*2)]
+top95freq <- pgram$freq[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*2)][order(top95pc, decreasing = T)]
+top95pc <- pgram$spec[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*2)][order(top95pc, decreasing = T)]
 top95seas.s <- 1/top95freq
+top95seas.s
 top95seas.h <- top95seas.s/3600
@@
 axis(side=3, at=top95freq, labels=as.integer(top95seas.h), padj = 1, las=2, cex.axis=0.8)
-abline(h=mean(pgram$spec) + sd(pgram$spec)*c(.5,1:3), lty="dotted", col=c("orange", "darkgreen", "blue", "red"))
+abline(h=mean(pgram$spec) + sd(pgram$spec)*c(.3,1:3), lty="dotted", col=c("orange", "darkgreen", "blue", "red"))
 for (i in 1:length(top95freq)) {
@@
 }
-axis(side=4, at=mean(pgram$spec) + sd(pgram$spec) * c(.5, 1:3), labels=c(".5", "1", "2", "3"), cex.axis=0.8)
+axis(side=4, at=mean(pgram$spec) + sd(pgram$spec) * c(.3, 1:3), labels=c(".3", "1", "2", "3"), cex.axis=0.8)
 mtext(side=4, "spectrum sd ratio", line=1)

### 2020-05-10 project.Rmd
edited scratchpad chapter1 and made more organic project.Rmd

```{r libraries, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(stringi) # %s+%
library(lubridate) # floor_date
library(xts) # zoo
library(fpp) # auto.arima
library(tidyverse) # map_df, read_csv, group_by, summarize
library(tsbox) # https://cran.r-project.org/web/packages/tsbox/vignettes/tsbox.html
print_dec <- function(x, decimals=2) trimws(format(round(x, decimals), nsmall=decimals))

setwd("/media/data/Documents/Uni/spatiotemp/villalobos/project/")
```

```{r import dataset, include=FALSE}
# UCCS
domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"
# MEMS
# domain <- c(-7200,7200); pattern <- "MEM.*.csv"; name <- "MEM"; description <- "Membach"
seis <-
  list.files(path="csv/", pattern=pattern) %>%
  map_df(~read_csv("csv/" %s+% .))
```

# Data preparation
```{r dataset description, echo=FALSE}
long_description <- description %s+% " (" %s+% name %s+% " station)"
long_description
"Domain values: " %s+% domain[1] %s+% " to " %s+% domain[2] %s+% " nm"
"Data from " %s+% format.POSIXct(min(seis$utc_second), "%Y-%m-%d %H:%M:%S") %s+% " to " %s+% format.POSIXct(max(seis$utc_second), "%Y-%m-%d %H:%M:%S")
```

```{r cleanup, echo=FALSE}
out_domain <- (seis$min < domain[1] | seis$max > domain[2])
invalid_values <- (seis$min == -1 & seis$max == 1) | (seis$min == 0 & seis$max == 0) | out_domain
"Removing " %s+% sum(out_domain) %s+% " values out of domain, " %s+% (sum(invalid_values) - sum(out_domain)) %s+% " explicit NA"
seis <-
  seis %>%
  mutate(min=replace(min, invalid_values, NA), max=replace(max, invalid_values, NA)) %>%
  drop_na()
# %>% mutate(belgian_second=with_tz(utc_second, "Europe/Brussels"))
remove(out_domain, invalid_values) # free ~350 Mb from RAM
```

```{r missing values analysis}
seis.complete <- seis %>% complete(utc_second=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = as.difftime("0:0:1")))
seis.first.second <- min(seis$utc_second)
seis.missing.ids <- which(is.na(seis.complete$min))

seis.present <- length(seis$utc_second)
seis.total <- length(seis.complete$utc_second)
seis.missing <- seis.total - seis.present
seis.present %s+% " present values, " %s+% seis.missing %s+% " missing points (" %s+% print_dec(seis.present / seis.total * 100, 2) %s+% "% / " %s+% print_dec(seis.missing / seis.total * 100, 2) %s+% "%)"

block_start <- NULL
block_end <- NULL
blocks <- data.frame(start=integer(), end=integer())
for (i in 1:length(seis.missing.ids)) {
  if (is.null(block_start)) {
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  } else if ((block_end + 1) == seis.missing.ids[i]) {
    block_end <- seis.missing.ids[i]
  } else {
    blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  }
}
blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
blocks$interval <- seconds(blocks$end - blocks$start + 1)
blocks$start <- seis.first.second + seconds(blocks$start)
blocks$end <- seis.first.second + seconds(blocks$end)
plot(blocks$start, blocks$interval,
     col=ifelse(blocks$interval<3600, "darkgreen", "darkred"),
     main="Consecutive seconds of missing values",
     ylab="Seconds", xlab="Interval start",
     xaxt="n", xlim=c(min(seis$utc_second), max(seis$utc_second))
     )
axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-%y", cex=0.8)
abline(h=3600, col="green", lty="dotted")
legend("topleft", legend=c("interval >= 1h", "interval < 1h", "1h"), col=c("darkred", "darkgreen", "green"), lty=c(NA,NA,3), pch=c(1,1,NA),cex=0.9, bty="n")
text(x=blocks$start[which(blocks$interval > 3600)], y=blocks$interval[which(blocks$interval > 3600)], labels = paste0(blocks$start[which(blocks$interval > 3600)], " - ", blocks$end[which(blocks$interval > 3600)]), cex=0.9, pos=4)
remove(seis.complete, seis.missing.ids)
```

```{r gc_post_missing_analysis, include=FALSE}
gc()
```

```{r group by hour, include=FALSE}
bruxelles <- function(d) with_tz(d, "Europe/Brussels")
force_bruxelles <- function(d) force_tz(d, "Europe/Brussels")

seis.h <- 
  seis %>%
  # mutate(utc_hour=as_datetime(date(utc_second)) + hour(utc_second) * 3600) %>% # this forces the hour to represent bruxelles local time in utc. tz-ed time is normally evaluated in utc by ts-related functions
  group_by(utc_hour=floor_date(force_tz(bruxelles(utc_second), "UTC"), "1 hour")) %>%
  summarize(mean_movement_s=mean(max-min)) # %>%
#  complete(utc_hour=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = as.difftime("1:0:0")))

seis.before <- 
  seis.h[which(is.na(force_bruxelles(seis.h$utc_hour) + hours(1))), ] %>%
  mutate(utc_hour=utc_hour + hours(1))

seis.after <- 
  seis.h[which(is.na(force_bruxelles(seis.h$utc_hour) - hours(1))), ] %>%
  mutate(utc_hour=utc_hour - hours(1))

seis.dst.fix <-
  merge(seis.before, seis.after, all=T) %>%
  group_by(utc_hour=utc_hour) %>%
  summarize(mean_movement_s=mean(mean_movement_s))

seis.h <- merge(seis.h, seis.dst.fix, all=T)
remove(seis.after, seis.before, seis.dst.fix)
gc()
```

```{r missing values hour analysis}
seis.complete <- seis.h %>% complete(utc_hour=seq.POSIXt(from=min(seis.h$utc_hour), to=max(seis.h$utc_hour), by = as.difftime("1:0:0")))
seis.first.second <- min(seis$utc_second)
seis.missing.ids <- which(is.na(seis.complete$mean_movement_s))

seis.present <- length(seis.h$utc_hour)
seis.total <- length(seis.complete$utc_hour)
seis.missing <- seis.total - seis.present
seis.present %s+% " present values, " %s+% seis.missing %s+% " missing points (" %s+% print_dec(seis.present / seis.total * 100, 2) %s+% "% / " %s+% print_dec(seis.missing / seis.total * 100, 2) %s+% "%)"

block_start <- NULL
block_end <- NULL
blocks <- data.frame(start=integer(), end=integer())
for (i in 1:length(seis.missing.ids)) {
  if (is.null(block_start)) {
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  } else if ((block_end + 1) == seis.missing.ids[i]) {
    block_end <- seis.missing.ids[i]
  } else {
    blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  }
}
blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
blocks$interval <- hours(blocks$end - blocks$start + 1)
blocks$start <- seis.first.second + hours(blocks$start)
blocks$end <- seis.first.second + hours(blocks$end)
plot(c(blocks$start, blocks$end), rep(hour(blocks$interval), 2),
     col=ifelse(blocks$interval<3600, "darkgreen", "darkred"),
     main="Consecutive hours of missing values",
     ylab="Hours", xlab="Interval start",
     type="l", lwd=10,
     xaxt="n", xlim=c(min(seis$utc_second), max(seis$utc_second))
     )
axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-%y", cex=0.8)
text(x=blocks$start, y=hour(blocks$interval), labels = paste0(format(blocks$start, "%y-%m-%d %H:%M")), cex=0.9, pos=2)
text(x=blocks$end, y=hour(blocks$interval), labels = paste0(format(blocks$end, "%y-%m-%d %H:%M")), cex=0.9, pos=4)
text(x=blocks$start + blocks$interval/2, y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=3)
remove(seis.complete, seis.missing.ids)

```

```{r gc_post_missing_analysis_bru, include=FALSE}
gc()
```

```{r build_ts function, include=FALSE}
build_ts <- function(dset, periods=NULL, start=NULL, end=NULL, force_ms=T) {
  dset.ts <- ts_zoo(dset)
  if (!is.null(start) | !is.null(end)) {
    dset.ts <- ts_span(dset.ts, start=start, end=end)
  }
  dset.ts <- ts_regular(dset.ts)
  if (is.null(periods)) {
    return(dset.ts)
  }
  if (is.vector(periods) & length(periods) == 1) {
    periods <- periods[1]
  }
  if (is.integer(periods)){
    if (force_ms) {
      periods <- c(periods)
    } else {
      frequency(dset.ts) <- periods
      return(dset.ts)
    }
  }
  return(msts(dset.ts, periods))
}
```

```{r}
seis.h.start <- as_datetime('2018-10-29T00:00:00')
seis.h.end <- as_datetime('2020-03-01T00:00:00')
seasonality.period <- 168
seis.h.ts <- build_ts(seis.h, periods = c(seasonality.period), start = seis.h.start, end = seis.h.end, force_ms = T)
autoplot(seis.h.ts, main = "Hourly seismic movement", xlab = "Weeks", ylab = "mean movement (nm)")
hist(seis.h.ts)
autoplot(acf(seis.h.ts, na.action=na.pass, plot=F))
autoplot(pacf(seis.h.ts, na.action=na.pass, plot=F))

```
Once notice the composition is additive (the fluctuations does not seem to grow with the trend)
```{r}
dec.addi <- decompose(na.aggregate(seis.h.ts, floor))
autoplot(dec.addi)
dec.multi <- decompose(log(na.aggregate(seis.h.ts, floor)))
autoplot(dec.multi)
autoplot(acf(dec.addi$random, na.action=na.omit, plot=F))
autoplot(acf(dec.multi$random, na.action=na.omit, plot=F))
autoplot(pacf(dec.addi$random, na.action=na.omit, plot=F))
autoplot(pacf(dec.multi$random, na.action=na.omit, plot=F))
```

We now extract the trend using a linear filter with smoothing parameter $p$ equal to 180 (almost half a year):

```{r}
p <- floor(seasonality.period/2)
weights <- rep(1/(2*p+1), times=2*p+1)
trend <- stats::filter(seis.h.ts, sides=2, filter = weights)
```

```{r show_overlap_and_diff, include=FALSE}
show_overlap_and_diff <- function(base, compare, base.desc, compare.desc, xlab, overlap.ylab, diff.title, diff.ylab, diff.ylab.sec, diff.ylab.sec.accuracy=.1) {
  overlap_plot <-
    autoplot(compare, xlab=xlab, ylab=overlap.ylab) +
      geom_line(aes(y=base, col="red"), show.legend = F)
  span <- max(compare, base, na.rm = T) - min(compare, base, na.rm = T)
  diff <- compare-base
  diff_plot <-
    autoplot(diff, ylab = paste0("Difference (", diff.ylab, ")"), xlab = xlab) +
      labs(title = diff.title, subtitle = paste0(base.desc, " - ", compare.desc)) +
      scale_y_continuous(sec.axis = sec_axis(~./span, name=paste0("Difference (", diff.ylab.sec, ")"), labels = scales::percent_format(accuracy = diff.ylab.sec.accuracy))) +
      geom_hline(yintercept=mean(diff, na.rm=T), linetype="dashed", color = "red")
  return(list(overlap_plot, diff_plot))
}
```

And they match:
```{r}
plots <- show_overlap_and_diff(
  dec.addi$trend, trend, xlab="Weeks",
  base.desc = "additive decomposition",
  compare.desc = paste0("linear filtered trend with p=", p),
  overlap.ylab = "Trend (nm)",
  diff.title = "Difference between trends",
  diff.ylab = "nm", diff.ylab.sec="% over trend span"
)
plots[[1]]
plots[[2]]

```

Then we use a filter able also to capture some seasonal effect so we will be able to find the residuals. We are going to choose $f=365$ so to have the moving average from the first day of the year to the last one and then shifted from the second day of the year to the first one of the following year and so, this is because we believe to have an yearly seasonality.

```{r}
f <- seasonality.period
weights.s <- c(0.5, rep(1,f-1), 0.5)/f
trend.s <- stats::filter(seis.h.ts, side=2, filter=weights.s)

plots <- show_overlap_and_diff(
  dec.addi$trend, trend.s, xlab="Weeks",
  base.desc = "additive decomposition",
  compare.desc = paste0("shift-filtered trend with f=", f),
  overlap.ylab = "Trend (nm)",
  diff.title = "Difference between trends",
  diff.ylab = "nm", diff.ylab.sec="% over trend span",
  diff.ylab.sec.accuracy = .001
)
plots[[1]]
plots[[2]]

```

and they match (with some error of course) with the ones found with the decompose function.

Finally we want to prove we can retrieve seasonality, in order to do so we can compute the average seasonality starting from the detrended series: to do so we put the time series into a matrix and we transform this matrix so each column contains elements of the same period and finally, we compute the mean of each column. We will check this with the element figure of the object decomposition i.e. the estimated seasonal figure not repeated for all the time series.

```{r}
detrended <- na.aggregate(seis.h.ts, floor)-trend.s
detrended <- window(detrended, end = c(end(detrended)[1], seasonality.period), extend=T)
matrix <- t(matrix(data = detrended, nrow = seasonality.period))
seasonality.figure <- as.ts(colMeans(matrix, na.rm = T))
seasonality <- rep(seasonality.figure, length(detrended)/length(seasonality.figure))[1:length(trend.s)]
```

and they almost match:

```{r}

plots <- show_overlap_and_diff(
  dec.addi$figure, seasonality.figure, xlab="Hours",
  base.desc = "additive decomposition",
  compare.desc = paste0("mean-over-period after removal of shift-filtered trend with f=", f),
  overlap.ylab = "Seasonality (nm)",
  diff.title = paste0("Difference between seasonalities (f=", seasonality.period, ")"),
  diff.ylab = "nm", diff.ylab.sec="% over seasonality span"
)
plots[[1]]
plots[[2]]

# difference is maybe given by na.aggregate used in decompose, while na.rm=T in manual seasonality?
```

```{r}
residuals <- seis.h.ts - trend.s - seasonality

plots <- show_overlap_and_diff(
  dec.addi$random, residuals, xlab="Weeks",
  base.desc = "additive decomposition",
  compare.desc = paste0("mean-over-period after removal of shift-filtered trend with f=", f),
  overlap.ylab = "Residuals (nm)",
  diff.title = paste0("Difference between residuals (f=", seasonality.period, ")"),
  diff.ylab = "nm", diff.ylab.sec="% over residuals span"
)
plots[[1]]
plots[[2]]

```

In fact we are able to reconstruct the original series, with of course some missing value at the beginning and at the end because of the methods applied previously:

```{r}
ts <- residuals+trend+seasonality

plots <- show_overlap_and_diff(
  seis.h.ts, ts, xlab="Weeks",
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span"
)
plots[[1]]
plots[[2]]

```

```{r}
acf(residuals, na.action = na.omit)
pacf(residuals, na.action = na.omit)
seis.h.ts.ar <- Arima(seis.h.ts, c(1,0,0))
tsdiag(seis.h.ts.ar)
```

```{r}
seis.h.start <- with_tz(as_datetime('2018-11-01T00:00:00'), "Europe/Brussels")
seis.h.end <- with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels")
seis.h.msts <- na.aggregate(build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end, force_ms = T))
seis.h.msts %>% mstl() %>% autoplot() + xlab("Week")
seis.h.msts.mstl <- seis.h.msts %>% mstl()
plot(seis.h.msts.mstl[,2], main="trend")
lines(trend.s, col="red")
plot(rowSums(seis.h.msts.mstl[,3:4]), main="seasonality",type="l")
lines(c(rep(seasonality,seasons), seasonality[1:(((length(detrended)/168-seasons)*168))]), col="red")
plot(residuals,  main="residuals")
lines(seis.h.msts.mstl[,5], col="red")
autoplot(seis.h.msts.mstl[,5])
acf(seis.h.msts.mstl[,5], na.action = na.omit)
bptests <- c()
for (i in 1:10)
  bptests<- c(bptests, Box.test(seis.h.msts.mstl[,5], lag = i, fitdf = 0, type = "Box-Pierce")$p.value)
plot(bptests, ylim = c(0,1))
abline(h=0.05,col="blue",lty="dotted")

```

### 2020-05-11 ProjectCode.Rmd
Nice working analysis

---
title: "Project Code"
author: "Diletta Olliaro"
date: "5/6/2020"
output: pdf_document
---

```{r}
library(zoo)
library(xts)
tides <- read.csv("dailymax.csv")
days <-seq(from=as.Date("1983-1-01"), length.out = nrow(tides), by="day")
ts.tides <- xts(tides$Level, order.by = days, frequency = 365)
#xts objects have frequency set to daily by default, so we change it
attr(ts.tides, 'frequency') <- 365
```
Once notice the composition is additive (the fluctuations does not seem to grow with the trend)
```{r}
decomposition <- decompose(as.ts(ts.tides), "additive")
plot(decomposition)
```

We now extract the trend using a linear filter with smoothing parameter $p$ equal to 180 (almost half a year):

```{r}
p <- 180
weights <- rep(1/(2*p+1), times=2*p+1)
trend <- filter(ts.tides, sides=2, filter = weights)
```

And they match:
```{r}
plot(trend)
lines(decomposition$trend, col="red")
```

Then we use a filter able also to capture some seasonal effect so we will be able to find the residuals. We are going to choose $f=365$ so to have the moving average from the first day of the year to the last one and then shifted from the second day of the year to the first one of the following year and so, this is because we believe to have an yearly seasonality.

```{r}
f <- 365
weights.s <- c(0.5, rep(1,f-1), 0.5)/f
trend.s <- filter(ts.tides, side=2, filter=weights.s)
residuals <- as.ts(ts.tides)-trend.s
```

and they match (with some error of course) with the ones found with the decompose function.

```{r}
plot(residuals)
lines(decomposition$random, col="red")
```
Finally we want to prove we can retrieve seasonality, in order to do so we can compute the average seasonality starting from the detrended series: to do so we put the time series into a matrix and we transform this matrix so each column contains elements of the same period and finally, we compute the mean of each column. We will check this with the element figure of the object decomposition i.e. the estimated seasonal figure not repeated for all the time series.

```{r}
detrended <- as.ts(ts.tides)-trend
#t takes the transpose
matrix <- t(matrix(data = detrended, nrow = 365))
seasonality = as.ts(colMeans(matrix, na.rm = T))
```

and they almost match:

```{r}
plot(seasonality)
lines(decomposition$figure, col="red")
```

In fact we are able to reconstruct the original series, with of course some missing value at the beginning and at the end because of the methods applied previously:

```{r}
ts <- residuals+trend+rep(seasonality,36)
plot(as.ts(ts.tides))
lines(ts, col="blue")
```

```{r}
matrix2 <- t(matrix(data = decomposition$random, nrow = 14))
seasonality2 = as.ts(colMeans(matrix2, na.rm = T))
plot(seasonality2)
```

```{r}
g <- residuals-rep(seasonality2, 469)
plot(g)
acf(na.omit(g))
```

### 2020-05-11 project2.Rmd
Nice working analysis

--- 2020-05-10 project.Rmd
@@
 library(fpp) # auto.arima
+library(gridExtra) # for gg-plotting multiple plots
 library(tidyverse) # map_df, read_csv, group_by, summarize
@@
 weights <- rep(1/(2*p+1), times=2*p+1)
-trend <- stats::filter(seis.h.ts, sides=2, filter = weights)
+trend <- stats::filter(na.aggregate(seis.h.ts, floor), sides=2, filter = weights)
 ```
@@
   diff <- compare-base
+  accu.lims <- diff.ylab.sec.accuracy / 100 * span * c(-3,3)
+
   diff_plot <-
     autoplot(diff, ylab = paste0("Difference (", diff.ylab, ")"), xlab = xlab) +
+      expand_limits(y=accu.lims) +
       labs(title = diff.title, subtitle = paste0(base.desc, " - ", compare.desc)) +
@@
 weights.s <- c(0.5, rep(1,f-1), 0.5)/f
-trend.s <- stats::filter(seis.h.ts, side=2, filter=weights.s)
+trend.s <- stats::filter(na.aggregate(seis.h.ts, floor), side=2, filter=weights.s)
 
@@
   diff.ylab = "nm", diff.ylab.sec="% over trend span",
-  diff.ylab.sec.accuracy = .001
+  diff.ylab.sec.accuracy = .01
 )
@@
   diff.title = paste0("Difference between seasonalities (f=", seasonality.period, ")"),
-  diff.ylab = "nm", diff.ylab.sec="% over seasonality span"
+  diff.ylab = "nm", diff.ylab.sec="% over seasonality span",
+  diff.ylab.sec.accuracy = .01
 )
@@
 ```{r}
-residuals <- seis.h.ts - trend.s - seasonality
+residuals <- na.aggregate(seis.h.ts, floor) - trend.s - seasonality
 
@@
   diff.title = paste0("Difference between residuals (f=", seasonality.period, ")"),
-  diff.ylab = "nm", diff.ylab.sec="% over residuals span"
+  diff.ylab = "nm", diff.ylab.sec="% over residuals span",
+  diff.ylab.sec.accuracy = .01
 )
@@
 ```{r}
-ts <- residuals+trend+seasonality
+ts <- residuals+trend.s+seasonality
 
@@
   diff.title = "Difference between ts",
-  diff.ylab = "nm", diff.ylab.sec="% over ts span"
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plots[[1]]
+plots[[2]]
+
+plots <- show_overlap_and_diff(
+  seis.h.ts, dec.addi$trend+dec.addi$seasonal+dec.addi$random, xlab="Weeks",
+  base.desc = "original series",
+  compare.desc = "reconstructed series",
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
 )
@@
 
-```{r}
-acf(residuals, na.action = na.omit)
-pacf(residuals, na.action = na.omit)
+```{r residuals cf arima, fig.height=960, fig.width=960, fig.path="/tmp/"}
+residuals.acf <- autoplot(acf(residuals, na.action = na.pass, plot=F))
+residuals.pacf <- autoplot(pacf(residuals, na.action = na.pass, plot=F))
+plot_grid(residuals.acf, residuals.pacf, labels = "AUTO", nrow=2)
+
 seis.h.ts.ar <- Arima(seis.h.ts, c(1,0,0))
 tsdiag(seis.h.ts.ar)
+library(cowplot)
+install.packages("cowplot")
 ```
@@
 plot(rowSums(seis.h.msts.mstl[,3:4]), main="seasonality",type="l")
-lines(c(rep(seasonality,seasons), seasonality[1:(((length(detrended)/168-seasons)*168))]), col="red")
+lines(seasonality, col="red")
 plot(residuals,  main="residuals")

### 2020-05-12 project.Rmd
more analysis

@@
 ```{r libraries, include=FALSE}
-knitr::opts_chunk$set(warning = F, message = F)
+knitr::opts_chunk$set(warning = F, message = F, fig.height = 360)
 library(stringi) # %s+%
@@
 library(fpp) # auto.arima
+# library(gridExtra) # for gg-plotting multiple plots
+library(cowplot) # for plot_grid
 library(tidyverse) # map_df, read_csv, group_by, summarize
@@
 ```{r import dataset, include=FALSE}
+# MEMS
+domain <- c(-7200,7200); pattern <- "MEM.*.csv"; name <- "MEM"; description <- "Membach"
 # UCCS
 domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"
-# MEMS
-# domain <- c(-7200,7200); pattern <- "MEM.*.csv"; name <- "MEM"; description <- "Membach"
 seis <-
@@
 
+Box.test(forec$model$residuals, type="Box-Pierce")
+
 ```
@@
 
-```{r}
+```{r plot data with (p)acf}
 seis.h.start <- as_datetime('2018-10-29T00:00:00')
 seis.h.end <- as_datetime('2020-03-01T00:00:00')
+seasonality.period <- 24
+seasonality.xlab <- "Days"
+seasonality.aggregate <- function(ds) na.aggregate(seis.h.ts, function(x) floor(x/7))
 seasonality.period <- 168
+seasonality.xlab <- "Weeks"
+seasonality.aggregate <- function(ds) na.aggregate(ds, floor)
 seis.h.ts <- build_ts(seis.h, periods = c(seasonality.period), start = seis.h.start, end = seis.h.end, force_ms = T)
-autoplot(seis.h.ts, main = "Hourly seismic movement", xlab = "Weeks", ylab = "mean movement (nm)")
-hist(seis.h.ts)
-autoplot(acf(seis.h.ts, na.action=na.pass, plot=F))
-autoplot(pacf(seis.h.ts, na.action=na.pass, plot=F))
+plot_grid(
+  autoplot(seis.h.ts, main = "Hourly seismic movement", xlab = seasonality.xlab, ylab = "mean movement (nm)"),
+  gghistogram(seis.h.ts) + xlab("mean movement (nm)"),
+  nrow = 2)
+plot_grid(
+  autoplot(acf(seis.h.ts, na.action=na.pass, plot=F, lag.max=seasonality.period)) + scale_x_continuous(),
+  autoplot(pacf(seis.h.ts, na.action=na.pass, plot=F, lag.max=seasonality.period)) + scale_x_continuous(),
+  nrow = 2)
 
@@
 Once notice the composition is additive (the fluctuations does not seem to grow with the trend)
-```{r}
-dec.addi <- decompose(na.aggregate(seis.h.ts, floor))
-autoplot(dec.addi)
-dec.multi <- decompose(log(na.aggregate(seis.h.ts, floor)))
-autoplot(dec.multi)
-autoplot(acf(dec.addi$random, na.action=na.omit, plot=F))
-autoplot(acf(dec.multi$random, na.action=na.omit, plot=F))
-autoplot(pacf(dec.addi$random, na.action=na.omit, plot=F))
-autoplot(pacf(dec.multi$random, na.action=na.omit, plot=F))
+```{r decompose}
+dec.addi <- decompose(seasonality.aggregate(seis.h.ts))
+autoplot(dec.addi) + labs(title="Decomposition of additive time series", subtitle = paste0("Seasonality ", seasonality.period))
+dec.multi <- decompose(log(seasonality.aggregate(seis.h.ts)))
+autoplot(dec.multi) + labs(title="Decomposition of multiplicative time series (log-scale additive)", subtitle = paste0("Seasonality ", seasonality.period))
+plot_grid(
+  autoplot(acf(dec.addi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
+  autoplot(acf(dec.multi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
+  nrow = 2)
+plot_grid(
+  autoplot(pacf(dec.addi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
+  autoplot(pacf(dec.multi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
+  nrow = 2)
 ```
@@
 
-```{r}
+```{r linear trend}
 p <- floor(seasonality.period/2)
 weights <- rep(1/(2*p+1), times=2*p+1)
-trend <- stats::filter(seis.h.ts, sides=2, filter = weights)
+trend <- stats::filter(seasonality.aggregate(seis.h.ts), sides=2, filter = weights)
 ```
@@
     autoplot(compare, xlab=xlab, ylab=overlap.ylab) +
-      geom_line(aes(y=base, col="red"), show.legend = F)
+      geom_line(aes(y=base, col="red"), show.legend = F) +
+      labs(title=paste0(base.desc, " vs ", compare.desc))
   span <- max(compare, base, na.rm = T) - min(compare, base, na.rm = T)
   diff <- compare-base
+  accu.lims <- diff.ylab.sec.accuracy / 100 * span * c(-3,3)
+
+  mse <- mean(diff^2)
+  mae <- mean(abs(diff))
+  rmse <- sqrt(mse)
+  R2 <- 1-(sum((diff)^2)/sum((base-mean(base))^2))
+  errorstring <- paste0("MSE=", mse, "    MAE=", mae, "    RMSE=", rmse, "    R2=", R2)
+  print(errorstring)
+
   diff_plot <-
     autoplot(diff, ylab = paste0("Difference (", diff.ylab, ")"), xlab = xlab) +
+      expand_limits(y=accu.lims) +
       labs(title = diff.title, subtitle = paste0(base.desc, " - ", compare.desc)) +
@@
 And they match:
-```{r}
+```{r linear trend plot}
 plots <- show_overlap_and_diff(
-  dec.addi$trend, trend, xlab="Weeks",
-  base.desc = "additive decomposition",
+  dec.addi$trend, trend, xlab=seasonality.xlab,
+  base.desc = "additive decomposition trend",
   compare.desc = paste0("linear filtered trend with p=", p),
@@
 )
-plots[[1]]
-plots[[2]]
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
 
@@
 
-```{r}
+```{r trend MA}
 f <- seasonality.period
 weights.s <- c(0.5, rep(1,f-1), 0.5)/f
-trend.s <- stats::filter(seis.h.ts, side=2, filter=weights.s)
+trend.s <- stats::filter(seasonality.aggregate(seis.h.ts), side=2, filter=weights.s)
 
 plots <- show_overlap_and_diff(
-  dec.addi$trend, trend.s, xlab="Weeks",
-  base.desc = "additive decomposition",
+  dec.addi$trend, trend.s, xlab=seasonality.xlab,
+  base.desc = "additive decomposition trend",
   compare.desc = paste0("shift-filtered trend with f=", f),
@@
   diff.ylab = "nm", diff.ylab.sec="% over trend span",
-  diff.ylab.sec.accuracy = .001
+  diff.ylab.sec.accuracy = .01
 )
-plots[[1]]
-plots[[2]]
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
 
@@
 
-```{r}
-detrended <- na.aggregate(seis.h.ts, floor)-trend.s
+```{r trend figure}
+detrended <- seasonality.aggregate(seis.h.ts)-trend.s
 detrended <- window(detrended, end = c(end(detrended)[1], seasonality.period), extend=T)
@@
 
-```{r}
+```{r difference of trend figure}
 
 plots <- show_overlap_and_diff(
-  dec.addi$figure, seasonality.figure, xlab="Hours",
+  dec.addi$figure, seasonality.figure, xlab=seasonality.xlab,
   base.desc = "additive decomposition",
@@
   diff.title = paste0("Difference between seasonalities (f=", seasonality.period, ")"),
-  diff.ylab = "nm", diff.ylab.sec="% over seasonality span"
+  diff.ylab = "nm", diff.ylab.sec="% over seasonality span",
+  diff.ylab.sec.accuracy = .01
 )
-plots[[1]]
-plots[[2]]
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
 
@@
 
-```{r}
-residuals <- seis.h.ts - trend.s - seasonality
+```{r residuals}
+residuals <- seasonality.aggregate(seis.h.ts) - trend.s - seasonality
 
 plots <- show_overlap_and_diff(
-  dec.addi$random, residuals, xlab="Weeks",
+  dec.addi$random, residuals, xlab=seasonality.xlab,
   base.desc = "additive decomposition",
@@
   diff.title = paste0("Difference between residuals (f=", seasonality.period, ")"),
-  diff.ylab = "nm", diff.ylab.sec="% over residuals span"
+  diff.ylab = "nm", diff.ylab.sec="% over residuals span",
+  diff.ylab.sec.accuracy = .01
 )
-plots[[1]]
-plots[[2]]
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+
+residuals.acf <- autoplot(acf(residuals, na.action = na.pass, plot=F)) + scale_x_continuous()
+residuals.pacf <- autoplot(pacf(residuals, na.action = na.pass, plot=F)) + scale_x_continuous()
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
 
@@
 
-```{r}
-ts <- residuals+trend+seasonality
+```{r difference of reconstruct}
+ts <- residuals+trend.s+seasonality
 
 plots <- show_overlap_and_diff(
-  seis.h.ts, ts, xlab="Weeks",
+  seis.h.ts, ts, xlab=seasonality.xlab,
   base.desc = "original series",
@@
   diff.title = "Difference between ts",
-  diff.ylab = "nm", diff.ylab.sec="% over ts span"
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
 )
-plots[[1]]
-plots[[2]]
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+
+plots <- show_overlap_and_diff(
+  seis.h.ts, dec.addi$trend+dec.addi$seasonal+dec.addi$random, xlab=seasonality.xlab,
+  base.desc = "original series",
+  compare.desc = "reconstructed series",
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+
+```
+
+```{r extract both 24 and 168}
+trend.s_compute <- function(ts.in, period) {
+  f <- period
+  weights.s <- c(0.5, rep(1,f-1), 0.5)/f
+  trend.s <- stats::filter(ts.in, side=2, filter=weights.s)
+  return(trend.s)
+}
+
+season_by_mean <- function(ts.in, period) {
+  freq <- frequency(ts.in)
+  if (freq == period) {
+    end <- c(end(ts.in)[1], period)
+  } else {
+    end <- c(((end(ts.in)[1] * freq) %/% period + 1) * (period/freq), freq)
+  }
+  extended <- window(ts.in, end = end, extend=T)
+  matrix <- t(matrix(data = extended, nrow = period))
+  seasonality.figure <- as.ts(colMeans(matrix, na.rm = T))
+  seasonality <- rep(seasonality.figure, length(extended)/length(seasonality.figure))[1:length(ts)]
+  return(list(figure=seasonality.figure, seasonal=seasonality))
+}
+
+trend_season <- function(ts.in, period) {
+  trend.s <- trend.s_compute(ts.in, period)
+  season <- season_by_mean(ts.in-trend.s, period)
+  season$trend <- trend.s
+  return(season)
+}
+
+aggregated <- seasonality.aggregate(seis.h.ts)
+season.24 <- trend_season(aggregated, 24)
+season.168 <- trend_season(aggregated-season.24$seasonal, 168)
+
+plot_grid(
+  autoplot(aggregated) + ylab("") + xlab(""),
+  autoplot(ts(season.24$seasonal, frequency=24)) + ylab("") + xlab(""),
+  autoplot(ts(season.168$seasonal, frequency=24)) + ylab("") + xlab(""),
+  autoplot(ts(season.168$trend, frequency=24)) + ylab("") + xlab(""),
+  autoplot(aggregated-season.24$seasonal-season.168$seasonal-season.168$trend) + ylab("") + xlab(""),
+  nrow = 5)
+
+season.24 <- trend_season(aggregated-season.168$seasonal, 24)
+season.168 <- trend_season(aggregated-season.24$seasonal, 168)
+
+season.24 <- trend_season(aggregated-season.168$seasonal, 24)
+season.168 <- trend_season(aggregated-season.24$seasonal, 168)
+
+plot_grid(
+  autoplot(aggregated) + ylab("") + xlab(""),
+  autoplot(ts(season.24$seasonal, frequency=24)) + ylab("") + xlab(""),
+  autoplot(ts(season.168$seasonal, frequency=24)) + ylab("") + xlab(""),
+  autoplot(ts(season.168$trend, frequency=24)) + ylab("") + xlab(""),
+  autoplot(aggregated-season.24$seasonal-season.168$seasonal-season.168$trend) + ylab("") + xlab(""),
+  nrow = 5)
+
+seasonbase.24 <- trend_season(aggregated, 24)
+seasonbase.168 <- trend_season(aggregated, 168)
+
+plot_grid(
+  autoplot(ts(seasonbase.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
+      geom_line(aes(y=season.24$figure, col="red"), show.legend = F),
+  autoplot(ts(seasonbase.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
+      geom_line(aes(y=season.168$figure, col="red")) +
+      geom_line(aes(y=season.168$figure + rep(season.24$figure, 7), col="blue")),
+  nrow = 2)
+
+plots <- show_overlap_and_diff(
+  seasonbase.168$figure, season.168$figure + rep(season.24$figure, 7), xlab="Hours",
+  base.desc = "168 from ds",
+  compare.desc = "168+24 from ds",
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between 168",
+  diff.ylab = "nm", diff.ylab.sec="% over season span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+
+```
+
+```{r test_ts}
+seis.test.start <- as_datetime('2019-05-20T00:00:00')
+seis.test.end <- as_datetime('2020-03-02T00:00:00')
+
+seis.h.test <- build_ts(seis.h, periods = c(seasonality.period), start = seis.test.start, end = seis.test.end, force_ms = T)
+```
+
+```{r diffs acf/pacf}
+seis.h.test %>% ggtsdisplay(lag.max = 168)
+seis.h.test %>% diff() %>% ggtsdisplay(lag.max = 168*5)
+# seis.h.test %>% diff(D=2) %>% ggtsdisplay(lag.max = 168)
+# seis.h.test %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
+# seis.h.test %>% diff() %>% diff(lag=24) %>% ggtsdisplay(lag.max = 24, main="Supposed model: ARIMA(24,1,0)(0,1,2)[24]") # ARIMA(0,1,0)(0,1,2)[24]
+# previous => ARIMA(5,1,0)(0,1,7)[24]
+# seis.h.test %>% diff() %>% diff(lag=24, D=2) %>% ggtsdisplay(lag.max = 168)
+seis.h.test %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,0,0)(0,1,1/2)[168]
+seis.h.test %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,1,0)(0,1,1)[168]
+# seis.h.test %>% diff() %>% diff(lag=168, D=2) %>% ggtsdisplay(lag.max = 168)
+seis.h.test %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168)
+```
+
+```{r arima}
+
+plot_forecast <- function(forec) {
+  BP <- Box.test(forec$model$residuals, type="Box-Pierce")
+  autoplot(forec) +
+    labs(title=forec$method, subtitle = paste0("Box-Pierce p-value=", print_dec(BP$p.value, decimals=4))) +
+    autolayer(forec$mean, series="Forecast") -> plotted
+  return(plotted)
+}
+
+to_test <- list(
+  #c(0,0,0,0,0,1),  # BAD, Box-Pierce = 0
+  #c(0,0,0,0,0,2),  # BAD, Box-Pierce = 0
+  #c(0,0,0,0,1,1),  # nice, Box-Pierce = 0
+  c(0,0,0,0,1,2),
+  #c(0,0,0,0,1,3),  # CAN'T BE DONE: 3*168 = 504 > 350 (limit of arima on r)
+  #c(0,1,0,0,0,1),  # BAD, Box-Pierce = 0
+  c(0,1,0,0,1,1),
+  c(0,1,0,0,1,2),
+  c(2,0,0,0,1,0),  # hint from auto.arima D=1
+  c(5,0,0,0,1,0),  # hint from auto.arima D=1
+  c(3,1,1,0,1,0),  # hint from auto.arima D=1 d=1
+  c(3,1,1,0,1,1),  # after hint then ACF + PACF
+)
+if (!exists("models")){
+  models <- list()
+}
+for (model in to_test) {
+  model <- c(3,1,1,0,1,1)
+  fit.name <- paste0("fit", paste0(model, collapse = ""))
+  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
+  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.test) %s+% "]")
+  models[[fit.name]] <- Arima(seis.h.test, model[1:3], model[4:6])
+  models[[forecast.name]] <- forecast(models[[fit.name]], h=168*10)
+}
+models[["fitauto"]] <- auto.arima(seis.h.test, D=1, approx=F, trace = T)
+models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*10)
+models[["fit311010"]] <- models[["fitauto"]]
+models[["forecast311010"]] <- forecast(models[["fit311010"]], h=168*10)
+
+# 000012 better than 000011, Box-Pierce = 0
+plotting.model <- models[["fit000012"]]
+plotting.forecast <- models[["forecast000012"]]
+# 010011 somehow better somehow worst than 000012, Box-Pierce = 0
+plotting.model <- models[["fit010011"]]
+plotting.forecast <- models[["forecast010011"]]
+# 010012 , Box-Pierce = 
+plotting.model <- models[["fit010012"]]
+plotting.forecast <- models[["forecast010012"]]
+# 311010 , Box-Pierce = 
+plotting.model <- models[["fit311010"]]
+plotting.forecast <- models[["forecast311010"]]
+# 311011 , Box-Pierce = 
+plotting.model <- models[["fit311011"]]
+plotting.forecast <- models[["forecast311011"]]
+
+checkresiduals(plotting.model)
+residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+plot_forecast(plotting.forecast)
+plots <- show_overlap_and_diff(
+  seis.h.test, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "original series",
+  compare.desc = "reconstructed series",
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+
+plotting.model <- models[["fit010012"]]
+plotting.forecast <- models[["forecast010012"]]
+checkresiduals(plotting.model)
+residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+plot_forecast(plotting.forecast)
+plots <- show_overlap_and_diff(
+  seis.h.test, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "original series",
+  compare.desc = "reconstructed series",
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
 
@@
 ```{r}
-acf(residuals, na.action = na.omit)
-pacf(residuals, na.action = na.omit)
-seis.h.ts.ar <- Arima(seis.h.ts, c(1,0,0))
-tsdiag(seis.h.ts.ar)
+(seis.test.start - seis.first.second)
+
+deseasoned24 <- seis.h.ts - seasonality
+autoplot(deseasoned24)
+frequency(deseasoned24)
+de24.test <- msts(window(deseasoned24, start=205), seasonal.periods = c(168))
+frequency(de24.test)
+autoplot(de24.test)
+de24.test %>% ggtsdisplay(lag.max = 168*2)
+de24.test %>% diff() %>% ggtsdisplay(lag.max = 168*2)
+de24.test %>% diff(D=2) %>% ggtsdisplay(lag.max = 168*2)
+de24.test %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*2)
+de24.test %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*2, main="Supposed model: ARIMA(0,1,0)(0,1,1)[168]")
+de24.test %>% diff() %>% diff(lag=168, D=2) %>% ggtsdisplay(lag.max = 168*2)
+
+if (!exists("fit010011de")){
+  fit010011de <- Arima(de24.test, c(0,1,0), c(0,1,1))
+}
+checkresiduals(fit010011de)
+residuals.acf <- autoplot(acf(fit010011de$residuals, na.action = na.pass, plot=F, lag.max=168*3)) + scale_x_continuous()
+residuals.pacf <- autoplot(pacf(fit010011de$residuals, na.action = na.pass, plot=F, lag.max=168*3)) + scale_x_continuous()
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+if (!exists("fit010011de.forec")){
+  fit010011de.forec <- forecast(fit010011de, h=168*10)
+}
+plot_forecast(forecast(fit010011de.forec, h=168*10))
+
+# NO  fit011011de <- Arima(de24.test, c(0,1,1), c(0,1,1))
+if (!exists("fit011011de")){
+  fit011011de <- Arima(de24.test, c(0,1,1), c(0,1,1))
+  arima
+}
+fit010011de$aic
+fit011011de$aic
+checkresiduals(fit011011de)
+residuals.acf <- autoplot(acf(fit011011de$residuals, na.action = na.pass, plot=F, lag.max=168*3)) + scale_x_continuous()
+residuals.pacf <- autoplot(pacf(fit011011de$residuals, na.action = na.pass, plot=F, lag.max=168*3)) + scale_x_continuous()
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+if (!exists("fit011011de.forec")){
+  fit011011de.forec <- forecast(fit011011de, h=168*10)
+}
+plot_forecast(forecast(fit011011de.forec, h=168*10))
+
+```
+
+```{r}
+
+seis.test.start <- as_datetime('2019-05-20T00:00:00')
+seis.test.end <- as_datetime('2020-03-02T00:00:00')
+
+seis.h.test <- build_ts(seis.h, periods = c(seasonality.period), start = seis.test.start, end = seis.test.end, force_ms = T)
+
+# First seasonal difference (quarterly)
+# seasonality.arima <- 24
+D<-1
+seis.h.test.ds<-diff(seis.h.test,differences=D)
+plot(seis.h.test.ds)
+abline(h=mean(seis.h.test.ds), col="red", lty="dotted")
+sd(seis.h.test.ds)
+tsdisplay(seis.h.test.ds, lag.max = 168*2)
+autoplot(acf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
+autoplot(pacf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
+seis.h.test.ds<-diff(seis.h.test,lag=seasonality.period,differences=D)
+plot(seis.h.test.ds)
+abline(h=mean(seis.h.test.ds), col="red", lty="dotted")
+sd(seis.h.test.ds)
+autoplot(acf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
+autoplot(pacf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
+tsdisplay(seis.h.test.ds, lag.max = 168*2)
+# pacf cuts off after p=168
+```
+
+```{r}
+fit <- auto.arima(seis.h.test, d=1, D=1, trace = T, parallel = T)
+```
+ Fitting models using approximations to speed things up...
+
+ ARIMA(2,0,2)(1,1,1)[24] with drift         : Inf
+ ARIMA(0,0,0)(0,1,0)[24] with drift         : 111462.7
+ ARIMA(1,0,0)(1,1,0)[24] with drift         : 101954.4
+ ARIMA(0,0,1)(0,1,1)[24] with drift         : Inf
+ ARIMA(0,0,0)(0,1,0)[24]                    : 111460.9
+ ARIMA(1,0,0)(0,1,0)[24] with drift         : 102426.5
+ ARIMA(1,0,0)(2,1,0)[24] with drift         : 101379
+ ARIMA(1,0,0)(2,1,1)[24] with drift         : Inf
+ ARIMA(1,0,0)(1,1,1)[24] with drift         : Inf
+ ARIMA(0,0,0)(2,1,0)[24] with drift         : 110647.6
+ ARIMA(2,0,0)(2,1,0)[24] with drift         : 101332.9
+ ARIMA(2,0,0)(1,1,0)[24] with drift         : 101878.7
+ ARIMA(2,0,0)(2,1,1)[24] with drift         : Inf
+ ARIMA(2,0,0)(1,1,1)[24] with drift         : Inf
+ ARIMA(3,0,0)(2,1,0)[24] with drift         : 101317.9
+ ARIMA(3,0,0)(1,1,0)[24] with drift         : 101858.5
+ ARIMA(3,0,0)(2,1,1)[24] with drift         : Inf
+ ARIMA(3,0,0)(1,1,1)[24] with drift         : Inf
+ ARIMA(4,0,0)(2,1,0)[24] with drift         : 101302.7
+ ARIMA(4,0,0)(1,1,0)[24] with drift         : 101845
+ ARIMA(4,0,0)(2,1,1)[24] with drift         : Inf
+ ARIMA(4,0,0)(1,1,1)[24] with drift         : Inf
+ ARIMA(5,0,0)(2,1,0)[24] with drift         : 101305.4
+ ARIMA(4,0,1)(2,1,0)[24] with drift         : 101304.5
+ ARIMA(3,0,1)(2,1,0)[24] with drift         : 101307.7
+ ARIMA(5,0,1)(2,1,0)[24] with drift         : 101307.7
+ ARIMA(4,0,0)(2,1,0)[24]                    : 101300.7
+ ARIMA(4,0,0)(1,1,0)[24]                    : 101843
+ ARIMA(4,0,0)(2,1,1)[24]                    : Inf
+ ARIMA(4,0,0)(1,1,1)[24]                    : Inf
+ ARIMA(3,0,0)(2,1,0)[24]                    : 101316
+ ARIMA(5,0,0)(2,1,0)[24]                    : 101303.4
+ ARIMA(4,0,1)(2,1,0)[24]                    : 101302.4
+ ARIMA(3,0,1)(2,1,0)[24]                    : 101305.8
+ ARIMA(5,0,1)(2,1,0)[24]                    : 101305.7
+
+ Now re-fitting the best model(s) without approximations...
+
+ ARIMA(4,0,0)(2,1,0)[24]                    : 101586.6
+
+ Best model: ARIMA(4,0,0)(2,1,0)[24]
+
+```{r}
+library(car)
+
+autoplot(acf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
+autoplot(pacf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
+
+qqPlot(fit$residuals)
+gghistogram(fit$residuals)
+Box.test(fit$residuals, type="Box-Pierce")
+fit.forec <- forecast(fit, h=24*10)
+plot(fit.forec, xlim=c(250, 300))
+
+frequency(seis.h.test)
+p<-168;d<-0;q<-0;P<-0;D<-1;Q<-0;
+paste0("ARIMA(", p, ",", d, ",", q, ")(", P, ",", D, ",", Q, ")_", frequency(seis.h.test))
+fit <- Arima(seis.h.test, order=c(p,d,q), seasonal=c(P,D,Q))
+fit
+#Diagnostics
+tsdisplay(residuals(fit), lag.max=168*2)
+
+p<-1;d<-0;q<-2;P<-1;D<-0;Q<-2;
+fit <- Arima(seis.h.test, order=c(p,d,q), seasonal=c(P,D,Q))
+#Diagnostics
+tsdisplay(residuals(fit))
+forecast()
+
 ```
@@
 seis.h.end <- with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels")
-seis.h.msts <- na.aggregate(build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end, force_ms = T))
-seis.h.msts %>% mstl() %>% autoplot() + xlab("Week")
-seis.h.msts.mstl <- seis.h.msts %>% mstl()
+seis.h.msts <- seasonality.aggregate(build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end, force_ms = T))
+seis.h.msts %>% mstl(s.window = "periodic") %>% autoplot() + xlab("Week")
+seis.h.msts.mstl <- seis.h.msts %>% mstl(s.window = "periodic")
 plot(seis.h.msts.mstl[,2], main="trend")
@@
 autoplot(seis.h.msts.mstl[,5])
-acf(seis.h.msts.mstl[,5], na.action = na.omit)
-bptests <- c()
-for (i in 1:10)
-  bptests<- c(bptests, Box.test(seis.h.msts.mstl[,5], lag = i, fitdf = 0, type = "Box-Pierce")$p.value)
-plot(bptests, ylim = c(0,1))
-abline(h=0.05,col="blue",lty="dotted")
+mstl.res.plot <- autoplot(acf(seis.h.msts.mstl[,5], na.action = na.omit, plot=F)) + scale_x_continuous()
+bptests <- data.frame(lag=integer(), `p-value`=numeric())
+for (i in 1:10){
+  bptests[nrow(bptests) + 1, ] <- c(i, Box.test(seis.h.msts.mstl[,5], lag = i, fitdf = 0, type = "Box-Pierce")$p.value)
+}
+bptests.plot <- 
+  ggplot(bptests, aes(lag, `p.value`), pch="o") + 
+    geom_point() + ylim(0,1) +
+    labs(title="Box-Pierce tests for lags 1 to 10") +
+    geom_hline(yintercept=0.05, linetype="dotted", color = "blue") +
+    scale_x_continuous(breaks=1:10)
+plot_grid(mstl.res.plot, bptests.plot, nrow = 2)
 

### 2020-05-22 project.Rmd
more analysis and better Rmd output

@@
 ```{r libraries, include=FALSE}
-knitr::opts_chunk$set(warning = F, message = F, fig.height = 360)
+library(knitr)
+opts_chunk$set(warning = F, message = F)
 library(stringi) # %s+%
@@
 library(cowplot) # for plot_grid
+library(car) # qqPlot
 library(tidyverse) # map_df, read_csv, group_by, summarize
@@
 ```{r import dataset, include=FALSE}
+if (!exists("domain")) {
+  # UCCS
+  domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
   # MEMS
-domain <- c(-7200,7200); pattern <- "MEM.*.csv"; name <- "MEM"; description <- "Membach"
-# UCCS
-domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"
+  domain <- c(-7200,7200); pattern <- "MEM.*\\.csv"; name <- "MEM"; description <- "Membach"; proj_file <- 'proj.mems.RData'
+}
+if (exists("proj_file")) {
+  load(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
+  # save.image(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
+}
 seis <-
@@
 blocks$end <- seis.first.second + hours(blocks$end)
-plot(c(blocks$start, blocks$end), rep(hour(blocks$interval), 2),
+plot(c(rbind(blocks$start, blocks$end, NA)), rep(hour(blocks$interval), each=3),
      col=ifelse(blocks$interval<3600, "darkgreen", "darkred"),
@@
 text(x=blocks$end, y=hour(blocks$interval), labels = paste0(format(blocks$end, "%y-%m-%d %H:%M")), cex=0.9, pos=4)
-text(x=blocks$start + blocks$interval/2, y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=3)
+text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=3)
 remove(seis.complete, seis.missing.ids)
 
-Box.test(forec$model$residuals, type="Box-Pierce")
+blocks$interval
+Box.test(seis.h$mean_movement_s, type="Box-Pierce")
 
@@
 
-```{r plot data with (p)acf}
+```{r test_train split ts, include=FALSE}
 seis.h.start <- as_datetime('2018-10-29T00:00:00')
-seis.h.end <- as_datetime('2020-03-01T00:00:00')
+seis.h.end <- as_datetime('2020-03-02T00:00:00') - seconds(1)
 seasonality.period <- 24
 seasonality.xlab <- "Days"
-seasonality.aggregate <- function(ds) na.aggregate(seis.h.ts, function(x) floor(x/7))
+seasonality.aggregate <- function(ds) na.aggregate(seis.h.ts, function(x) floor(x/14))
 seasonality.period <- 168
@@
 seis.h.ts <- build_ts(seis.h, periods = c(seasonality.period), start = seis.h.start, end = seis.h.end, force_ms = T)
+
+seis.test.split <- seis.h.start + days(52*7)
+
+seis.train.start <- seis.h.start
+seis.train.end <- seis.test.split - seconds(1)
+seis.h.train <- build_ts(seis.h, periods = c(seasonality.period), start = seis.train.start, end = seis.train.end, force_ms = T)
+
+seis.test.start <- seis.test.split
+seis.test.end <- as_datetime('2020-05-04T00:00:00') - seconds(1)
+seis.h.test <- build_ts(seis.h, periods = c(seasonality.period), start = seis.h.start, end = seis.test.end, force_ms = T)
+seis.h.test <- window(seis.h.test, start = c(end(seis.h.train)[1] + 1, 1))
+seis.test.length <- length(seis.h.test)
+```
+
+```{r plot data with (p)acf}
+seis.h.24 <- build_ts(seis.h, periods = c(24), start = seis.h.start, end = seis.h.end)
+seasonplot(seis.h.24, type="l", col=rainbow(7))
+legend("topleft", legend=weekdays(seis.h.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
+
 plot_grid(
   autoplot(seis.h.ts, main = "Hourly seismic movement", xlab = seasonality.xlab, ylab = "mean movement (nm)"),
-  gghistogram(seis.h.ts) + xlab("mean movement (nm)"),
+  gghistogram(seis.h.ts, add.normal = T) + xlab("mean movement (nm)"),
   nrow = 2)
@@
 
-  mse <- mean(diff^2)
-  mae <- mean(abs(diff))
+  mse <- mean(diff^2, na.rm = T)
+  mae <- mean(abs(diff), na.rm = T)
   rmse <- sqrt(mse)
-  R2 <- 1-(sum((diff)^2)/sum((base-mean(base))^2))
+  R2 <- 1-(sum((diff)^2, na.rm = T)/sum((base-mean(base))^2, na.rm = T))
   errorstring <- paste0("MSE=", mse, "    MAE=", mae, "    RMSE=", rmse, "    R2=", R2)
@@
 
-plot_grid(
-  autoplot(aggregated) + ylab("") + xlab(""),
-  autoplot(ts(season.24$seasonal, frequency=24)) + ylab("") + xlab(""),
-  autoplot(ts(season.168$seasonal, frequency=24)) + ylab("") + xlab(""),
-  autoplot(ts(season.168$trend, frequency=24)) + ylab("") + xlab(""),
-  autoplot(aggregated-season.24$seasonal-season.168$seasonal-season.168$trend) + ylab("") + xlab(""),
-  nrow = 5)
-
 season.24 <- trend_season(aggregated-season.168$seasonal, 24)
@@
 
+# this third iteration isn't actually needed
 season.24 <- trend_season(aggregated-season.168$seasonal, 24)
@@
 plot_grid(
-  autoplot(aggregated) + ylab("") + xlab(""),
-  autoplot(ts(season.24$seasonal, frequency=24)) + ylab("") + xlab(""),
-  autoplot(ts(season.168$seasonal, frequency=24)) + ylab("") + xlab(""),
-  autoplot(ts(season.168$trend, frequency=24)) + ylab("") + xlab(""),
-  autoplot(aggregated-season.24$seasonal-season.168$seasonal-season.168$trend) + ylab("") + xlab(""),
-  nrow = 5)
+  autoplot(aggregated) + labs(title="Timeseries", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(b=-6, unit="pt")),
+  autoplot(ts(season.24$figure, frequency=1)) + labs(title="Daily seasonality", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(r=10, b=-6, unit="pt")),
+  autoplot(ts(season.168$trend, frequency=168)) + labs(title="Trend", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(l=4, b=-6, unit="pt")),
+  autoplot(ts(season.168$figure, frequency=1)) + labs(title="Weekly seasonality", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(r=10, b=-6, unit="pt")),
+  autoplot(aggregated-season.24$seasonal-season.168$seasonal-season.168$trend) + labs(title="Residuals", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(t=-6, b=-6, unit="pt")),
+  nrow = 3,
+  greedy = T)
 
@@
 plot_grid(
-  autoplot(ts(seasonbase.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
-      geom_line(aes(y=season.24$figure, col="red"), show.legend = F),
-  autoplot(ts(seasonbase.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
-      geom_line(aes(y=season.168$figure, col="red")) +
-      geom_line(aes(y=season.168$figure + rep(season.24$figure, 7), col="blue")),
+  autoplot(ts(seasonbase.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
+    geom_line(aes(y=season.24$figure, col="Multi daily")) +
+    scale_colour_manual("", values=c("purple", "yellow"), breaks=c("Direct daily", "Multi daily")),
+  autoplot(ts(seasonbase.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
+    geom_line(aes(y=season.168$figure + rep(season.24$figure, 7), col="Multi w+d")) +
+    #geom_line(aes(y=rep(seasonbase.24$figure, 7), col="Direct daily"), size=1.5) +
+    #geom_line(aes(y=rep(season.24$figure, 7), col="Multi daily")) +
+    geom_line(aes(y=season.168$figure, col="Multi weekly")) +
+    scale_colour_manual("", values=c("red", "purple", "blue", "yellow", "green"), breaks=c("Direct weekly", "Direct daily", "Multi" %s+% c(" weekly", " daily", " w+d"))),
   nrow = 2)
@@
 
-```{r test_ts}
-seis.test.start <- as_datetime('2019-05-20T00:00:00')
-seis.test.end <- as_datetime('2020-03-02T00:00:00')
-
-seis.h.test <- build_ts(seis.h, periods = c(seasonality.period), start = seis.test.start, end = seis.test.end, force_ms = T)
+```{r diffs acf/pacf}
+seis.h.train %>% ggtsdisplay(lag.max = 168*5)
+seis.h.train %>% diff() %>% ggtsdisplay(lag.max = 168*5)
+# seis.h.train %>% diff(D=2) %>% ggtsdisplay(lag.max = 168)
+# seis.h.train %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
+# seis.h.train %>% diff() %>% diff(lag=24) %>% ggtsdisplay(lag.max = 24, main="Supposed model: ARIMA(24,1,0)(0,1,2)[24]") # ARIMA(0,1,0)(0,1,2)[24]
+# previous => ARIMA(5,1,0)(0,1,7)[24]
+# seis.h.train %>% diff() %>% diff(lag=24, D=2) %>% ggtsdisplay(lag.max = 168)
+seis.h.train %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,0,0)(0,1,1/2)[168]
+# acf variable and partially sinusoidal, pacf shows peaks each 168
+seis.h.train %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,1,0)(0,1,1)[168]
+# acf cuts off after p=168, pacf shows peaks each 168
+# seis.h.train %>% diff() %>% diff(lag=168, D=2) %>% ggtsdisplay(lag.max = 168)
+seis.h.train %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)
 ```
 
-```{r diffs acf/pacf}
-seis.h.test %>% ggtsdisplay(lag.max = 168)
-seis.h.test %>% diff() %>% ggtsdisplay(lag.max = 168*5)
-# seis.h.test %>% diff(D=2) %>% ggtsdisplay(lag.max = 168)
-# seis.h.test %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
-# seis.h.test %>% diff() %>% diff(lag=24) %>% ggtsdisplay(lag.max = 24, main="Supposed model: ARIMA(24,1,0)(0,1,2)[24]") # ARIMA(0,1,0)(0,1,2)[24]
-# previous => ARIMA(5,1,0)(0,1,7)[24]
-# seis.h.test %>% diff() %>% diff(lag=24, D=2) %>% ggtsdisplay(lag.max = 168)
-seis.h.test %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,0,0)(0,1,1/2)[168]
-seis.h.test %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,1,0)(0,1,1)[168]
-# seis.h.test %>% diff() %>% diff(lag=168, D=2) %>% ggtsdisplay(lag.max = 168)
-seis.h.test %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168)
-```
-
-```{r arima}
-
+```{r plot_forecast func}
 plot_forecast <- function(forec) {
@@
 }
+```
 
+```{r arima}
 to_test <- list(
@@
   c(3,1,1,0,1,1),  # after hint then ACF + PACF
+  c(4,1,2,0,1,0),  # hint from auto.arima d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10
+  c(4,1,2,0,1,1)  # after hint then ACF + PACF
 )
@@
 for (model in to_test) {
-  model <- c(3,1,1,0,1,1)
+  # model <- c(4,1,2,0,1,0)
   fit.name <- paste0("fit", paste0(model, collapse = ""))
   forecast.name <- paste0("forecast", paste0(model, collapse = ""))
-  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.test) %s+% "]")
-  models[[fit.name]] <- Arima(seis.h.test, model[1:3], model[4:6])
-  models[[forecast.name]] <- forecast(models[[fit.name]], h=168*10)
+  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
+  if (!(fit.name %in% names(models))) {
+    models[[fit.name]] <- Arima(seis.h.train, model[1:3], model[4:6])
   }
-models[["fitauto"]] <- auto.arima(seis.h.test, D=1, approx=F, trace = T)
-models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*10)
-models[["fit311010"]] <- models[["fitauto"]]
-models[["forecast311010"]] <- forecast(models[["fit311010"]], h=168*10)
+  if (!(forecast.name %in% names(models))) {
+    models[[forecast.name]] <- forecast(models[[fit.name]], h=168*18)
+  }
+}
+#models[["fitauto"]] <- auto.arima(seis.h.train, D=1, approx=F, trace = T)
+#models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*18)  
+#models[["fit311010"]] <- models[["fitauto"]]
+#models[["fit412010"]] <- fit
+#models[["forecast412010"]] <- forecast(models[["fit412010"]], h=168*18)
 
@@
 plot_grid(residuals.acf, residuals.pacf, nrow=2)
-plot_forecast(plotting.forecast)
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Original")
 plots <- show_overlap_and_diff(
-  seis.h.test, plotting.forecast$fitted, xlab=seasonality.xlab,
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
   base.desc = "original series",
@@
 plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
 
@@
 plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
 
+plotting.model <- models[["fit412010"]]
+plotting.forecast <- models[["forecast412010"]]
+checkresiduals(plotting.model)
+residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Original") 
+plots <- show_overlap_and_diff(
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "original series",
+  compare.desc = "reconstructed series",
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
+
+plotting.model <- models[["fit500010"]]
+plotting.forecast <- models[["forecast500010"]]
+checkresiduals(plotting.model)
+residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Original") 
+plots <- show_overlap_and_diff(
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "original series",
+  compare.desc = "reconstructed series",
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
+
+for (model in to_test) {
+  # model <- c(4,1,2,0,1,0)
+  fit.name <- paste0("fit", paste0(model, collapse = ""))
+  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
+  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
+  if ((fit.name %in% names(models))) {  if ((forecast.name %in% names(models))) {
+    plotting.model <- models[[fit.name]]
+    plotting.forecast <- models[[forecast.name]]
+    checkresiduals(plotting.model)
+    residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+    residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+    plot_grid(residuals.acf, residuals.pacf, nrow=2)
+    plot_forecast(plotting.forecast) + 
+      autolayer(seis.h.test, series="Original") 
+    plots <- show_overlap_and_diff(
+      seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+      base.desc = "original series",
+      compare.desc = paste0("refit with ", plotting.forecast$method),
+      overlap.ylab = "Mean hourly movement (nm)",
+      diff.title = "Difference between ts",
+      diff.ylab = "nm", diff.ylab.sec="% over ts span",
+      diff.ylab.sec.accuracy = .01
+    )
+    plot_grid(plots[[1]], plots[[2]], nrow = 2)
+    accuracy(plotting.forecast, seis.h.test)
+  }}
+}
+qqPlot(models[["fit500010"]]$residuals)
 ```
@@
 ```{r}
-(seis.test.start - seis.first.second)
-
-deseasoned24 <- seis.h.ts - seasonality
-autoplot(deseasoned24)
-frequency(deseasoned24)
-de24.test <- msts(window(deseasoned24, start=205), seasonal.periods = c(168))
-frequency(de24.test)
-autoplot(de24.test)
-de24.test %>% ggtsdisplay(lag.max = 168*2)
-de24.test %>% diff() %>% ggtsdisplay(lag.max = 168*2)
-de24.test %>% diff(D=2) %>% ggtsdisplay(lag.max = 168*2)
-de24.test %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*2)
-de24.test %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*2, main="Supposed model: ARIMA(0,1,0)(0,1,1)[168]")
-de24.test %>% diff() %>% diff(lag=168, D=2) %>% ggtsdisplay(lag.max = 168*2)
-
-if (!exists("fit010011de")){
-  fit010011de <- Arima(de24.test, c(0,1,0), c(0,1,1))
-}
-checkresiduals(fit010011de)
-residuals.acf <- autoplot(acf(fit010011de$residuals, na.action = na.pass, plot=F, lag.max=168*3)) + scale_x_continuous()
-residuals.pacf <- autoplot(pacf(fit010011de$residuals, na.action = na.pass, plot=F, lag.max=168*3)) + scale_x_continuous()
-plot_grid(residuals.acf, residuals.pacf, nrow=2)
-if (!exists("fit010011de.forec")){
-  fit010011de.forec <- forecast(fit010011de, h=168*10)
-}
-plot_forecast(forecast(fit010011de.forec, h=168*10))
-
-# NO  fit011011de <- Arima(de24.test, c(0,1,1), c(0,1,1))
-if (!exists("fit011011de")){
-  fit011011de <- Arima(de24.test, c(0,1,1), c(0,1,1))
-  arima
-}
-fit010011de$aic
-fit011011de$aic
-checkresiduals(fit011011de)
-residuals.acf <- autoplot(acf(fit011011de$residuals, na.action = na.pass, plot=F, lag.max=168*3)) + scale_x_continuous()
-residuals.pacf <- autoplot(pacf(fit011011de$residuals, na.action = na.pass, plot=F, lag.max=168*3)) + scale_x_continuous()
-plot_grid(residuals.acf, residuals.pacf, nrow=2)
-if (!exists("fit011011de.forec")){
-  fit011011de.forec <- forecast(fit011011de, h=168*10)
-}
-plot_forecast(forecast(fit011011de.forec, h=168*10))
-
+#fit <- auto.arima(seis.h.train, d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10, trace = T, parallel = T)
 ```
-
-```{r}
-
-seis.test.start <- as_datetime('2019-05-20T00:00:00')
-seis.test.end <- as_datetime('2020-03-02T00:00:00')
-
-seis.h.test <- build_ts(seis.h, periods = c(seasonality.period), start = seis.test.start, end = seis.test.end, force_ms = T)
-
-# First seasonal difference (quarterly)
-# seasonality.arima <- 24
-D<-1
-seis.h.test.ds<-diff(seis.h.test,differences=D)
-plot(seis.h.test.ds)
-abline(h=mean(seis.h.test.ds), col="red", lty="dotted")
-sd(seis.h.test.ds)
-tsdisplay(seis.h.test.ds, lag.max = 168*2)
-autoplot(acf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
-autoplot(pacf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
-seis.h.test.ds<-diff(seis.h.test,lag=seasonality.period,differences=D)
-plot(seis.h.test.ds)
-abline(h=mean(seis.h.test.ds), col="red", lty="dotted")
-sd(seis.h.test.ds)
-autoplot(acf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
-autoplot(pacf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
-tsdisplay(seis.h.test.ds, lag.max = 168*2)
-# pacf cuts off after p=168
-```
-
-```{r}
-fit <- auto.arima(seis.h.test, d=1, D=1, trace = T, parallel = T)
-```
+# D=1 m=24
  Fitting models using approximations to speed things up...
@@
 
-```{r}
-library(car)
+# d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10 m=168
+ Fitting models using approximations to speed things up...
 
-autoplot(acf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
-autoplot(pacf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
+ ARIMA(2,1,2)(1,1,1)[168]                    : Inf
+ ARIMA(0,1,0)(0,1,0)[168]                    : 122379.8
+ ARIMA(1,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(0,1,1)(0,1,1)[168]                    : Inf
+ ARIMA(0,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(0,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(0,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(1,1,0)(0,1,0)[168]                    : 121638
+ ARIMA(1,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(1,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(2,1,0)(0,1,0)[168]                    : 121468.5
+ ARIMA(2,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(2,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(2,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(3,1,0)(0,1,0)[168]                    : 121452.9
+ ARIMA(3,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(3,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(3,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(4,1,0)(0,1,0)[168]                    : 121403.9
+ ARIMA(4,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(4,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(4,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(5,1,0)(0,1,0)[168]                    : 121361.7
+ ARIMA(5,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(5,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(5,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(5,1,1)(0,1,0)[168]                    : 118580.2
+ ARIMA(5,1,1)(1,1,0)[168]                    : Inf
+ ARIMA(5,1,1)(0,1,1)[168]                    : Inf
+ ARIMA(5,1,1)(1,1,1)[168]                    : Inf
+ ARIMA(4,1,1)(0,1,0)[168]                    : 118580.3
+ ARIMA(5,1,2)(0,1,0)[168]                    : 118581.5
+ ARIMA(4,1,2)(0,1,0)[168]                    : 118553.1
+ ARIMA(4,1,2)(1,1,0)[168]                    : Inf
+ ARIMA(4,1,2)(0,1,1)[168]                    : Inf
+ ARIMA(4,1,2)(1,1,1)[168]                    : Inf
+ ARIMA(3,1,2)(0,1,0)[168]                    : 118642.8
+ ARIMA(3,1,1)(0,1,0)[168]                    : 119155
+
+ Now re-fitting the best model(s) without approximations...
+
+ ARIMA(4,1,2)(0,1,0)[168]                    : 122838.5
+
+ Best model: ARIMA(4,1,2)(0,1,0)[168]
+
+```{r eval=FALSE, include=FALSE}
+
+autoplot(acf(seis.h.train.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
+autoplot(pacf(seis.h.train.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
 
@@
 
-frequency(seis.h.test)
+frequency(seis.h.train)
 p<-168;d<-0;q<-0;P<-0;D<-1;Q<-0;
-paste0("ARIMA(", p, ",", d, ",", q, ")(", P, ",", D, ",", Q, ")_", frequency(seis.h.test))
-fit <- Arima(seis.h.test, order=c(p,d,q), seasonal=c(P,D,Q))
+paste0("ARIMA(", p, ",", d, ",", q, ")(", P, ",", D, ",", Q, ")_", frequency(seis.h.train))
+fit <- Arima(seis.h.train, order=c(p,d,q), seasonal=c(P,D,Q))
 fit
@@
 p<-1;d<-0;q<-2;P<-1;D<-0;Q<-2;
-fit <- Arima(seis.h.test, order=c(p,d,q), seasonal=c(P,D,Q))
+fit <- Arima(seis.h.train, order=c(p,d,q), seasonal=c(P,D,Q))
 #Diagnostics
@@
 
-```{r}
-seis.h.start <- with_tz(as_datetime('2018-11-01T00:00:00'), "Europe/Brussels")
-seis.h.end <- with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels")
-seis.h.msts <- seasonality.aggregate(build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end, force_ms = T))
-seis.h.msts %>% mstl(s.window = "periodic") %>% autoplot() + xlab("Week")
-seis.h.msts.mstl <- seis.h.msts %>% mstl(s.window = "periodic")
-plot(seis.h.msts.mstl[,2], main="trend")
+```{r mstl on msts}
+seis.h.train.m <- msts(seis.h.train, c(24, 168))
+seis.mstl <- seis.h.train.m %>% mstl(s.window="periodic") #c(15, 7))
+seis.mstl.fc <- forecast(seis.mstl, h=168*18)
+
+autoplot(seis.mstl) + xlab("Week")
+plot(seis.mstl[,2], main="trend")
 lines(trend.s, col="red")
-plot(rowSums(seis.h.msts.mstl[,3:4]), main="seasonality",type="l")
+plot(rowSums(seis.mstl[,3:4]), main="seasonality",type="l")
 lines(seasonality, col="red")
-plot(residuals,  main="residuals")
-lines(seis.h.msts.mstl[,5], col="red")
-autoplot(seis.h.msts.mstl[,5])
-mstl.res.plot <- autoplot(acf(seis.h.msts.mstl[,5], na.action = na.omit, plot=F)) + scale_x_continuous()
+plot(seis.mstl[,5],  main="residuals")
+lines(residuals, col="red")
+
+plotting.model <- seis.mstl
+plotting.forecast <- seis.mstl.fc
+checkresiduals(plotting.model[,5])
+Box.test(plotting.forecast$model$residuals, type="L")
+residuals.acf <- autoplot(acf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Original") +
+  xlim(52, 70) + ylim(0, 8750)
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Original")
+plots <- show_overlap_and_diff(
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "train series",
+  compare.desc = paste0("refit with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+plots <- show_overlap_and_diff(
+  seis.h.test[1:length(plotting.forecast$mean)], plotting.forecast$mean, xlab=seasonality.xlab,
+  base.desc = "test series",
+  compare.desc = paste0("refit with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
+qqPlot(plotting.model[,5])
+
 bptests <- data.frame(lag=integer(), `p-value`=numeric())
 for (i in 1:10){
-  bptests[nrow(bptests) + 1, ] <- c(i, Box.test(seis.h.msts.mstl[,5], lag = i, fitdf = 0, type = "Box-Pierce")$p.value)
+  bptests[nrow(bptests) + 1, ] <- c(i, Box.test(plotting.model[,5], lag = i, fitdf = 0, type = "Box-Pierce")$p.value)
 }
@@
     scale_x_continuous(breaks=1:10)
-plot_grid(mstl.res.plot, bptests.plot, nrow = 2)
+bptests.plot
 
 ```
+
+```{r eval=FALSE, include=FALSE}
+gc()
+if (name == "MEM") {
+  proj_file <- 'proj.mems.RData'
+  save.image(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
+  remove(proj_file)
+}
+```

### 2020-05-22 project.mems.Rmd
more analysis and better Rmd output

---
title: "Project Membach"
author: "Marco Ciotola"
date: "16/05/2020"
output:
  pdf_document:
    keep_tex: yes
    fig_width: 6.5
    fig_height: 4
#  html_document:
#    mathjax: local
#    self_contained: no
#    fig_width: 6.5
#    fig_height: 4
---

```{r tentatives, eval=FALSE, child='chapter1.Rmd', include=FALSE}
```

```{r set mems, include=FALSE}
# MEMS
domain <- c(-7200,7200); pattern <- "MEM.*\\.csv"; name <- "MEM"; description <- "Membach"; proj_file <- 'proj.mems.RData'
```

```{r project, child='project.Rmd'}
```

### 2020-05-22 project.uccs.Rmd
more analysis and better Rmd output

---
title: "Project Uccle"
author: "Marco Ciotola"
output:
  pdf_document:
    keep_tex: yes
    fig_width: 6.5
    fig_height: 4
#  html_document:
#    mathjax: local
#    self_contained: no
#    fig_width: 6.5
#    fig_height: 4
---

```{r tentatives, eval=FALSE, child='chapter1.Rmd', include=FALSE}
```

```{r set uccs, include=FALSE}
# UCCS
domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
```

```{r project, child='project.Rmd'}
```

### 2020-05-22 crawl.py

@@ -118,3 +118,4 @@ def download_station(station: str):
 if __name__ == '__main__':
     download_station('UCCS')
     download_station('MEMS')
+    download_station('OSTB')

### 2020-05-23 chapter1.Rmd
divided not-common analysis and some more analysis

@@
 
-library(spec)
 pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2018-11-01T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)

### 2020-05-23 project.Rmd
divided not-common analysis and some more analysis

@@
-```{r libraries, include=FALSE}
+```{r "libraries", include=FALSE}
 library(knitr)
@@
 library(fpp) # auto.arima
+library(ggplot2)
 # library(gridExtra) # for gg-plotting multiple plots
 library(cowplot) # for plot_grid
-library(car) # qqPlot
+# library(car) # qqPlot
+library(qqplotr) # qqplot with ggplot
 library(tidyverse) # map_df, read_csv, group_by, summarize
 library(tsbox) # https://cran.r-project.org/web/packages/tsbox/vignettes/tsbox.html
+```
+
+```{r "initial setup", include=FALSE}
 print_dec <- function(x, decimals=2) trimws(format(round(x, decimals), nsmall=decimals))
@@
 setwd("/media/data/Documents/Uni/spatiotemp/villalobos/project/")
-```
-
-```{r import dataset, include=FALSE}
 if (!exists("domain")) {
@@
 if (exists("proj_file")) {
-  load(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
+  #load(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
   # save.image(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
 }
+
+gg_qqplot <- function(dataset) {
+  if (is.ts(dataset))
+    dataset <- as.numeric(dataset)
+  ggplot(data.frame(values=dataset), aes(sample=values)) +
+    stat_qq_band() + stat_qq_line(col="red") + stat_qq_point(shape = 1, size = 3) +
+    xlab("Theoretical quantiles (95% conf.)")
+}
+```
+
+```{r "import dataset", include=FALSE}
 seis <-
@@
   map_df(~read_csv("csv/" %s+% .))
+
 ```
@@
 text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=3)
+text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=1)
 remove(seis.complete, seis.missing.ids)
@@
 seis.test.length <- length(seis.h.test)
+```
+
+```{r periodgram}
+library(spec)
+pgram <- spec.pgram(seis.h.train, main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
+topfreq <- pgram$freq[order(pgram$spec, decreasing = T)][1:100]
+topseas <- 1/topfreq
+topseas.h <- topseas/1
+
+filter_freq <- (pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*2)) 
+top95pc <- pgram$spec[filter_freq]
+top95freq <- pgram$freq[filter_freq][order(top95pc, decreasing = T)]
+top95pc <- pgram$spec[filter_freq][order(top95pc, decreasing = T)]
+top95seas.s <- 1/top95freq
+top95seas.s
+top95seas.h <- top95seas.s/1
+top95pc/sd(pgram$spec)
+top95seas.h
+
+ylims <- 10^par("usr")[3:4]
+axis(side=3, at=top95freq, labels=as.integer(top95seas.h), padj = 1, las=2, cex.axis=0.8)
+abline(h=mean(pgram$spec) + sd(pgram$spec)*c(.3,1:3), lty="dotted", col=c("orange", "darkgreen", "blue", "red"))
+for (i in 1:length(top95freq)) {
+  color <- ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*3), "red", ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*2), "blue", ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*1), "darkgreen", "orange")))
+  lines(x=c(top95freq[i], top95freq[i]), y=c(top95pc[i],ylims[2]), lty="dotted", col=color)
+  points(x=top95freq[i], y=top95pc[i], col=color)
+}
+axis(side=4, at=mean(pgram$spec) + sd(pgram$spec) * c(.3, 1:3), labels=c(".3", "1", "2", "3"), cex.axis=0.8)
+mtext(side=4, "spectrum sd ratio", line=1)
+gc()
 ```
@@
 ```
-
-```{r arima}
-to_test <- list(
-  #c(0,0,0,0,0,1),  # BAD, Box-Pierce = 0
-  #c(0,0,0,0,0,2),  # BAD, Box-Pierce = 0
-  #c(0,0,0,0,1,1),  # nice, Box-Pierce = 0
-  c(0,0,0,0,1,2),
-  #c(0,0,0,0,1,3),  # CAN'T BE DONE: 3*168 = 504 > 350 (limit of arima on r)
-  #c(0,1,0,0,0,1),  # BAD, Box-Pierce = 0
-  c(0,1,0,0,1,1),
-  c(0,1,0,0,1,2),
-  c(2,0,0,0,1,0),  # hint from auto.arima D=1
-  c(5,0,0,0,1,0),  # hint from auto.arima D=1
-  c(3,1,1,0,1,0),  # hint from auto.arima D=1 d=1
-  c(3,1,1,0,1,1),  # after hint then ACF + PACF
-  c(4,1,2,0,1,0),  # hint from auto.arima d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10
-  c(4,1,2,0,1,1)  # after hint then ACF + PACF
-)
-if (!exists("models")){
-  models <- list()
-}
-for (model in to_test) {
-  # model <- c(4,1,2,0,1,0)
-  fit.name <- paste0("fit", paste0(model, collapse = ""))
-  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
-  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
-  if (!(fit.name %in% names(models))) {
-    models[[fit.name]] <- Arima(seis.h.train, model[1:3], model[4:6])
-  }
-  if (!(forecast.name %in% names(models))) {
-    models[[forecast.name]] <- forecast(models[[fit.name]], h=168*18)
-  }
-}
-#models[["fitauto"]] <- auto.arima(seis.h.train, D=1, approx=F, trace = T)
-#models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*18)  
-#models[["fit311010"]] <- models[["fitauto"]]
-#models[["fit412010"]] <- fit
-#models[["forecast412010"]] <- forecast(models[["fit412010"]], h=168*18)
-
-# 000012 better than 000011, Box-Pierce = 0
-plotting.model <- models[["fit000012"]]
-plotting.forecast <- models[["forecast000012"]]
-# 010011 somehow better somehow worst than 000012, Box-Pierce = 0
-plotting.model <- models[["fit010011"]]
-plotting.forecast <- models[["forecast010011"]]
-# 010012 , Box-Pierce = 
-plotting.model <- models[["fit010012"]]
-plotting.forecast <- models[["forecast010012"]]
-# 311010 , Box-Pierce = 
-plotting.model <- models[["fit311010"]]
-plotting.forecast <- models[["forecast311010"]]
-# 311011 , Box-Pierce = 
-plotting.model <- models[["fit311011"]]
-plotting.forecast <- models[["forecast311011"]]
-
-checkresiduals(plotting.model)
-residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-plot_grid(residuals.acf, residuals.pacf, nrow=2)
-plot_forecast(plotting.forecast) + 
-  autolayer(seis.h.test, series="Original")
-plots <- show_overlap_and_diff(
-  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
-  base.desc = "original series",
-  compare.desc = "reconstructed series",
-  overlap.ylab = "Mean hourly movement (nm)",
-  diff.title = "Difference between ts",
-  diff.ylab = "nm", diff.ylab.sec="% over ts span",
-  diff.ylab.sec.accuracy = .01
-)
-plot_grid(plots[[1]], plots[[2]], nrow = 2)
-accuracy(plotting.forecast, seis.h.test)
-
-plotting.model <- models[["fit010012"]]
-plotting.forecast <- models[["forecast010012"]]
-checkresiduals(plotting.model)
-residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-plot_grid(residuals.acf, residuals.pacf, nrow=2)
-plot_forecast(plotting.forecast) + 
-  autolayer(seis.h.test, series="Original")
-plots <- show_overlap_and_diff(
-  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
-  base.desc = "original series",
-  compare.desc = "reconstructed series",
-  overlap.ylab = "Mean hourly movement (nm)",
-  diff.title = "Difference between ts",
-  diff.ylab = "nm", diff.ylab.sec="% over ts span",
-  diff.ylab.sec.accuracy = .01
-)
-plot_grid(plots[[1]], plots[[2]], nrow = 2)
-accuracy(plotting.forecast, seis.h.test)
-
-plotting.model <- models[["fit412010"]]
-plotting.forecast <- models[["forecast412010"]]
-checkresiduals(plotting.model)
-residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-plot_grid(residuals.acf, residuals.pacf, nrow=2)
-plot_forecast(plotting.forecast) + 
-  autolayer(seis.h.test, series="Original") 
-plots <- show_overlap_and_diff(
-  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
-  base.desc = "original series",
-  compare.desc = "reconstructed series",
-  overlap.ylab = "Mean hourly movement (nm)",
-  diff.title = "Difference between ts",
-  diff.ylab = "nm", diff.ylab.sec="% over ts span",
-  diff.ylab.sec.accuracy = .01
-)
-plot_grid(plots[[1]], plots[[2]], nrow = 2)
-accuracy(plotting.forecast, seis.h.test)
-
-plotting.model <- models[["fit500010"]]
-plotting.forecast <- models[["forecast500010"]]
-checkresiduals(plotting.model)
-residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-plot_grid(residuals.acf, residuals.pacf, nrow=2)
-plot_forecast(plotting.forecast) + 
-  autolayer(seis.h.test, series="Original") 
-plots <- show_overlap_and_diff(
-  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
-  base.desc = "original series",
-  compare.desc = "reconstructed series",
-  overlap.ylab = "Mean hourly movement (nm)",
-  diff.title = "Difference between ts",
-  diff.ylab = "nm", diff.ylab.sec="% over ts span",
-  diff.ylab.sec.accuracy = .01
-)
-plot_grid(plots[[1]], plots[[2]], nrow = 2)
-accuracy(plotting.forecast, seis.h.test)
-
-for (model in to_test) {
-  # model <- c(4,1,2,0,1,0)
-  fit.name <- paste0("fit", paste0(model, collapse = ""))
-  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
-  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
-  if ((fit.name %in% names(models))) {  if ((forecast.name %in% names(models))) {
-    plotting.model <- models[[fit.name]]
-    plotting.forecast <- models[[forecast.name]]
-    checkresiduals(plotting.model)
-    residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-    residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-    plot_grid(residuals.acf, residuals.pacf, nrow=2)
-    plot_forecast(plotting.forecast) + 
-      autolayer(seis.h.test, series="Original") 
-    plots <- show_overlap_and_diff(
-      seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
-      base.desc = "original series",
-      compare.desc = paste0("refit with ", plotting.forecast$method),
-      overlap.ylab = "Mean hourly movement (nm)",
-      diff.title = "Difference between ts",
-      diff.ylab = "nm", diff.ylab.sec="% over ts span",
-      diff.ylab.sec.accuracy = .01
-    )
-    plot_grid(plots[[1]], plots[[2]], nrow = 2)
-    accuracy(plotting.forecast, seis.h.test)
-  }}
-}
-qqPlot(models[["fit500010"]]$residuals)
-```
-
-```{r}
-#fit <- auto.arima(seis.h.train, d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10, trace = T, parallel = T)
-```
-# D=1 m=24
- Fitting models using approximations to speed things up...
-
- ARIMA(2,0,2)(1,1,1)[24] with drift         : Inf
- ARIMA(0,0,0)(0,1,0)[24] with drift         : 111462.7
- ARIMA(1,0,0)(1,1,0)[24] with drift         : 101954.4
- ARIMA(0,0,1)(0,1,1)[24] with drift         : Inf
- ARIMA(0,0,0)(0,1,0)[24]                    : 111460.9
- ARIMA(1,0,0)(0,1,0)[24] with drift         : 102426.5
- ARIMA(1,0,0)(2,1,0)[24] with drift         : 101379
- ARIMA(1,0,0)(2,1,1)[24] with drift         : Inf
- ARIMA(1,0,0)(1,1,1)[24] with drift         : Inf
- ARIMA(0,0,0)(2,1,0)[24] with drift         : 110647.6
- ARIMA(2,0,0)(2,1,0)[24] with drift         : 101332.9
- ARIMA(2,0,0)(1,1,0)[24] with drift         : 101878.7
- ARIMA(2,0,0)(2,1,1)[24] with drift         : Inf
- ARIMA(2,0,0)(1,1,1)[24] with drift         : Inf
- ARIMA(3,0,0)(2,1,0)[24] with drift         : 101317.9
- ARIMA(3,0,0)(1,1,0)[24] with drift         : 101858.5
- ARIMA(3,0,0)(2,1,1)[24] with drift         : Inf
- ARIMA(3,0,0)(1,1,1)[24] with drift         : Inf
- ARIMA(4,0,0)(2,1,0)[24] with drift         : 101302.7
- ARIMA(4,0,0)(1,1,0)[24] with drift         : 101845
- ARIMA(4,0,0)(2,1,1)[24] with drift         : Inf
- ARIMA(4,0,0)(1,1,1)[24] with drift         : Inf
- ARIMA(5,0,0)(2,1,0)[24] with drift         : 101305.4
- ARIMA(4,0,1)(2,1,0)[24] with drift         : 101304.5
- ARIMA(3,0,1)(2,1,0)[24] with drift         : 101307.7
- ARIMA(5,0,1)(2,1,0)[24] with drift         : 101307.7
- ARIMA(4,0,0)(2,1,0)[24]                    : 101300.7
- ARIMA(4,0,0)(1,1,0)[24]                    : 101843
- ARIMA(4,0,0)(2,1,1)[24]                    : Inf
- ARIMA(4,0,0)(1,1,1)[24]                    : Inf
- ARIMA(3,0,0)(2,1,0)[24]                    : 101316
- ARIMA(5,0,0)(2,1,0)[24]                    : 101303.4
- ARIMA(4,0,1)(2,1,0)[24]                    : 101302.4
- ARIMA(3,0,1)(2,1,0)[24]                    : 101305.8
- ARIMA(5,0,1)(2,1,0)[24]                    : 101305.7
-
- Now re-fitting the best model(s) without approximations...
-
- ARIMA(4,0,0)(2,1,0)[24]                    : 101586.6
-
- Best model: ARIMA(4,0,0)(2,1,0)[24]
-
-# d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10 m=168
- Fitting models using approximations to speed things up...
-
- ARIMA(2,1,2)(1,1,1)[168]                    : Inf
- ARIMA(0,1,0)(0,1,0)[168]                    : 122379.8
- ARIMA(1,1,0)(1,1,0)[168]                    : Inf
- ARIMA(0,1,1)(0,1,1)[168]                    : Inf
- ARIMA(0,1,0)(1,1,0)[168]                    : Inf
- ARIMA(0,1,0)(0,1,1)[168]                    : Inf
- ARIMA(0,1,0)(1,1,1)[168]                    : Inf
- ARIMA(1,1,0)(0,1,0)[168]                    : 121638
- ARIMA(1,1,0)(0,1,1)[168]                    : Inf
- ARIMA(1,1,0)(1,1,1)[168]                    : Inf
- ARIMA(2,1,0)(0,1,0)[168]                    : 121468.5
- ARIMA(2,1,0)(1,1,0)[168]                    : Inf
- ARIMA(2,1,0)(0,1,1)[168]                    : Inf
- ARIMA(2,1,0)(1,1,1)[168]                    : Inf
- ARIMA(3,1,0)(0,1,0)[168]                    : 121452.9
- ARIMA(3,1,0)(1,1,0)[168]                    : Inf
- ARIMA(3,1,0)(0,1,1)[168]                    : Inf
- ARIMA(3,1,0)(1,1,1)[168]                    : Inf
- ARIMA(4,1,0)(0,1,0)[168]                    : 121403.9
- ARIMA(4,1,0)(1,1,0)[168]                    : Inf
- ARIMA(4,1,0)(0,1,1)[168]                    : Inf
- ARIMA(4,1,0)(1,1,1)[168]                    : Inf
- ARIMA(5,1,0)(0,1,0)[168]                    : 121361.7
- ARIMA(5,1,0)(1,1,0)[168]                    : Inf
- ARIMA(5,1,0)(0,1,1)[168]                    : Inf
- ARIMA(5,1,0)(1,1,1)[168]                    : Inf
- ARIMA(5,1,1)(0,1,0)[168]                    : 118580.2
- ARIMA(5,1,1)(1,1,0)[168]                    : Inf
- ARIMA(5,1,1)(0,1,1)[168]                    : Inf
- ARIMA(5,1,1)(1,1,1)[168]                    : Inf
- ARIMA(4,1,1)(0,1,0)[168]                    : 118580.3
- ARIMA(5,1,2)(0,1,0)[168]                    : 118581.5
- ARIMA(4,1,2)(0,1,0)[168]                    : 118553.1
- ARIMA(4,1,2)(1,1,0)[168]                    : Inf
- ARIMA(4,1,2)(0,1,1)[168]                    : Inf
- ARIMA(4,1,2)(1,1,1)[168]                    : Inf
- ARIMA(3,1,2)(0,1,0)[168]                    : 118642.8
- ARIMA(3,1,1)(0,1,0)[168]                    : 119155
-
- Now re-fitting the best model(s) without approximations...
-
- ARIMA(4,1,2)(0,1,0)[168]                    : 122838.5
-
- Best model: ARIMA(4,1,2)(0,1,0)[168]
-
-```{r eval=FALSE, include=FALSE}
-
-autoplot(acf(seis.h.train.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
-autoplot(pacf(seis.h.train.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
-
-qqPlot(fit$residuals)
-gghistogram(fit$residuals)
-Box.test(fit$residuals, type="Box-Pierce")
-fit.forec <- forecast(fit, h=24*10)
-plot(fit.forec, xlim=c(250, 300))
-
-frequency(seis.h.train)
-p<-168;d<-0;q<-0;P<-0;D<-1;Q<-0;
-paste0("ARIMA(", p, ",", d, ",", q, ")(", P, ",", D, ",", Q, ")_", frequency(seis.h.train))
-fit <- Arima(seis.h.train, order=c(p,d,q), seasonal=c(P,D,Q))
-fit
-#Diagnostics
-tsdisplay(residuals(fit), lag.max=168*2)
-
-p<-1;d<-0;q<-2;P<-1;D<-0;Q<-2;
-fit <- Arima(seis.h.train, order=c(p,d,q), seasonal=c(P,D,Q))
-#Diagnostics
-tsdisplay(residuals(fit))
-forecast()
-
-```
-
-```{r mstl on msts}
-seis.h.train.m <- msts(seis.h.train, c(24, 168))
-seis.mstl <- seis.h.train.m %>% mstl(s.window="periodic") #c(15, 7))
-seis.mstl.fc <- forecast(seis.mstl, h=168*18)
-
-autoplot(seis.mstl) + xlab("Week")
-plot(seis.mstl[,2], main="trend")
-lines(trend.s, col="red")
-plot(rowSums(seis.mstl[,3:4]), main="seasonality",type="l")
-lines(seasonality, col="red")
-plot(seis.mstl[,5],  main="residuals")
-lines(residuals, col="red")
-
-plotting.model <- seis.mstl
-plotting.forecast <- seis.mstl.fc
-checkresiduals(plotting.model[,5])
-Box.test(plotting.forecast$model$residuals, type="L")
-residuals.acf <- autoplot(acf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-residuals.pacf <- autoplot(pacf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-plot_grid(residuals.acf, residuals.pacf, nrow=2)
-
-plot_forecast(plotting.forecast) + 
-  autolayer(seis.h.test, series="Original") +
-  xlim(52, 70) + ylim(0, 8750)
-plot_forecast(plotting.forecast) + 
-  autolayer(seis.h.test, series="Original")
-plots <- show_overlap_and_diff(
-  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
-  base.desc = "train series",
-  compare.desc = paste0("refit with ", plotting.forecast$method),
-  overlap.ylab = "Mean hourly movement (nm)",
-  diff.title = "Difference between ts",
-  diff.ylab = "nm", diff.ylab.sec="% over ts span",
-  diff.ylab.sec.accuracy = .01
-)
-plot_grid(plots[[1]], plots[[2]], nrow = 2)
-plots <- show_overlap_and_diff(
-  seis.h.test[1:length(plotting.forecast$mean)], plotting.forecast$mean, xlab=seasonality.xlab,
-  base.desc = "test series",
-  compare.desc = paste0("refit with ", plotting.forecast$method),
-  overlap.ylab = "Mean hourly movement (nm)",
-  diff.title = "Difference between ts",
-  diff.ylab = "nm", diff.ylab.sec="% over ts span",
-  diff.ylab.sec.accuracy = .01
-)
-plot_grid(plots[[1]], plots[[2]], nrow = 2)
-accuracy(plotting.forecast, seis.h.test)
-qqPlot(plotting.model[,5])
-
-bptests <- data.frame(lag=integer(), `p-value`=numeric())
-for (i in 1:10){
-  bptests[nrow(bptests) + 1, ] <- c(i, Box.test(plotting.model[,5], lag = i, fitdf = 0, type = "Box-Pierce")$p.value)
-}
-bptests.plot <- 
-  ggplot(bptests, aes(lag, `p.value`), pch="o") + 
-    geom_point() + ylim(0,1) +
-    labs(title="Box-Pierce tests for lags 1 to 10") +
-    geom_hline(yintercept=0.05, linetype="dotted", color = "blue") +
-    scale_x_continuous(breaks=1:10)
-bptests.plot
-
-```
-
-```{r eval=FALSE, include=FALSE}
-gc()
-if (name == "MEM") {
-  proj_file <- 'proj.mems.RData'
-  save.image(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
-  remove(proj_file)
-}
-```

### 2020-05-23 project.mems.Rmd
divided not-common analysis and some more analysis

@@
 ```
+
+```{r}
+plot(seis.h, xaxt="n", type="l")
+axis.POSIXct(1, at=seq.POSIXt(from=seis.h.start, to=seis.h.end, by = "1 month"), format="%m-%y", cex=0.8)
+
+hist(seis.h.train)
+plot(log(seis.h.train))
+hist(log(seis.h.train))
+train.bc.lambda <- BoxCox.lambda(seis.h.train)
+train.bc <- BoxCox(seis.h.train, train.bc.lambda)
+plot(train.bc)
+gghistogram(train.bc, add.normal = T)
+gg_qqplot(seis.h.train) + 
+        labs(title="Q-Q plot for training dataset") + ylab("Sample quantiles")
+gg_qqplot(log(seis.h.train)) + 
+        labs(title="Q-Q plot for log(training) dataset") + ylab("Sample quantiles")
+gg_qqplot(train.bc) + 
+        labs(title="Q-Q plot for BoxCox(training) dataset") + ylab("Sample quantiles")
+
+train.bc %>% ggtsdisplay(lag.max = 168*5)
+train.bc %>% diff() %>% ggtsdisplay(lag.max = 168*5)
+train.bc %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
+train.bc %>% diff() %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
+train.bc %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)
+train.bc %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,1,0)(0,1,1)[168]
+train.bc %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168) # arima (0,1,0)(0,1,1)[168]
+# acf cuts off after p=168, pacf shows peaks each 168
+train.bc %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)
+```
+
+```{r arima}
+to_test <- list(
+  c(0,1,0,0,1,1)
+)
+
+if (!exists("models")){
+  models <- list()
+}
+for (model in to_test) {
+  # model <- c(4,1,2,0,1,0)
+  fit.name <- paste0("fit", paste0(model, collapse = ""))
+  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
+  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
+  if (!(fit.name %in% names(models))) {
+    models[[fit.name]] <- Arima(seis.h.train, model[1:3], model[4:6], lambda=train.bc.lambda)
+  }
+  if (!(forecast.name %in% names(models))) {
+    models[[forecast.name]] <- forecast(models[[fit.name]], h=168*18, lambda=train.bc.lambda)
+  }
+}
+#models[["fitauto"]] <- auto.arima(train.bc, D=1, approx=F, trace = T)
+#models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*18)  
+#models[["fit311010"]] <- models[["fitauto"]]
+#models[["fit412010"]] <- fit
+#models[["forecast412010"]] <- forecast(models[["fit412010"]], h=168*18)
+
+for (model in to_test) {
+  # model <- c(4,1,2,0,1,0)
+  fit.name <- paste0("fit", paste0(model, collapse = ""))
+  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
+  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
+  if ((fit.name %in% names(models))) {  if ((forecast.name %in% names(models))) {
+    plotting.model <- models[[fit.name]]
+    plotting.forecast <- models[[forecast.name]]
+    checkresiduals(plotting.model)
+    residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+    residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+    print(plot_grid(residuals.acf, residuals.pacf, nrow=2))
+    print(plot_forecast(plotting.forecast) + 
+      autolayer(BoxCox(seis.h.test, train.bc.lambda), series="Original"))
+    plots <- show_overlap_and_diff(
+      seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+      base.desc = "original series",
+      compare.desc = paste0("refit with ", plotting.forecast$method),
+      overlap.ylab = "Mean hourly movement (nm)",
+      diff.title = "Difference between ts",
+      diff.ylab = "nm", diff.ylab.sec="% over ts span",
+      diff.ylab.sec.accuracy = .01
+    )
+
+    print(plot_grid(plots[[1]], plots[[2]], nrow = 2))
+    accuracy(plotting.forecast, seis.h.test)
+    # qqPlot(plotting.model$residuals, main = paste0("Q-Q plot for ", plotting.forecast$method), xlab = "residuals")
+    print(gg_qqplot(plotting.model$residuals) + 
+        labs(title=paste0("Q-Q plot for ", plotting.forecast$method, " residuals")) + ylab("Residuals quantiles")
+    )
+  }}
+}
+
+```
+
+```{r}
+#fit <- auto.arima(seis.h.train, d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10, trace = T, parallel = T)
+```
+
+```{r mstl on msts}
+seis.h.train.m <- msts(seis.h.train, c(24, 168))
+seis.mstl <- seis.h.train.m %>% mstl(s.window="periodic", lambda=train.bc.lambda) #c(15, 7))
+seis.mstl.fc <- forecast(seis.mstl, h=168*18, lambda=train.bc.lambda)
+
+autoplot(seis.mstl) + xlab("Week")
+plot(seis.mstl[,2], main="trend")
+lines(trend.s, col="red")
+plot(rowSums(seis.mstl[,3:4]), main="seasonality",type="l")
+lines(seasonality, col="red")
+plot(seis.mstl[,5],  main="residuals")
+lines(residuals, col="red")
+gg_qqplot(seis.mstl[,5]) + 
+  labs(title=paste0("Q-Q plot for ", seis.mstl.fc$method, " residuals")) + ylab("Residuals quantiles") + xlab("Theoretical quantiles")
+
+plotting.model <- seis.mstl
+plotting.forecast <- seis.mstl.fc
+checkresiduals(plotting.model[,5])
+Box.test(plotting.forecast$model$residuals, type="L")
+residuals.acf <- autoplot(acf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Test") + 
+  autolayer(seis.h.train, series="Train")
+plots <- show_overlap_and_diff(
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "train series",
+  compare.desc = paste0("refit with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+plots <- show_overlap_and_diff(
+  seis.h.test[1:length(plotting.forecast$mean)], plotting.forecast$mean, xlab=seasonality.xlab,
+  base.desc = "test series",
+  compare.desc = paste0("refit with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
+gg_qqplot(plotting.model[,5])
+
+bptests <- data.frame(lag=integer(), `p-value`=numeric())
+for (i in 1:10){
+  bptests[nrow(bptests) + 1, ] <- c(i, Box.test(plotting.model[,5], lag = i, fitdf = 0, type = "Box-Pierce")$p.value)
+}
+bptests.plot <- 
+  ggplot(bptests, aes(lag, `p.value`), pch="o") + 
+    geom_point() + ylim(0,1) +
+    labs(title="Box-Pierce tests for lags 1 to 10") +
+    geom_hline(yintercept=0.05, linetype="dotted", color = "blue") +
+    scale_x_continuous(breaks=1:10)
+bptests.plot
+
+```
+
+```{r eval=FALSE, include=FALSE}
+gc()
+if (name == "MEM") {
+  proj_file <- 'proj.mems.RData'
+  save.image(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
+  remove(proj_file)
+}
+```

### 2020-05-23 project.uccs.Rmd
divided not-common analysis and some more analysis

@@
 ```
+
+```{r arima}
+to_test <- list(
+  #c(0,0,0,0,0,1),  # BAD, Box-Pierce = 0
+  #c(0,0,0,0,0,2),  # BAD, Box-Pierce = 0
+  #c(0,0,0,0,1,1),  # nice, Box-Pierce = 0
+  c(0,0,0,0,1,2),
+  #c(0,0,0,0,1,3),  # CAN'T BE DONE: 3*168 = 504 > 350 (limit of arima on r)
+  #c(0,1,0,0,0,1),  # BAD, Box-Pierce = 0
+  c(0,1,0,0,1,1),
+  c(0,1,0,0,1,2),
+  c(2,0,0,0,1,0),  # hint from auto.arima D=1
+  c(5,0,0,0,1,0),  # hint from auto.arima D=1
+  c(3,1,1,0,1,0),  # hint from auto.arima D=1 d=1
+  c(3,1,1,0,1,1),  # after hint then ACF + PACF
+  c(4,1,2,0,1,0),  # hint from auto.arima d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10
+  c(4,1,2,0,1,1)  # after hint then ACF + PACF
+)
+if (!exists("models")){
+  models <- list()
+}
+for (model in to_test) {
+  # model <- c(4,1,2,0,1,0)
+  fit.name <- paste0("fit", paste0(model, collapse = ""))
+  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
+  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
+  if (!(fit.name %in% names(models))) {
+    models[[fit.name]] <- Arima(seis.h.train, model[1:3], model[4:6])
+  }
+  if (!(forecast.name %in% names(models))) {
+    models[[forecast.name]] <- forecast(models[[fit.name]], h=168*18)
+  }
+}
+#models[["fitauto"]] <- auto.arima(seis.h.train, D=1, approx=F, trace = T)
+#models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*18)  
+#models[["fit311010"]] <- models[["fitauto"]]
+#models[["fit412010"]] <- fit
+#models[["forecast412010"]] <- forecast(models[["fit412010"]], h=168*18)
+
+# 000012 better than 000011, Box-Pierce = 0
+plotting.model <- models[["fit000012"]]
+plotting.forecast <- models[["forecast000012"]]
+# 010011 somehow better somehow worst than 000012, Box-Pierce = 0
+plotting.model <- models[["fit010011"]]
+plotting.forecast <- models[["forecast010011"]]
+# 010012 , Box-Pierce = 
+plotting.model <- models[["fit010012"]]
+plotting.forecast <- models[["forecast010012"]]
+# 311010 , Box-Pierce = 
+plotting.model <- models[["fit311010"]]
+plotting.forecast <- models[["forecast311010"]]
+# 311011 , Box-Pierce = 
+plotting.model <- models[["fit311011"]]
+plotting.forecast <- models[["forecast311011"]]
+
+checkresiduals(plotting.model)
+residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Original")
+plots <- show_overlap_and_diff(
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "original series",
+  compare.desc = "reconstructed series",
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
+
+plotting.model <- models[["fit010012"]]
+plotting.forecast <- models[["forecast010012"]]
+checkresiduals(plotting.model)
+residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Original")
+plots <- show_overlap_and_diff(
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "original series",
+  compare.desc = "reconstructed series",
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
+
+plotting.model <- models[["fit412010"]]
+plotting.forecast <- models[["forecast412010"]]
+checkresiduals(plotting.model)
+residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Original") 
+plots <- show_overlap_and_diff(
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "original series",
+  compare.desc = "reconstructed series",
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
+
+plotting.model <- models[["fit500010"]]
+plotting.forecast <- models[["forecast500010"]]
+checkresiduals(plotting.model)
+residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Original") 
+plots <- show_overlap_and_diff(
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "original series",
+  compare.desc = "reconstructed series",
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
+
+for (model in to_test) {
+  # model <- c(4,1,2,0,1,0)
+  fit.name <- paste0("fit", paste0(model, collapse = ""))
+  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
+  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
+  if ((fit.name %in% names(models))) {  if ((forecast.name %in% names(models))) {
+    plotting.model <- models[[fit.name]]
+    plotting.forecast <- models[[forecast.name]]
+    checkresiduals(plotting.model)
+    residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+    residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+    print(plot_grid(residuals.acf, residuals.pacf, nrow=2))
+    print(plot_forecast(plotting.forecast) + 
+      autolayer(seis.h.test, series="Original") )
+    plots <- show_overlap_and_diff(
+      seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+      base.desc = "original series",
+      compare.desc = paste0("refit with ", plotting.forecast$method),
+      overlap.ylab = "Mean hourly movement (nm)",
+      diff.title = "Difference between ts",
+      diff.ylab = "nm", diff.ylab.sec="% over ts span",
+      diff.ylab.sec.accuracy = .01
+    )
+    print(plot_grid(plots[[1]], plots[[2]], nrow = 2))
+    accuracy(plotting.forecast, seis.h.test)
+  }}
+}
+gg_qqplot(models[["fit500010"]]$residuals)
+```
+
+```{r}
+#fit <- auto.arima(seis.h.train, d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10, trace = T, parallel = T)
+```
+# D=1 m=24
+ Fitting models using approximations to speed things up...
+
+ ARIMA(2,0,2)(1,1,1)[24] with drift         : Inf
+ ARIMA(0,0,0)(0,1,0)[24] with drift         : 111462.7
+ ARIMA(1,0,0)(1,1,0)[24] with drift         : 101954.4
+ ARIMA(0,0,1)(0,1,1)[24] with drift         : Inf
+ ARIMA(0,0,0)(0,1,0)[24]                    : 111460.9
+ ARIMA(1,0,0)(0,1,0)[24] with drift         : 102426.5
+ ARIMA(1,0,0)(2,1,0)[24] with drift         : 101379
+ ARIMA(1,0,0)(2,1,1)[24] with drift         : Inf
+ ARIMA(1,0,0)(1,1,1)[24] with drift         : Inf
+ ARIMA(0,0,0)(2,1,0)[24] with drift         : 110647.6
+ ARIMA(2,0,0)(2,1,0)[24] with drift         : 101332.9
+ ARIMA(2,0,0)(1,1,0)[24] with drift         : 101878.7
+ ARIMA(2,0,0)(2,1,1)[24] with drift         : Inf
+ ARIMA(2,0,0)(1,1,1)[24] with drift         : Inf
+ ARIMA(3,0,0)(2,1,0)[24] with drift         : 101317.9
+ ARIMA(3,0,0)(1,1,0)[24] with drift         : 101858.5
+ ARIMA(3,0,0)(2,1,1)[24] with drift         : Inf
+ ARIMA(3,0,0)(1,1,1)[24] with drift         : Inf
+ ARIMA(4,0,0)(2,1,0)[24] with drift         : 101302.7
+ ARIMA(4,0,0)(1,1,0)[24] with drift         : 101845
+ ARIMA(4,0,0)(2,1,1)[24] with drift         : Inf
+ ARIMA(4,0,0)(1,1,1)[24] with drift         : Inf
+ ARIMA(5,0,0)(2,1,0)[24] with drift         : 101305.4
+ ARIMA(4,0,1)(2,1,0)[24] with drift         : 101304.5
+ ARIMA(3,0,1)(2,1,0)[24] with drift         : 101307.7
+ ARIMA(5,0,1)(2,1,0)[24] with drift         : 101307.7
+ ARIMA(4,0,0)(2,1,0)[24]                    : 101300.7
+ ARIMA(4,0,0)(1,1,0)[24]                    : 101843
+ ARIMA(4,0,0)(2,1,1)[24]                    : Inf
+ ARIMA(4,0,0)(1,1,1)[24]                    : Inf
+ ARIMA(3,0,0)(2,1,0)[24]                    : 101316
+ ARIMA(5,0,0)(2,1,0)[24]                    : 101303.4
+ ARIMA(4,0,1)(2,1,0)[24]                    : 101302.4
+ ARIMA(3,0,1)(2,1,0)[24]                    : 101305.8
+ ARIMA(5,0,1)(2,1,0)[24]                    : 101305.7
+
+ Now re-fitting the best model(s) without approximations...
+
+ ARIMA(4,0,0)(2,1,0)[24]                    : 101586.6
+
+ Best model: ARIMA(4,0,0)(2,1,0)[24]
+
+# d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10 m=168
+ Fitting models using approximations to speed things up...
+
+ ARIMA(2,1,2)(1,1,1)[168]                    : Inf
+ ARIMA(0,1,0)(0,1,0)[168]                    : 122379.8
+ ARIMA(1,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(0,1,1)(0,1,1)[168]                    : Inf
+ ARIMA(0,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(0,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(0,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(1,1,0)(0,1,0)[168]                    : 121638
+ ARIMA(1,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(1,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(2,1,0)(0,1,0)[168]                    : 121468.5
+ ARIMA(2,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(2,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(2,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(3,1,0)(0,1,0)[168]                    : 121452.9
+ ARIMA(3,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(3,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(3,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(4,1,0)(0,1,0)[168]                    : 121403.9
+ ARIMA(4,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(4,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(4,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(5,1,0)(0,1,0)[168]                    : 121361.7
+ ARIMA(5,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(5,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(5,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(5,1,1)(0,1,0)[168]                    : 118580.2
+ ARIMA(5,1,1)(1,1,0)[168]                    : Inf
+ ARIMA(5,1,1)(0,1,1)[168]                    : Inf
+ ARIMA(5,1,1)(1,1,1)[168]                    : Inf
+ ARIMA(4,1,1)(0,1,0)[168]                    : 118580.3
+ ARIMA(5,1,2)(0,1,0)[168]                    : 118581.5
+ ARIMA(4,1,2)(0,1,0)[168]                    : 118553.1
+ ARIMA(4,1,2)(1,1,0)[168]                    : Inf
+ ARIMA(4,1,2)(0,1,1)[168]                    : Inf
+ ARIMA(4,1,2)(1,1,1)[168]                    : Inf
+ ARIMA(3,1,2)(0,1,0)[168]                    : 118642.8
+ ARIMA(3,1,1)(0,1,0)[168]                    : 119155
+
+ Now re-fitting the best model(s) without approximations...
+
+ ARIMA(4,1,2)(0,1,0)[168]                    : 122838.5
+
+ Best model: ARIMA(4,1,2)(0,1,0)[168]
+
+```{r eval=FALSE, include=FALSE}
+
+autoplot(acf(seis.h.train.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
+autoplot(pacf(seis.h.train.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
+
+gg_qqplot(fit$residuals)
+gghistogram(fit$residuals)
+Box.test(fit$residuals, type="Box-Pierce")
+fit.forec <- forecast(fit, h=24*10)
+plot(fit.forec, xlim=c(250, 300))
+
+frequency(seis.h.train)
+p<-168;d<-0;q<-0;P<-0;D<-1;Q<-0;
+paste0("ARIMA(", p, ",", d, ",", q, ")(", P, ",", D, ",", Q, ")_", frequency(seis.h.train))
+fit <- Arima(seis.h.train, order=c(p,d,q), seasonal=c(P,D,Q))
+fit
+#Diagnostics
+tsdisplay(residuals(fit), lag.max=168*2)
+
+p<-1;d<-0;q<-2;P<-1;D<-0;Q<-2;
+fit <- Arima(seis.h.train, order=c(p,d,q), seasonal=c(P,D,Q))
+#Diagnostics
+tsdisplay(residuals(fit))
+forecast()
+
+```
+
+```{r mstl on msts}
+seis.h.train.m <- msts(seis.h.train, c(24, 168))
+seis.mstl <- seis.h.train.m %>% mstl(s.window="periodic") #c(15, 7))
+seis.mstl.fc <- forecast(seis.mstl, h=168*18)
+
+autoplot(seis.mstl) + xlab("Week")
+plot(seis.mstl[,2], main="trend")
+lines(trend.s, col="red")
+plot(rowSums(seis.mstl[,3:4]), main="seasonality",type="l")
+lines(seasonality, col="red")
+plot(seis.mstl[,5],  main="residuals")
+lines(residuals, col="red")
+
+plotting.model <- seis.mstl
+plotting.forecast <- seis.mstl.fc
+checkresiduals(plotting.model[,5])
+Box.test(plotting.forecast$model$residuals, type="L")
+residuals.acf <- autoplot(acf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Original") +
+  xlim(52, 70) + ylim(0, 8750)
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Original")
+plots <- show_overlap_and_diff(
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "train series",
+  compare.desc = paste0("refit with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+plots <- show_overlap_and_diff(
+  seis.h.test[1:length(plotting.forecast$mean)], plotting.forecast$mean, xlab=seasonality.xlab,
+  base.desc = "test series",
+  compare.desc = paste0("refit with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
+gg_qqplot(plotting.model[,5])
+
+bptests <- data.frame(lag=integer(), `p-value`=numeric())
+for (i in 1:10){
+  bptests[nrow(bptests) + 1, ] <- c(i, Box.test(plotting.model[,5], lag = i, fitdf = 0, type = "Box-Pierce")$p.value)
+}
+bptests.plot <- 
+  ggplot(bptests, aes(lag, `p.value`), pch="o") + 
+    geom_point() + ylim(0,1) +
+    labs(title="Box-Pierce tests for lags 1 to 10") +
+    geom_hline(yintercept=0.05, linetype="dotted", color = "blue") +
+    scale_x_continuous(breaks=1:10)
+bptests.plot
+
+```
+
+```{r eval=FALSE, include=FALSE}
+gc()
+if (name == "MEM") {
+  proj_file <- 'proj.mems.RData'
+  save.image(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
+  remove(proj_file)
+}
+```

### 2020-05-26 project.Rmd
correctly use train set

@@
 if (exists("proj_file")) {
-  #load(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
+  load(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
   # save.image(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
@@
 remove(seis.after, seis.before, seis.dst.fix)
+
+# invalid_values <- seis.h$mean_movement_s > 200
+# seis.h <-
+#   seis.h %>%
+#   mutate(mean_movement_s=replace(mean_movement_s, invalid_values, NA)) %>%
+#   drop_na()
+# remove(invalid_values)
+
 gc()
@@
 seasonality.xlab <- "Days"
-seasonality.aggregate <- function(ds) na.aggregate(seis.h.ts, function(x) floor(x/14))
+seasonality.aggregate <- function(ds) na.aggregate(ds, function(x) floor(x/14))
 seasonality.period <- 168
@@
 seis.test.start <- seis.test.split
-seis.test.end <- as_datetime('2020-05-04T00:00:00') - seconds(1)
-seis.h.test <- build_ts(seis.h, periods = c(seasonality.period), start = seis.h.start, end = seis.test.end, force_ms = T)
+seis.test.end <- seis.test.start + days(7*18) - seconds(1)
+seis.h.test <- build_ts(seis.h, periods = c(seasonality.period), start = seis.train.start, end = seis.test.end, force_ms = T)
 seis.h.test <- window(seis.h.test, start = c(end(seis.h.train)[1] + 1, 1))
@@
 library(spec)
-pgram <- spec.pgram(seis.h.train, main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
+pgram <- spec.pgram(build_ts(seis.h, start = seis.train.start, end = seis.train.end), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
 topfreq <- pgram$freq[order(pgram$spec, decreasing = T)][1:100]
 topseas <- 1/topfreq
-topseas.h <- topseas/1
+topseas.h <- topseas/3600
 
@@
 top95pc <- pgram$spec[filter_freq][order(top95pc, decreasing = T)]
+cat("SD ratio\n")
+(top95pc - mean(pgram$spec))/sd(pgram$spec)
 top95seas.s <- 1/top95freq
+cat("Raw frequency by unit\n")
 top95seas.s
-top95seas.h <- top95seas.s/1
-top95pc/sd(pgram$spec)
+top95seas.h <- top95seas.s/3600
+cat("Frequency by unit\n")
 top95seas.h
@@
 ```{r plot data with (p)acf}
-seis.h.24 <- build_ts(seis.h, periods = c(24), start = seis.h.start, end = seis.h.end)
+seis.h.24 <- build_ts(seis.h.train, periods = c(24))
 seasonplot(seis.h.24, type="l", col=rainbow(7))
-legend("topleft", legend=weekdays(seis.h.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
+legend("topleft", legend=weekdays(seis.train.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
 
 plot_grid(
-  autoplot(seis.h.ts, main = "Hourly seismic movement", xlab = seasonality.xlab, ylab = "mean movement (nm)"),
-  gghistogram(seis.h.ts, add.normal = T) + xlab("mean movement (nm)"),
+  autoplot(seis.h.train, main = "Hourly seismic movement", xlab = seasonality.xlab, ylab = "mean movement (nm)"),
+  gghistogram(seis.h.train, add.normal = T) + xlab("mean movement (nm)"),
   nrow = 2)
 plot_grid(
-  autoplot(acf(seis.h.ts, na.action=na.pass, plot=F, lag.max=seasonality.period)) + scale_x_continuous(),
-  autoplot(pacf(seis.h.ts, na.action=na.pass, plot=F, lag.max=seasonality.period)) + scale_x_continuous(),
+  autoplot(acf(seis.h.train, na.action=na.pass, plot=F, lag.max=seasonality.period)) + scale_x_continuous(),
+  autoplot(pacf(seis.h.train, na.action=na.pass, plot=F, lag.max=seasonality.period)) + scale_x_continuous(),
   nrow = 2)
@@
 ```{r decompose}
-dec.addi <- decompose(seasonality.aggregate(seis.h.ts))
+dec.addi <- decompose(seasonality.aggregate(seis.h.train))
 autoplot(dec.addi) + labs(title="Decomposition of additive time series", subtitle = paste0("Seasonality ", seasonality.period))
-dec.multi <- decompose(log(seasonality.aggregate(seis.h.ts)))
+dec.multi <- decompose(log(seasonality.aggregate(seis.h.train)))
 autoplot(dec.multi) + labs(title="Decomposition of multiplicative time series (log-scale additive)", subtitle = paste0("Seasonality ", seasonality.period))
@@
 weights <- rep(1/(2*p+1), times=2*p+1)
-trend <- stats::filter(seasonality.aggregate(seis.h.ts), sides=2, filter = weights)
+trend <- stats::filter(seasonality.aggregate(seis.h.train), sides=2, filter = weights)
 ```
@@
   errorstring <- paste0("MSE=", mse, "    MAE=", mae, "    RMSE=", rmse, "    R2=", R2)
-  print(errorstring)
+  cat("Errors: ", compare.desc, "\n")
+  cat("\t","MSE=", mse, "    MAE=", mae, "    RMSE=", rmse, "    R2=", R2, "\n")
 
@@
 weights.s <- c(0.5, rep(1,f-1), 0.5)/f
-trend.s <- stats::filter(seasonality.aggregate(seis.h.ts), side=2, filter=weights.s)
+trend.s <- stats::filter(seasonality.aggregate(seis.h.train), side=2, filter=weights.s)
 
@@
 ```{r trend figure}
-detrended <- seasonality.aggregate(seis.h.ts)-trend.s
+detrended <- seasonality.aggregate(seis.h.train)-trend.s
 detrended <- window(detrended, end = c(end(detrended)[1], seasonality.period), extend=T)
@@
 ```{r residuals}
-residuals <- seasonality.aggregate(seis.h.ts) - trend.s - seasonality
+residuals <- seasonality.aggregate(seis.h.train) - trend.s - seasonality
 
@@
 plots <- show_overlap_and_diff(
-  seis.h.ts, ts, xlab=seasonality.xlab,
+  seis.h.train, ts, xlab=seasonality.xlab,
   base.desc = "original series",
-  compare.desc = "reconstructed series",
+  compare.desc = "reconstructed serie (manual)",
   overlap.ylab = "Mean hourly movement (nm)",
@@
 plots <- show_overlap_and_diff(
-  seis.h.ts, dec.addi$trend+dec.addi$seasonal+dec.addi$random, xlab=seasonality.xlab,
+  seis.h.train, dec.addi$trend+dec.addi$seasonal+dec.addi$random, xlab=seasonality.xlab,
   base.desc = "original series",
-  compare.desc = "reconstructed series",
+  compare.desc = "reconstructed serie (decompose additive)",
   overlap.ylab = "Mean hourly movement (nm)",
@@
 
-aggregated <- seasonality.aggregate(seis.h.ts)
+aggregated <- seasonality.aggregate(seis.h.train)
 season.24 <- trend_season(aggregated, 24)

### 2020-05-26 project.mems.Rmd
correctly use train set

--- 2020-05-22 project.mems.Rmd
@@
 ```
+
+```{r transformation}
+plot(seis.h, xaxt="n", type="l")
+axis.POSIXct(1, at=seq.POSIXt(from=seis.h.start, to=seis.h.end, by = "1 month"), format="%m-%y", cex=0.8)
+
+hist(seis.h.train)
+plot(log(seis.h.train))
+hist(log(seis.h.train))
+train.bc.lambda <- BoxCox.lambda(seis.h.train)
+train.bc <- BoxCox(seis.h.train, train.bc.lambda)
+plot(train.bc)
+gghistogram(train.bc, add.normal = T)
+gg_qqplot(seis.h.train) + 
+        labs(title="Q-Q plot for training dataset") + ylab("Sample quantiles")
+gg_qqplot(log(seis.h.train)) + 
+        labs(title="Q-Q plot for log(training) dataset") + ylab("Sample quantiles")
+gg_qqplot(train.bc) + 
+        labs(title="Q-Q plot for BoxCox(training) dataset") + ylab("Sample quantiles")
+```
+
+```{r acf-pacf}
+train.bc %>% ggtsdisplay(lag.max = 168*5)
+train.bc %>% diff() %>% ggtsdisplay(lag.max = 168*5)
+train.bc %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
+train.bc %>% diff() %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
+train.bc %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)
+train.bc %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,1,0)(0,1,1)[168]
+train.bc %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168) # arima (0,1,0)(0,1,1)[168]
+# acf cuts off after p=168, pacf shows peaks each 168
+train.bc %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)
+```
+
+```{r arima}
+to_test <- list(
+  c(0,1,0,0,1,1)
+)
+
+if (!exists("models")){
+  models <- list()
+}
+for (model in to_test) {
+  # model <- c(4,1,2,0,1,0)
+  fit.name <- paste0("fit", paste0(model, collapse = ""))
+  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
+  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
+  if (!(fit.name %in% names(models))) {
+    models[[fit.name]] <- Arima(seis.h.train, model[1:3], model[4:6], lambda=train.bc.lambda)
+  }
+  if (!(forecast.name %in% names(models))) {
+    models[[forecast.name]] <- forecast(models[[fit.name]], h=168*18, lambda=train.bc.lambda)
+  }
+}
+#models[["fitauto"]] <- auto.arima(train.bc, D=1, approx=F, trace = T)
+#models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*18)  
+#models[["fit311010"]] <- models[["fitauto"]]
+#models[["fit412010"]] <- fit
+#models[["forecast412010"]] <- forecast(models[["fit412010"]], h=168*18)
+
+for (model in to_test) {
+  # model <- c(4,1,2,0,1,0)
+  fit.name <- paste0("fit", paste0(model, collapse = ""))
+  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
+  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
+  if ((fit.name %in% names(models))) {  if ((forecast.name %in% names(models))) {
+    plotting.model <- models[[fit.name]]
+    plotting.forecast <- models[[forecast.name]]
+    checkresiduals(plotting.model)
+    residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+    residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+    print(plot_grid(residuals.acf, residuals.pacf, nrow=2))
+    print(plot_forecast(plotting.forecast) + 
+      autolayer(seis.h.test, series="Original"))
+    plots <- show_overlap_and_diff(
+      seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+      base.desc = "original series",
+      compare.desc = paste0("refit with ", plotting.forecast$method),
+      overlap.ylab = "Mean hourly movement (nm)",
+      diff.title = "Difference between ts",
+      diff.ylab = "nm", diff.ylab.sec="% over ts span",
+      diff.ylab.sec.accuracy = .01
+    )
+
+    print(plot_grid(plots[[1]], plots[[2]], nrow = 2))
+    accuracy(plotting.forecast, seis.h.test)
+    # qqPlot(plotting.model$residuals, main = paste0("Q-Q plot for ", plotting.forecast$method), xlab = "residuals")
+    print(gg_qqplot(plotting.model$residuals) + 
+        labs(title=paste0("Q-Q plot for ", plotting.forecast$method, " residuals")) + ylab("Residuals quantiles")
+    )
+  }}
+}
+
+```
+
+```{r arima on seasonality}
+autoplot(residuals+trend.s)
+residuals.bc.lambda <- BoxCox.lambda(residuals+trend.s)
+residuals.bc <- BoxCox(residuals+trend.s, residuals.bc.lambda)
+residuals.bc %>% ggtsdisplay(lag.max = 168*5)
+residuals.bc %>% diff() %>% ggtsdisplay(lag.max = 168*5)
+residuals.bc %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
+residuals.bc %>% diff() %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
+residuals.bc %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)
+residuals.bc %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,1,0)(0,1,1)[168]
+residuals.bc %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168) # arima (0,1,0)(0,1,1)[168]
+# acf cuts off after p=168, pacf shows peaks each 168
+residuals.bc %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)
+
+gg_qqplot(residuals) + 
+  labs(title=paste0("Q-Q plot for ", plotting.forecast$method, " residuals")) + ylab("Residuals quantiles")
+
+a <- Arima(residuals, order = c(1,0,0), seasonal = c(0,0,0), lambda=residuals.bc.lambda)
+afor <- forecast(a, h=168*18, lambda=residuals.bc.lambda)
+
+plotting.model <- a
+plotting.forecast <- afor
+
+autoplot(plotting.model) + xlab("Week")
+
+gg_qqplot(plotting.model$residuals) + 
+  labs(title=paste0("Q-Q plot for ", plotting.forecast$method, " residuals")) + ylab("Residuals quantiles")
+checkresiduals(plotting.model$residuals)
+Box.test(plotting.forecast$model$residuals, type="L")
+residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Test") + 
+  autolayer(seis.h.train, series="Train")
+plots <- show_overlap_and_diff(
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "train series",
+  compare.desc = paste0("refit with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+plots <- show_overlap_and_diff(
+  seis.h.test[1:length(plotting.forecast$mean)], plotting.forecast$mean, xlab=seasonality.xlab,
+  base.desc = "test series",
+  compare.desc = paste0("refit with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
+
+```
+
+```{r auto arima}
+#fit <- auto.arima(seis.h.train, d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10, trace = T, parallel = T)
+```
+
+```{r mstl on msts}
+seis.h.train.m <- msts(seis.h.train, c(24, 168))
+seis.mstl <- seis.h.train.m %>% mstl(s.window="periodic", lambda=train.bc.lambda) #c(15, 7))
+seis.mstl.fc <- forecast(seis.mstl, h=168*18, lambda=train.bc.lambda)
+
+plot(InvBoxCox(seis.mstl[,2], lambda=train.bc.lambda), main="trend")
+lines(trend.s, col="red")
+plot(InvBoxCox(rowSums(seis.mstl[,3:4]), lambda=train.bc.lambda), main="seasonality",type="l")
+lines(seasonality, col="red")
+plot(InvBoxCox(seis.mstl[,5], lambda=train.bc.lambda),  main="residuals")
+lines(residuals, col="red")
+
+plotting.model <- seis.mstl
+plotting.forecast <- seis.mstl.fc
+
+autoplot(plotting.model) + xlab("Week")
+
+gg_qqplot(plotting.model[,5]) + 
+  labs(title=paste0("Q-Q plot for ", seis.mstl.fc$method, " residuals")) + ylab("Residuals quantiles")
+checkresiduals(plotting.model[,5])
+Box.test(plotting.forecast$model$residuals, type="L")
+residuals.acf <- autoplot(acf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Test") + 
+  autolayer(seis.h.train, series="Train")
+plots <- show_overlap_and_diff(
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "train series",
+  compare.desc = paste0("refit with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+plots <- show_overlap_and_diff(
+  seis.h.test[1:length(plotting.forecast$mean)], plotting.forecast$mean, xlab=seasonality.xlab,
+  base.desc = "test series",
+  compare.desc = paste0("refit with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
+
+bptests <- data.frame(lag=integer(), `p-value`=numeric())
+for (i in 1:10){
+  bptests[nrow(bptests) + 1, ] <- c(i, Box.test(plotting.model[,5], lag = i, fitdf = 0, type = "Box-Pierce")$p.value)
+}
+bptests.plot <- 
+  ggplot(bptests, aes(lag, `p.value`), pch="o") + 
+    geom_point() + ylim(0,1) +
+    labs(title="Box-Pierce tests for lags 1 to 10") +
+    geom_hline(yintercept=0.05, linetype="dotted", color = "blue") +
+    scale_x_continuous(breaks=1:10)
+bptests.plot
+
+```
+
+```{r save RData, include=FALSE}
+gc()
+proj_file <- 'proj.mems.RData'
+save.image(proj_file)
+```

### 2020-05-29 draft.tex
working on draft

\documentclass[12pt]{article}

% BY MINE REPORTS
\usepackage{fancyhdr}
\usepackage{lastpage} % last page ref
\usepackage{array}
\usepackage[table]{xcolor}
\usepackage{scrextend}
\usepackage{enumitem}
\usepackage{caption} % for image/tables/... captions
\usepackage{fancyvrb} % for centered verbatim
%\usepackage{hyperref} % for links in refs
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[round, sort, numbers]{natbib}
\bibliographystyle{unsrtnat}
% /BY MINE REPORTS

% BY VARIN
\usepackage[sc]{mathpazo}
\usepackage{amsthm,amsmath,amssymb}
\usepackage{titlesec}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
\setcounter{secnumdepth}{2}
\setcounter{tocdepth}{2}
\usepackage{url}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=true,bookmarksopen=true,bookmarksopenlevel=2,
 breaklinks=false,pdfborder={0 0 1},backref=false,colorlinks=false]
 {hyperref}
\hypersetup{
 pdfstartview={XYZ null null 1}}
\usepackage{breakurl}
% /BY VARIN

% BY MINE REPORTS
\renewcommand{\labelenumii}{\theenumi\theenumii.}
\renewcommand{\theenumii}{.\arabic{enumii}}
\renewcommand{\labelenumiii}{\theenumi\theenumii\theenumiii.}
\renewcommand{\theenumiii}{.\arabic{enumiii}}

\newtheorem*{theorem}{Theorem}
\newtheorem*{proposition}{Proposition}

\pagestyle{fancy}
\fancyhf{}
% /BY MINE REPORTS

% FOR KNITR CHILD
%\usepackage[]{graphicx} %% used before
\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
	\ifdim\Gin@nat@width>\linewidth
	\linewidth
	\else
	\Gin@nat@width
	\fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
	\def\at@end@of@kframe{}%
	\ifinner\ifhmode%
	\def\at@end@of@kframe{\end{minipage}}%
\begin{minipage}{\columnwidth}%
	\fi\fi%
	\def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
		\colorbox{shadecolor}{##1}\hskip-\fboxsep
		% There is no \\@totalrightmargin, so:
		\hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
	\MakeFramed {\advance\hsize-\width
		\@totalleftmargin\z@ \linewidth\hsize
		\@setminipage}}%
{\par\unskip\endMakeFramed%
	\at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% / FOR KNITR CHILD

\renewcommand{\theequation}{\roman{equation}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\p}{p}
\DeclareMathOperator*{\cov}{cov}
\DeclareMathOperator*{\var}{var}
\DeclareMathOperator*{\diag}{diag}
\DeclareMathOperator{\Bernoulli}{Bernoulli}
\DeclareMathOperator{\given}{\vert}
\newcommand*\mean[1]{\overline{#1}}
\DeclareMathOperator*{\simf}{sim}

\title{Seismic Time Series Analysis\\Statistics for Spatio-Temporal Data\\\textit{-- project draft --}}
\author{Marco Ciotola, 848222}
\date{\today}

\begin{document}
\pagenumbering{gobble}
\maketitle
\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=100px]{ca_foscari_logo.png}
	\end{center}
\end{figure}
\vfill Prof. Isadora Antoniano Villalobos - CM0477
\hfill A.Y. 2019/2020
\newpage
\tableofcontents \clearpage
\fancyfoot[RO,LE]{Page \thepage \hspace{1pt} of \pageref{LastPage}}
\pagenumbering{arabic}
% /BY MINE REPORTS

% BY VARIN
%\setlength{\parindent}{0pt}
% /BY VARIN

\section{Introduction}

\paragraph{Motivation}
Thanks to the recent coronavirus outbreak worldwide, we currently have data about many different phoenomena from periods with standard human activity, and from a period with reduced human activity.

Ispired by an article published during the initial phases of the Belgium-wide lockdown \cite{NatureCoronavirusSeismic}, we want to assess whether the lockdown period can be correlated with a visible change in seismic activity.
It is important to notice that some industrial events, such as mine explosions, are already mapped by seismologic institutions since they interfere with seismic surveys \cite{OtherSeismicEvents}, sometimes comparable with natural seismic events, depending on the distance from the seismograph.

\paragraph{Data source and description}
The Royal Observatory of Belgium (ROB) gathers and publishes \cite{RoyalDataPolicy,Data20200402} data about vertical ground displacement from 9 stations through the country, linking them to seismic events whenever possible \cite{DataEvents20200402}. They also offer a data visualization tool for the same data \cite{DataVisualization20200402}.

The vertical ground displacement is represented in by the minimum and maximum displacement of the ground in the vertical axis, with $nm$ as unit of measure. This can easily be transformed in total ground displacement for that second by their difference.
For each day, a pair per second is available (86400 per day, around 30 million per year).

ROB collects data from 9 stations. Of those, the following 6 have interesting positions, related to possible correlations with human activity:
\begin{itemize}
	\item \textbf{Uccle}, \textbf{Sart-Tilman}, and \textbf{Ében-Émael} are near a city (Bruxelles, Liege, and Maastricht).
	\item \textbf{Ostenda} This station is near the city of Ostenda, that borders with the sea.
	\item \textbf{Membach} and \textbf{Dourbes} are near a natural park.
\end{itemize}

Data collected from near a city and from near a natural park could be selected, in order to compare possible seasonality results and differences from before and after the lockdown. %Also data from \textbf{Ostenda} station could be interesting, given the possible sea influence.
Considered these factors, we selected the \textbf{Uccle} and \textbf{Membach} stations.

It is important to notice that the \textit{day}ly data in each file is considered in UTC time, an optimal choice to publish data in a machine-readable way, even if this could cause some issues while interpreting the results.

\subsection{Data cleanup and transformation}
\paragraph{Data gathering and cleanup}
The data is provided with a CSV per day, with only two columns related to the min-max pair \cite{Data20200402}. This causes two problems:
\begin{enumerate}
	\item Before merging different days, we need to add a column with the complete datetime for each row
	\item Days with some \texttt{NA} values cannot skip rows, so we will find placeholder values that indicate them $(\min, \max) \in \{(0,0), (-1,1)\}$
\end{enumerate}

Data from each station have different domains, as distinct equipment can register a specific amplitude of the seismic movements. In fact, we found some values outside the domains, transformed in \texttt{NA}.

In particular, we find that out of $\approx50$ million seconds, from 27-10-2018 to 28-05-2020, only 0.67\% (Uccle) and 0.37\% (Membach) have invalid values. We will see later how they are distributed for each station.

% Uccle, Bruxelles (UCCS station)
% Domain values: -350000 to 350000 nm
% Data from 2018-10-27 00:00:00 to 2020-05-29 00:12:07
% Removing 7 values out of domain, 28169 explicit NA
% 49779920 present values, 334345 missing points ( 99.33 % / 0.67 % )

% Membach (MEM station)
% Domain values: -7200 to 7200 nm
% Data from 2018-10-27 00:00:00 to 2020-05-29 00:09:55
% Removing 8 values out of domain, 7018 explicit NA
% 49929756 present values, 184421 missing points ( 99.63 % / 0.37 % )

\paragraph{Data aggregation and initial transformation}
Since we are not interested in second-precise analysis, and it would be computationally expensive to deal with the whole dataset, we are going to aggregate data by hour. The aggregation will get us around 14 thousand points, that can still result in interpretable models.

While performing this operation, we should consider the correct timezone: any seasonal analysis dealing with UTC data would be out of phase between the half-year when Belgium follows the UTC+01 timezone, and the half-year when it follows the UTC+02 timezone. With the help of the \texttt{lubridate} R package we converted to the actual timezones, and then we kept the actual objects without being timezone explicit, because all the methods we are going to use are smart enough to work on the UTC conversion of datetimes. This transformation made it possible to have, for example, the time 6:00 in the timeseries to represent the local time in Belgium (\textit{Europe/Bruxelles}), without being influenced by the actual timezone on that specific day.

The aggregation was performed considering the mean max-min difference, for the following reasons:
\begin{itemize}
	\item When an hour contains some missing seconds, these do not heavily affect the aggregation result, in contrast with summing the differences
	\item Seismic events, such as quakes and mine explosions, normally last some minutes at most, and will affect the result only slightly, rather than keeping the hourly maximum
	\item Normally, maximum values in defined period of times do not follow the Normal distribution, while some models that we are going to use assume the data follow a Normal distribution
\end{itemize}

\section{Analysis and decomposition}
\paragraph{Missing values analysis}

% Uccle, Bruxelles (UCCS station)
% Domain values: -350000 to 350000 nm
% Data from 2018-10-27 00:00:00 to 2020-05-29 00:12:07
% Removing 7 values out of domain, 28169 explicit NA
% 49779920 present values, 334345 missing points ( 99.33 % / 0.67 % )
% 13831 present values, 90 missing points (99.35% / 0.65%)

% Membach (MEM station)
% Domain values: -7200 to 7200 nm
% Data from 2018-10-27 00:00:00 to 2020-05-29 00:09:55
% Removing 8 values out of domain, 7018 explicit NA
% 49929756 present values, 184421 missing points ( 99.63 % / 0.37 % )
% 13872 present values, 49 missing points (99.65% / 0.35%)

After aggregating by hour, the percentages of missing values remain quite stable. As can be seen from figures \ref{missing-values-seconds:uccs} and \ref{missing-values-seconds:mems}, in both stations missing values are grouped in few consecutive periods.
To deal with them when R methods need complete timeseries, we will define different completion strategies depending on the seasonality we will be studying, always based on the R function \texttt{na.aggregate}.

\begin{figure}
	\caption{Missing values representation}
	\label{missing-values-seconds}
	\begin{subfigure}{.5\linewidth}
		\caption{Uccle station}
		\label{missing-values-seconds:uccs}
		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/missing values hour analysis-1.pdf}
	\end{subfigure}
	\begin{subfigure}{.5\linewidth}
		\caption{Membach station}
		\label{missing-values-seconds:mems}
		\includegraphics[width=\linewidth]{project.mems_files/figure-latex/missing values hour analysis-1.pdf}
	\end{subfigure}
\end{figure}

In particular, after a visual analysis after decomposition, we can suppose the 90 hours gap from the Uccle station is related to some changes to the seismic equipment, that resulted in slightly more precise measurements.

\paragraph{Dataset division}
As best practices suggest, we are going to divide the dataset into training and testing datasets. 
Since one of our objectives is to assess any significant difference from before and after the Belgium lockdown (on March, 14$^{th}$ the `soft' lockdown, while on March, 18$^{th}$ the complete lockdown)

\paragraph*{Hypotesis of analysis}

\begin{enumerate}
	\item Download data from start of March 2019 to present day.
	\item Aggregate the data in slots of 30 minutes or 1 hour, keeping the maximum ground displacements in both negative and positive directions to maintain the same variables meaning.
	\item Compute the total ground displacement for each result point.
	\item Verify that we have a stationary time series (we expect that for the meaning that it brings: ground is not moving up or down).
	\item Extract through a linear filter (probably MA) different seasonalities and residuals.
\end{enumerate}

The extraction of seasonalities is the focus point of this work: as \cite{NatureCoronavirusSeismic} seems to conclude, there is a clear weekly seasonality in at least the \textbf{Uccle} data. We should test against different seasonalities through the stations, extracting more at once if the residuals suggest their presence.

Fitted models with these seasonalities, only from data before the Belgium lockdowns (on March, 14th the "soft" lockdown, while on March, 18th the complete lockdown), should let us forecast data for after the lockdown and test for statistical difference.

Also, comparison between models and seasonalities from chosen stations could be done, to assess any statistical difference between apparent human impact registered from places with different standard human activity.

Eventually, some attention could be given to events registered through the analysed time.

\newpage
\bibliography{project}

\end{document}

### 2020-05-29 project.Rmd
working on draft

@@
 setwd("/media/data/Documents/Uni/spatiotemp/villalobos/project/")
-if (!exists("domain")) {
-  # UCCS
-  domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
-  # MEMS
-  domain <- c(-7200,7200); pattern <- "MEM.*\\.csv"; name <- "MEM"; description <- "Membach"; proj_file <- 'proj.mems.RData'
-}
 if (exists("proj_file")) {
@@
 # Data preparation
-```{r dataset description, echo=FALSE}
+```{r "dataset description", echo=FALSE}
 long_description <- description %s+% " (" %s+% name %s+% " station)"
-long_description
-"Domain values: " %s+% domain[1] %s+% " to " %s+% domain[2] %s+% " nm"
-"Data from " %s+% format.POSIXct(min(seis$utc_second), "%Y-%m-%d %H:%M:%S") %s+% " to " %s+% format.POSIXct(max(seis$utc_second), "%Y-%m-%d %H:%M:%S")
+cat(long_description, "\n")
+cat("Domain values:", domain[1], "to", domain[2], "nm\n")
+cat("Data from", format.POSIXct(min(seis$utc_second), "%Y-%m-%d %H:%M:%S"), "to", format.POSIXct(max(seis$utc_second), "%Y-%m-%d %H:%M:%S"), "\n")
 ```
@@
 invalid_values <- (seis$min == -1 & seis$max == 1) | (seis$min == 0 & seis$max == 0) | out_domain
-"Removing " %s+% sum(out_domain) %s+% " values out of domain, " %s+% (sum(invalid_values) - sum(out_domain)) %s+% " explicit NA"
+cat("Removing", sum(out_domain), "values out of domain,", (sum(invalid_values) - sum(out_domain)), "explicit NA\n")
 seis <-
@@
 seis.missing <- seis.total - seis.present
-  seis.present %s+% " present values, " %s+% seis.missing %s+% " missing points (" %s+% print_dec(seis.present / seis.total * 100, 2) %s+% "% / " %s+% print_dec(seis.missing / seis.total * 100, 2) %s+% "%)"
+cat(seis.present, "present values,", seis.missing, "missing points (", print_dec(seis.present / seis.total * 100, 2), "% /", print_dec(seis.missing / seis.total * 100, 2), "% )\n")
 
@@
      main="Consecutive hours of missing values",
-     ylab="Hours", xlab="Interval start",
+     ylab="Hours", xlab="Interval start (mm-'yy)",
      type="l", lwd=10,
@@
      )
-axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-%y", cex=0.8)
+axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-'%y", cex=0.8)
 text(x=blocks$start, y=hour(blocks$interval), labels = paste0(format(blocks$start, "%y-%m-%d %H:%M")), cex=0.9, pos=2)
@@
 
+```{r training data transformation plots}
+month.starts <- seq.POSIXt(from=floor_date(seis.train.start, 'month'), to=ceiling_date(seis.train.end, 'month'), by = "1 month")
+hours.from.train.start <- as.numeric(month.starts-seis.train.start)*24+168
+
+autoplot(seis.h.train, type="l") +
+  scale_x_continuous(breaks=hours.from.train.start/168, labels=format(month.starts, "%m-%y")) +
+  ylab("mean hourly movement (nm)") + xlab(NULL) + ggtitle("Training set") + labs(caption = "Data in original scale")
+gghistogram(seis.h.train, add.normal = T) + xlab("mean hourly movement (nm)") + labs(caption = "Data in original scale")
+gg_qqplot(seis.h.train) + 
+  labs(title="Q-Q plot for training dataset") + ylab("Sample quantiles") + labs(caption = "Data in original scale")
+
+autoplot(log(seis.h.train), type="l") +
+  scale_x_continuous(breaks=hours.from.train.start/168, labels=format(month.starts, "%m-%y")) +
+  ylab("mean hourly movement (nm)") + xlab(NULL) + ggtitle("Training set") + labs(caption = "Data in log-scale")
+gghistogram(log(seis.h.train), add.normal = T) + xlab("mean hourly movement (nm)") + labs(caption = "Data in log-scale")
+gg_qqplot(log(seis.h.train)) + 
+  labs(title="Q-Q plot for training dataset") + ylab("Sample quantiles") + labs(caption = "Data in log-scale")
+
+train.bc.lambda <- BoxCox.lambda(seis.h.train)
+train.bc <- BoxCox(seis.h.train, train.bc.lambda)
+autoplot(train.bc, type="l") +
+  scale_x_continuous(breaks=hours.from.train.start/168, labels=format(month.starts, "%m-%y")) +
+  ylab("mean hourly movement (nm)") + xlab(NULL) + ggtitle("Training set") + labs(caption = paste0("Data Box-Cox transformed with lambda = ", train.bc.lambda))
+gghistogram(train.bc, add.normal = T) + xlab("mean hourly movement (nm)") + labs(caption = paste0("Data Box-Cox transformed with lambda = ", train.bc.lambda))
+gg_qqplot(train.bc) + 
+  labs(title="Q-Q plot for training dataset") + ylab("Sample quantiles") + labs(caption = paste0("Data Box-Cox transformed with lambda = ", train.bc.lambda))
+```
+
 ```{r periodgram}
 library(spec)
-pgram <- spec.pgram(build_ts(seis.h, start = seis.train.start, end = seis.train.end), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
+pgram <- spec.pgram(build_ts(seis.h, start = seis.train.start, end = seis.train.end), main="Periodgram", detrend=F, demean=T, na.action=na.aggregate, plot=F, xlim=c(0,1/3600))
+pgram.data <- data.frame(freq=pgram$freq, spec=pgram$spec)
+ggplot(pgram.data) +
+  geom_line(aes(freq, log(spec))) +
+  xlab("frequency") + ylab("log(spectrum)") +
+  labs(title = "Periodgram", caption = paste0("bandwidth = ", pgram$bandwidth, "\tdf = ", pgram$df)) ->
+  periodgram
+
 topfreq <- pgram$freq[order(pgram$spec, decreasing = T)][1:100]
@@
 
-filter_freq <- (pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*2)) 
+start_ratio <- 2
+filter_freq <- (pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*start_ratio)) 
 top95pc <- pgram$spec[filter_freq]
@@
 
-ylims <- 10^par("usr")[3:4]
-axis(side=3, at=top95freq, labels=as.integer(top95seas.h), padj = 1, las=2, cex.axis=0.8)
-abline(h=mean(pgram$spec) + sd(pgram$spec)*c(.3,1:3), lty="dotted", col=c("orange", "darkgreen", "blue", "red"))
-for (i in 1:length(top95freq)) {
-  color <- ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*3), "red", ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*2), "blue", ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*1), "darkgreen", "orange")))
-  lines(x=c(top95freq[i], top95freq[i]), y=c(top95pc[i],ylims[2]), lty="dotted", col=color)
-  points(x=top95freq[i], y=top95pc[i], col=color)
-}
-axis(side=4, at=mean(pgram$spec) + sd(pgram$spec) * c(.3, 1:3), labels=c(".3", "1", "2", "3"), cex.axis=0.8)
-mtext(side=4, "spectrum sd ratio", line=1)
+ratio.color <- function(x) ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*3), "red", ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*2), "blue", ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*1), "darkgreen", "orange")))
+
+library(ggrepel)
+periodgram +
+  scale_x_continuous(sec.axis = sec_axis(~1/./3600, breaks=c(1, 6*c(1,2,4,4*7)), labels = as.integer, name="hours")) +
+  scale_y_continuous(sec.axis = sec_axis(~exp(.), breaks=mean(pgram$spec) + sd(pgram$spec) * c(1:3), labels=c("1", "2", "3")[1:3], name="spectrum/sd ratio")) +
+  geom_hline(yintercept=log(mean(pgram$spec) + sd(pgram$spec) * c(1:3)), linetype="dotted", color = ratio.color(mean(pgram$spec) + sd(pgram$spec) * c(1:3))) +
+  geom_point(data=data.frame(freq=top95freq, spec=top95pc), aes(freq, log(spec), color=ratio.color(spec)), shape="o", size=4, show.legend = F) +
+  geom_label_repel(data=data.frame(freq=top95freq, spec=top95pc)[1:3,], aes(freq, log(spec), label=round(1/freq/3600), color=ratio.color(spec)), segment.color="grey", show.legend=F) +
+  scale_color_manual(values = c("darkgreen", "blue", "red")[start_ratio:3])
+
 gc()
@@
 autoplot(dec.multi) + labs(title="Decomposition of multiplicative time series (log-scale additive)", subtitle = paste0("Seasonality ", seasonality.period))
+dec.multi.plain <- decompose(seasonality.aggregate(seis.h.train), type="multip")
+autoplot(dec.multi.plain) + labs(title="Decomposition of multiplicative time series", subtitle = paste0("Seasonality ", seasonality.period))
 plot_grid(
@@
   autoplot(acf(dec.multi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
-  nrow = 2)
+  autoplot(acf(dec.multi.plain$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
+  nrow = 3)
 plot_grid(
@@
   autoplot(pacf(dec.multi$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
-  nrow = 2)
+  autoplot(pacf(dec.multi.plain$random, na.action=na.omit, plot=F)) + scale_x_continuous(),
+  nrow = 3)
 ```

### 2020-05-29 project.mems.Rmd
working on draft

@@
 author: "Marco Ciotola"
-date: "16/05/2020"
 output:
@@
 
-```{r transformation}
-plot(seis.h, xaxt="n", type="l")
-axis.POSIXct(1, at=seq.POSIXt(from=seis.h.start, to=seis.h.end, by = "1 month"), format="%m-%y", cex=0.8)
-
-hist(seis.h.train)
-plot(log(seis.h.train))
-hist(log(seis.h.train))
-train.bc.lambda <- BoxCox.lambda(seis.h.train)
-train.bc <- BoxCox(seis.h.train, train.bc.lambda)
-plot(train.bc)
-gghistogram(train.bc, add.normal = T)
-gg_qqplot(seis.h.train) + 
-        labs(title="Q-Q plot for training dataset") + ylab("Sample quantiles")
-gg_qqplot(log(seis.h.train)) + 
-        labs(title="Q-Q plot for log(training) dataset") + ylab("Sample quantiles")
-gg_qqplot(train.bc) + 
-        labs(title="Q-Q plot for BoxCox(training) dataset") + ylab("Sample quantiles")
-```
-
-```{r acf-pacf}
+```{r bc acf-pacf}
 train.bc %>% ggtsdisplay(lag.max = 168*5)
@@
 gc()
-proj_file <- 'proj.mems.RData'
 save.image(proj_file)

### 2020-05-30 project.Rmd
cleanup project files

@@
 library(ggplot2)
+library(ggrepel) # for nicer labels on ggplot
 # library(gridExtra) # for gg-plotting multiple plots
@@
 library(qqplotr) # qqplot with ggplot
-library(tidyverse) # map_df, read_csv, group_by, summarize
+library(spec) # for periodgram
+library(tidyverse) # map_df, read_csv, group_by, summarize...
 library(tsbox) # https://cran.r-project.org/web/packages/tsbox/vignettes/tsbox.html
@@
 ```
-
+## Missing values analysis (seconds)
 ```{r missing values analysis}
@@
 
+## Missing values analysis (hours)
 ```{r missing values hour analysis}
@@
 
+# Analysis
+```{r whole data plot}
+month.starts <- seq.POSIXt(from=floor_date(seis.h.start, 'month'), to=ceiling_date(seis.h.end, 'month'), by = "1 month")
+hours.from.start <- as.numeric(month.starts-seis.h.start)*24+168
+autoplot(seis.h.ts, type="l") +
+  scale_x_continuous(breaks=hours.from.start/168, labels=format(month.starts, "%m-%y")) +
+  ylab("mean hourly movement (nm)") + xlab(NULL) + ggtitle("Training set") + labs(caption = "Data in original scale")
+dec.addi <- decompose(seasonality.aggregate(seis.h.ts))
+autoplot(dec.addi) + labs(title="Decomposition of additive time series", subtitle = paste0("Seasonality ", seasonality.period)) + xlab("Weeks")
+```
+
 ```{r training data transformation plots}
@@
 
+## Periodgram
 ```{r periodgram}
-library(spec)
 pgram <- spec.pgram(build_ts(seis.h, start = seis.train.start, end = seis.train.end), main="Periodgram", detrend=F, demean=T, na.action=na.aggregate, plot=F, xlim=c(0,1/3600))
@@
 ratio.color <- function(x) ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*3), "red", ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*2), "blue", ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*1), "darkgreen", "orange")))
-
-library(ggrepel)
+shown_labels <- 4
 periodgram +
@@
   geom_point(data=data.frame(freq=top95freq, spec=top95pc), aes(freq, log(spec), color=ratio.color(spec)), shape="o", size=4, show.legend = F) +
-  geom_label_repel(data=data.frame(freq=top95freq, spec=top95pc)[1:3,], aes(freq, log(spec), label=round(1/freq/3600), color=ratio.color(spec)), segment.color="grey", show.legend=F) +
+  geom_label_repel(data=data.frame(freq=top95freq, spec=top95pc)[1:shown_labels,], aes(freq, log(spec), label=round(1/freq/3600), color=ratio.color(spec)), segment.color="grey", show.legend=F) +
   scale_color_manual(values = c("darkgreen", "blue", "red")[start_ratio:3])
@@
 ```
+
+## Decomposition
 
@@
 Once notice the composition is additive (the fluctuations does not seem to grow with the trend)
+
 ```{r decompose}
@@
 
+### Manual filtering/decomposition
 We now extract the trend using a linear filter with smoothing parameter $p$ equal to 180 (almost half a year):
@@
 
+### Multi trend extraction
+
 ```{r extract both 24 and 168}
@@
 
+# Forecast
+
 ```{r diffs acf/pacf}

### 2020-05-30 project.mems.Rmd
cleanup project files

@@
 ---
-
-```{r tentatives, eval=FALSE, child='chapter1.Rmd', include=FALSE}
-```
 
@@
 
+## Arima
 ```{r arima}
@@
 
+## Arima without seasonality
 ```{r arima on seasonality}
@@
 ```
+
+## MSTL
 

### 2020-05-30 project.uccs.Rmd
working on draft
cleanup project files

@@
 ```
+
+## Arima
 
@@
 
+## MSTL
 ```{r mstl on msts}
@@
 
-```{r eval=FALSE, include=FALSE}
+```{r save RData, include=FALSE}
 gc()
-if (name == "MEM") {
-  proj_file <- 'proj.mems.RData'
-  save.image(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
-  remove(proj_file)
-}
+save.image(proj_file)
 ```

### 2020-05-30 draft.tex
Working on draft

@@
 \usepackage{geometry}
-\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
+\geometry{a4paper,verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
 \setcounter{secnumdepth}{2}
@@
 
-\paragraph{Motivation}
+\subsubsection{Motivation and objective}
 Thanks to the recent coronavirus outbreak worldwide, we currently have data about many different phoenomena from periods with standard human activity, and from a period with reduced human activity.
 
-Ispired by an article published during the initial phases of the Belgium-wide lockdown \cite{NatureCoronavirusSeismic}, we want to assess whether the lockdown period can be correlated with a visible change in seismic activity.
-It is important to notice that some industrial events, such as mine explosions, are already mapped by seismologic institutions since they interfere with seismic surveys \cite{OtherSeismicEvents}, sometimes comparable with natural seismic events, depending on the distance from the seismograph.
+Inspired by an article published during the initial phases of the Belgium-wide lockdown \cite{NatureCoronavirusSeismic}, we want to assess whether the lockdown period can be correlated with a visible change in seismic activity.
+It is important to notice that some industrial events, such as mine explosions, are already mapped by seismologic institutions since they interfere with seismic surveys \cite{OtherSeismicEvents}.
+Such events are sometimes comparable with natural seismic events, depending on the distance from the seismograph.
 
-\paragraph{Data source and description}
+\subsubsection{Data source and description}
 The Royal Observatory of Belgium (ROB) gathers and publishes \cite{RoyalDataPolicy,Data20200402} data about vertical ground displacement from 9 stations through the country, linking them to seismic events whenever possible \cite{DataEvents20200402}.
@@
 
-\subsection{Data cleanup and transformation}
-\paragraph{Data gathering and cleanup}
+\subsection{Data cleanup and transformation}\label{sec:cleanup-transform}
+\subsubsection{Data gathering and cleanup}
 The data is provided with a CSV per day, with only two columns related to the min-max pair \cite{Data20200402}. This causes two problems:
@@
 
-\paragraph{Data aggregation and initial transformation}
+\subsubsection{Data aggregation and initial transformation}
 Since we are not interested in second-precise analysis, and it would be computationally expensive to deal with the whole dataset, we are going to aggregate data by hour.
@@
 \section{Analysis and decomposition}
-\paragraph{Missing values analysis}
+\subsubsection{Missing values analysis}
 
@@
 After aggregating by hour, the percentages of missing values remain quite stable.
-As can be seen from figures \ref{missing-values-seconds:uccs} and \ref{missing-values-seconds:mems}, in both stations missing values are grouped in few consecutive periods.
+As can be seen from \figurename{s} \ref{missing-values-seconds:uccs} and \ref{missing-values-seconds:mems}, both stations missing values are grouped in few consecutive periods.
 To deal with them when R methods need complete timeseries, we will define different completion strategies depending on the seasonality we will be studying, always based on the R function \texttt{na.aggregate}.
-
-\begin{figure}
-	\caption{Missing values representation}
-	\label{missing-values-seconds}
-	\begin{subfigure}{.5\linewidth}
-		\caption{Uccle station}
-		\label{missing-values-seconds:uccs}
-		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/missing values hour analysis-1.pdf}
-	\end{subfigure}
-	\begin{subfigure}{.5\linewidth}
-		\caption{Membach station}
-		\label{missing-values-seconds:mems}
-		\includegraphics[width=\linewidth]{project.mems_files/figure-latex/missing values hour analysis-1.pdf}
-	\end{subfigure}
-\end{figure}
 
@@
 
-\paragraph{Dataset division}
-As best practices suggest, we are going to divide the dataset into training and testing datasets. 
-Since one of our objectives is to assess any significant difference from before and after the Belgium lockdown (on March, 14$^{th}$ the `soft' lockdown, while on March, 18$^{th}$ the complete lockdown)
+\begin{figure}[h]
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/missing values hour analysis-1.pdf}
+		\caption{Uccle station}
+		\label{missing-values-seconds:uccs}
+	\end{subfigure}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{project.mems_files/figure-latex/missing values hour analysis-1.pdf}
+		\caption{Membach station}
+		\label{missing-values-seconds:mems}
+	\end{subfigure}
+	\caption{Missing values representation}
+	\label{missing-values-seconds}
+\end{figure}
 
-\paragraph*{Hypotesis of analysis}
+\subsubsection{Seasonality assumptions}
+The already cited article \cite{NatureCoronavirusSeismic} suggests the existence of a weekly seasonality. Looking at the data it seems confirmed, in addition to a daily seasonality.
+
+Since our main focus is on the seasonalities themselves, in order to prevent assumptions based on the objective, so that searching for correlation with human activity becomes imposing human-based seasonality periods, we will use an analytical method to extract seasonalities that the data itself suggests to us.
+Still, we recognize that the already performed transformations impose an human-based time alteration to the data itself (Section \ref{sec:cleanup-transform}).
+
+The method we are going to apply is based on the Fourier Transform, and generates a \textit{periodgram} highlighting the most interesting frequencies that might correspond to a seasonality component.
+Applying the FT for any frequency, we transform our 1-dimensional data into a 2-dimensional data mapped around a circle.
+Then, we can sum all the resulting vectors (that are our new data points) to obtain a vector for each frequency, whose absolute value represents the power level of the same frequency.
+An high power level represents a significant periodicity on that frequency in our original data.
+
+A common operation to increase the difference between significant and non-significant periodicities is to apply a MA smoothing filter with a smaller window than the interesting ones.
+We will use such a filter with $f=2$.
+
+On this note we must specify that, when a significant periodicity is found, it influences smaller periodicities given by $f*\{2,3,\dots\}$ the significant frequency.
+Also, if any periodicity is found that is longer than the original dataset, it must be considered as a \textit{coincidence}, rather than a surprising serendipity.
+
+Since we will be applying this method on the dataset indexed by a datetime object, which uses the second as the unit, the transformation of frequency $f$ into hourly periodicity $p$ is obtained by $p=(1/f)/3600$.
+We can analyse the resulting periodgrams for Uccle and Membach stations in \figurename{s} \ref{fig:periodgram:uccs} and \ref{fig:periodgram:mems}.
+
+\begin{figure}[h]
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/periodgram-1"}
+		\caption{Uccle station}
+		\label{fig:periodgram:uccs}
+	\end{subfigure}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/periodgram-1"}
+		\caption{Membach station}
+		\label{fig:periodgram:mems}
+	\end{subfigure}
+	\caption{Periodgram results}
+	\label{fig:periodgram}
+\end{figure}
+
+Both highlight multiple 24-hour periodicities.
+This is given by approximation factors, since we round the divisions result to the nearest natural number.
+It also highlights that using a single 24-hour seasonality could leave some information on the remaining components.
+Still, in both stations it is the most significant periodicity.
+
+Both stations highlight a significant 168-hour (1-week) periodicity, that also confirms our initial assumptions.
+
+Both stations show a possibly significant 12-hour seasonality, but since it corresponds to double the frequency of the 24-hour seasonality, we can safely ignore it.
+
+In particular from the Uccle station, we can see a great number of spikes, that will probably undermine our models.
+
+% From the Membach station, we can see an anomalous spike at a 6-hour periodicity.
+
+\subsubsection{Dataset division}
+As best practices suggest, we are going to divide the dataset into training and testing datasets with a ratio around $85\%/15\% - 75\%/25\%$.
+Since one of our objectives is to assess any significant difference from before and after the Belgium lockdown (on March, $14^{th}$ the \textit{soft} lockdown, while on March, $18^{th}$ the complete lockdown), we selected data from the start of our timeseries until 20-01-2020 (64 weeks $\approx 82\%$) as training, while the test follows and ends on 27-04-2020 (14 weeks $\approx 18\%$).
+
+This initial division demostrated to be computationally heavy while fitting some models with weekly seasonality, and we noticed a possible yearly seasonality on Uccle data (\figurename~\ref{fig:whole-data-decomposition}).
+Even if it is not demonstrable given the short period our data comes from, applying a 24-hour MA smoothing filter on the whole dataset and then producing the periodgram as previously done, it suggests a 8748-hours significant periodicity (364.5 days).
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.7\linewidth]{"project.uccs_files/figure-latex/whole data plot-2"}
+	\caption{Whole Uccle data decomposition}{\footnotesize Weeks 10 and 62 include January, $1^{st}~2019/2020$}
+	\label{fig:whole-data-decomposition}
+\end{figure}
+
+Given those issues, we choose the training/testing sets to hold 52/18 weeks: this is the greater train-to-test ratio we deem usable ($\approx 75\%/25\%$), while maintaining computational feasibility, and we hope it will render the possible yearly seasonality less intrusive.
+
+\subsubsection{Decomposition}
+In order to decompose our dataset into trend and seasonality(-ties) components, we followed these steps:
+\begin{enumerate}
+	\item Apply a smoothing filter to extract the trend component
+	\item Remove the trend from the dataset and extract the seasonal figure using the local mean method
+	\item Remove the trend and the seasonal components to obtain the residuals
+\end{enumerate}
+While doing so, we tested different smoothing filters: simple filter (p=seasonality), Spencer's 15-point filter and MA smoothing filter (f=seasonality).
+Since they do not produce particularly different seasonalities, we choose to keep the results obtained by applying the MA smoothing filter, for its significance in terms of seasonally adjusted data.
+
+Since the periodgrams suggested the presence of two significative periodicities, we also tested the extraction of both corresponding seasonalities:
+\begin{enumerate}
+	\item Apply the MA smoothing filter (f=24) on the dataset to extract the trend component
+	\item Remove the trend from the dataset and extract the 24-hour seasonal figure using the local mean method
+	\item Remove the 24-hour seasonal components to obtain a deseasoned dataset
+	\item Apply the MA smoothing filter (f=168) on the deseasoned dataset to extract the updated trend component
+	\item Remove the trend from the already deseasoned dataset and extract the 168-hour seasonal figure using the local mean method
+	\item Repeat steps 3-5 steps alternating the 168/24 seasonalities until convergence (2 to 3 cycles should be enough)
+	\item Removing the last fitted trend and the two seasonalities we obtain the final residuals
+\end{enumerate}
+
+\begin{figure}[h]
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/extract both 24 and 168-2.pdf}
+		\caption{Uccle station}
+		\label{fig:direct-multi-seasonality:uccs}
+	\end{subfigure}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{project.mems_files/figure-latex/extract both 24 and 168-2.pdf}
+		\caption{Membach station}
+		\label{fig:direct-multi-seasonality:mems}
+	\end{subfigure}
+	\begin{center}
+		\caption{Seasonality figures}{\small\textit{Direct} refers to single seasonality extraction\\\textit{Multi} refers to both seasonalities extracted at the same time}
+		\label{fig:direct-multi-seasonality}
+	\end{center}
+\end{figure}
+
+In \figurename~\ref{fig:direct-multi-seasonality} we see both approaches applied to both stations. Those plots show an interesting property of our seasonalities:
+\begin{itemize}
+	\item Both ways of extraction result in the same daily seasonality extracted
+	\item The sum of the daily and weekly seasonalities extracted together is the same as the weekly seasonality extracted directly
+\end{itemize}
+Given in particular the second property, we will only use the 168-hours seasonality while fitting our models, for both the stations.
+
+In \figurename~\ref{fig:decomposed} we see the decomposition results.
+While data from Uccle station have a clear significativity of both trend and seasonalities, data from Membach could seem not significant.
+Still, if we notice that the main timeseries lies between values of 0 and 100, we can find out that around half of the value will be composed by the trend, a 30\% by the joined seasonalities component, and the remaining value by the residuals.
+We will need to better handle the spikes from around half our dataset.
+
+\begin{figure}[h]
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/extract both 24 and 168-1.pdf}
+		\caption{Uccle station}
+		\label{fig:decomposed:uccs}
+	\end{subfigure}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{project.mems_files/figure-latex/extract both 24 and 168-1.pdf}
+		\caption{Membach station}
+		\label{fig:decomposed:mems}
+	\end{subfigure}
+	\begin{center}
+		\caption{Timeseries decomposition}{\small Only the second method results are shown for clarity}
+		\label{fig:decomposed}
+	\end{center}
+\end{figure}
+
+\newpage
+\section{Parameter estimation}
+\subsection{Uccle station}
+Our training set, despite the precautions taken during the initial data cleanup and aggregation, doesn't satisfy the normality assumption, even after applying the \textit{log} or \textit{Box-Cox} transformations.
+Thus, we will estimate models to fit our data starting with the original scale.
+\subsubsection{ARIMA}
+The first step to fit an ARIMA will be to check the ACF/PACF plots to estimate the number of parameters.
+As we can see in \figurename~\ref{fig:pacf-4}, applying the diff operator both on the data and on the seasonal data, result in interpretable plots.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\linewidth]{"project.uccs_files/figure-latex/diffs acf/pacf-4"}
+	\caption{ACF and PACF after diff and diff with lag 168}
+	\label{fig:pacf-4}
+\end{figure}
+
+\subsubsection{MSTL}
+
+\subsection{Membach station}
+\subsubsection{Transformation and normality assumption}
+\subsubsection{ARIMA}
+\subsubsection{MSTL}
+
+\subsection{Seasonplots}
+
+\subsubsection*{Hypotesis of analysis}
 

### 2020-05-30 project.Rmd
Working on draft

@@
 setwd("/media/data/Documents/Uni/spatiotemp/villalobos/project/")
-if (exists("proj_file")) {
-  load(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
-  # save.image(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
+if (exists("proj_file") & file.exists(proj_file)) {
+  load(proj_file)
+  # save.image(proj_file)
 }
@@
 ```{r periodgram}
-pgram <- spec.pgram(build_ts(seis.h, start = seis.train.start, end = seis.train.end), main="Periodgram", detrend=F, demean=T, na.action=na.aggregate, plot=F, xlim=c(0,1/3600))
+min_h_seasonality <- 24
+min_h_seasonality <- 2
+pgram <- spec.pgram(stats::filter(build_ts(seis.h, start = seis.train.start, end = seis.train.end), side=2, filter=c(0.5, rep(1,min_h_seasonality-1), 0.5)/min_h_seasonality),
+#`pgram <- spec.pgram(build_ts(seis.h, start = seis.train.start, end = seis.train.end),
+                    detrend=F, demean=T, na.action=na.aggregate, plot=F)
 pgram.data <- data.frame(freq=pgram$freq, spec=pgram$spec)
@@
   xlab("frequency") + ylab("log(spectrum)") +
-  labs(title = "Periodgram", caption = paste0("bandwidth = ", pgram$bandwidth, "\tdf = ", pgram$df)) ->
+  labs(title = "Periodgram", caption = paste0("bandwidth = ", pgram$bandwidth, "    df = ", pgram$df)) ->
   periodgram
-
-topfreq <- pgram$freq[order(pgram$spec, decreasing = T)][1:100]
-topseas <- 1/topfreq
-topseas.h <- topseas/3600
 
@@
 periodgram +
-  scale_x_continuous(sec.axis = sec_axis(~1/./3600, breaks=c(1, 6*c(1,2,4,4*7)), labels = as.integer, name="hours")) +
-  scale_y_continuous(sec.axis = sec_axis(~exp(.), breaks=mean(pgram$spec) + sd(pgram$spec) * c(1:3), labels=c("1", "2", "3")[1:3], name="spectrum/sd ratio")) +
+  scale_x_continuous(sec.axis = sec_axis(~1/./3600, breaks=c(1, 3, 2*c(1,2,3,4,6,12), 24*7), labels = as.integer, name="hours")) +
+  scale_y_continuous(sec.axis = sec_axis(~exp(.), breaks=mean(pgram$spec) + sd(pgram$spec) * c(1,3), labels=c("1", "2", "3")[c(1,3)], name="spectrum/sd ratio")) +
   geom_hline(yintercept=log(mean(pgram$spec) + sd(pgram$spec) * c(1:3)), linetype="dotted", color = ratio.color(mean(pgram$spec) + sd(pgram$spec) * c(1:3))) +

### 2020-06-01 draft.tex
working on draft

@@
 \usepackage{geometry}
-\geometry{a4paper,verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
+\geometry{a4paper,verbose,tmargin=2.5cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
 \setcounter{secnumdepth}{2}
@@
 ROB collects data from 9 stations. Of those, the following 6 have interesting positions, related to possible correlations with human activity:
-\begin{itemize}
+\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
 	\item \textbf{Uccle}, \textbf{Sart-Tilman}, and \textbf{Ében-Émael} are near a city (Bruxelles, Liege, and Maastricht).
@@
 The data is provided with a CSV per day, with only two columns related to the min-max pair \cite{Data20200402}. This causes two problems:
-\begin{enumerate}
+\begin{enumerate}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
 	\item Before merging different days, we need to add a column with the complete datetime for each row
@@
 The aggregation was performed considering the mean max-min difference, for the following reasons:
-\begin{itemize}
+\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
 	\item When an hour contains some missing seconds, these do not heavily affect the aggregation result, in contrast with summing the differences
@@
 In order to decompose our dataset into trend and seasonality(-ties) components, we followed these steps:
-\begin{enumerate}
+\begin{enumerate}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
 	\item Apply a smoothing filter to extract the trend component
@@
 Since the periodgrams suggested the presence of two significative periodicities, we also tested the extraction of both corresponding seasonalities:
-\begin{enumerate}
+\begin{enumerate}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
 	\item Apply the MA smoothing filter (f=24) on the dataset to extract the trend component
@@
 In \figurename~\ref{fig:direct-multi-seasonality} we see both approaches applied to both stations. Those plots show an interesting property of our seasonalities:
-\begin{itemize}
+\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
 	\item Both ways of extraction result in the same daily seasonality extracted
@@
 
-\newpage
 \section{Parameter estimation}
+Having two different datasets, we need to fit two separate models that represents those stations.
+During this estimation, our aim is to obtain a model that generalises our training dataset in such a way that forecasting becomes effective, also assessing errors analytically.
+Obtaining such a reliable model will be the first step to verify any significant difference forecasting over the lockdown period.
+
+The first model we are going to estimate is the ARIMA model, including the seasonality components.
+To identify the order of the different components described by ARIMA, we will analyse the autocorrelation and partial autocorrelation of our datasets.
+As a reminder:
+\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
+	\item non-seasonal AR process of order $p$ is described by an exponentially decaying ACF, and by a cut-off of significant PACF values at lag $p$
+	\item non-seasonal MA process of order $q$ is described by an exponentially decaying PACF, and by a cut-off of significant ACF values at lag $q$
+	\item seasonal AR process of order $P$ is described by an exponentially decay at seasonal lags on ACF values, and by a cut-off of significant PACF values at each seasonal lag until the $P^{th}$
+	\item seasonal MA process of order $Q$ is described by an exponentially decay at seasonal lags on PACF values, and by a cut-off of significant ACF values at each seasonal lag until the $Q^{th}$
+	\item given how significance is assessed for the ACF/PACF spikes, one over 20 spikes could be \textit{coincidentally significant}
+\end{itemize}
+
 \subsection{Uccle station}
-Our training set, despite the precautions taken during the initial data cleanup and aggregation, doesn't satisfy the normality assumption, even after applying the \textit{log} or \textit{Box-Cox} transformations.
-Thus, we will estimate models to fit our data starting with the original scale.
 \subsubsection{ARIMA}
-The first step to fit an ARIMA will be to check the ACF/PACF plots to estimate the number of parameters.
-As we can see in \figurename~\ref{fig:pacf-4}, applying the diff operator both on the data and on the seasonal data, result in interpretable plots.
+Our training set, despite the precautions taken during the initial data cleanup and aggregation, doesn't satisfy the Normality assumption, even after applying the \textit{log} or \textit{Box-Cox} transformations.
+Thus, we will estimate the models using the original scale.
+
+Since our dataset shows a relevant trend component, resulting in non-constant mean, we consider the usage of the \textit{diff} operator to smooth the mean.
+After applying the \textit{diff} operator with lag 168, in conjunction with the simple \textit{diff} or not, we come up with two possible models by checking the ACF and PACF plots:
+\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
+	\item from using both \textit{diff}s, the ACF and PACF on the long run hint to $\mathrm{ARIMA}(0,1,0)(0,1,1)_{168}$
+	\item from using just the lagged \textit{diff}, the PACF on the first lags and the ACF/PACF combination on the long run hint to $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$
+\end{itemize}
+We also employ the \textit{auto.arima} stepwise methodology, locking the \textit{diff} values to 0 or 1, to get some hints on the possible models.
+
+Following both the analytical ways and the \textit{auto.arima} hints, and refining the models through further inspection of ACF/PACF plots of the residuals, we end up with three characteristic models.
+We selected them through the Akaike Information Criterion (AIC), being the only three out of the fitted models with $\mathrm{AIC} < 120.000$: $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$, $\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$, and $\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$.
+
+Also checking the Mean Absolute Percentage Error (MAPE) on the training set, they actually have the lowest three values ($\mathrm{MAPE} < 6.2\%$).
+Nevertheless, MAPE is not a single valid criterion for selection, since a low error predicting the training set itself does not necessarily mean a good forecast power.
+
 \begin{figure}[h]
 	\centering
-	\includegraphics[width=1\linewidth]{"project.uccs_files/figure-latex/diffs acf/pacf-4"}
-	\caption{ACF and PACF after diff and diff with lag 168}
-	\label{fig:pacf-4}
+	\includegraphics[width=.7\linewidth]{project.uccs_files/figure-latex/arima-26}
+	\caption{ACF and PACF plots for $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$}
+	\label{fig:arima-26}
 \end{figure}
 
-\subsubsection{MSTL}
+The ACF and PACF of the residuals for all three models are really similar (\figurename~\ref{fig:arima-26}), even if the models themselves have great differences in meaning and in number of parameters: they highlight a clear seasonal ARMA process with periodicity 24.
+This shows how our previous conclusion that the 168 hours seasonality could include the 24 hours one is flawed.
+Unfortunately, the parameters estimator we are using does not permit to estimate multiple seasonalities\footnote{\textit{Arima} method from the \textit{forecast} R package, only specifying the seasonal and non-seasonal orders}.
+
+Probably related to the 24 hours periodicity that remains on the residuals, together with the non-Normality of the original data, all Q-Q plots show clear divergence on both sides of the distribution.
+The Box-Pierce test statistics accept the independence hypothesis between residuals.
+
+Then, we assess how the forecasts behave on the test set.
+The MAPE for all three models more than doubles to around 17\% for $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$ and around 15\% for the other two.
+This is expected, since normally the model is more fitted to the training data than to unknown future data.
+Also, it must be noted that a January $1^{st}$ period lies in our test dataset, and without being able to include a yearly seasonality, that leads to greater errors just from those couple of weeks (\figurename~\ref{fig:arima-forecast:200011}).
+
+\begin{figure}[h]
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-27}
+		\caption{$\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$}
+		\label{fig:arima-forecast:200011}
+	\end{subfigure}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-51}
+		\caption{$\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$}
+		\label{fig:arima-forecast:412011}
+	\end{subfigure}
+	\begin{center}
+		\caption{ARIMA forecasts}{\footnotesize$\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$ behaves like $\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$ with faster increasing variablity}
+		\label{fig:arima-forecast}
+	\end{center}
+\end{figure}
+
+We can see in \figurename~\ref{fig:arima-forecast} the forecast behaviour for our models.
+We can appreciate the constant span of the confidence intervals in \figurename~\ref{fig:arima-forecast:200011}, in constrast with the more complex models that have increasing intervals further away from the training set.
+It must be noted that the increasing intervals include negative values, that are actually impossible to obtain given the data meaning (difference between maximum and minimum measurements).
+
+The increasing variability also make the percentage of test values falling inside the confidence intervals to be higher than normal:
+\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
+	\item $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$: $\approx81\%$ of data is in the $80\%$ CI and $\approx91\%$ in the $95\%$ CI
+	\item $\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$: $\approx96\%$ of data is in the $80\%$ CI and $\approx98\%$ in the $95\%$ CI
+	\item $\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$: $\approx98\%$ of data is in the $80\%$ CI and $\approx99\%$ in the $95\%$ CI
+\end{itemize}
+
+Given all these factors, we are confident choosing the $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$ model, being aware of the missed 24 hours seasonality.
+
+%\subsubsection{MSTL}
+%\subsubsection{Fourier}
 
 \subsection{Membach station}
-\subsubsection{Transformation and normality assumption}
 \subsubsection{ARIMA}
-\subsubsection{MSTL}
+Similarly to data from Uccle, data from Membach doesn't respect the Normality assumption.
+We tested if the Q-Q plot showed significantly better approximation of the Normal distribution after applying the \textit{log} or \textit{Box-Cox} transformations, and the latter makes the data approach Normality with $\lambda\approx-1$ (\figurename~\ref{fig:mems:box-cox}).
+This solves the high peaks problem we already noticed while decomposing to highlight the trend and seasonalities.
 
-\subsection{Seasonplots}
+Since the $\lambda$ value approximately simplifies the Box-Cox transformation formula from $(y^\lambda-1)/\lambda$ into $(y-1)/y$, and since we only have positive values, all transformed values fall into the higher $0-1$ spectrum, in particular $0.95-1$ (\figurename~\ref{fig:mems:box-cox:data}).
+This will probably cause some approximation errors that could rig the final results.
+
+\begin{figure}[h]
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=1\linewidth]{"project.mems_files/figure-latex/training data transformation plots-7"}
+		\caption{Transformed training set}
+		\label{fig:mems:box-cox:data}
+	\end{subfigure}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=1\linewidth]{"project.mems_files/figure-latex/training data transformation plots-9"}
+		\caption{Normal vs. transformed data distribution}
+		\label{fig:mems:box-cox:qq}
+	\end{subfigure}
+	\begin{center}
+		\caption{Results of Box-Cox transformation}
+		\label{fig:mems:box-cox}
+	\end{center}
+\end{figure}
+
+The transformed data still shows no constant mean, even if it has a way better shape than the original data.
+Thus, we consider the usage of the \textit{diff} operator, simple and at lag 168.
+Interestingly, the analysis of ACF and PACF plots hints to similar models to the Uccle station:
+$\mathrm{ARIMA}(0,1,0)(0,1,1)_{168}$ and $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$.
+
+In this case, the employment of \textit{auto.arima} stepwise methodology does not lead to any useful model, since it obtains $\mathrm{AICc}=\inf$ for the two manually-chosen models, that have an AICc value around -84.500, while the best model found by that methodology has an AICc around -81.000.
+This can be caused by approximation errors\footnote{The \textit{auto.arima} methodology was used only with initial approximation, due to high computing time}, as we supposed earlier.
+
+The analysis of ACF/PACF plots of the residuals from both models do not lead to any new model but, as with the Uccle station, we can notice a clear 24 hours periodicity.
+
+Both models have valid p-value results for the Box-Pierce test, similar AICc values, and similar MAPE on the training set.
+Thus, we can assess the forecasts for both models in \figurename~\ref{fig:mems:arima}.
+
+\begin{figure}[h]
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=1\linewidth]{"project.mems_files/figure-latex/arima-15"} %18
+		\caption{$\mathrm{ARIMA}(0,1,1)(0,1,1)_{168}$}
+		\label{fig:mems:arima:011011}
+	\end{subfigure}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=1\linewidth]{"project.mems_files/figure-latex/arima-9"} %11
+		\caption{$\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$}
+		\label{fig:mems:arima:200011}
+	\end{subfigure}
+	\begin{center}
+		\caption{ARIMA forecasts}
+		\label{fig:mems:arima}
+	\end{center}
+\end{figure}
+
+The first model catches a rapidly-increasing trend, while the latter predicts more stable values moving the most part of the increasing trend into increasing variability.
+Visually, both could be acceptable, even if the latter seems to better behave on \textit{normal} days losing precision on high-valued ones, while the first tries to better accomodate the high-valued days losing precision on the \textit{normal} ones.
+
+Analytically, we have quite a difference between the two models in terms of adherence of the confidence intervals.
+\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
+	\item $\mathrm{ARIMA}(0,1,1)(0,1,1)_{168}$: $\approx98\%$ of data is in the $80\%$ CI and $\approx99\%$ in the $95\%$ CI
+	\item $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$: $\approx73\%$ of data is in the $80\%$ CI and $\approx89\%$ in the $95\%$ CI
+\end{itemize}
+As we see, the first model over-generalises the variability, while the latter seems more precise, even if losing some information.
+Also considering the MAPE on the test set, the latter model fits the data better with a $16\%$ error, instead of a $18\%$.
+
+Thus, we can select the $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$ model on the Box-Cox($\lambda\approx-1$)-transformed data, still being aware of the missed 24 hours seasonality, similarly to the Uccle station.
+
+%\subsubsection{MSTL}
+
+\subsection{Seasonality comparison}
+\subsection{Lockdown forecast divergence}
+\section{Conclusion}
+\subsubsection{Limitations}
+\subsubsection{Future Works}
 

### 2020-06-01 project.Rmd
working on draft

@@
   mae <- mean(abs(diff), na.rm = T)
+  mape <- mean(abs(diff/base) * 100, na.rm = T)
   rmse <- sqrt(mse)
   R2 <- 1-(sum((diff)^2, na.rm = T)/sum((base-mean(base))^2, na.rm = T))
-  errorstring <- paste0("MSE=", mse, "    MAE=", mae, "    RMSE=", rmse, "    R2=", R2)
+  errorstring <- paste0("MSE=", mse, "  MAE=", mae, "  MAPE=", mape, "  RMSE=", rmse, "  R2=", R2)
   cat("Errors: ", compare.desc, "\n")
-  cat("\t","MSE=", mse, "    MAE=", mae, "    RMSE=", rmse, "    R2=", R2, "\n")
+  cat("\t","MSE=", mse, "  MAE=", mae, "  MAPE=", mape, "  RMSE=", rmse, "  R2=", R2, "\n")
 
@@
 trend.s <- stats::filter(seasonality.aggregate(seis.h.train), side=2, filter=weights.s)
+stats::filter(seasonality.aggregate(seis.h.train), sides=2, filter = weights.s) %>% autoplot()
+stats::filter(seasonality.aggregate(seis.h.train) %>% diff(), sides=2, filter = weights.s) %>% autoplot()
+stats::filter(seasonality.aggregate(seis.h.train) %>% diff(lag=168), sides=2, filter = weights.s) %>% autoplot()
+stats::filter(seasonality.aggregate(seis.h.train) %>% diff() %>% diff(lag=168), sides=2, filter = weights.s) %>% autoplot()
+mean((seasonality.aggregate(seis.h.train) %>% diff(lag=168))[(1+(168*4)):(1+(168*6))])
 
@@
 ```{r diffs acf/pacf}
-seis.h.train %>% ggtsdisplay(lag.max = 168*5)
-seis.h.train %>% diff() %>% ggtsdisplay(lag.max = 168*5)
+seis.h.train %>% ggtsdisplay(lag.max = 168*4)
+seis.h.train %>% diff() %>% ggtsdisplay(lag.max = 168*4)
 # seis.h.train %>% diff(D=2) %>% ggtsdisplay(lag.max = 168)
@@
 # seis.h.train %>% diff() %>% diff(lag=24, D=2) %>% ggtsdisplay(lag.max = 168)
-seis.h.train %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,0,0)(0,1,1/2)[168]
+seis.h.train %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*4) # arima (0,0,0)(0,1,1/2)[168]
 # acf variable and partially sinusoidal, pacf shows peaks each 168
-seis.h.train %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,1,0)(0,1,1)[168]
+seis.h.train %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*4) # arima (0,1,0)(0,1,1)[168]
+seis.h.train %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 172) # arima (0,1,0)(0,1,1)[168]
 # acf cuts off after p=168, pacf shows peaks each 168   
 # seis.h.train %>% diff() %>% diff(lag=168, D=2) %>% ggtsdisplay(lag.max = 168)
-seis.h.train %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)
+seis.h.train %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*4)
+```
+
+```{r diffs/d168-acf-pacf}
+plot_grid(
+  seis.h.train %>% diff(lag=168) %>%  acf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  seis.h.train %>% diff(lag=168) %>%  acf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  seis.h.train %>% diff(lag=168) %>% pacf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  seis.h.train %>% diff(lag=168) %>% pacf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  nrow = 2
+)
+```
+
+```{r diffs/d-d168-acf-pacf}
+plot_grid(
+  seis.h.train %>% diff() %>% diff(lag=168) %>%  acf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  seis.h.train %>% diff() %>% diff(lag=168) %>%  acf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  seis.h.train %>% diff() %>% diff(lag=168) %>% pacf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  seis.h.train %>% diff() %>% diff(lag=168) %>% pacf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  nrow = 2
+)
 ```

### 2020-06-01 project.mems.Rmd
working on draft

@@
 ```{r bc acf-pacf}
-train.bc %>% ggtsdisplay(lag.max = 168*5)
-train.bc %>% diff() %>% ggtsdisplay(lag.max = 168*5)
+train.bc %>% ggtsdisplay(lag.max = 168*4)
+train.bc %>% diff() %>% ggtsdisplay(lag.max = 168*4)
 train.bc %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
 train.bc %>% diff() %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
-train.bc %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)
-train.bc %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,1,0)(0,1,1)[168]
-train.bc %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168) # arima (0,1,0)(0,1,1)[168]
+train.bc %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*4)
+train.bc %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*4) # arima (0,1,0)(0,1,1)[168]
 # acf cuts off after p=168, pacf shows peaks each 168
-train.bc %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)
+train.bc %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*4)
+```
+
+```{r diffs/bc-d168-acf-pacf}
+plot_grid(
+  train.bc %>% diff(lag=168) %>%  acf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  train.bc %>% diff(lag=168) %>%  acf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  train.bc %>% diff(lag=168) %>% pacf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  train.bc %>% diff(lag=168) %>% pacf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  nrow = 2
+)
+```
+
+```{r diffs/bc-d-d168-acf-pacf}
+plot_grid(
+  train.bc %>% diff() %>% diff(lag=168) %>%  acf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  train.bc %>% diff() %>% diff(lag=168) %>%  acf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  train.bc %>% diff() %>% diff(lag=168) %>% pacf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  train.bc %>% diff() %>% diff(lag=168) %>% pacf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
+  ,
+  nrow = 2
+)
 ```
@@
 to_test <- list(
-  c(0,1,0,0,1,1)
+  c(0,1,0,0,1,1),
+  c(2,0,0,0,1,1),
+  #c(3,0,0,0,1,1),
+  c(0,1,1,0,1,1) #,
+  #c(0,1,2,0,1,1)
 )
@@
 }
-for (model in to_test) {
-  # model <- c(4,1,2,0,1,0)
-  fit.name <- paste0("fit", paste0(model, collapse = ""))
-  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
-  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
-  if (!(fit.name %in% names(models))) {
-    models[[fit.name]] <- Arima(seis.h.train, model[1:3], model[4:6], lambda=train.bc.lambda)
-  }
-  if (!(forecast.name %in% names(models))) {
-    models[[forecast.name]] <- forecast(models[[fit.name]], h=168*18, lambda=train.bc.lambda)
-  }
-}
-#models[["fitauto"]] <- auto.arima(train.bc, D=1, approx=F, trace = T)
-#models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*18)  
-#models[["fit311010"]] <- models[["fitauto"]]
-#models[["fit412010"]] <- fit
-#models[["forecast412010"]] <- forecast(models[["fit412010"]], h=168*18)
 
@@
   forecast.name <- paste0("forecast", paste0(model, collapse = ""))
-  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
-  if ((fit.name %in% names(models))) {  if ((forecast.name %in% names(models))) {
+  cat("SARIMA(", paste0(model[1:3], collapse = ","), ")(", paste0(model[4:6], collapse = ","), ")[", frequency(seis.h.train), "]", sep = "")
+  if (!(fit.name %in% names(models))) {
+    cat(" fitting...")
+    models[[fit.name]] <- Arima(seis.h.train, model[1:3], model[4:6], lambda=train.bc.lambda)
+    cat("DONE")
+  }
+  if (!(forecast.name %in% names(models))) {
+    cat(" forecasting...")
+    models[[forecast.name]] <- forecast(models[[fit.name]], h=168*18, lambda=train.bc.lambda)
+    cat("DONE")
+  }
+  cat("\n")
   plotting.model <- models[[fit.name]]
   plotting.forecast <- models[[forecast.name]]
+  cat("AIC =",plotting.model$aic, "\n")
+  cat("AICc =",plotting.model$aicc, "\n")
   checkresiduals(plotting.model)
-    residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-    residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+  residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*4)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
+  residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*4)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
   print(plot_grid(residuals.acf, residuals.pacf, nrow=2))
-    print(plot_forecast(plotting.forecast) + 
-      autolayer(seis.h.test, series="Original"))
+  print(plot_grid(
+    autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
+    ,
+    residuals.acf
+    ,
+    autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
+    ,
+    residuals.pacf
+    ,
+    nrow = 2
+  ))
+  print(plot_forecast(plotting.forecast) + autolayer(seis.h.test, series="Original") + ylab("Mean hourly movement (nm)"))
   plots <- show_overlap_and_diff(
     seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
-      base.desc = "original series",
+    base.desc = "training set",
     compare.desc = paste0("refit with ", plotting.forecast$method),
@@
     diff.title = "Difference between ts",
-      diff.ylab = "nm", diff.ylab.sec="% over ts span",
+    diff.ylab = "nm", diff.ylab.sec="%",
     diff.ylab.sec.accuracy = .01
   )
+  print(plot_grid(plots[[1]], plots[[2]], nrow = 2))
 
+  cat("Test set inside the CI\n")
+  cat("80% CI \t ", print_dec(mean(plotting.forecast$lower[,1] < seis.h.test & seis.h.test < plotting.forecast$upper[,1], na.rm = T)*100), "%\n")
+  cat("95% CI \t ", print_dec(mean(plotting.forecast$lower[,2] < seis.h.test & seis.h.test < plotting.forecast$upper[,2], na.rm = T)*100), "%\n")
+
+  plots <- show_overlap_and_diff(
+    seis.h.test, plotting.forecast$mean, xlab=seasonality.xlab,
+    base.desc = "test set",
+    compare.desc = paste0("forecast with ", plotting.forecast$method),
+    overlap.ylab = "Mean hourly movement (nm)",
+    diff.title = "Difference between ts",
+    diff.ylab = "nm", diff.ylab.sec="%",
+    diff.ylab.sec.accuracy = .01
+  )
   print(plot_grid(plots[[1]], plots[[2]], nrow = 2))
-    accuracy(plotting.forecast, seis.h.test)
+
+  print(accuracy(plotting.forecast, seis.h.test))
   # qqPlot(plotting.model$residuals, main = paste0("Q-Q plot for ", plotting.forecast$method), xlab = "residuals")
@@
   )
-  }}
 }
 
+```
+
+```{r auto arima, eval=FALSE, include=FALSE}
+fit <- auto.arima(seis.h.train, lambda=train.bc.lambda, d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10, trace = T)
+models[["forecast011011"]]$mean
 ```
@@
 
-```
-
-```{r auto arima}
-#fit <- auto.arima(seis.h.train, d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10, trace = T, parallel = T)
 ```

### 2020-06-01 project.uccs.Rmd
Working on draft
working on draft

--- 2020-05-22 project.uccs.Rmd
@@
 ```
+
+## Arima
+
+```{r arima}
+to_test <- list(
+  #c(0,0,0,0,0,1),  # BAD, Box-Pierce = 0
+  #c(0,0,0,0,0,2),  # BAD, Box-Pierce = 0
+  #c(0,0,0,0,1,1),  # nice, Box-Pierce = 0
+  c(0,0,0,0,1,2),
+  #c(0,0,0,0,1,3),  # CAN'T BE DONE: 3*168 = 504 > 350 (limit of arima on r)
+  #c(0,1,0,0,0,1),  # BAD, Box-Pierce = 0
+  c(0,1,0,0,1,1),
+  c(0,1,0,0,1,2),
+  c(2,0,0,0,1,0),  # hint from auto.arima D=1
+  c(2,0,0,0,1,1),  # after hint from auto.arima D=1
+  c(2,1,0,0,1,1),  # after hint from auto.arima D=1 
+  c(5,0,0,0,1,0),  # hint from auto.arima D=1
+  c(3,1,1,0,1,0),  # hint from auto.arima D=1 d=1
+  c(3,1,1,0,1,1),  # after hint then ACF + PACF
+  c(4,1,2,0,1,0),  # hint from auto.arima d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10
+  c(4,1,2,0,1,1)  # after hint then ACF + PACF
+)
+if (!exists("models")){
+  models <- list()
+}
+for (model in to_test) {
+  # model <- c(4,1,2,0,1,0)
+  fit.name <- paste0("fit", paste0(model, collapse = ""))
+  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
+  cat("SARIMA(", paste0(model[1:3], collapse = ","), ")(", paste0(model[4:6], collapse = ","), ")[", frequency(seis.h.train), "]", sep = "")
+  if (!(fit.name %in% names(models))) {
+    cat(" fitting...")
+    models[[fit.name]] <- Arima(seis.h.train, model[1:3], model[4:6])
+    cat("DONE")
+  }
+  if (!(forecast.name %in% names(models))) {
+    cat(" forecasting...")
+    models[[forecast.name]] <- forecast(models[[fit.name]], h=168*18)
+    cat("DONE")
+  }
+  cat("\n")
+}
+#models[["fitauto"]] <- auto.arima(seis.h.train, D=1, approx=F, trace = T)
+#models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*18)  
+#models[["fit311010"]] <- models[["fitauto"]]
+#models[["fit412010"]] <- fit
+#models[["forecast412010"]] <- forecast(models[["fit412010"]], h=168*18)
+
+# 000012 better than 000011, Box-Pierce = 0
+# 010011 somehow better somehow worst than 000012, Box-Pierce = 0
+# 010012 , Box-Pierce = 
+# 311010 , Box-Pierce = 
+# 311011 , Box-Pierce =
+
+for (model in to_test) {
+  # model <- c(4,1,2,0,1,0)
+  fit.name <- paste0("fit", paste0(model, collapse = ""))
+  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
+  cat("SARIMA(", paste0(model[1:3], collapse = ","), ")(", paste0(model[4:6], collapse = ","), ")[", frequency(seis.h.train), "]\n", sep = "")
+    if ((fit.name %in% names(models))) {  if ((forecast.name %in% names(models))) {
+    plotting.model <- models[[fit.name]]
+    plotting.forecast <- models[[forecast.name]]
+    cat("AIC =",plotting.model$aic, "\n")
+    checkresiduals(plotting.model)
+    residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+    residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+    print(plot_grid(residuals.acf, residuals.pacf, nrow=2))
+    print(plot_forecast(plotting.forecast) + autolayer(seis.h.test, series="Original"))
+    plots <- show_overlap_and_diff(
+      seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+      base.desc = "training set",
+      compare.desc = paste0("refit with ", plotting.forecast$method),
+      overlap.ylab = "Mean hourly movement (nm)",
+      diff.title = "Difference between ts",
+      diff.ylab = "nm", diff.ylab.sec="%",
+      diff.ylab.sec.accuracy = .01
+    )
+    print(plot_grid(plots[[1]], plots[[2]], nrow = 2))
+
+    cat("Test set inside the CI\n")
+    cat("80% CI \t ", print_dec(mean(plotting.forecast$lower[,1] < seis.h.test & seis.h.test < plotting.forecast$upper[,1], na.rm = T)*100), "%\n")
+    cat("95% CI \t ", print_dec(mean(plotting.forecast$lower[,2] < seis.h.test & seis.h.test < plotting.forecast$upper[,2], na.rm = T)*100), "%\n")
+
+    plots <- show_overlap_and_diff(
+      seis.h.test, plotting.forecast$mean, xlab=seasonality.xlab,
+      base.desc = "test set",
+      compare.desc = paste0("forecast with ", plotting.forecast$method),
+      overlap.ylab = "Mean hourly movement (nm)",
+      diff.title = "Difference between ts",
+      diff.ylab = "nm", diff.ylab.sec="%",
+      diff.ylab.sec.accuracy = .01
+    )
+    print(plot_grid(plots[[1]], plots[[2]], nrow = 2))
+
+    print(accuracy(plotting.forecast, seis.h.test))
+    # qqPlot(plotting.model$residuals, main = paste0("Q-Q plot for ", plotting.forecast$method), xlab = "residuals")
+    print(gg_qqplot(plotting.model$residuals) + 
+        labs(title=paste0("Q-Q plot for ", plotting.forecast$method, " residuals")) + ylab("Residuals quantiles")
+    )
+  }}
+}
+
+```
+
+```{r}
+#fit <- auto.arima(seis.h.train, d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10, trace = T, parallel = T)
+```
+ # D=1 m=24
+ Fitting models using approximations to speed things up...
+
+ ARIMA(2,0,2)(1,1,1)[24] with drift         : Inf
+ ARIMA(0,0,0)(0,1,0)[24] with drift         : 111462.7
+ ARIMA(1,0,0)(1,1,0)[24] with drift         : 101954.4
+ ARIMA(0,0,1)(0,1,1)[24] with drift         : Inf
+ ARIMA(0,0,0)(0,1,0)[24]                    : 111460.9
+ ARIMA(1,0,0)(0,1,0)[24] with drift         : 102426.5
+ ARIMA(1,0,0)(2,1,0)[24] with drift         : 101379
+ ARIMA(1,0,0)(2,1,1)[24] with drift         : Inf
+ ARIMA(1,0,0)(1,1,1)[24] with drift         : Inf
+ ARIMA(0,0,0)(2,1,0)[24] with drift         : 110647.6
+ ARIMA(2,0,0)(2,1,0)[24] with drift         : 101332.9
+ ARIMA(2,0,0)(1,1,0)[24] with drift         : 101878.7
+ ARIMA(2,0,0)(2,1,1)[24] with drift         : Inf
+ ARIMA(2,0,0)(1,1,1)[24] with drift         : Inf
+ ARIMA(3,0,0)(2,1,0)[24] with drift         : 101317.9
+ ARIMA(3,0,0)(1,1,0)[24] with drift         : 101858.5
+ ARIMA(3,0,0)(2,1,1)[24] with drift         : Inf
+ ARIMA(3,0,0)(1,1,1)[24] with drift         : Inf
+ ARIMA(4,0,0)(2,1,0)[24] with drift         : 101302.7
+ ARIMA(4,0,0)(1,1,0)[24] with drift         : 101845
+ ARIMA(4,0,0)(2,1,1)[24] with drift         : Inf
+ ARIMA(4,0,0)(1,1,1)[24] with drift         : Inf
+ ARIMA(5,0,0)(2,1,0)[24] with drift         : 101305.4
+ ARIMA(4,0,1)(2,1,0)[24] with drift         : 101304.5
+ ARIMA(3,0,1)(2,1,0)[24] with drift         : 101307.7
+ ARIMA(5,0,1)(2,1,0)[24] with drift         : 101307.7
+ ARIMA(4,0,0)(2,1,0)[24]                    : 101300.7
+ ARIMA(4,0,0)(1,1,0)[24]                    : 101843
+ ARIMA(4,0,0)(2,1,1)[24]                    : Inf
+ ARIMA(4,0,0)(1,1,1)[24]                    : Inf
+ ARIMA(3,0,0)(2,1,0)[24]                    : 101316
+ ARIMA(5,0,0)(2,1,0)[24]                    : 101303.4
+ ARIMA(4,0,1)(2,1,0)[24]                    : 101302.4
+ ARIMA(3,0,1)(2,1,0)[24]                    : 101305.8
+ ARIMA(5,0,1)(2,1,0)[24]                    : 101305.7
+
+ Now re-fitting the best model(s) without approximations...
+
+ ARIMA(4,0,0)(2,1,0)[24]                    : 101586.6
+
+ Best model: ARIMA(4,0,0)(2,1,0)[24]
+
+ # d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10 m=168
+ Fitting models using approximations to speed things up...
+
+ ARIMA(2,1,2)(1,1,1)[168]                    : Inf
+ ARIMA(0,1,0)(0,1,0)[168]                    : 122379.8
+ ARIMA(1,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(0,1,1)(0,1,1)[168]                    : Inf
+ ARIMA(0,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(0,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(0,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(1,1,0)(0,1,0)[168]                    : 121638
+ ARIMA(1,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(1,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(2,1,0)(0,1,0)[168]                    : 121468.5
+ ARIMA(2,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(2,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(2,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(3,1,0)(0,1,0)[168]                    : 121452.9
+ ARIMA(3,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(3,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(3,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(4,1,0)(0,1,0)[168]                    : 121403.9
+ ARIMA(4,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(4,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(4,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(5,1,0)(0,1,0)[168]                    : 121361.7
+ ARIMA(5,1,0)(1,1,0)[168]                    : Inf
+ ARIMA(5,1,0)(0,1,1)[168]                    : Inf
+ ARIMA(5,1,0)(1,1,1)[168]                    : Inf
+ ARIMA(5,1,1)(0,1,0)[168]                    : 118580.2
+ ARIMA(5,1,1)(1,1,0)[168]                    : Inf
+ ARIMA(5,1,1)(0,1,1)[168]                    : Inf
+ ARIMA(5,1,1)(1,1,1)[168]                    : Inf
+ ARIMA(4,1,1)(0,1,0)[168]                    : 118580.3
+ ARIMA(5,1,2)(0,1,0)[168]                    : 118581.5
+ ARIMA(4,1,2)(0,1,0)[168]                    : 118553.1
+ ARIMA(4,1,2)(1,1,0)[168]                    : Inf
+ ARIMA(4,1,2)(0,1,1)[168]                    : Inf
+ ARIMA(4,1,2)(1,1,1)[168]                    : Inf
+ ARIMA(3,1,2)(0,1,0)[168]                    : 118642.8
+ ARIMA(3,1,1)(0,1,0)[168]                    : 119155
+
+ Now re-fitting the best model(s) without approximations...
+
+ ARIMA(4,1,2)(0,1,0)[168]                    : 122838.5
+
+ Best model: ARIMA(4,1,2)(0,1,0)[168]
+
+```{r eval=FALSE, include=FALSE}
+
+autoplot(acf(seis.h.train.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
+autoplot(pacf(seis.h.train.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
+
+gg_qqplot(fit$residuals)
+gghistogram(fit$residuals)
+Box.test(fit$residuals, type="Box-Pierce")
+fit.forec <- forecast(fit, h=24*10)
+plot(fit.forec, xlim=c(250, 300))
+
+frequency(seis.h.train)
+p<-168;d<-0;q<-0;P<-0;D<-1;Q<-0;
+paste0("ARIMA(", p, ",", d, ",", q, ")(", P, ",", D, ",", Q, ")_", frequency(seis.h.train))
+fit <- Arima(seis.h.train, order=c(p,d,q), seasonal=c(P,D,Q))
+fit
+#Diagnostics
+tsdisplay(residuals(fit), lag.max=168*2)
+
+p<-1;d<-0;q<-2;P<-1;D<-0;Q<-2;
+fit <- Arima(seis.h.train, order=c(p,d,q), seasonal=c(P,D,Q))
+#Diagnostics
+tsdisplay(residuals(fit))
+forecast()
+
+```
+
+## MSTL
+```{r mstl on msts}
+seis.h.train.m <- msts(seis.h.train, c(24, 168))
+seis.mstl <- seis.h.train.m %>% mstl(s.window="periodic") #c(15, 7))
+seis.mstl.fc <- forecast(seis.mstl, h=168*18)
+
+autoplot(seis.mstl) + xlab("Week")
+plot(seis.mstl[,2], main="trend")
+lines(trend.s, col="red")
+plot(rowSums(seis.mstl[,3:4]), main="seasonality",type="l")
+lines(seasonality, col="red")
+plot(seis.mstl[,5],  main="residuals")
+lines(residuals, col="red")
+
+plotting.model <- seis.mstl
+plotting.forecast <- seis.mstl.fc
+checkresiduals(plotting.model[,5])
+Box.test(plotting.forecast$model$residuals, type="L")
+residuals.acf <- autoplot(acf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+residuals.pacf <- autoplot(pacf(plotting.model[,5], na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+plot_grid(residuals.acf, residuals.pacf, nrow=2)
+
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Original") +
+  xlim(52, 70) + ylim(0, 8750)
+plot_forecast(plotting.forecast) + 
+  autolayer(seis.h.test, series="Original")
+plots <- show_overlap_and_diff(
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
+  base.desc = "train series",
+  compare.desc = paste0("refit with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+plots <- show_overlap_and_diff(
+  seis.h.test[1:length(plotting.forecast$mean)], plotting.forecast$mean, xlab=seasonality.xlab,
+  base.desc = "test series",
+  compare.desc = paste0("refit with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="% over ts span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+accuracy(plotting.forecast, seis.h.test)
+gg_qqplot(plotting.model[,5])
+
+bptests <- data.frame(lag=integer(), `p-value`=numeric())
+for (i in 1:10){
+  bptests[nrow(bptests) + 1, ] <- c(i, Box.test(plotting.model[,5], lag = i, fitdf = 0, type = "Box-Pierce")$p.value)
+}
+bptests.plot <- 
+  ggplot(bptests, aes(lag, `p.value`), pch="o") + 
+    geom_point() + ylim(0,1) +
+    labs(title="Box-Pierce tests for lags 1 to 10") +
+    geom_hline(yintercept=0.05, linetype="dotted", color = "blue") +
+    scale_x_continuous(breaks=1:10)
+bptests.plot
+
+```
+
+```{r save RData, include=FALSE}
+gc()
+save.image(proj_file)
+```

### 2020-06-01 draft.tex
delivered draft

@@
 This initial division demostrated to be computationally heavy while fitting some models with weekly seasonality, and we noticed a possible yearly seasonality on Uccle data (\figurename~\ref{fig:whole-data-decomposition}).
-Even if it is not demonstrable given the short period our data comes from, applying a 24-hour MA smoothing filter on the whole dataset and then producing the periodgram as previously done, it suggests a 8748-hours significant periodicity (364.5 days).
+Even if it is not demonstrable given the short period our data comes from, applying a 24-hour MA smoothing filter on the whole dataset (546 days) and then producing the periodgram as previously done, it suggests a 8748-hours significant periodicity (364.5 days).
 
@@
 	\item seasonal MA process of order $Q$ is described by an exponentially decay at seasonal lags on PACF values, and by a cut-off of significant ACF values at each seasonal lag until the $Q^{th}$
-	\item given how significance is assessed for the ACF/PACF spikes, one over 20 spikes could be \textit{coincidentally significant}
+	\item given how significance is assessed for the ACF/PACF spikes, one over 20 spikes (5\%) could be \textit{coincidentally significant} or \textit{coincidentally \textbf{non}-significant}
 \end{itemize}
@@
 	\centering
-	\includegraphics[width=.7\linewidth]{project.uccs_files/figure-latex/arima-26}
+	\includegraphics[width=.7\linewidth]{project.uccs_files/figure-latex/arima-30} %26
 	\caption{ACF and PACF plots for $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$}
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-27}
+		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-32} %27
 		\caption{$\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$}
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-51}
+		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-60} %51
 		\caption{$\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$}
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=1\linewidth]{"project.mems_files/figure-latex/arima-15"} %18
+		\includegraphics[width=1\linewidth]{"project.mems_files/figure-latex/arima-18"} %15
 		\caption{$\mathrm{ARIMA}(0,1,1)(0,1,1)_{168}$}
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=1\linewidth]{"project.mems_files/figure-latex/arima-9"} %11
+		\includegraphics[width=1\linewidth]{"project.mems_files/figure-latex/arima-11"} %9
 		\caption{$\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$}
@@
 
-Analytically, we have quite a difference between the two models in terms of adherence of the confidence intervals.
+Analytically, we have quite a difference between the two models in terms of adherence of the confidence intervals:
 \begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
@@
 
-\subsection{Seasonality comparison}
-\subsection{Lockdown forecast divergence}
 \section{Conclusion}
-\subsubsection{Limitations}
-\subsubsection{Future Works}
+\subsection{Lockdown influence}
+Even if the two models have the same order, they cannot be directly compared, for the Membach model works after a transformation pass.
 
-\subsubsection*{Hypotesis of analysis}
+We can still compare how the seasonalities differ with respect with each original dataset and between each other.
+We extract the seasonalities for each station highlighting the standard deviation confidence band for an estimated error of 67\%.
+This estimation lies on the assumption that values for each single hour in the hourly/weekly seasonality is Normally distributed, a slightly different assumption that is not necessarily refuted by the already seen Q-Q plots.
 
-\begin{enumerate}
-	\item Download data from start of March 2019 to present day.
-	\item Aggregate the data in slots of 30 minutes or 1 hour, keeping the maximum ground displacements in both negative and positive directions to maintain the same variables meaning.
-	\item Compute the total ground displacement for each result point.
-	\item Verify that we have a stationary time series (we expect that for the meaning that it brings: ground is not moving up or down).
-	\item Extract through a linear filter (probably MA) different seasonalities and residuals.
-\end{enumerate}
+\begin{figure}[h]
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/seasonalities/before-after-compare-1"}
+		\caption{Uccle station}
+		\label{fig:season.lockdown:uccs}
+	\end{subfigure}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/before-after-compare-1"}
+		\caption{Membach station}
+		\label{fig:season.lockdown:mems}
+	\end{subfigure}
+	\begin{center}
+		\caption{Seasonalities computed before and after the lockdown}{\small Seasonalities are extracted together}
+		\label{fig:season.lockdown}
+	\end{center}
+\end{figure}
 
-The extraction of seasonalities is the focus point of this work: as \cite{NatureCoronavirusSeismic} seems to conclude, there is a clear weekly seasonality in at least the \textbf{Uccle} data.
-We should test against different seasonalities through the stations, extracting more at once if the residuals suggest their presence.
+As we can see in \figurename~\ref{fig:season.lockdown}, even if the seasonalities shapes seem very different, in particular regarding the weekly seasonality, the differences are still included in the highlighted confidence bands, apart from a few hours.
+This is true for both stations.
+Yet, it is useful to notice that the weekly seasonalities include a possibly interesting daily component, influenced by the weekend drops in values.
 
-Fitted models with these seasonalities, only from data before the Belgium lockdowns (on March, 14th the "soft" lockdown, while on March, 18th the complete lockdown), should let us forecast data for after the lockdown and test for statistical difference.
+%\subsection{Lockdown forecast divergence}
+%Uccle: remodel on 52+18 (until 1st March 2020), forecast 2 months (8 weeks) => diff in MAPE
+%Membach: remodel on 52+18 (until 1st March 2020), forecast 2 months (8 weeks) => diff in MAPE
 
-Also, comparison between models and seasonalities from chosen stations could be done, to assess any statistical difference between apparent human impact registered from places with different standard human activity.
+\subsection{Limitations and future works}
+The main limitations of this projects that we can identify are: multiple seasonalities and Normality assumption.
 
-Eventually, some attention could be given to events registered through the analysed time.
+The tested models do not effectively include both daily and weekly seasonalities, even if initially we thought so.
+Also they cannot include the apparent yearly seasonality, mainly because of the short timeseries.
+To overcome this limitation, we could try different models, also ARIMA-based, that can estimate multiple seasonalities.
+The first that we tested models seasonalites through Fourier terms.
+The main problem that lead us not to go further with that approach is that our seasonalities cannot be expressed by a combination of few sine and cosine curves.
+This requires a high number of curves to estimate, that requires an heavy computational power.
+
+To compensate those difficulties, in the literature it is suggested to test TBATS models, which combines Fourier terms, Box-Cox transformation, and exponential smoothing.
+This seems a suitable approach given the analyses we performed until now, that can possibly address the non-Normal distribution.
+
+The final suggestion for future work would be to better analyse the differences between stations, in such a way to better correlate different influences of the lockdown on the areas near each station.
 

### 2020-06-01 project.Rmd
delivered draft

@@
   seasonality.figure <- as.ts(colMeans(matrix, na.rm = T))
-  seasonality <- rep(seasonality.figure, length(extended)/length(seasonality.figure))[1:length(ts)]
-  return(list(figure=seasonality.figure, seasonal=seasonality))
+  seasonality.sd <- as.ts(apply(matrix, 2, sd, na.rm = T))
+  seasonality <- rep(seasonality.figure, length(extended)/length(seasonality.figure))[1:length(ts.in)]
+  return(list(figure=seasonality.figure, seasonal=seasonality, sd=seasonality.sd))
 }

### 2020-06-01 project.after.Rmd
delivered draft

```{r final-model}
# chosen.model must be c(p,d,q,P,D,Q)
# seasonality.period a number
# chosen.lambda NULL or Box-Cox lambda

seis.test.final.start <- seis.test.end + seconds(1)
seis.test.final.end <- seis.test.final.start + days(7*8) - seconds(1)
c(seis.train.start, seis.test.final.start, seis.test.final.end)

final.training <- build_ts(seis.h, periods = c(seasonality.period), start = seis.train.start, end = seis.test.end, force_ms = T)
final.testing <- build_ts(seis.h, periods = c(seasonality.period), start = seis.train.start, end = seis.test.final.end, force_ms = T)
final.testing <- window(final.testing, start = c(end(final.training)[1] + 1, 1))
c(chosen.lambda, BoxCox.lambda(final.training))

fit.name <- paste0("fit", paste0(chosen.model, collapse = ""))
forecast.name <- paste0("forecast", paste0(chosen.model, collapse = ""))
cat("SARIMA(", paste0(chosen.model[1:3], collapse = ","), ")(", paste0(chosen.model[4:6], collapse = ","), ")[", frequency(final.training), "]", sep = "")
if (!exists("final.model")) {
  cat(" fitting...")
  final.model <- Arima(final.training, chosen.model[1:3], chosen.model[4:6], lambda=chosen.lambda)
  cat("DONE")
}
if (!exists("final.forecast")) {
  cat(" forecasting...")
  final.forecast <- forecast(final.model, h=168*8, lambda=chosen.lambda)
  cat("DONE")
}
cat("\n")
plotting.model <- final.model
plotting.forecast <- final.forecast
cat("AIC =",plotting.model$aic, "\n")
cat("AICc =",plotting.model$aicc, "\n")
checkresiduals(plotting.model)
residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*4)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*4)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
print(plot_grid(residuals.acf, residuals.pacf, nrow=2))
print(plot_grid(
  autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
  ,
  residuals.acf
  ,
  autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
  ,
  residuals.pacf
  ,
  nrow = 2
))
print(plot_forecast(plotting.forecast) + autolayer(final.testing, series="Original") + ylab("Mean hourly movement (nm)"))
plots <- show_overlap_and_diff(
  final.training, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "training set",
  compare.desc = paste0("refit with ", plotting.forecast$method),
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="%",
  diff.ylab.sec.accuracy = .01
)
print(plot_grid(plots[[1]], plots[[2]], nrow = 2))

cat("Test set inside the CI\n")
cat("80% CI \t ", print_dec(mean(plotting.forecast$lower[,1] < final.testing & final.testing < plotting.forecast$upper[,1], na.rm = T)*100), "%\n")
cat("95% CI \t ", print_dec(mean(plotting.forecast$lower[,2] < final.testing & final.testing < plotting.forecast$upper[,2], na.rm = T)*100), "%\n")

plots <- show_overlap_and_diff(
  final.testing, plotting.forecast$mean, xlab=seasonality.xlab,
  base.desc = "test set",
  compare.desc = paste0("forecast with ", plotting.forecast$method),
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="%",
  diff.ylab.sec.accuracy = .01
)
print(plot_grid(plots[[1]], plots[[2]], nrow = 2))

print(accuracy(plotting.forecast, final.testing))
# qqPlot(plotting.model$residuals, main = paste0("Q-Q plot for ", plotting.forecast$method), xlab = "residuals")
print(gg_qqplot(plotting.model$residuals) + 
    labs(title=paste0("Q-Q plot for ", plotting.forecast$method, " residuals")) + ylab("Residuals quantiles")
)

```

```{r seasonalities/seasonplot}
final.training.24 <- build_ts(final.training, periods = c(24))
seasonplot(final.training.24/sd(final.training.24, na.rm = T), type="l", col=rainbow(7))
legend("topleft", legend=weekdays(seis.train.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)

final.testing.24 <- build_ts(final.testing, periods = c(24))
seasonplot(final.testing.24/sd(final.training.24, na.rm = T), type="l", col=rainbow(7))
legend("topleft", legend=weekdays(seis.test.final.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)

```

```{r seasonalities/before}
aggregated.final <- seasonality.aggregate(final.training)
season.final.24 <- trend_season(aggregated.final, 24)
season.final.168 <- trend_season(aggregated.final-season.final.24$seasonal, 168)

season.final.24 <- trend_season(aggregated.final-season.final.168$seasonal, 24)
season.final.168 <- trend_season(aggregated.final-season.final.24$seasonal, 168)

# this third iteration isn't actually needed
season.final.24 <- trend_season(aggregated.final-season.final.168$seasonal, 24)
season.final.168 <- trend_season(aggregated.final-season.final.24$seasonal, 168)

residuals.final <- aggregated.final-season.final.24$seasonal-season.final.168$seasonal-season.final.168$trend

seasonbase.final.24 <- trend_season(aggregated.final, 24)
seasonbase.final.168 <- trend_season(aggregated.final, 168)

plot_grid(
  autoplot(ts(seasonbase.final.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
    geom_line(aes(y=season.final.24$figure, col="Multi daily")) +
    scale_colour_manual("", values=c("purple", "yellow"), breaks=c("Direct daily", "Multi daily")),
  autoplot(ts(seasonbase.final.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
    geom_line(aes(y=season.final.168$figure + rep(season.final.24$figure, 7), col="Multi w+d")) +
    #geom_line(aes(y=rep(seasonbase.final.24$figure, 7), col="Direct daily"), size=1.5) +
    #geom_line(aes(y=rep(season.final.24$figure, 7), col="Multi daily")) +
    geom_line(aes(y=season.final.168$figure, col="Multi weekly")) +
    scale_colour_manual("", values=c("red", "purple", "blue", "yellow", "green"), breaks=c("Direct weekly", "Direct daily", "Multi" %s+% c(" weekly", " daily", " w+d"))),
  nrow = 2)

```

```{r seasonalities/lockdown}
aggregated.lockdown <- seasonality.aggregate(final.testing)
season.lockdown.24 <- trend_season(aggregated.lockdown, 24)
season.lockdown.168 <- trend_season(aggregated.lockdown-season.lockdown.24$seasonal, 168)

season.lockdown.24 <- trend_season(aggregated.lockdown-season.lockdown.168$seasonal, 24)
season.lockdown.168 <- trend_season(aggregated.lockdown-season.lockdown.24$seasonal, 168)

# this third iteration isn't actually needed
season.lockdown.24 <- trend_season(aggregated.lockdown-season.lockdown.168$seasonal, 24)
season.lockdown.168 <- trend_season(aggregated.lockdown-season.lockdown.24$seasonal, 168)

residuals.lockdown <- aggregated.lockdown-season.lockdown.24$seasonal-season.lockdown.168$seasonal-season.lockdown.168$trend

seasonbase.lockdown.24 <- trend_season(aggregated.lockdown, 24)
seasonbase.lockdown.168 <- trend_season(aggregated.lockdown, 168)

plot_grid(
  autoplot(ts(seasonbase.lockdown.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
    geom_line(aes(y=season.lockdown.24$figure, col="Multi daily")) +
    scale_colour_manual("", values=c("purple", "yellow"), breaks=c("Direct daily", "Multi daily")),
  autoplot(ts(seasonbase.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
    geom_line(aes(y=season.lockdown.168$figure + rep(season.lockdown.24$figure, 7), col="Multi w+d")) +
    #geom_line(aes(y=rep(seasonbase.lockdown.24$figure, 7), col="Direct daily"), size=1.5) +
    #geom_line(aes(y=rep(season.lockdown.24$figure, 7), col="Multi daily")) +
    geom_line(aes(y=season.lockdown.168$figure, col="Multi weekly")) +
    scale_colour_manual("", values=c("red", "purple", "blue", "yellow", "green"), breaks=c("Direct weekly", "Direct daily", "Multi" %s+% c(" weekly", " daily", " w+d"))),
  nrow = 2)

```

```{r seasonalities/before-after-compare}

plot_grid(
  autoplot(ts(season.lockdown.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
    geom_ribbon(aes(ymin = season.lockdown.24$figure - season.lockdown.24$sd, ymax = season.lockdown.24$figure + season.lockdown.24$sd), fill = "mediumpurple1", alpha=0.5) +
    geom_line(aes(col="Daily lockdown")) +
    geom_ribbon(aes(ymin = season.final.24$figure - season.final.24$sd, ymax = season.final.24$figure + season.final.24$sd), fill = "lightcoral", alpha=0.5) +
    geom_line(aes(y=season.final.24$figure, col="Daily before")) +
    scale_colour_manual("", values=c("purple", "red"), breaks=c("Daily lockdown", "Daily before")),
  autoplot(ts(season.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
    geom_ribbon(aes(ymin = season.lockdown.168$figure - season.lockdown.168$sd, ymax = season.lockdown.168$figure + season.lockdown.168$sd), fill = "mediumpurple1", alpha=0.5) +
    geom_line(aes(col="Weekly lockdown")) +
    geom_ribbon(aes(ymin = season.final.168$figure - season.final.168$sd, ymax = season.final.168$figure + season.final.168$sd), fill = "lightcoral", alpha=0.5) +
    geom_line(aes(y=season.final.168$figure, col="Weekly before")) +
    scale_colour_manual("", values=c("purple", "red"), breaks=c("Weekly lockdown", "Weekly before")),
nrow = 2)

autoplot(ts(seasonbase.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
  geom_ribbon(aes(ymin = seasonbase.lockdown.168$figure - seasonbase.lockdown.168$sd, ymax = seasonbase.lockdown.168$figure + seasonbase.lockdown.168$sd), fill = "mediumpurple1", alpha=0.5) +
  geom_line(aes(col="Weekly lockdown")) +
  geom_ribbon(aes(ymin = seasonbase.final.168$figure - seasonbase.final.168$sd, ymax = seasonbase.final.168$figure + seasonbase.final.168$sd), fill = "lightcoral", alpha=0.5) +
  geom_line(aes(y=seasonbase.final.168$figure, col="Weekly before")) +
  scale_colour_manual("", values=c("purple", "red"), breaks=c("Weekly lockdown", "Weekly before"))
```

### 2020-06-01 project.mems.Rmd
delivered draft

@@
 fit <- auto.arima(seis.h.train, lambda=train.bc.lambda, d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10, trace = T)
-models[["forecast011011"]]$mean
+models[["forecast011011"]]$upper-models[["forecast011011"]]$mean
 ```
@@
 
+```{r final model selection}
+chosen.model <- c(2,0,0,0,1,1)
+chosen.lambda <- train.bc.lambda
+```
+
+```{r project, child='project.after.Rmd'}
+```
+
 ```{r save RData, include=FALSE}

### 2020-06-02 draft.tex
delivered draft errata corrige

@@
 	\centering
-	\includegraphics[width=.7\linewidth]{project.uccs_files/figure-latex/arima-30} %26
+	\includegraphics[width=.7\linewidth]{project.uccs_files/figure-latex/arima-26} %30
 	\caption{ACF and PACF plots for $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$}
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-32} %27
+		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-27} %32
 		\caption{$\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$}
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-60} %51
+		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-51} %60
 		\caption{$\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$}

### 2020-06-02 project.Rmd
start moving all R code into project.R for delivery

@@
 blocks$interval <- seconds(blocks$end - blocks$start + 1)
-blocks$start <- seis.first.second + seconds(blocks$start)
-blocks$end <- seis.first.second + seconds(blocks$end)
+blocks$start <- seis.first.second + seconds(blocks$start - 1)
+blocks$end <- seis.first.second + seconds(blocks$end - 1)
 plot(blocks$start, blocks$interval,
@@
 force_bruxelles <- function(d) force_tz(d, "Europe/Brussels")
+force_utc <- function(d) force_tz(d, "UTC")
 
@@
 tz-ed time is normally evaluated in utc by ts-related functions
-  group_by(utc_hour=floor_date(force_tz(bruxelles(utc_second), "UTC"), "1 hour")) %>%
+  group_by(utc_hour=floor_date(force_utc(bruxelles(utc_second)), "1 hour")) %>%
   summarize(mean_movement_s=mean(max-min)) # %>%
@@
 seis.missing <- seis.total - seis.present
-seis.present %s+% " present values, " %s+% seis.missing %s+% " missing points (" %s+% print_dec(seis.present / seis.total * 100, 2) %s+% "% / " %s+% print_dec(seis.missing / seis.total * 100, 2) %s+% "%)"
+cat(seis.present, "present values,", seis.missing, "missing points (", print_dec(seis.present / seis.total * 100, 2), "% /", print_dec(seis.missing / seis.total * 100, 2), "% )\n")
 
@@
 blocks$interval <- hours(blocks$end - blocks$start + 1)
-blocks$start <- seis.first.second + hours(blocks$start)
-blocks$end <- seis.first.second + hours(blocks$end)
+blocks$start <- seis.first.second + seconds(blocks$start - 1)
+blocks$end <- seis.first.second + seconds(blocks$end - 1)
 plot(c(rbind(blocks$start, blocks$end, NA)), rep(hour(blocks$interval), each=3),
@@
 blocks$interval
-Box.test(seis.h$mean_movement_s, type="Box-Pierce")
 
@@
       labs(title=paste0(base.desc, " vs ", compare.desc))
-  span <- max(compare, base, na.rm = T) - min(compare, base, na.rm = T)
   diff <- compare-base
-  accu.lims <- diff.ylab.sec.accuracy / 100 * span * c(-3,3)
 
@@
   R2 <- 1-(sum((diff)^2, na.rm = T)/sum((base-mean(base))^2, na.rm = T))
-  errorstring <- paste0("MSE=", mse, "  MAE=", mae, "  MAPE=", mape, "  RMSE=", rmse, "  R2=", R2)
   cat("Errors: ", compare.desc, "\n")
@@
 
+  span <- max(compare, base, na.rm = T) - min(compare, base, na.rm = T)
+  accu.lims <- diff.ylab.sec.accuracy / 100 * span * c(-3,3)
   diff_plot <-

### 2020-06-02 project.R

@@ -1,0 +1,356 @@
+
+library(stringi) # %s+%
+library(lubridate) # floor_date
+library(xts) # zoo
+library(fpp) # auto.arima
+library(ggplot2) # plots
+library(ggrepel) # for nicer labels on ggplot
+library(cowplot) # for plot_grid
+library(qqplotr) # qqplot with ggplot
+library(spec) # for periodgram
+library(tidyverse) # map_df, read_csv, group_by, summarize...
+library(tsbox) # https://cran.r-project.org/web/packages/tsbox/vignettes/tsbox.html
+
+# Before starting we define this helper function for printing formatted decimals
+print_dec <- function(x, decimals=2) trimws(format(round(x, decimals), nsmall=decimals))
+
+
+
+# If needed, set working directory. The folder structure inside it should contain
+# - csv/
+#   - UCCS*.csv files
+#   - MEMS*.csv files
+# - proj.*.RData (optional)
+# setwd("/media/data/Documents/Uni/spatiotemp/villalobos/project/")
+
+
+
+# We define some variables that identify the dataset. Only the row about the wanted dataset should be run
+# UCCS
+domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
+# MEMS
+domain <- c(-7200,7200); pattern <- "MEMS.*\\.csv"; name <- "MEMS"; description <- "Membach"; proj_file <- 'proj.mems.RData'
+
+# If the corresponding RData file is present, we load it mainly to avoid refitting all models
+if (exists("proj_file") & file.exists(proj_file)) {
+  load(proj_file)
+}
+
+
+
+# Load all csv files in the folder corresponding to the defined pattern, station-dependant
+seis <-
+  list.files(path="csv/", pattern=pattern) %>%
+  map_df(~read_csv("csv/" %s+% .))
+
+
+
+# Prepare a description for plots and see print some information about the data
+long_description <- description %s+% " (" %s+% name %s+% " station)"
+cat(long_description, "\n")
+cat("Domain values:", domain[1], "to", domain[2], "nm\n")
+cat("Data from", format.POSIXct(min(seis$utc_second), "%Y-%m-%d %H:%M:%S"), "to", format.POSIXct(max(seis$utc_second), "%Y-%m-%d %H:%M:%S"), "\n")
+
+
+# As a first step, we remove all values out of domain, and all invalid values: (-1,1) and (0,0) pairs
[…]

### 2020-06-06 draft.tex
fix draft as Diletta and Fabio suggested

@@
 \subsubsection{Motivation and objective}
-Thanks to the recent coronavirus outbreak worldwide, we currently have data about many different phoenomena from periods with standard human activity, and from a period with reduced human activity.
+Thanks to the recent COVID-19 outbreak worldwide, we currently have data about many different phoenomena from periods with standard human activity, and from a period with reduced human activity.
 
@@
 
-It is important to notice that the \textit{day}ly data in each file is considered in UTC time, an optimal choice to publish data in a machine-readable way, even if this could cause some issues while interpreting the results.
+It is important to notice that the daily data in each file refers to the UTC time instead of the local Europe/Brussels time, an optimal choice to publish data in a machine-readable way, even if this could cause some issues while interpreting the results.
 
@@
 \subsubsection{Data aggregation and initial transformation}
-Since we are not interested in second-precise analysis, and it would be computationally expensive to deal with the whole dataset, we are going to aggregate data by hour.
+Since we are not interested in a second-precise analysis, and it would be computationally expensive to deal with the whole dataset, we are going to aggregate data by hour.
 The aggregation will get us around 14 thousand points, that can still result in interpretable models.
@@
 
-In particular, after a visual analysis after decomposition, we can suppose the 90 hours gap from the Uccle station is related to some changes to the seismic equipment, that resulted in slightly more precise measurements.
+In particular, after a visual analysis of the decomposition results, we can suppose the 90 hours gap from the Uccle station is related to some changes to the seismic equipment, that resulted in slightly more precise measurements.
 
@@
 
-Since our main focus is on the seasonalities themselves, in order to prevent assumptions based on the objective, so that searching for correlation with human activity becomes imposing human-based seasonality periods, we will use an analytical method to extract seasonalities that the data itself suggests to us.
+Since our main focus is on the seasonalities themselves, we aim to prevent assumptions based on this objective.
+In other words, we want to avoid searching for correlation with human activity while imposing human-based seasonality periods.
+Thus, we will use an analytical method to extract seasonalities that the data itself suggests to us.
 Still, we recognize that the already performed transformations impose an human-based time alteration to the data itself (Section \ref{sec:cleanup-transform}).
@@
 On this note we must specify that, when a significant periodicity is found, it influences smaller periodicities given by $f*\{2,3,\dots\}$ the significant frequency.
-Also, if any periodicity is found that is longer than the original dataset, it must be considered as a \textit{coincidence}, rather than a surprising serendipity.
+Also, if any periodicity is found that is longer than the original dataset, it must be considered as simple chance, rather than serendipity.
 
@@
 Still, if we notice that the main timeseries lies between values of 0 and 100, we can find out that around half of the value will be composed by the trend, a 30\% by the joined seasonalities component, and the remaining value by the residuals.
-We will need to better handle the spikes from around half our dataset.
+We will need to better handle the spikes starting from the middle of our dataset.
 
@@
 \end{figure}
+
+Some R commands cannot handle timeseries with missing values, but as we have seen both stations datasets include missing values.
+Also, please note that we are going to have a timeseries object indexed by weeks, so that it contains the 168 hours between each integer.
+Since we know that the longest missing values series is 90 hours long, we can apply a simple completion strategy for serving those commands a filled timeseries: we will substitute missing values with the mean of the week they are in; in R terms \texttt{na.aggregate(x, floor)}.
 
@@
 
-Since our dataset shows a relevant trend component, resulting in non-constant mean, we consider the usage of the \textit{diff} operator to smooth the mean.
+Since our dataset shows a relevant trend component, resulting in non-constant mean, we consider the usage of the \textit{diff}\footnote{The \textit{diff} operator applies the \textit{d}-order differences, by default on lag 1 (\textit{simple}).
+All \textit{diff} operators we will apply use $d=1$.} operator to smooth the mean.
 After applying the \textit{diff} operator with lag 168, in conjunction with the simple \textit{diff} or not, we come up with two possible models by checking the ACF and PACF plots:
@@
 \end{itemize}
-We also employ the \textit{auto.arima} stepwise methodology, locking the \textit{diff} values to 0 or 1, to get some hints on the possible models.
+We also employ the \textit{auto.arima} stepwise methodology\footnote{The \textit{auto.arima} stepwise methodology fits different ARIMA models within a range of orders, and based on each result's information criteria it will try to minimize it by changing each time one of the \textit{p}, \textit{d}, \textit{q}, \textit{P}, \textit{D}, or \textit{Q} values, depending on the configured limits.}, locking the \textit{diff} values to 0 or 1, to get some hints on the possible models.
 
@@
 \begin{figure}[h]
-	\centering
-	\includegraphics[width=.7\linewidth]{project.uccs_files/figure-latex/arima-26} %30
-	\caption{ACF and PACF plots for $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/arima-26} %30
+		\caption{ACF and PACF plots}
 		\label{fig:arima-26}
+	\end{subfigure}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/arima-24} %28
+		\caption{Q-Q plot}
+		\label{fig:arima-24}
+	\end{subfigure}
+	\begin{center}
+		\caption{Plots for $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$}
+		\label{fig:arima}
+	\end{center}
 \end{figure}
@@
 This shows how our previous conclusion that the 168 hours seasonality could include the 24 hours one is flawed.
-Unfortunately, the parameters estimator we are using does not permit to estimate multiple seasonalities\footnote{\textit{Arima} method from the \textit{forecast} R package, only specifying the seasonal and non-seasonal orders}.
+Unfortunately, the parameters estimator we are using does not permit to estimate multiple seasonalities\footnote{\textit{Arima} method from the \textit{forecast} R package, only specifying the seasonal and non-seasonal orders.}.
 
-Probably related to the 24 hours periodicity that remains on the residuals, together with the non-Normality of the original data, all Q-Q plots show clear divergence on both sides of the distribution.
+Probably related to the 24 hours periodicity that remains on the residuals, together with the non-Normality of the original data, all Q-Q plots show clear divergence on both sides of the distribution (\figurename~\ref{fig:arima-24}).
 The Box-Pierce test statistics accept the independence hypothesis between residuals.
@@
 In this case, the employment of \textit{auto.arima} stepwise methodology does not lead to any useful model, since it obtains $\mathrm{AICc}=\inf$ for the two manually-chosen models, that have an AICc value around -84.500, while the best model found by that methodology has an AICc around -81.000.
-This can be caused by approximation errors\footnote{The \textit{auto.arima} methodology was used only with initial approximation, due to high computing time}, as we supposed earlier.
+This can be caused by approximation errors\footnote{The \textit{auto.arima} methodology was only used with initial approximation, due to high computing time.}, as we supposed earlier.
 
@@
 As we see, the first model over-generalises the variability, while the latter seems more precise, even if losing some information.
-Also considering the MAPE on the test set, the latter model fits the data better with a $16\%$ error, instead of a $18\%$.
+Also considering the MAPE on the test set, the latter model fits the data better with a $16\%$ error, instead of the $18\%$ for the first model.
 
-Thus, we can select the $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$ model on the Box-Cox($\lambda\approx-1$)-transformed data, still being aware of the missed 24 hours seasonality, similarly to the Uccle station.
+Thus, we can select the $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$ model on the Box-Cox($\lambda\approx-1$) transformed data, still being aware of the missed 24 hours seasonality, similarly to the Uccle station.
 
@@
 \subsection{Lockdown influence}
-Even if the two models have the same order, they cannot be directly compared, for the Membach model works after a transformation pass.
+Even if the two models have the same order, they cannot be directly compared using the Information Criteria, for the Membach model works after a transformation pass, and they are fitted from different data.
 
@@
 \subsection{Limitations and future works}
-The main limitations of this projects that we can identify are: multiple seasonalities and Normality assumption.
+The main problems of this projects that we identified are multiple seasonalities and the Normality assumption.
 

### 2020-06-06 project.Rmd
add spencer filter

@@
 
+```{r trend Spencer}
+weights.s <- c(-3,-6,-5,3,21,46,67)
+weights.s <- c(weights.s, 74, rev(weights.s))
+weights.s <- weights.s/sum(weights.s)
+trend.s <- stats::filter(seasonality.aggregate(seis.h.train), side=2, filter=weights.s)
+
+plots <- show_overlap_and_diff(
+  dec.addi$trend, trend.s, xlab=seasonality.xlab,
+  base.desc = "additive decomposition trend",
+  compare.desc = "Spencer's 15-point MA",
+  overlap.ylab = "Trend (nm)",
+  diff.title = "Difference between trends",
+  diff.ylab = "nm", diff.ylab.sec="% over trend span",
+  diff.ylab.sec.accuracy = .01
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+
+```
+
 ```{r trend MA}

### 2020-06-06 project.mems.Rmd
add - in file pattern

@@
 # MEMS
-domain <- c(-7200,7200); pattern <- "MEM.*\\.csv"; name <- "MEM"; description <- "Membach"; proj_file <- 'proj.mems.RData'
+domain <- c(-7200,7200); pattern <- "MEMS-.*\\.csv"; name <- "MEM"; description <- "Membach"; proj_file <- 'proj.mems.RData'
 ```

### 2020-06-06 project.uccs.Rmd
delivered draft
add - in file pattern

@@
 # UCCS
-domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
+domain <- c(-350000,350000); pattern <- "UCCS-.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
 ```
@@
 
+```{r final model selection}
+chosen.model <- c(2,0,0,0,1,1)
+chosen.lambda <- NULL
+```
+
+```{r project.after, child='project.after.Rmd'}
+```
+
 ```{r save RData, include=FALSE}

### 2020-06-06 project.R

@@ -27,9 +27,9 @@ print_dec <- function(x, decimals=2) trimws(format(round(x, decimals), nsmall=de
 
 # We define some variables that identify the dataset. Only the row about the wanted dataset should be run
 # UCCS
-domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
+domain <- c(-350000,350000); pattern <- "UCCS-.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
 # MEMS
-domain <- c(-7200,7200); pattern <- "MEMS.*\\.csv"; name <- "MEMS"; description <- "Membach"; proj_file <- 'proj.mems.RData'
+domain <- c(-7200,7200); pattern <- "MEMS-.*\\.csv"; name <- "MEMS"; description <- "Membach"; proj_file <- 'proj.mems.RData'
 
 # If the corresponding RData file is present, we load it mainly to avoid refitting all models
 if (exists("proj_file") & file.exists(proj_file)) {

### 2020-06-06 settings.py
actually, data starts from 2018-10-27

@@ -4,5 +4,5 @@ LAST_DAY_FILE = 'last_day'
 START_DATE = {
   'year': 2018,
   'month': 10,
-  'day': 25,
+  'day': 27,
 }

### 2020-06-06 crawl.py
add compressor to crawler

@@ -8,6 +8,8 @@ from requests_session import SESSION
 TZ_BELGIAN = pytz.timezone('Europe/Brussels')
 TZ_UTC = pytz.utc
 TODAY = datetime.date.today()
+DELTA_SECOND = datetime.timedelta(seconds=1)
+DELTA_DAY = datetime.timedelta(days=1)
 
 
 def get_csv_url_for(station_code: str, day: datetime.date):
@@ -31,12 +33,11 @@ def iter_data_for(station_code: str, day: datetime.date):
     if len(day_data) > 0:
         utc_second = TZ_UTC.localize(datetime.datetime(year=day.year, month=day.month, day=day.day))
         # belgian_second = TZ_BELGIAN.normalize(utc_second)
-        second = datetime.timedelta(seconds=1)
         for row in day_data[1:]:
             data = [utc_second.isoformat()]  # , belgian_second.isoformat()]
             data.extend(row.split(','))
             yield data
-            utc_second += second
+            utc_second += DELTA_SECOND
             # belgian_second += second
 
 
@@ -52,15 +53,19 @@ def get_events_on(day: datetime.date):
         return {}
 
 
+def get_month_file(station: str, month: datetime.date):
+    filename = f"{station}-{month:%Y-%m}.csv"
+    return os.path.join(CSV_FOLDER, filename)
+
+
 def save_month(station_code: str, month: datetime.date, force: bool = False):
     day = month.replace(day=1)
-    filename = f"{station_code}-{day:%Y-%m}.csv"
-    path = os.path.join(CSV_FOLDER, filename)
+    path = get_month_file(station_code, day)
     if os.path.exists(path) and not force:
         while day.month == month.month and day < TODAY:
-            day += datetime.timedelta(days=1)
+            day += DELTA_DAY
         return day
-    print(f"Loading {day:%Y-%m}", end='... ')
+    print(f"Loading {day:%Y-%m}", end='... ', flush=True)
     try:
         with open(path, 'w') as f:
             # f.write(f"utc_second,belgian_second,min,max\n")
@@ -70,7 +75,7 @@ def save_month(station_code: str, month: datetime.date, force: bool = False):
                     f.write(','.join(row))
                     f.write('\n')
                 get_events_on(day)
-                day += datetime.timedelta(days=1)
+                day += DELTA_DAY
     except (InterruptedError, KeyboardInterrupt) as _:
         print('ERROR')
         os.remove(path)
[…]

### 2020-06-06 project.R
add filters and manual single decomposition

@@ -52,6 +52,7 @@ cat("Domain values:", domain[1], "to", domain[2], "nm\n")
 cat("Data from", format.POSIXct(min(seis$utc_second), "%Y-%m-%d %H:%M:%S"), "to", format.POSIXct(max(seis$utc_second), "%Y-%m-%d %H:%M:%S"), "\n")
 
 
+
 # As a first step, we remove all values out of domain, and all invalid values: (-1,1) and (0,0) pairs
 # This is done following these steps:
 # - locate the values through boolean arrays
@@ -354,3 +355,91 @@ dec.addi <- decompose(seasonality.aggregate(seis.h.ts))
 autoplot(dec.addi) + labs(title="Decomposition of additive time series", subtitle = paste0("Seasonality ", seasonality.period)) + xlab(seasonality.xlab)
 
 
+
+# We now apply three different filters to check which one behaves better
+# since we are going to compare the results with the decomposed trend, we will apply the filters on the missing values-free dataset
+
+# First we apply a simple linear MA filter
+p <- floor(seasonality.period/2)
+weights <- rep(1/(2*p+1), times=2*p+1)
+trend <- stats::filter(seasonality.aggregate(seis.h.train), sides=2, filter = weights)
+plots <- show_overlap_and_diff(
+  dec.addi$trend, trend, xlab=seasonality.xlab,
+  base.desc = "additive decomposition trend",
+  compare.desc = paste0("linear filtered trend with p=", p),
+  overlap.ylab = "Trend (nm)",
+  diff.title = "Difference between trends",
+  diff.ylab = "nm", diff.ylab.sec="% over trend span"
+)
+plot_grid(plots[[1]], plots[[2]], nrow = 2)
+
+# Then we apply Spencer's 15-points weights MA filter
+weights.s <- c(-3,-6,-5,3,21,46,67)
+weights.s <- c(weights.s, 74, rev(weights.s))
+weights.s <- weights.s/sum(weights.s)
+trend.s <- stats::filter(seasonality.aggregate(seis.h.train), side=2, filter=weights.s)
+plots <- show_overlap_and_diff(
+  dec.addi$trend, trend.s, xlab=seasonality.xlab,
+  base.desc = "additive decomposition trend",
+  compare.desc = "Spencer's 15-point MA",
+  overlap.ylab = "Trend (nm)",
+  diff.title = "Difference between trends",
[…]

### 2020-06-06 project.R
add multi seasonalities extraction

@@ -442,4 +442,86 @@ plot_grid(plots[[1]], plots[[2]], nrow = 2)
 
 
 
-# After that first extraction, we 
+# After that single extraction, we try extracting two seasonalities at once.
+# Since the procedure is fixed, we define a couple of functions to better handle it.
+# trend.s_compute only apply the seasonality-gnostic symmetric MA filter
+trend.s_compute <- function(ts.in, period) {
+  f <- period
+  weights.s <- c(0.5, rep(1,f-1), 0.5)/f
+  trend.s <- stats::filter(ts.in, side=2, filter=weights.s)
+  return(trend.s)
+}
+# season_by_mean computes the seasonality figure using the mean-over-period methodology
+# it also computes the standard deviation and produces the complete seasonality on all ts length
+season_by_mean <- function(ts.in, period) {
+  freq <- frequency(ts.in)
+  if (freq == period) {
+    end <- c(end(ts.in)[1], period)
+  } else {
+    end <- c(((end(ts.in)[1] * freq) %/% period + 1) * (period/freq), freq)
+  }
+  extended <- window(ts.in, end = end, extend=T)
+  matrix <- t(matrix(data = extended, nrow = period))
+  seasonality.figure <- as.ts(colMeans(matrix, na.rm = T))
+  seasonality.sd <- as.ts(apply(matrix, 2, sd, na.rm = T))
+  seasonality <- rep(seasonality.figure, length(extended)/length(seasonality.figure))[1:length(ts.in)]
+  return(list(figure=seasonality.figure, seasonal=seasonality, sd=seasonality.sd))
+}
+# trend_season applies the two functions to compute both trend and seasonality with one call
+trend_season <- function(ts.in, period) {
+  trend.s <- trend.s_compute(ts.in, period)
+  season <- season_by_mean(ts.in-trend.s, period)
+  season$trend <- trend.s
+  return(season)
+}
+
+
+
+# We now use those functions to extract the 24/168 hours seasonalities together
+aggregated <- seasonality.aggregate(seis.h.train)
+# first iteration
+season.24 <- trend_season(aggregated, 24)
+season.168 <- trend_season(aggregated-season.24$seasonal, 168)
+# second iteration
+season.24 <- trend_season(aggregated-season.168$seasonal, 24)
[…]

### 2020-06-06 project.Rmd
add composed weekly seasonality figure

@@
   autoplot(aggregated-season.24$seasonal-season.168$seasonal-season.168$trend) + labs(title="Residuals", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(t=-6, b=-6, unit="pt")),
+  autoplot(season.168$figure + rep(season.24$figure, 7)) + labs(title="Weekly seasonality (sum)", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(r=10, b=-6, unit="pt")),
   nrow = 3,

### 2020-06-06 draft.tex
more specific intro and reword/reorder some pieces

@@
 Such events are sometimes comparable with natural seismic events, depending on the distance from the seismograph.
+
+The correct estimation of a background noise, correlated or not with human activity, could help institutions to better isolate less prominent earthquakes and similar events from the seismic data.
+This could result in better understanding of the frequency of earthquakes and by studying their more-precise evolution over time.
 
@@
 
+Some R commands cannot handle timeseries with missing values, but as we have seen both stations datasets include missing values.
+Also, please note that we are going to have a timeseries object indexed by weeks, so that it contains the 168 hours between each integer index.
+Since we know that the longest missing values series is 90 hours long, we can apply a simple completion strategy for serving those commands a filled timeseries: we will substitute missing values with the mean of the week they are in; in R terms \texttt{na.aggregate(x, floor)}.
+
 \subsubsection{Decomposition}
@@
 
+\begin{figure}[h]
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/extract both 24 and 168-1.pdf}
+		\caption{Uccle station}
+		\label{fig:decomposed:uccs}
+	\end{subfigure}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{project.mems_files/figure-latex/extract both 24 and 168-1.pdf}
+		\caption{Membach station}
+		\label{fig:decomposed:mems}
+	\end{subfigure}
+	\begin{center}
+		\caption{Timeseries decomposition}{\small Only the second method results are shown for clarity}
+		\label{fig:decomposed}
+	\end{center}
+\end{figure}
+
 Since the periodgrams suggested the presence of two significative periodicities, we also tested the extraction of both corresponding seasonalities:
@@
 	\item Remove the trend from the already deseasoned dataset and extract the 168-hour seasonal figure using the local mean method
-	\item Repeat steps 3-5 steps alternating the 168/24 seasonalities until convergence (2 to 3 cycles should be enough)
+	\item Repeat steps 1-5 using the last deseasoned dataset until convergence (2-3 cycles)
 	\item Removing the last fitted trend and the two seasonalities we obtain the final residuals
@@
 
-In \figurename~\ref{fig:decomposed} we see the decomposition results.
+In \figurename~\ref{fig:decomposed} we see the double decomposition results.
 While data from Uccle station have a clear significativity of both trend and seasonalities, data from Membach could seem not significant.
 Still, if we notice that the main timeseries lies between values of 0 and 100, we can find out that around half of the value will be composed by the trend, a 30\% by the joined seasonalities component, and the remaining value by the residuals.
-We will need to better handle the spikes starting from the middle of our dataset.
-
-\begin{figure}[h]
-	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/extract both 24 and 168-1.pdf}
-		\caption{Uccle station}
-		\label{fig:decomposed:uccs}
-	\end{subfigure}
-	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=\linewidth]{project.mems_files/figure-latex/extract both 24 and 168-1.pdf}
-		\caption{Membach station}
-		\label{fig:decomposed:mems}
-	\end{subfigure}
-	\begin{center}
-		\caption{Timeseries decomposition}{\small Only the second method results are shown for clarity}
-		\label{fig:decomposed}
-	\end{center}
-\end{figure}
-
-Some R commands cannot handle timeseries with missing values, but as we have seen both stations datasets include missing values.
-Also, please note that we are going to have a timeseries object indexed by weeks, so that it contains the 168 hours between each integer.
-Since we know that the longest missing values series is 90 hours long, we can apply a simple completion strategy for serving those commands a filled timeseries: we will substitute missing values with the mean of the week they are in; in R terms \texttt{na.aggregate(x, floor)}.
+We will need to better handle the spikes that start from the middle of that dataset.
 
@@
 
-The transformed data still shows no constant mean, even if it has a way better shape than the original data.
+The transformed data still shows no constant mean, even if it has a way better shape than the original data, confirmed by the Q-Q plot (\figurename~\ref{fig:mems:box-cox:qq}).
 Thus, we consider the usage of the \textit{diff} operator, simple and at lag 168.
@@
 We extract the seasonalities for each station highlighting the standard deviation confidence band for an estimated error of 67\%.
-This estimation lies on the assumption that values for each single hour in the hourly/weekly seasonality is Normally distributed, a slightly different assumption that is not necessarily refuted by the already seen Q-Q plots.
+This interval estimation lies on the assumption that values for each single hour in the hourly/weekly seasonality is Normally distributed, a slightly different assumption that is not necessarily refuted by the already seen Q-Q plots.
 

### 2020-06-06 project.Rmd
better R code and add roots plot

@@
 plot_grid(
-  seis.h.train %>% diff(lag=168) %>%  acf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
-  seis.h.train %>% diff(lag=168) %>%  acf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
-  seis.h.train %>% diff(lag=168) %>% pacf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
-  seis.h.train %>% diff(lag=168) %>% pacf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
+  seis.h.train %>% diff(lag=168) %>%  acf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
+  seis.h.train %>% diff(lag=168) %>%  acf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
+  seis.h.train %>% diff(lag=168) %>% pacf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
+  seis.h.train %>% diff(lag=168) %>% pacf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
   nrow = 2
@@
 plot_grid(
-  seis.h.train %>% diff() %>% diff(lag=168) %>%  acf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
-  seis.h.train %>% diff() %>% diff(lag=168) %>%  acf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
-  seis.h.train %>% diff() %>% diff(lag=168) %>% pacf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
-  seis.h.train %>% diff() %>% diff(lag=168) %>% pacf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
+  seis.h.train %>% diff() %>% diff(lag=168) %>%  acf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
+  seis.h.train %>% diff() %>% diff(lag=168) %>%  acf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
+  seis.h.train %>% diff() %>% diff(lag=168) %>% pacf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
+  seis.h.train %>% diff() %>% diff(lag=168) %>% pacf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
   nrow = 2

### 2020-06-06 project.after.Rmd
better R code and add roots plot

@@
 plotting.forecast <- final.forecast
-cat("AIC =",plotting.model$aic, "\n")
-cat("AICc =",plotting.model$aicc, "\n")
+print(plotting.model)
+print(autoplot(plotting.model) + labs(subtitle = plotting.forecast$method))
 checkresiduals(plotting.model)

### 2020-06-06 project.mems.Rmd
better R code and add roots plot

--- 2020-06-01 project.mems.Rmd
@@
 # MEMS
-domain <- c(-7200,7200); pattern <- "MEM.*\\.csv"; name <- "MEM"; description <- "Membach"; proj_file <- 'proj.mems.RData'
+domain <- c(-7200,7200); pattern <- "MEMS-.+\\.csv"; name <- "MEM"; description <- "Membach"; proj_file <- 'proj.mems.RData'
 ```
@@
   plotting.forecast <- models[[forecast.name]]
-  cat("AIC =",plotting.model$aic, "\n")
-  cat("AICc =",plotting.model$aicc, "\n")
+  print(plotting.model)
+  print(autoplot(plotting.model) + labs(subtitle = plotting.forecast$method))
   checkresiduals(plotting.model)
@@
   print(plot_grid(
-    autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
-    ,
-    residuals.acf
-    ,
-    autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
-    ,
-    residuals.pacf
-    ,
+    autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method)),
+    residuals.acf,
+    autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method)),
+    residuals.pacf,
     nrow = 2
@@
   print(accuracy(plotting.forecast, seis.h.test))
-  # qqPlot(plotting.model$residuals, main = paste0("Q-Q plot for ", plotting.forecast$method), xlab = "residuals")
   print(gg_qqplot(plotting.model$residuals) + 

### 2020-06-06 project.uccs.Rmd
better R code and add roots plot

@@
 # UCCS
-domain <- c(-350000,350000); pattern <- "UCCS-.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
+domain <- c(-350000,350000); pattern <- "UCCS-.+\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
 ```
@@
 }
+
+#models[["fitauto"]] <- auto.arima(seis.h.train, D=1, approx=F, trace = T)
+#models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*18)  
+#models[["fit311010"]] <- models[["fitauto"]]
+#models[["fit412010"]] <- fit
+#models[["forecast412010"]] <- forecast(models[["fit412010"]], h=168*18)
+
+# 000012 better than 000011, Box-Pierce = 0
+# 010011 somehow better somehow worst than 000012, Box-Pierce = 0
+# 010012 , Box-Pierce = 
+# 311010 , Box-Pierce = 
+# 311011 , Box-Pierce =
+
 for (model in to_test) {
@@
   cat("\n")
-}
-#models[["fitauto"]] <- auto.arima(seis.h.train, D=1, approx=F, trace = T)
-#models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*18)  
-#models[["fit311010"]] <- models[["fitauto"]]
-#models[["fit412010"]] <- fit
-#models[["forecast412010"]] <- forecast(models[["fit412010"]], h=168*18)
-
-# 000012 better than 000011, Box-Pierce = 0
-# 010011 somehow better somehow worst than 000012, Box-Pierce = 0
-# 010012 , Box-Pierce = 
-# 311010 , Box-Pierce = 
-# 311011 , Box-Pierce =
-
-for (model in to_test) {
-  # model <- c(4,1,2,0,1,0)
-  fit.name <- paste0("fit", paste0(model, collapse = ""))
-  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
-  cat("SARIMA(", paste0(model[1:3], collapse = ","), ")(", paste0(model[4:6], collapse = ","), ")[", frequency(seis.h.train), "]\n", sep = "")
-    if ((fit.name %in% names(models))) {  if ((forecast.name %in% names(models))) {
   plotting.model <- models[[fit.name]]
   plotting.forecast <- models[[forecast.name]]
-    cat("AIC =",plotting.model$aic, "\n")
+  print(plotting.model)
+  print(autoplot(plotting.model) + labs(subtitle = plotting.forecast$method))
   checkresiduals(plotting.model)
-    residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
-    residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
+  residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*4)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
+  residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*4)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
   print(plot_grid(residuals.acf, residuals.pacf, nrow=2))
-    print(plot_forecast(plotting.forecast) + autolayer(seis.h.test, series="Original"))
+  print(plot_grid(
+    autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method)),
+    residuals.acf,
+    autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method)),
+    residuals.pacf,
+    nrow = 2
+  ))
+  print(plot_forecast(plotting.forecast) + autolayer(seis.h.test, series="Original") + ylab("Mean hourly movement (nm)"))
   plots <- show_overlap_and_diff(
@@
   print(accuracy(plotting.forecast, seis.h.test))
-    # qqPlot(plotting.model$residuals, main = paste0("Q-Q plot for ", plotting.forecast$method), xlab = "residuals")
   print(gg_qqplot(plotting.model$residuals) + 
@@
   )
-  }}
 }

### 2020-06-06 project.R
add Membac/Uccle specific R code

@@ -524,4 +524,334 @@ plot_grid(plots[[1]], plots[[2]], nrow = 2)
 
 
 
-# 
+# We now proceed to find the needed transformations for our dataset to choose the order of the model and estimate the parameters
+# Since this must be done specifically for each of the two datasets, we start with what we need for the Uccle station, then we proceed with the Membach station.
+# Please note that the variable names are the same, so the Membach-specific code will overwrite and clash with the Uccle-specific code.
+# If at the start the Uccle data (UCCS) was chosen, then follow the code delimited for that station and skip the Membach-specific.
+# Follow and skip the other code when Membach data (MEMS) was chosen at the start.
+
+
+
+# To plot the forecasts, we will use this shorthand to see in the plot the Box-Pierce test result on the residuals
+plot_forecast <- function(forec) {
+  BP <- Box.test(forec$model$residuals, type="Box-Pierce")
+  autoplot(forec) +
+    labs(title=forec$method, subtitle = paste0("Box-Pierce p-value=", print_dec(BP$p.value, decimals=4))) +
+    autolayer(forec$mean, series="Forecast") -> plotted
+  return(plotted)
+}
+# To plot the Q-Q plot with ggplot we will use this shorthand
+gg_qqplot <- function(dataset) {
+  if (is.ts(dataset))
+    dataset <- as.numeric(dataset)
+  ggplot(data.frame(values=dataset), aes(sample=values)) +
+    stat_qq_band() + stat_qq_line(col="red") + stat_qq_point(shape = 1, size = 3) +
+    xlab("Theoretical quantiles (95% conf.)")
+}
+
+
+
+#################### Uccle station START ####################
+
+
+
+# We check possible transformations, without obtaining significant better distributions
+month.starts <- seq.POSIXt(from=floor_date(seis.train.start, 'month'), to=ceiling_date(seis.train.end, 'month'), by = "1 month")
+hours.from.train.start <- as.numeric(month.starts-seis.train.start)*24+168
+# base dataset
+autoplot(seis.h.train, type="l") +
+  scale_x_continuous(breaks=hours.from.train.start/168, labels=format(month.starts, "%m-%y")) +
[…]

### 2020-06-06 project.mems.Rmd
cleaner Membach Rmd code

@@
 plot_grid(
-  train.bc %>% diff(lag=168) %>%  acf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
-  train.bc %>% diff(lag=168) %>%  acf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
-  train.bc %>% diff(lag=168) %>% pacf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
-  train.bc %>% diff(lag=168) %>% pacf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
+  train.bc %>% diff(lag=168) %>%  acf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
+  train.bc %>% diff(lag=168) %>%  acf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
+  train.bc %>% diff(lag=168) %>% pacf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
+  train.bc %>% diff(lag=168) %>% pacf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
   nrow = 2
@@
 plot_grid(
-  train.bc %>% diff() %>% diff(lag=168) %>%  acf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
-  train.bc %>% diff() %>% diff(lag=168) %>%  acf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
-  train.bc %>% diff() %>% diff(lag=168) %>% pacf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
-  train.bc %>% diff() %>% diff(lag=168) %>% pacf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous()
-  ,
+  train.bc %>% diff() %>% diff(lag=168) %>%  acf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
+  train.bc %>% diff() %>% diff(lag=168) %>%  acf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
+  train.bc %>% diff() %>% diff(lag=168) %>% pacf(lag.max = 24,    plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
+  train.bc %>% diff() %>% diff(lag=168) %>% pacf(lag.max = 168*4, plot = F, na.action = na.pass) %>% autoplot() + labs(title = NULL) + scale_x_continuous(),
   nrow = 2

### 2020-06-06 draft.tex
fix plot numbers

@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/arima-26} %30
+		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/arima-35}
 		\caption{ACF and PACF plots}
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/arima-24} %28
+		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/arima-32}
 		\caption{Q-Q plot}
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-27} %32
+		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-37}
 		\caption{$\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$}
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-51} %60
+		\includegraphics[width=1\linewidth]{project.uccs_files/figure-latex/arima-69}
 		\caption{$\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$}
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=1\linewidth]{"project.mems_files/figure-latex/arima-18"} %15
+		\includegraphics[width=1\linewidth]{"project.mems_files/figure-latex/arima-21"}
 		\caption{$\mathrm{ARIMA}(0,1,1)(0,1,1)_{168}$}
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=1\linewidth]{"project.mems_files/figure-latex/arima-11"} %9
+		\includegraphics[width=1\linewidth]{"project.mems_files/figure-latex/arima-13"}
 		\caption{$\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$}

### 2020-06-07 project.R
reduce duplicated code

@@ -525,8 +525,8 @@ plot_grid(plots[[1]], plots[[2]], nrow = 2)
 
 
 # We now proceed to find the needed transformations for our dataset to choose the order of the model and estimate the parameters
-# Since this must be done specifically for each of the two datasets, we start with what we need for the Uccle station, then we proceed with the Membach station.
-# Please note that the variable names are the same, so the Membach-specific code will overwrite and clash with the Uccle-specific code.
+# When specific choices and checks are done for any station, we will highlight it.
+# Please note that the variable names are normally the same, so the Membach-specific code will overwrite and clash with the Uccle-specific code.
 # If at the start the Uccle data (UCCS) was chosen, then follow the code delimited for that station and skip the Membach-specific.
 # Follow and skip the other code when Membach data (MEMS) was chosen at the start.
 
@@ -551,11 +551,7 @@ gg_qqplot <- function(dataset) {
 
 
 
-#################### Uccle station START ####################
-
-
-
-# We check possible transformations, without obtaining significant better distributions
+# We check possible transformations
 month.starts <- seq.POSIXt(from=floor_date(seis.train.start, 'month'), to=ceiling_date(seis.train.end, 'month'), by = "1 month")
 hours.from.train.start <- as.numeric(month.starts-seis.train.start)*24+168
 # base dataset
@@ -582,6 +578,15 @@ gghistogram(train.bc, add.normal = T) + xlab("mean hourly movement (nm)") + labs
 gg_qqplot(train.bc) + 
   labs(title="Q-Q plot for training dataset") + ylab("Sample quantiles") + labs(caption = paste0("Data Box-Cox transformed with lambda = ", train.bc.lambda))
 
+
+
+
+
+
+#################### Uccle station START ####################
+# For Uccle dataset, no transformation have significant benefit. So we proceed with non transformed data
+# We set the chosen Box-Cox transformation lambda to NULL to have more common code between stations
+chosen.lambda <- NULL
 
 
 # We check the behaviour of ACF and PACF by applying different diff operators
@@ -606,7 +611,6 @@ plot_grid(
[…]

### 2020-06-07 project.after.Rmd
minor edit

--- 2020-06-01 project.after.Rmd
@@
 print(plot_grid(
-  autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
-  ,
-  residuals.acf
-  ,
-  autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
-  ,
-  residuals.pacf
-  ,
+  autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method)),
+  residuals.acf,
+  autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method)),
+  residuals.pacf,
   nrow = 2

### 2020-06-07 project.R
use project.data.RData, comments

@@ -1,12 +1,12 @@
-
+# Inlcuding all used libraries
 library(stringi) # %s+%
-library(lubridate) # floor_date
+library(lubridate) # date shorthands (floor_date, ...)
 library(xts) # zoo
 library(fpp) # auto.arima
 library(ggplot2) # plots
 library(ggrepel) # for nicer labels on ggplot
 library(cowplot) # for plot_grid
-library(qqplotr) # qqplot with ggplot
+library(qqplotr) # Q-Q plot with ggplot
 library(spec) # for periodgram
 library(tidyverse) # map_df, read_csv, group_by, summarize...
 library(tsbox) # https://cran.r-project.org/web/packages/tsbox/vignettes/tsbox.html
@@ -17,31 +17,28 @@ print_dec <- function(x, decimals=2) trimws(format(round(x, decimals), nsmall=de
 
 
 # If needed, set working directory. The folder structure inside it should contain
-# - csv/
-#   - UCCS*.csv files
-#   - MEMS*.csv files
-# - proj.*.RData (optional)
+# - proj.data.RData
+# - proj.mems.RData (optional)
+# - proj.uccs.RData (optional)
 # setwd("/media/data/Documents/Uni/spatiotemp/villalobos/project/")
+# Load the datasets. proj.data.RData contains seis.uccs and seis.mems data.frame objects
+load('proj.data.RData')
 
 
 
 # We define some variables that identify the dataset. Only the row about the wanted dataset should be run
 # UCCS
-domain <- c(-350000,350000); pattern <- "UCCS-.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
+seis <- seis.uccs; domain <- c(-350000,350000); name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
 # MEMS
-domain <- c(-7200,7200); pattern <- "MEMS-.*\\.csv"; name <- "MEMS"; description <- "Membach"; proj_file <- 'proj.mems.RData'
-
+seis <- seis.mems; domain <- c(-7200,7200); name <- "MEMS"; description <- "Membach"; proj_file <- 'proj.mems.RData'
 # If the corresponding RData file is present, we load it mainly to avoid refitting all models
 if (exists("proj_file") & file.exists(proj_file)) {
   load(proj_file)
 }
+# If high RAM usage is noticed, it is possible to remove the unused station data
[…]

### 2020-06-07 project.R
add small final testing

@@ -765,3 +765,19 @@ chosen.model <- c(2,0,0,0,1,1)
 # they seem equal, but the chosen.lambda is actually different
 
 
+
+#################### Final testing ####################
+# Now we will bring the chosen model to cover all training+test data (ending on 2020-03-01 23:59:59 UTC), and test it for the next 8 weeks (until 2020-04-26 23:59:59 UTC)
+
+
+seis.test.final.start <- seis.test.end + seconds(1)
+seis.test.final.end <- seis.test.final.start + days(7*8) - seconds(1)
+c(seis.train.start, seis.test.final.start, seis.test.final.end)
+
+final.training <- build_ts(seis.h, periods = c(seasonality.period), start = seis.train.start, end = seis.test.end, force_ms = T)
+final.testing <- build_ts(seis.h, periods = c(seasonality.period), start = seis.train.start, end = seis.test.final.end, force_ms = T)
+final.testing <- window(final.testing, start = c(end(final.training)[1] + 1, 1))
+if (!is.null(chosen.lambda)) {
+  cat(chosen.lambda, BoxCox.lambda(final.training), chosen.lambda == BoxCox.lambda(final.training))
+}
+

### 2020-06-08 project.after.Rmd
complete porting and commenting to project.R

@@
 final.testing <- window(final.testing, start = c(end(final.training)[1] + 1, 1))
-c(chosen.lambda, BoxCox.lambda(final.training))
+if (!is.null(chosen.lambda)) {
+  cat(chosen.lambda, BoxCox.lambda(final.training), chosen.lambda == BoxCox.lambda(final.training), "\n")
+}
 
@@
 final.testing.24 <- build_ts(final.testing, periods = c(24))
-seasonplot(final.testing.24/sd(final.training.24, na.rm = T), type="l", col=rainbow(7))
+seasonplot(final.testing.24/sd(final.testing.24, na.rm = T), type="l", col=rainbow(7))
 legend("topleft", legend=weekdays(seis.test.final.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)

### 2020-06-08 project.R

@@ -766,18 +766,121 @@ chosen.model <- c(2,0,0,0,1,1)
 
 
 
-#################### Final testing ####################
+#################### Final assessing ####################
 # Now we will bring the chosen model to cover all training+test data (ending on 2020-03-01 23:59:59 UTC), and test it for the next 8 weeks (until 2020-04-26 23:59:59 UTC)
+# If a similar procedure as the above is followed, we will find that the same model (and the same Box-Cox lambda for Membach) are the best choices
+# This is probably due to both datasets being quite stable over time
 
 
+
+# We make our final train/test split date on 2nd of March
 seis.test.final.start <- seis.test.end + seconds(1)
 seis.test.final.end <- seis.test.final.start + days(7*8) - seconds(1)
-c(seis.train.start, seis.test.final.start, seis.test.final.end)
-
 final.training <- build_ts(seis.h, periods = c(seasonality.period), start = seis.train.start, end = seis.test.end, force_ms = T)
 final.testing <- build_ts(seis.h, periods = c(seasonality.period), start = seis.train.start, end = seis.test.final.end, force_ms = T)
 final.testing <- window(final.testing, start = c(end(final.training)[1] + 1, 1))
 if (!is.null(chosen.lambda)) {
-  cat(chosen.lambda, BoxCox.lambda(final.training), chosen.lambda == BoxCox.lambda(final.training))
+  cat(chosen.lambda, BoxCox.lambda(final.training), chosen.lambda == BoxCox.lambda(final.training), "\n")
 }
+# Then, we fit the chosen model on the new training set
+fit.name <- paste0("fit", paste0(chosen.model, collapse = ""))
+forecast.name <- paste0("forecast", paste0(chosen.model, collapse = ""))
+cat("SARIMA(", paste0(chosen.model[1:3], collapse = ","), ")(", paste0(chosen.model[4:6], collapse = ","), ")[", frequency(final.training), "]", sep = "")
+if (!exists("final.model")) {
+  cat(" fitting...")
+  final.model <- Arima(final.training, chosen.model[1:3], chosen.model[4:6], lambda=chosen.lambda)
+  cat("DONE")
+}
+if (!exists("final.forecast")) {
+  cat(" forecasting...")
[…]

### 2020-06-08 project.R
fix project.R so that it is actually executable

@@ -36,8 +36,8 @@ if (exists("proj_file") & file.exists(proj_file)) {
   load(proj_file)
 }
 # If high RAM usage is noticed, it is possible to remove the unused station data
-# remove(seis.uccs)
-# remove(seis.mems)
+remove(seis.uccs)
+remove(seis.mems)
 
 
 
@@ -354,6 +354,12 @@ autoplot(dec.addi) + labs(title="Decomposition of additive time series", subtitl
 
 
 #################### Decomposition ####################
+
+
+
+# We first use the stats::decompose method with default parameters
+dec.addi <- decompose(seasonality.aggregate(seis.h.train))
+autoplot(dec.addi) + labs(title="Decomposition of additive time series", subtitle = paste0("Seasonality ", seasonality.period)) + xlab(seasonality.xlab)
 
 
 
@@ -801,14 +807,14 @@ cat("\n")
 plotting.model <- final.model
 plotting.forecast <- final.forecast
 # plot forecasts
-print(plot_forecast(plotting.forecast) + autolayer(seis.h.test, series="Original") + ylab("Mean hourly movement (nm)"))
+print(plot_forecast(plotting.forecast) + autolayer(final.testing, series="Original") + ylab("Mean hourly movement (nm)"))
 # print stats about data falling inside 80% and 90% CIs
 cat("Test set inside the CI\n")
-cat("80% CI \t ", print_dec(mean(plotting.forecast$lower[,1] < seis.h.test & seis.h.test < plotting.forecast$upper[,1], na.rm = T)*100), "%\n")
-cat("95% CI \t ", print_dec(mean(plotting.forecast$lower[,2] < seis.h.test & seis.h.test < plotting.forecast$upper[,2], na.rm = T)*100), "%\n")
+cat("80% CI \t ", print_dec(mean(plotting.forecast$lower[,1] < final.testing & final.testing < plotting.forecast$upper[,1], na.rm = T)*100), "%\n")
+cat("95% CI \t ", print_dec(mean(plotting.forecast$lower[,2] < final.testing & final.testing < plotting.forecast$upper[,2], na.rm = T)*100), "%\n")
 # plot overlap/difference with test data
 plots <- show_overlap_and_diff(
-  seis.h.test, plotting.forecast$mean, xlab=seasonality.xlab,
[…]

### 2020-06-08 project.after.Rmd
renaming variables

@@
 ```{r seasonalities/before}
-aggregated.final <- seasonality.aggregate(final.training)
-season.final.24 <- trend_season(aggregated.final, 24)
-season.final.168 <- trend_season(aggregated.final-season.final.24$seasonal, 168)
-
-season.final.24 <- trend_season(aggregated.final-season.final.168$seasonal, 24)
-season.final.168 <- trend_season(aggregated.final-season.final.24$seasonal, 168)
-
+aggregated.before.lockdown <- seasonality.aggregate(before.lockdown)
+season.before.lockdown.24 <- trend_season(aggregated.before.lockdown, 24)
+season.before.lockdown.168 <- trend_season(aggregated.before.lockdown-season.before.lockdown.24$seasonal, 168)
+season.before.lockdown.24 <- trend_season(aggregated.before.lockdown-season.before.lockdown.168$seasonal, 24)
+season.before.lockdown.168 <- trend_season(aggregated.before.lockdown-season.before.lockdown.24$seasonal, 168)
 # this third iteration isn't actually needed
-season.final.24 <- trend_season(aggregated.final-season.final.168$seasonal, 24)
-season.final.168 <- trend_season(aggregated.final-season.final.24$seasonal, 168)
-
-residuals.final <- aggregated.final-season.final.24$seasonal-season.final.168$seasonal-season.final.168$trend
-
-seasonbase.final.24 <- trend_season(aggregated.final, 24)
-seasonbase.final.168 <- trend_season(aggregated.final, 168)
+season.before.lockdown.24 <- trend_season(aggregated.before.lockdown-season.before.lockdown.168$seasonal, 24)
+season.before.lockdown.168 <- trend_season(aggregated.before.lockdown-season.before.lockdown.24$seasonal, 168)
+seasonbase.before.lockdown.24 <- trend_season(aggregated.before.lockdown, 24)
+seasonbase.before.lockdown.168 <- trend_season(aggregated.before.lockdown, 168)
 
 plot_grid(
-  autoplot(ts(seasonbase.final.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
-    geom_line(aes(y=season.final.24$figure, col="Multi daily")) +
+  autoplot(ts(seasonbase.before.lockdown.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
+    geom_line(aes(y=season.before.lockdown.24$figure, col="Multi daily")) +
     scale_colour_manual("", values=c("purple", "yellow"), breaks=c("Direct daily", "Multi daily")),
-  autoplot(ts(seasonbase.final.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
-    geom_line(aes(y=season.final.168$figure + rep(season.final.24$figure, 7), col="Multi w+d")) +
-    #geom_line(aes(y=rep(seasonbase.final.24$figure, 7), col="Direct daily"), size=1.5) +
-    #geom_line(aes(y=rep(season.final.24$figure, 7), col="Multi daily")) +
-    geom_line(aes(y=season.final.168$figure, col="Multi weekly")) +
+  autoplot(ts(seasonbase.before.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
+    geom_line(aes(y=season.before.lockdown.168$figure + rep(season.before.lockdown.24$figure, 7), col="Multi w+d")) +
+    #geom_line(aes(y=rep(seasonbase.before.lockdown.24$figure, 7), col="Direct daily"), size=1.5) +
+    #geom_line(aes(y=rep(season.before.lockdown.24$figure, 7), col="Multi daily")) +
+    geom_line(aes(y=season.before.lockdown.168$figure, col="Multi weekly")) +
     scale_colour_manual("", values=c("red", "purple", "blue", "yellow", "green"), breaks=c("Direct weekly", "Direct daily", "Multi" %s+% c(" weekly", " daily", " w+d"))),
@@
 ```{r seasonalities/lockdown}
-aggregated.lockdown <- seasonality.aggregate(final.testing)
-season.lockdown.24 <- trend_season(aggregated.lockdown, 24)
-season.lockdown.168 <- trend_season(aggregated.lockdown-season.lockdown.24$seasonal, 168)
-
-season.lockdown.24 <- trend_season(aggregated.lockdown-season.lockdown.168$seasonal, 24)
-season.lockdown.168 <- trend_season(aggregated.lockdown-season.lockdown.24$seasonal, 168)
-
-# this third iteration isn't actually needed
-season.lockdown.24 <- trend_season(aggregated.lockdown-season.lockdown.168$seasonal, 24)
-season.lockdown.168 <- trend_season(aggregated.lockdown-season.lockdown.24$seasonal, 168)
-
-residuals.lockdown <- aggregated.lockdown-season.lockdown.24$seasonal-season.lockdown.168$seasonal-season.lockdown.168$trend
-
-seasonbase.lockdown.24 <- trend_season(aggregated.lockdown, 24)
-seasonbase.lockdown.168 <- trend_season(aggregated.lockdown, 168)
+aggregated.after.lockdown <- seasonality.aggregate(after.lockdown)
+season.after.lockdown.24 <- trend_season(aggregated.after.lockdown, 24)
+season.after.lockdown.168 <- trend_season(aggregated.after.lockdown-season.after.lockdown.24$seasonal, 168)
+season.after.lockdown.24 <- trend_season(aggregated.after.lockdown-season.after.lockdown.168$seasonal, 24)
+season.after.lockdown.168 <- trend_season(aggregated.after.lockdown-season.after.lockdown.24$seasonal, 168)
+season.after.lockdown.24 <- trend_season(aggregated.after.lockdown-season.after.lockdown.168$seasonal, 24)
+season.after.lockdown.168 <- trend_season(aggregated.after.lockdown-season.after.lockdown.24$seasonal, 168)
+seasonbase.after.lockdown.24 <- trend_season(aggregated.after.lockdown, 24)
+seasonbase.after.lockdown.168 <- trend_season(aggregated.after.lockdown, 168)
 
 plot_grid(
-  autoplot(ts(seasonbase.lockdown.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
-    geom_line(aes(y=season.lockdown.24$figure, col="Multi daily")) +
+  autoplot(ts(seasonbase.after.lockdown.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
+    geom_line(aes(y=season.after.lockdown.24$figure, col="Multi daily")) +
     scale_colour_manual("", values=c("purple", "yellow"), breaks=c("Direct daily", "Multi daily")),
-  autoplot(ts(seasonbase.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
-    geom_line(aes(y=season.lockdown.168$figure + rep(season.lockdown.24$figure, 7), col="Multi w+d")) +
-    #geom_line(aes(y=rep(seasonbase.lockdown.24$figure, 7), col="Direct daily"), size=1.5) +
-    #geom_line(aes(y=rep(season.lockdown.24$figure, 7), col="Multi daily")) +
-    geom_line(aes(y=season.lockdown.168$figure, col="Multi weekly")) +
+  autoplot(ts(seasonbase.after.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
+    geom_line(aes(y=season.after.lockdown.168$figure + rep(season.after.lockdown.24$figure, 7), col="Multi w+d")) +
+    #geom_line(aes(y=rep(seasonbase.after.lockdown.24$figure, 7), col="Direct daily"), size=1.5) +
+    #geom_line(aes(y=rep(season.after.lockdown.24$figure, 7), col="Multi daily")) +
+    geom_line(aes(y=season.after.lockdown.168$figure, col="Multi weekly")) +
     scale_colour_manual("", values=c("red", "purple", "blue", "yellow", "green"), breaks=c("Direct weekly", "Direct daily", "Multi" %s+% c(" weekly", " daily", " w+d"))),
@@
 plot_grid(
-  autoplot(ts(season.lockdown.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
-    geom_ribbon(aes(ymin = season.lockdown.24$figure - season.lockdown.24$sd, ymax = season.lockdown.24$figure + season.lockdown.24$sd), fill = "mediumpurple1", alpha=0.5) +
+  autoplot(ts(season.after.lockdown.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
+    geom_ribbon(aes(ymin = season.after.lockdown.24$figure - season.after.lockdown.24$sd, ymax = season.after.lockdown.24$figure + season.after.lockdown.24$sd), fill = "mediumpurple1", alpha=0.5) +
     geom_line(aes(col="Daily lockdown")) +
-    geom_ribbon(aes(ymin = season.final.24$figure - season.final.24$sd, ymax = season.final.24$figure + season.final.24$sd), fill = "lightcoral", alpha=0.5) +
-    geom_line(aes(y=season.final.24$figure, col="Daily before")) +
+    geom_ribbon(aes(ymin = season.before.lockdown.24$figure - season.before.lockdown.24$sd, ymax = season.before.lockdown.24$figure + season.before.lockdown.24$sd), fill = "lightcoral", alpha=0.5) +
+    geom_line(aes(y=season.before.lockdown.24$figure, col="Daily before")) +
     scale_colour_manual("", values=c("purple", "red"), breaks=c("Daily lockdown", "Daily before")),
-  autoplot(ts(season.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
-    geom_ribbon(aes(ymin = season.lockdown.168$figure - season.lockdown.168$sd, ymax = season.lockdown.168$figure + season.lockdown.168$sd), fill = "mediumpurple1", alpha=0.5) +
+  autoplot(ts(season.after.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
+    geom_ribbon(aes(ymin = season.after.lockdown.168$figure - season.after.lockdown.168$sd, ymax = season.after.lockdown.168$figure + season.after.lockdown.168$sd), fill = "mediumpurple1", alpha=0.5) +
     geom_line(aes(col="Weekly lockdown")) +
-    geom_ribbon(aes(ymin = season.final.168$figure - season.final.168$sd, ymax = season.final.168$figure + season.final.168$sd), fill = "lightcoral", alpha=0.5) +
-    geom_line(aes(y=season.final.168$figure, col="Weekly before")) +
+    geom_ribbon(aes(ymin = season.before.lockdown.168$figure - season.before.lockdown.168$sd, ymax = season.before.lockdown.168$figure + season.before.lockdown.168$sd), fill = "lightcoral", alpha=0.5) +
+    geom_line(aes(y=season.before.lockdown.168$figure, col="Weekly before")) +
     scale_colour_manual("", values=c("purple", "red"), breaks=c("Weekly lockdown", "Weekly before")),
@@
 
-autoplot(ts(seasonbase.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
-  geom_ribbon(aes(ymin = seasonbase.lockdown.168$figure - seasonbase.lockdown.168$sd, ymax = seasonbase.lockdown.168$figure + seasonbase.lockdown.168$sd), fill = "mediumpurple1", alpha=0.5) +
+autoplot(ts(seasonbase.after.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
+  geom_ribbon(aes(ymin = seasonbase.after.lockdown.168$figure - seasonbase.after.lockdown.168$sd, ymax = seasonbase.after.lockdown.168$figure + seasonbase.after.lockdown.168$sd), fill = "mediumpurple1", alpha=0.5) +
   geom_line(aes(col="Weekly lockdown")) +
-  geom_ribbon(aes(ymin = seasonbase.final.168$figure - seasonbase.final.168$sd, ymax = seasonbase.final.168$figure + seasonbase.final.168$sd), fill = "lightcoral", alpha=0.5) +
-  geom_line(aes(y=seasonbase.final.168$figure, col="Weekly before")) +
+  geom_ribbon(aes(ymin = seasonbase.before.lockdown.168$figure - seasonbase.before.lockdown.168$sd, ymax = seasonbase.before.lockdown.168$figure + seasonbase.before.lockdown.168$sd), fill = "lightcoral", alpha=0.5) +
+  geom_line(aes(y=seasonbase.before.lockdown.168$figure, col="Weekly before")) +
   scale_colour_manual("", values=c("purple", "red"), breaks=c("Weekly lockdown", "Weekly before"))

### 2020-06-08 project.after.Rmd
renaming variables

@@
 ```{r seasonalities/seasonplot}
-final.training.24 <- build_ts(final.training, periods = c(24))
-seasonplot(final.training.24/sd(final.training.24, na.rm = T), type="l", col=rainbow(7))
-legend("topleft", legend=weekdays(seis.train.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
-
-final.testing.24 <- build_ts(final.testing, periods = c(24))
-seasonplot(final.testing.24/sd(final.testing.24, na.rm = T), type="l", col=rainbow(7))
-legend("topleft", legend=weekdays(seis.test.final.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
+before.lockdown.start <- seis.train.start
+before.lockdown.end <- as_datetime('2020-03-14T00:00:00') - seconds(1)
+after.lockdown.start <- as_datetime('2020-03-14T00:00:00')
+after.lockdown.end <- as_datetime('2020-04-26T00:00:00') - seconds(1)
+before.lockdown <- build_ts(seis.h, periods = c(24), start = before.lockdown.start, end = before.lockdown.end, force_ms = T)
+after.lockdown <- build_ts(seis.h, periods = c(24), start = after.lockdown.start, end = after.lockdown.end, force_ms = T)
+# We first check the seasonplots, highlighting different weekdays
+seasonplot(before.lockdown, type="l", col=rainbow(7))
+legend("topleft", legend=weekdays(before.lockdown.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
+seasonplot(after.lockdown, type="l", col=rainbow(7)[c(6:7,1:5)])
+legend("topleft", legend=weekdays(after.lockdown.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7)[c(6:7,1:5)], lty=1, cex=0.8)
 

### 2020-06-08 project.after.Rmd
add mean/sd seasonplots

--- 2020-06-08 project.after.Rmd
@@
 ```{r seasonalities/seasonplot}
-final.training.24 <- build_ts(final.training, periods = c(24))
-seasonplot(final.training.24/sd(final.training.24, na.rm = T), type="l", col=rainbow(7))
-legend("topleft", legend=weekdays(seis.train.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
-
-final.testing.24 <- build_ts(final.testing, periods = c(24))
-seasonplot(final.testing.24/sd(final.testing.24, na.rm = T), type="l", col=rainbow(7))
-legend("topleft", legend=weekdays(seis.test.final.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
+before.lockdown.start <- seis.train.start
+before.lockdown.end <- as_datetime('2020-03-14T00:00:00') - seconds(1)
+after.lockdown.start <- as_datetime('2020-03-14T00:00:00')
+after.lockdown.end <- as_datetime('2020-04-26T00:00:00') - seconds(1)
+before.lockdown <- build_ts(seis.h, periods = c(24), start = before.lockdown.start, end = before.lockdown.end, force_ms = T)
+after.lockdown <- build_ts(seis.h, periods = c(24), start = after.lockdown.start, end = after.lockdown.end, force_ms = T)
+# We first check the seasonplots, highlighting different weekdays. We divide by sd to account for spikes
+seasonplot(before.lockdown/sd(before.lockdown, na.rm=T), type="l", col=rainbow(7), main="Seasonal plot: before lockdown")
+legend("topleft", legend=weekdays(before.lockdown.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
+seasonplot(after.lockdown/sd(after.lockdown, na.rm=T), type="l", col=rainbow(7)[c(6:7,1:5)], main="Seasonal plot: after lockdown")
+legend("topleft", legend=weekdays(after.lockdown.start + (as.difftime(c(0:6), units="days")))[c(3:7,1:2)], col=rainbow(7), lty=1, cex=0.8)
 
@@
     scale_colour_manual("", values=c("purple", "yellow"), breaks=c("Direct daily", "Multi daily")),
-  autoplot(ts(seasonbase.after.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
-    geom_line(aes(y=season.after.lockdown.168$figure + rep(season.after.lockdown.24$figure, 7), col="Multi w+d")) +
+  autoplot(ts(seasonbase.after.lockdown.168$figure[c(49:168,1:48)], frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
+    geom_line(aes(y=season.after.lockdown.168$figure[c(49:168,1:48)] + rep(season.after.lockdown.24$figure, 7), col="Multi w+d")) +
     #geom_line(aes(y=rep(seasonbase.after.lockdown.24$figure, 7), col="Direct daily"), size=1.5) +
     #geom_line(aes(y=rep(season.after.lockdown.24$figure, 7), col="Multi daily")) +
-    geom_line(aes(y=season.after.lockdown.168$figure, col="Multi weekly")) +
+    geom_line(aes(y=season.after.lockdown.168$figure[c(49:168,1:48)], col="Multi weekly")) +
     scale_colour_manual("", values=c("red", "purple", "blue", "yellow", "green"), breaks=c("Direct weekly", "Direct daily", "Multi" %s+% c(" weekly", " daily", " w+d"))),
@@
     scale_colour_manual("", values=c("purple", "red"), breaks=c("Daily lockdown", "Daily before")),
-  autoplot(ts(season.after.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
-    geom_ribbon(aes(ymin = season.after.lockdown.168$figure - season.after.lockdown.168$sd, ymax = season.after.lockdown.168$figure + season.after.lockdown.168$sd), fill = "mediumpurple1", alpha=0.5) +
+  autoplot(ts(season.after.lockdown.168$figure[c(49:168,1:48)], frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
+    geom_ribbon(aes(ymin = (season.after.lockdown.168$figure - season.after.lockdown.168$sd)[c(49:168,1:48)], ymax = (season.after.lockdown.168$figure + season.after.lockdown.168$sd)[c(49:168,1:48)]), fill = "mediumpurple1", alpha=0.5) +
     geom_line(aes(col="Weekly lockdown")) +
@@
 
-autoplot(ts(seasonbase.after.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
-  geom_ribbon(aes(ymin = seasonbase.after.lockdown.168$figure - seasonbase.after.lockdown.168$sd, ymax = seasonbase.after.lockdown.168$figure + seasonbase.after.lockdown.168$sd), fill = "mediumpurple1", alpha=0.5) +
+autoplot(ts(seasonbase.after.lockdown.168$figure[c(49:168,1:48)], frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
+  geom_ribbon(aes(ymin = (seasonbase.after.lockdown.168$figure - seasonbase.after.lockdown.168$sd)[c(49:168,1:48)], ymax = (seasonbase.after.lockdown.168$figure + seasonbase.after.lockdown.168$sd)[c(49:168,1:48)]), fill = "mediumpurple1", alpha=0.5) +
   geom_line(aes(col="Weekly lockdown")) +
@@
 ```
+```{r seasonalities/seasonplot-mean-sd}
+weekdays <- c("Monday", "Tuesday", "Wednseday", "Thursday", "Friday", "Saturday", "Sunday")
+
+seasonbase.before.lockdown.168.df <- data.frame(hour=rep(1:24, 7), mean=seasonbase.before.lockdown.168$figure[1:168], sd=seasonbase.before.lockdown.168$sd[1:168], Weekday=rep(weekdays, each=24))
+ggplot(data=seasonbase.before.lockdown.168.df) +
+  geom_line(aes(hour, mean - sd, col=Weekday), alpha=.1) +
+  geom_line(aes(hour, mean + sd, col=Weekday), alpha=.1) +
+  geom_ribbon(aes(hour, ymin = mean - sd, ymax = mean + sd, fill = Weekday), alpha=0.1, show.legend=F) +
+  geom_line(aes(hour, mean, col=Weekday)) +
+  ylab("Mean hourly movement (nm)") +
+  scale_color_manual(values = rainbow(7), breaks=weekdays, aesthetics = c("colour", "fill"))
+
+seasonbase.after.lockdown.168.df <- data.frame(hour=rep(1:24, 7), mean=seasonbase.after.lockdown.168$figure[c(49:168,1:48)], sd=seasonbase.after.lockdown.168$sd[c(49:168,1:48)], Weekday=rep(weekdays, each=24))
+ggplot(data=seasonbase.after.lockdown.168.df) +
+  geom_line(aes(hour, mean - sd, col=Weekday), alpha=.1) +
+  geom_line(aes(hour, mean + sd, col=Weekday), alpha=.1) +
+  geom_ribbon(aes(hour, ymin = mean - sd, ymax = mean + sd, fill = Weekday), alpha=0.1, show.legend=F) +
+  geom_line(aes(hour, mean, col=Weekday)) +
+  ylab("Mean hourly movement (nm)") +
+  scale_color_manual(values = rainbow(7), breaks=weekdays, aesthetics = c("colour", "fill"))
+```

### 2020-06-08 project.R

@@ -835,11 +835,11 @@ after.lockdown.start <- as_datetime('2020-03-14T00:00:00')
 after.lockdown.end <- as_datetime('2020-04-26T00:00:00') - seconds(1)
 before.lockdown <- build_ts(seis.h, periods = c(24), start = before.lockdown.start, end = before.lockdown.end, force_ms = T)
 after.lockdown <- build_ts(seis.h, periods = c(24), start = after.lockdown.start, end = after.lockdown.end, force_ms = T)
-# We first check the seasonplots, highlighting different weekdays
-seasonplot(before.lockdown, type="l", col=rainbow(7))
+# We first check the seasonplots, highlighting different weekdays. We divide by sd to account for spikes
+seasonplot(before.lockdown/sd(before.lockdown, na.rm=T), type="l", col=rainbow(7), main="Seasonal plot: before lockdown")
 legend("topleft", legend=weekdays(before.lockdown.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
-seasonplot(after.lockdown, type="l", col=rainbow(7)[c(6:7,1:5)])
-legend("topleft", legend=weekdays(after.lockdown.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7)[c(6:7,1:5)], lty=1, cex=0.8)
+seasonplot(after.lockdown/sd(after.lockdown, na.rm=T), type="l", col=rainbow(7)[c(6:7,1:5)], main="Seasonal plot: after lockdown")
+legend("topleft", legend=weekdays(after.lockdown.start + (as.difftime(c(0:6), units="days")))[c(3:7,1:2)], col=rainbow(7), lty=1, cex=0.8)
 
 
 
@@ -890,3 +890,21 @@ autoplot(ts(seasonbase.after.lockdown.168$figure[c(49:168,1:48)], frequency=1)) 
   geom_ribbon(aes(ymin = seasonbase.before.lockdown.168$figure - seasonbase.before.lockdown.168$sd, ymax = seasonbase.before.lockdown.168$figure + seasonbase.before.lockdown.168$sd), fill = "lightcoral", alpha=0.5) +
   geom_line(aes(y=seasonbase.before.lockdown.168$figure, col="Weekly before")) +
[…]

### 2020-06-09 draft.tex
add a couple of rows to draft
yesterday last work

@@
 
-\section{Parameter estimation}
+\section{Methodology}
 Having two different datasets, we need to fit two separate models that represents those stations.
@@
 
+Also, ARIMA parameters roots have values within the unit circle for the first two models, while for $\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$ two of the roots fall outside the unit circle, with values of $\approx1.49$ and $\approx1.80$.
+
 Given all these factors, we are confident choosing the $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$ model, being aware of the missed 24 hours seasonality.
@@
 
+Parameters roots for both models have values within the unit circle.
+
 Thus, we can select the $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$ model on the Box-Cox($\lambda\approx-1$) transformed data, still being aware of the missed 24 hours seasonality, similarly to the Uccle station.
@@
 
-\section{Conclusion}
 \subsection{Lockdown influence}
@@
 
-We can still compare how the seasonalities differ with respect with each original dataset and between each other.
-We extract the seasonalities for each station highlighting the standard deviation confidence band for an estimated error of 67\%.
+\subsubsection{Seasonalities divergence}
+We can still compare how the seasonalities differ with respect with each original dataset and between each other, by extracting the seasonalities for each station from before and after the lockdown (March, $14^{th}$).
+
+We will highlight the confidence bands for an estimated error of 67\%.
 This interval estimation lies on the assumption that values for each single hour in the hourly/weekly seasonality is Normally distributed, a slightly different assumption that is not necessarily refuted by the already seen Q-Q plots.
@@
 	\begin{center}
-		\caption{Seasonalities computed before and after the lockdown}{\small Seasonalities are extracted together}
+		\caption{Seasonalities computed before and after the lockdown}{\small Seasonalities extracted together, confidence bands $\pm\mathrm{sd}$}
 		\label{fig:season.lockdown}
@@
 
-As we can see in \figurename~\ref{fig:season.lockdown}, even if the seasonalities shapes seem very different, in particular regarding the weekly seasonality, the differences are still included in the highlighted confidence bands, apart from a few hours.
+As we can see in \figurename~\ref{fig:season.lockdown}, the seasonalities shapes seem different, in particular comparing the weekly seasonality.
+The seasonalities from before and after the lockdown are still included in the highlighted confidence bands, apart from a few hours.
 This is true for both stations.
-Yet, it is useful to notice that the weekly seasonalities include a possibly interesting daily component, influenced by the weekend drops in values.
 
-%\subsection{Lockdown forecast divergence}
+\begin{figure}[h]
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/seasonplot-mean-sd-1"}
+		\caption{Before lockdown}
+		\label{fig:season.lockdown:msb:mems}
+	\end{subfigure}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/seasonplot-mean-sd-2"}
+		\caption{After lockdown}
+		\label{fig:season.lockdown:msa:mems}
+	\end{subfigure}
+	\begin{center}
+		\caption{Seasonal plots for Membach station}
+		\label{fig:season.lockdown:ms:mems}
+	\end{center}
+\end{figure}
+
+The differences in shape are more highlighted checking the seasonal plots \cite{hyndman2018forecasting} in \figurename~\ref{fig:season.lockdown:ms:mems}.
+It is possible to notice a couple of changes: the Monday to Friday shapes are heavily flattened almost on the weekend values, and all the values are flattened towards 0.
+Both flattenings can be noticed checking the Uccle station, not showed for brevity.
+
+\subsubsection{Forecast divergence}
+\begin{figure}[h]
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-7"}
+		\caption{Uccle station}
+		\label{fig:forecast.lockdown:uccs}
+	\end{subfigure}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-7"}
+		\caption{Membach station}
+		\label{fig:forecast.lockdown:mems}
+	\end{subfigure}
+	\begin{center}
+		\caption{Forecasts from March, $2^{nd}$.}{\small In red the actual values, in black the forecasts. March, $14^{th}$ at midnight is highlighted.}
+		\label{fig:forecast.lockdown}
+	\end{center}
+\end{figure}
+
+By fitting the model\footnote{Data is different, because we are fitting the model on the original train+test dataset, but if we follow similar steps for both stations we can see that the final model we shall choose is the same.
+Also, the $\lambda$ parameter for the Box-Cox transformation of Membach data is estimated equally.} on data until March, $1^{st}$ we can compute forecasts for 8 weeks ahead, until April, $26^{th}$.
+As we can see in \figurename~\ref{fig:forecast.lockdown}, both stations overestimate starting after the lockdown.
+Uccle data in particular show a greater overestimation that can be noticed also by looking at the difference between the forecasted and measured values.
+
+Analytically, the MAPE for Uccle doubles with respect to the previous test set ($\approx34\%$), and only 58\% and 78\% of the test data falls inside the confidence intevals (80\% and 95\%).
+Analysing the Membach station, we find those measures have comparable values to the previous test set: the MAPE remains at $16.6\%$, while 74\% and 90\% of data falls inside the confidence intervals.
+
+%MEMS
+%Test set inside the CI
+%80% CI 	  74.18 %
+%95% CI 	  89.58 %
+%Errors:  forecast with ARIMA(2,0,0)(0,1,1)[168] 
+%MSE= 117.9258   MAE= 7.709243   MAPE= 16.60103   RMSE= 10.85937   R2= 0.4777765
+
+%UCCS
+%Test set inside the CI
+%80% CI 	  57.51 %
+%95% CI 	  78.42 %
+%Errors:  forecast with ARIMA(2,0,0)(0,1,1)[168] 
+%MSE= 670897.8   MAE= 664.5903   MAPE= 33.94793   RMSE= 819.0835   R2= 0.5188538
+
 %Uccle: remodel on 52+18 (until 1st March 2020), forecast 2 months (8 weeks) => diff in MAPE
@@
 
+\section{Conclusions}
+\subsubsection{Lockdown influence}
 \subsection{Limitations and future works}

### 2020-06-09 project.after.Rmd
yesterday last work

@@
 ))
-print(plot_forecast(plotting.forecast) + autolayer(final.testing, series="Original") + ylab("Mean hourly movement (nm)"))
+print(plot_forecast(plotting.forecast) + autolayer(final.testing, series="Original") + ylab("Mean hourly movement (nm)") + xlab("Weeks") + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2))
 plots <- show_overlap_and_diff(
@@
 )
-print(plot_grid(plots[[1]], plots[[2]], nrow = 2))
+print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
+                plots[[2]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
+                nrow = 2
+                ))
 
@@
 
+weekdays <- c("Monday", "Tuesday", "Wednseday", "Thursday", "Friday", "Saturday", "Sunday")
 ```
@@
 season.before.lockdown.168 <- trend_season(aggregated.before.lockdown-season.before.lockdown.24$seasonal, 168)
+season.before.lockdown.168.df <- data.frame(Hour=rep(1:24, 7), figure=season.before.lockdown.168$figure[1:168], sd=season.before.lockdown.168$sd[1:168], Weekday=rep(weekdays, each=24))
 seasonbase.before.lockdown.24 <- trend_season(aggregated.before.lockdown, 24)
+seasonbase.before.lockdown.24.df <- data.frame(Hour=1:24, figure=seasonbase.before.lockdown.24$figure, sd=seasonbase.before.lockdown.24$sd)
 seasonbase.before.lockdown.168 <- trend_season(aggregated.before.lockdown, 168)
+seasonbase.before.lockdown.168.df <- data.frame(Hour=rep(1:24, 7), figure=seasonbase.before.lockdown.168$figure[1:168], sd=seasonbase.before.lockdown.168$sd[1:168], Weekday=rep(weekdays, each=24))
 
 plot_grid(
-  autoplot(ts(seasonbase.before.lockdown.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
-    geom_line(aes(y=season.before.lockdown.24$figure, col="Multi daily")) +
+  autoplot(ts(seasonbase.before.lockdown.24.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
+    geom_line(aes(y=season.before.lockdown.24.df$figure, col="Multi daily")) +
     scale_colour_manual("", values=c("purple", "yellow"), breaks=c("Direct daily", "Multi daily")),
-  autoplot(ts(seasonbase.before.lockdown.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
-    geom_line(aes(y=season.before.lockdown.168$figure + rep(season.before.lockdown.24$figure, 7), col="Multi w+d")) +
+  autoplot(ts(seasonbase.before.lockdown.168.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
+    geom_line(aes(y=season.before.lockdown.168.df$figure + rep(season.before.lockdown.24.df$figure, 7), col="Multi w+d")) +
     #geom_line(aes(y=rep(seasonbase.before.lockdown.24$figure, 7), col="Direct daily"), size=1.5) +
     #geom_line(aes(y=rep(season.before.lockdown.24$figure, 7), col="Multi daily")) +
-    geom_line(aes(y=season.before.lockdown.168$figure, col="Multi weekly")) +
+    geom_line(aes(y=season.before.lockdown.168.df$figure, col="Multi weekly")) +
     scale_colour_manual("", values=c("red", "purple", "blue", "yellow", "green"), breaks=c("Direct weekly", "Direct daily", "Multi" %s+% c(" weekly", " daily", " w+d"))),
@@
 season.after.lockdown.168 <- trend_season(aggregated.after.lockdown-season.after.lockdown.24$seasonal, 168)
+season.after.lockdown.168.df <- data.frame(Hour=rep(1:24, 7), figure=season.after.lockdown.168$figure[c(49:168,1:48)], sd=season.after.lockdown.168$sd[c(49:168,1:48)], Weekday=rep(weekdays, each=24))
 seasonbase.after.lockdown.24 <- trend_season(aggregated.after.lockdown, 24)
+seasonbase.after.lockdown.24 <- trend_season(aggregated.after.lockdown, 24)
+seasonbase.after.lockdown.24.df <- data.frame(Hour=1:24, figure=seasonbase.after.lockdown.24$figure, sd=seasonbase.after.lockdown.24$sd)
 seasonbase.after.lockdown.168 <- trend_season(aggregated.after.lockdown, 168)
+seasonbase.after.lockdown.168.df <- data.frame(Hour=rep(1:24, 7), figure=seasonbase.after.lockdown.168$figure[c(49:168,1:48)], sd=seasonbase.after.lockdown.168$sd[c(49:168,1:48)], Weekday=rep(weekdays, each=24))
 
 plot_grid(
-  autoplot(ts(seasonbase.after.lockdown.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
+  autoplot(ts(seasonbase.after.lockdown.24.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
     geom_line(aes(y=season.after.lockdown.24$figure, col="Multi daily")) +
     scale_colour_manual("", values=c("purple", "yellow"), breaks=c("Direct daily", "Multi daily")),
-  autoplot(ts(seasonbase.after.lockdown.168$figure[c(49:168,1:48)], frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
-    geom_line(aes(y=season.after.lockdown.168$figure[c(49:168,1:48)] + rep(season.after.lockdown.24$figure, 7), col="Multi w+d")) +
+  autoplot(ts(seasonbase.after.lockdown.168.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
+    geom_line(aes(y=season.after.lockdown.168.df$figure + rep(season.after.lockdown.24$figure, 7), col="Multi w+d")) +
     #geom_line(aes(y=rep(seasonbase.after.lockdown.24$figure, 7), col="Direct daily"), size=1.5) +
     #geom_line(aes(y=rep(season.after.lockdown.24$figure, 7), col="Multi daily")) +
-    geom_line(aes(y=season.after.lockdown.168$figure[c(49:168,1:48)], col="Multi weekly")) +
+    geom_line(aes(y=season.after.lockdown.168.df$figure, col="Multi weekly")) +
     scale_colour_manual("", values=c("red", "purple", "blue", "yellow", "green"), breaks=c("Direct weekly", "Direct daily", "Multi" %s+% c(" weekly", " daily", " w+d"))),
@@
   autoplot(ts(season.after.lockdown.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
-    geom_ribbon(aes(ymin = season.after.lockdown.24$figure - season.after.lockdown.24$sd, ymax = season.after.lockdown.24$figure + season.after.lockdown.24$sd), fill = "mediumpurple1", alpha=0.5) +
-    geom_line(aes(col="Daily lockdown")) +
-    geom_ribbon(aes(ymin = season.before.lockdown.24$figure - season.before.lockdown.24$sd, ymax = season.before.lockdown.24$figure + season.before.lockdown.24$sd), fill = "lightcoral", alpha=0.5) +
-    geom_line(aes(y=season.before.lockdown.24$figure, col="Daily before")) +
-    scale_colour_manual("", values=c("purple", "red"), breaks=c("Daily lockdown", "Daily before")),
-  autoplot(ts(season.after.lockdown.168$figure[c(49:168,1:48)], frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
-    geom_ribbon(aes(ymin = (season.after.lockdown.168$figure - season.after.lockdown.168$sd)[c(49:168,1:48)], ymax = (season.after.lockdown.168$figure + season.after.lockdown.168$sd)[c(49:168,1:48)]), fill = "mediumpurple1", alpha=0.5) +
-    geom_line(aes(col="Weekly lockdown")) +
-    geom_ribbon(aes(ymin = season.before.lockdown.168$figure - season.before.lockdown.168$sd, ymax = season.before.lockdown.168$figure + season.before.lockdown.168$sd), fill = "lightcoral", alpha=0.5) +
-    geom_line(aes(y=season.before.lockdown.168$figure, col="Weekly before")) +
-    scale_colour_manual("", values=c("purple", "red"), breaks=c("Weekly lockdown", "Weekly before")),
+    geom_ribbon(aes(ymin = season.after.lockdown.24$figure  - season.after.lockdown.24$sd,  ymax = season.after.lockdown.24$figure  + season.after.lockdown.24$sd),  alpha=0.5, fill = "mediumpurple1") +
+    geom_line(aes(col="After lockdown")) +
+    geom_ribbon(aes(ymin = season.before.lockdown.24$figure - season.before.lockdown.24$sd, ymax = season.before.lockdown.24$figure + season.before.lockdown.24$sd), alpha=0.5, fill = "lightcoral") +
+    geom_line(aes(y=season.before.lockdown.24$figure, col="Before lockdown")) +
+    scale_colour_manual("", values=c("purple", "red"), breaks=c("After lockdown", "Before lockdown")),
+  autoplot(ts(season.after.lockdown.168.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
+    geom_ribbon(aes(ymin = season.after.lockdown.168.df$figure  - season.after.lockdown.168.df$sd,  ymax = season.after.lockdown.168.df$figure  + season.after.lockdown.168.df$sd),  alpha=0.5, fill = "mediumpurple1") +
+    geom_line(aes(col="After lockdown")) +
+    geom_ribbon(aes(ymin = season.before.lockdown.168.df$figure - season.before.lockdown.168.df$sd, ymax = season.before.lockdown.168.df$figure + season.before.lockdown.168.df$sd), alpha=0.5, fill = "lightcoral") +
+    geom_line(aes(y=season.before.lockdown.168.df$figure, col="Before lockdown")) +
+    scale_colour_manual("", values=c("purple", "red"), breaks=c("After lockdown", "Before lockdown")),
   nrow = 2)
 
-autoplot(ts(seasonbase.after.lockdown.168$figure[c(49:168,1:48)], frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
-  geom_ribbon(aes(ymin = (seasonbase.after.lockdown.168$figure - seasonbase.after.lockdown.168$sd)[c(49:168,1:48)], ymax = (seasonbase.after.lockdown.168$figure + seasonbase.after.lockdown.168$sd)[c(49:168,1:48)]), fill = "mediumpurple1", alpha=0.5) +
-  geom_line(aes(col="Weekly lockdown")) +
-  geom_ribbon(aes(ymin = seasonbase.before.lockdown.168$figure - seasonbase.before.lockdown.168$sd, ymax = seasonbase.before.lockdown.168$figure + seasonbase.before.lockdown.168$sd), fill = "lightcoral", alpha=0.5) +
-  geom_line(aes(y=seasonbase.before.lockdown.168$figure, col="Weekly before")) +
-  scale_colour_manual("", values=c("purple", "red"), breaks=c("Weekly lockdown", "Weekly before"))
+autoplot(ts(seasonbase.after.lockdown.168.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
+  geom_ribbon(aes(ymin = (seasonbase.after.lockdown.168.df$figure - seasonbase.after.lockdown.168.df$sd), ymax = (seasonbase.after.lockdown.168.df$figure + seasonbase.after.lockdown.168.df$sd)), fill = "mediumpurple1", alpha=0.5) +
+  geom_line(aes(col="After lockdown")) +
+  geom_ribbon(aes(ymin = seasonbase.before.lockdown.168.df$figure - seasonbase.before.lockdown.168.df$sd, ymax = seasonbase.before.lockdown.168.df$figure + seasonbase.before.lockdown.168.df$sd), fill = "lightcoral", alpha=0.5) +
+  geom_line(aes(y=seasonbase.before.lockdown.168.df$figure, col="Before lockdown")) +
+  scale_colour_manual("", values=c("purple", "red"), breaks=c("After lockdown", "Before lockdown"))
 ```
 ```{r seasonalities/seasonplot-mean-sd}
-weekdays <- c("Monday", "Tuesday", "Wednseday", "Thursday", "Friday", "Saturday", "Sunday")
-
-seasonbase.before.lockdown.168.df <- data.frame(hour=rep(1:24, 7), mean=seasonbase.before.lockdown.168$figure[1:168], sd=seasonbase.before.lockdown.168$sd[1:168], Weekday=rep(weekdays, each=24))
 ggplot(data=seasonbase.before.lockdown.168.df) +
-  geom_line(aes(hour, mean - sd, col=Weekday), alpha=.1) +
-  geom_line(aes(hour, mean + sd, col=Weekday), alpha=.1) +
-  geom_ribbon(aes(hour, ymin = mean - sd, ymax = mean + sd, fill = Weekday), alpha=0.1, show.legend=F) +
-  geom_line(aes(hour, mean, col=Weekday)) +
+  geom_line(aes(Hour, figure - sd, col=Weekday), alpha=.1) +
+  geom_line(aes(Hour, figure + sd, col=Weekday), alpha=.1) +
+  geom_ribbon(aes(Hour, ymin = figure - sd, ymax = figure + sd, fill = Weekday), alpha=0.1, show.legend=F) +
+  geom_line(aes(Hour, figure, col=Weekday)) +
   ylab("Mean hourly movement (nm)") +
+  labs(title="Seasonal plot: before lockdown") +
   scale_color_manual(values = rainbow(7), breaks=weekdays, aesthetics = c("colour", "fill"))
 
-seasonbase.after.lockdown.168.df <- data.frame(hour=rep(1:24, 7), mean=seasonbase.after.lockdown.168$figure[c(49:168,1:48)], sd=seasonbase.after.lockdown.168$sd[c(49:168,1:48)], Weekday=rep(weekdays, each=24))
 ggplot(data=seasonbase.after.lockdown.168.df) +
-  geom_line(aes(hour, mean - sd, col=Weekday), alpha=.1) +
-  geom_line(aes(hour, mean + sd, col=Weekday), alpha=.1) +
-  geom_ribbon(aes(hour, ymin = mean - sd, ymax = mean + sd, fill = Weekday), alpha=0.1, show.legend=F) +
-  geom_line(aes(hour, mean, col=Weekday)) +
+  geom_line(aes(Hour, figure - sd, col=Weekday), alpha=.1) +
+  geom_line(aes(Hour, figure + sd, col=Weekday), alpha=.1) +
+  geom_ribbon(aes(Hour, ymin = figure - sd, ymax = figure + sd, fill = Weekday), alpha=0.1, show.legend=F) +
+  geom_line(aes(Hour, figure, col=Weekday)) +
   ylab("Mean hourly movement (nm)") +
+  labs(title="Seasonal plot: after lockdown") +
   scale_color_manual(values = rainbow(7), breaks=weekdays, aesthetics = c("colour", "fill"))

### 2020-06-09 project.R

@@ -769,6 +769,7 @@ chosen.model <- c(2,0,0,0,1,1)
 # we finally set the chosen model for Membach
 chosen.model <- c(2,0,0,0,1,1)
 # they seem equal, but the chosen.lambda is actually different
+#################### Parameters Estimation END ####################
 
 
 
@@ -807,7 +808,7 @@ cat("\n")
 plotting.model <- final.model
 plotting.forecast <- final.forecast
 # plot forecasts
-print(plot_forecast(plotting.forecast) + autolayer(final.testing, series="Original") + ylab("Mean hourly movement (nm)"))
+print(plot_forecast(plotting.forecast) + autolayer(final.testing, series="Original") + ylab("Mean hourly movement (nm)") + xlab("Weeks"))
 # print stats about data falling inside 80% and 90% CIs
 cat("Test set inside the CI\n")
 cat("80% CI \t ", print_dec(mean(plotting.forecast$lower[,1] < final.testing & final.testing < plotting.forecast$upper[,1], na.rm = T)*100), "%\n")
@@ -843,6 +844,7 @@ legend("topleft", legend=weekdays(after.lockdown.start + (as.difftime(c(0:6), un
 
 
 
+weekdays <- c("Monday", "Tuesday", "Wednseday", "Thursday", "Friday", "Saturday", "Sunday")
 # Then, we proceed with the mean-over-period method, with 3 cycles for concurrent extraction, before lockdown
 aggregated.before.lockdown <- seasonality.aggregate(before.lockdown)
 season.before.lockdown.24 <- trend_season(aggregated.before.lockdown, 24)
@@ -851,9 +853,9 @@ season.before.lockdown.24 <- trend_season(aggregated.before.lockdown-season.befo
 season.before.lockdown.168 <- trend_season(aggregated.before.lockdown-season.before.lockdown.24$seasonal, 168)
 season.before.lockdown.24 <- trend_season(aggregated.before.lockdown-season.before.lockdown.168$seasonal, 24)
 season.before.lockdown.168 <- trend_season(aggregated.before.lockdown-season.before.lockdown.24$seasonal, 168)
[…]

### 2020-06-09 draft.tex
first draft fix from IAV

@@
 
-The vertical ground displacement is represented in by the minimum and maximum displacement of the ground in the vertical axis, with $nm$ as unit of measure.
-This can easily be transformed in total ground displacement for that second by their difference.
+The vertical ground displacement is represented by the minimum and maximum displacements of the ground in the vertical axis, with $nm$ as unit of measure.
+This can easily be transformed to total ground displacement for that second by their difference.
 For each day, a pair per second is available (86400 per day, around 30 million per year).
 
-ROB collects data from 9 stations. Of those, the following 6 have interesting positions, related to possible correlations with human activity:
+Out of the 9 stations, the following 6 have interesting positions, related to possible correlations with human activity:
 \begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
@@
 %Also data from \textbf{Ostenda} station could be interesting, given the possible sea influence.
-Considered these factors, we selected the \textbf{Uccle} and \textbf{Membach} stations.
+Considering these factors, we selected the \textbf{Uccle} and \textbf{Membach} stations.
 
@@
 \subsubsection{Data gathering and cleanup}
-The data is provided with a CSV per day, with only two columns related to the min-max pair \cite{Data20200402}. This causes two problems:
+The data is provided in CSV format, one file per day, with only two columns related to the min-max pair \cite{Data20200402} and no column related to time. This causes two problems:
 \begin{enumerate}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
@@
 \subsubsection{Data aggregation and initial transformation}
-Since we are not interested in a second-precise analysis, and it would be computationally expensive to deal with the whole dataset, we are going to aggregate data by hour.
-The aggregation will get us around 14 thousand points, that can still result in interpretable models.
+Since we are not interested in a second-precise analysis, and it would be computationally expensive to deal with the whole dataset, we aggregate data by hour.
+The aggregation will result in around 14 thousand points, that can still be captured by interpretable models.
 
-While performing this operation, we should consider the correct timezone: any seasonal analysis dealing with UTC data would be out of phase between the half-year when Belgium follows the UTC+01 timezone, and the half-year when it follows the UTC+02 timezone.
-With the help of the \texttt{lubridate} R package we converted to the actual timezones, and then we kept the actual objects without being timezone explicit, because all the methods we are going to use are smart enough to work on the UTC conversion of datetimes.
-This transformation made it possible to have, for example, the time 6:00 in the timeseries to represent the local time in Belgium (\textit{Europe/Bruxelles}), without being influenced by the actual timezone on that specific day.
+While performing this operation, we should consider the correct time-zone: any seasonal analysis dealing with UTC data would be out of phase between the half-year when Belgium follows the UTC+01 timezone, and the half-year when it follows the UTC+02 timezone.
+With the help of the \texttt{lubridate} R package we converted to the actual timezones, and then we kept the actual objects without being timezone explicit, because all the methods we use are smart enough to work on the UTC conversion of datetimes.
+This transformation made it possible to have, for example, the time 6:00 in the timeseries to represent the local time in Belgium (\textit{Europe/Bruxelles}), without being influenced by the actual time-zone on that specific day.
 
-The aggregation was performed considering the mean max-min difference, for the following reasons:
+The aggregation was performed considering the mean ground displacement, for the following reasons:
 \begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
 	\item When an hour contains some missing seconds, these do not heavily affect the aggregation result, in contrast with summing the differences
-	\item Seismic events, such as quakes and mine explosions, normally last some minutes at most, and will affect the result only slightly, rather than keeping the hourly maximum
-	\item Normally, maximum values in defined period of times do not follow the Normal distribution, while some models that we are going to use assume the data follow a Normal distribution
+	\item Seismic events, such as quakes and mine explosions, normally last some minutes at most, and will affect the result only slightly, as opposed to the hourly maximum
+	\item Usually, maximum value aggregation does not follow the Normal distribution, while some models that we are going to use assume the data follow a Normal distribution
 \end{itemize}
@@
 After aggregating by hour, the percentages of missing values remain quite stable.
-As can be seen from \figurename{s} \ref{missing-values-seconds:uccs} and \ref{missing-values-seconds:mems}, both stations missing values are grouped in few consecutive periods.
-To deal with them when R methods need complete timeseries, we will define different completion strategies depending on the seasonality we will be studying, always based on the R function \texttt{na.aggregate}.
+As can be seen from \figurename{s} \ref{missing-values-seconds:uccs} and \ref{missing-values-seconds:mems}, missing values at both stations are grouped in a few consecutive periods.
+To deal with them when R methods need complete timeseries, we define different completion strategies, depending on the seasonality, using the R function \texttt{na.aggregate}.
 
@@
 
-Since we will be applying this method on the dataset indexed by a datetime object, which uses the second as the unit, the transformation of frequency $f$ into hourly periodicity $p$ is obtained by $p=(1/f)/3600$.
+Since we apply this method on a dataset indexed by seconds, the transformation of the frequency $f$ into hourly periodicity $p$ is obtained by $p=(1/f)/3600$.
 We can analyse the resulting periodgrams for Uccle and Membach stations in \figurename{s} \ref{fig:periodgram:uccs} and \ref{fig:periodgram:mems}.
@@
 As best practices suggest, we are going to divide the dataset into training and testing datasets with a ratio around $85\%/15\% - 75\%/25\%$.
-Since one of our objectives is to assess any significant difference from before and after the Belgium lockdown (on March, $14^{th}$ the \textit{soft} lockdown, while on March, $18^{th}$ the complete lockdown), we selected data from the start of our timeseries until 20-01-2020 (64 weeks $\approx 82\%$) as training, while the test follows and ends on 27-04-2020 (14 weeks $\approx 18\%$).
+Since one of our objectives is to assess any significant difference from before and after the Belgium lockdown (on March $14^{th}$ the \textit{soft} lockdown, while on March $18^{th}$ the complete lockdown), we selected data from the start of our timeseries until 20-01-2020 (64 weeks $\approx 82\%$) as training, while the test follows and ends on 27-04-2020 (14 weeks $\approx 18\%$).
 
 This initial division demostrated to be computationally heavy while fitting some models with weekly seasonality, and we noticed a possible yearly seasonality on Uccle data (\figurename~\ref{fig:whole-data-decomposition}).
-Even if it is not demonstrable given the short period our data comes from, applying a 24-hour MA smoothing filter on the whole dataset (546 days) and then producing the periodgram as previously done, it suggests a 8748-hours significant periodicity (364.5 days).
+Even if it is not demonstrable given the short period our data covers, applying a 24-hour MA smoothing filter on the whole dataset (546 days) and then producing the periodgram as previously done, it suggests a 8748-hours significant periodicity (364.5 days).
 
@@
 	\includegraphics[width=0.7\linewidth]{"project.uccs_files/figure-latex/whole data plot-2"}
-	\caption{Whole Uccle data decomposition}{\footnotesize Weeks 10 and 62 include January, $1^{st}~2019/2020$}
+	\caption{Whole Uccle data decomposition}{\footnotesize Weeks 10 and 62 include January $1^{st}~2019/2020$}
 	\label{fig:whole-data-decomposition}
@@
 While data from Uccle station have a clear significativity of both trend and seasonalities, data from Membach could seem not significant.
-Still, if we notice that the main timeseries lies between values of 0 and 100, we can find out that around half of the value will be composed by the trend, a 30\% by the joined seasonalities component, and the remaining value by the residuals.
+Still, if we notice that the main timeseries lies between values of 0 and 100, we can find out that around half of the value will be composed by the trend, a 30\% by the joined seasonality components, and the remaining value by the residuals.
 We will need to better handle the spikes that start from the middle of that dataset.
@@
 \section{Methodology}
-Having two different datasets, we need to fit two separate models that represents those stations.
+Having two different datasets, we need to fit two separate models that represent the two stations.
 During this estimation, our aim is to obtain a model that generalises our training dataset in such a way that forecasting becomes effective, also assessing errors analytically.
@@
 \begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
-	\item non-seasonal AR process of order $p$ is described by an exponentially decaying ACF, and by a cut-off of significant PACF values at lag $p$
-	\item non-seasonal MA process of order $q$ is described by an exponentially decaying PACF, and by a cut-off of significant ACF values at lag $q$
-	\item seasonal AR process of order $P$ is described by an exponentially decay at seasonal lags on ACF values, and by a cut-off of significant PACF values at each seasonal lag until the $P^{th}$
-	\item seasonal MA process of order $Q$ is described by an exponentially decay at seasonal lags on PACF values, and by a cut-off of significant ACF values at each seasonal lag until the $Q^{th}$
+	\item non-seasonal AR process of order $p$ is described by an exponentially or sigmoidally decaying ACF, and by a cut-off of significant PACF values at lag $p$
+	\item non-seasonal MA process of order $q$ is described by an exponentially or sigmoidally decaying PACF, and by a cut-off of significant ACF values at lag $q$
+	\item seasonal AR process of order $P$ is described by an exponentially or sigmoidally decay at seasonal lags on ACF values, and by a cut-off of significant PACF values at each seasonal lag until the $P^{th}$
+	\item seasonal MA process of order $Q$ is described by an exponentially or sigmoidally decay at seasonal lags on PACF values, and by a cut-off of significant ACF values at each seasonal lag until the $Q^{th}$
 	\item given how significance is assessed for the ACF/PACF spikes, one over 20 spikes (5\%) could be \textit{coincidentally significant} or \textit{coincidentally \textbf{non}-significant}
@@
 \subsubsection{ARIMA}
-Our training set, despite the precautions taken during the initial data cleanup and aggregation, doesn't satisfy the Normality assumption, even after applying the \textit{log} or \textit{Box-Cox} transformations.
+Our training set, despite the precautions taken during the initial data cleanup and aggregation, does not satisfy the Normality assumption, even after applying the \textit{log} or \textit{Box-Cox} transformations.
 Thus, we will estimate the models using the original scale.
 
-Since our dataset shows a relevant trend component, resulting in non-constant mean, we consider the usage of the \textit{diff}\footnote{The \textit{diff} operator applies the \textit{d}-order differences, by default on lag 1 (\textit{simple}).
+Since our dataset shows a relevant trend component, resulting in non-constant mean, we consider the use of the \textit{diff}\footnote{The \textit{diff} operator applies the \textit{d}-order differences, by default on lag 1.
 All \textit{diff} operators we will apply use $d=1$.} operator to smooth the mean.
-After applying the \textit{diff} operator with lag 168, in conjunction with the simple \textit{diff} or not, we come up with two possible models by checking the ACF and PACF plots:
+After applying the \textit{diff} operator with lag 168, in conjunction with the lag 1 \textit{diff} or not, we come up with two possible models by checking the ACF and PACF plots:
 \begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
 	\item from using both \textit{diff}s, the ACF and PACF on the long run hint to $\mathrm{ARIMA}(0,1,0)(0,1,1)_{168}$
-	\item from using just the lagged \textit{diff}, the PACF on the first lags and the ACF/PACF combination on the long run hint to $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$
+	\item from using just the lag 168 \textit{diff}, the PACF on the first lags and the ACF/PACF combination on the long run hint to $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$
 \end{itemize}
@@
 
-The ACF and PACF of the residuals for all three models are really similar (\figurename~\ref{fig:arima-26}), even if the models themselves have great differences in meaning and in number of parameters: they highlight a clear seasonal ARMA process with periodicity 24.
+The ACF and PACF of the residuals for all three models are very similar (\figurename~\ref{fig:arima-26}), even if the models themselves have great differences in meaning and in number of parameters: they highlight a clear seasonal ARMA process with periodicity 24.
 This shows how our previous conclusion that the 168 hours seasonality could include the 24 hours one is flawed.
@@
 The transformed data still shows no constant mean, even if it has a way better shape than the original data, confirmed by the Q-Q plot (\figurename~\ref{fig:mems:box-cox:qq}).
-Thus, we consider the usage of the \textit{diff} operator, simple and at lag 168.
+Thus, we consider the usage of the \textit{diff} operator, at lag 1 and 168.
 Interestingly, the analysis of ACF and PACF plots hints to similar models to the Uccle station:
@@
 \subsubsection{Seasonalities divergence}
-We can still compare how the seasonalities differ with respect with each original dataset and between each other, by extracting the seasonalities for each station from before and after the lockdown (March, $14^{th}$).
+We can still compare how the seasonalities differ with respect with each original dataset and between each other, by extracting the seasonalities for each station from before and after the lockdown (March $14^{th}$).
 
@@
 
-As we can see in \figurename~\ref{fig:season.lockdown}, the seasonalities shapes seem different, in particular comparing the weekly seasonality.
+As we can see in \figurename~\ref{fig:season.lockdown}, the seasonalities' shapes seem different, in particular comparing the weekly seasonality.
 The seasonalities from before and after the lockdown are still included in the highlighted confidence bands, apart from a few hours.
@@
 	\begin{center}
-		\caption{Forecasts from March, $2^{nd}$.}{\small In red the actual values, in black the forecasts. March, $14^{th}$ at midnight is highlighted.}
+		\caption{Forecasts from March $2^{nd}$.}{\small In red the actual values, in black the forecasts. March $14^{th}$ at midnight is highlighted.}
 		\label{fig:forecast.lockdown}
@@
 By fitting the model\footnote{Data is different, because we are fitting the model on the original train+test dataset, but if we follow similar steps for both stations we can see that the final model we shall choose is the same.
-Also, the $\lambda$ parameter for the Box-Cox transformation of Membach data is estimated equally.} on data until March, $1^{st}$ we can compute forecasts for 8 weeks ahead, until April, $26^{th}$.
+Also, the $\lambda$ parameter for the Box-Cox transformation of Membach data is estimated equally.} on data until March $1^{st}$ we can compute forecasts for 8 weeks ahead, until April $26^{th}$.
 As we can see in \figurename~\ref{fig:forecast.lockdown}, both stations overestimate starting after the lockdown.

### 2020-06-09 draft.tex
more fixes on the draft and project code

@@
 
-Inspired by an article published during the initial phases of the Belgium-wide lockdown \cite{NatureCoronavirusSeismic}, we want to assess whether the lockdown period can be correlated with a visible change in seismic activity.
+Inspired by an article published during the initial phases of the Belgium-wide lockdown \cite{NatureCoronavirusSeismic}, we want to assess whether the lockdown period can be correlated with a visible change in background seismic activity.
+
 It is important to notice that some industrial events, such as mine explosions, are already mapped by seismologic institutions since they interfere with seismic surveys \cite{OtherSeismicEvents}.
@@
 
-The vertical ground displacement is represented by the minimum and maximum displacements of the ground in the vertical axis, with $nm$ as unit of measure.
-This can easily be transformed to total ground displacement for that second by their difference.
-For each day, a pair per second is available (86400 per day, around 30 million per year).
+The vertical ground displacement is represented by the minimum and maximum displacements of the ground in the vertical axis registered in a second, with $nm$ as unit of measure.
+This can easily be transformed to total ground displacement for each second by taking their difference.
+There are 86400 measurements per day, around 30 million per year.
 
@@
 \begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
-	\item \textbf{Uccle}, \textbf{Sart-Tilman}, and \textbf{Ében-Émael} are near a city (Bruxelles, Liege, and Maastricht).
+	\item \textbf{Uccle}, \textbf{Sart-Tilman}, and \textbf{Ében-Émael} are near a city (Brussels, Liege, and Maastricht).
 	\item \textbf{Ostenda} This station is near the city of Ostenda, that borders with the sea.
@@
 
-It is important to notice that the daily data in each file refers to the UTC time instead of the local Europe/Brussels time, an optimal choice to publish data in a machine-readable way, even if this could cause some issues while interpreting the results.
+It is important to notice that the daily data in each file refers to the UTC time instead of the local Europe/Brussels time, an optimal choice to publish data in a machine-readable way.
+This causes an issue fitting a fixed seasonality, when it depends on the legal time rather than on the absolute time.
 
@@
 \begin{enumerate}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
-	\item Before merging different days, we need to add a column with the complete datetime for each row
-	\item Days with some \texttt{NA} values cannot skip rows, so we will find placeholder values that indicate them $(\min, \max) \in \{(0,0), (-1,1)\}$
+	\item For merging days, we need to add a column with the second-precise datetime
+	\item In days with missing values there are placeholder values instead a missing row, that would make the following rows misinterpreted for measurements on the wrong second.
+The placeholder values are one of these two min-max pairs: $(0,0)\, (-1,1)$
 \end{enumerate}
@@
 Data from each station have different domains, as distinct equipment can register a specific amplitude of the seismic movements.
-In fact, we found some values outside the domains, transformed in \texttt{NA}.
+In fact, we found some values\footnote{Values out of domain for both stations are less than 10 out of $\approx50$ million.} outside the domains, and we considered them as missing values.
 
-In particular, we find that out of $\approx50$ million seconds, from 27-10-2018 to 28-05-2020, only 0.67\% (Uccle) and 0.37\% (Membach) have invalid values.
+In particular, we find that out of $\approx50$ million seconds, from 27-10-2018 to 28-05-2020, only 0.67\% (Uccle) and 0.37\% (Membach) have missing values.
 We will see later how they are distributed for each station.
@@
 Since we are not interested in a second-precise analysis, and it would be computationally expensive to deal with the whole dataset, we aggregate data by hour.
-The aggregation will result in around 14 thousand points, that can still be captured by interpretable models.
+The aggregation results in around 14 thousand points, that can still be captured by interpretable models.
 
-While performing this operation, we should consider the correct time-zone: any seasonal analysis dealing with UTC data would be out of phase between the half-year when Belgium follows the UTC+01 timezone, and the half-year when it follows the UTC+02 timezone.
-With the help of the \texttt{lubridate} R package we converted to the actual timezones, and then we kept the actual objects without being timezone explicit, because all the methods we use are smart enough to work on the UTC conversion of datetimes.
-This transformation made it possible to have, for example, the time 6:00 in the timeseries to represent the local time in Belgium (\textit{Europe/Bruxelles}), without being influenced by the actual time-zone on that specific day.
+While performing this operation, we should consider the correct time-zone: any seasonal analysis dealing with UTC dates would be out of phase for the half-year of legal time, with respect to the half-year of solar time.
+With the help of the \texttt{lubridate} R package we converted times to local time-zones, and then we kept them as absolute values, because all the R functions we use work on the UTC conversion of datetimes by default.
+This transformation made it possible to have each hour in the timeseries to represent the local time in Belgium (\textit{Europe/Brussels}), without being influenced by the legal time.
 
-The aggregation was performed considering the mean ground displacement, for the following reasons:
+The aggregation was performed considering the mean total ground displacement, for the following reasons:
 \begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
-	\item When an hour contains some missing seconds, these do not heavily affect the aggregation result, in contrast with summing the differences
+	\item When an hour contains some missing seconds, these do not heavily affect the aggregation result, in contrast with summing the values
 	\item Seismic events, such as quakes and mine explosions, normally last some minutes at most, and will affect the result only slightly, as opposed to the hourly maximum
@@
 
-In particular, after a visual analysis of the decomposition results, we can suppose the 90 hours gap from the Uccle station is related to some changes to the seismic equipment, that resulted in slightly more precise measurements.
-
-\begin{figure}[h]
+\begin{figure}[t]
 	\begin{subfigure}{.5\linewidth}
@@
 \subsubsection{Seasonality assumptions}
-The already cited article \cite{NatureCoronavirusSeismic} suggests the existence of a weekly seasonality. Looking at the data it seems confirmed, in addition to a daily seasonality.
+The already cited article \cite{NatureCoronavirusSeismic} suggests the existence of a weekly seasonality.
+Looking at the data it seems confirmed, in addition to a daily seasonality (\figurename~\ref{fig:zoom}).
 
@@
 In other words, we want to avoid searching for correlation with human activity while imposing human-based seasonality periods.
-Thus, we will use an analytical method to extract seasonalities that the data itself suggests to us.
+Thus, we use an analytical method to extract seasonalities that the data itself suggests to us.
 Still, we recognize that the already performed transformations impose an human-based time alteration to the data itself (Section \ref{sec:cleanup-transform}).
 
-The method we are going to apply is based on the Fourier Transform, and generates a \textit{periodgram} highlighting the most interesting frequencies that might correspond to a seasonality component.
+\begin{figure}[h]
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/whole data plot-2"}
+		\caption{Uccle station}
+		\label{fig:zoom:uccs}
+	\end{subfigure}
+	\begin{subfigure}{.5\linewidth}
+		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/whole data plot-2"}
+		\caption{Membach station}
+		\label{fig:zoom:mems}
+	\end{subfigure}
+	\caption{Zoom from January $6^{th}$ to February $3^{rd}$ 2019}
+	\label{fig:zoom}
+\end{figure}
+
+The method we are going to apply is based on the Fourier Transform (FT), and generates a \textit{periodgram} highlighting the most interesting frequencies that might correspond to a seasonality component.
 Applying the FT for any frequency, we transform our 1-dimensional data into a 2-dimensional data mapped around a circle.
@@
 A common operation to increase the difference between significant and non-significant periodicities is to apply a MA smoothing filter with a smaller window than the interesting ones.
-We will use such a filter with $f=2$.
+We will use such a filter with $p=2$.
 
@@
 
-% From the Membach station, we can see an anomalous spike at a 6-hour periodicity.
-
 \subsubsection{Dataset division}
@@
 
-This initial division demostrated to be computationally heavy while fitting some models with weekly seasonality, and we noticed a possible yearly seasonality on Uccle data (\figurename~\ref{fig:whole-data-decomposition}).
-Even if it is not demonstrable given the short period our data covers, applying a 24-hour MA smoothing filter on the whole dataset (546 days) and then producing the periodgram as previously done, it suggests a 8748-hours significant periodicity (364.5 days).
+Having the test set span from before to after the lockdown enables us to assess significant changes in the forecast performance on the two different periods.
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.7\linewidth]{"project.uccs_files/figure-latex/whole data plot-2"}
-	\caption{Whole Uccle data decomposition}{\footnotesize Weeks 10 and 62 include January $1^{st}~2019/2020$}
-	\label{fig:whole-data-decomposition}
-\end{figure}
-
-Given those issues, we choose the training/testing sets to hold 52/18 weeks: this is the greater train-to-test ratio we deem usable ($\approx 75\%/25\%$), while maintaining computational feasibility, and we hope it will render the possible yearly seasonality less intrusive.
+This initial division demostrated to be computationally heavy while fitting some models with weekly seasonality, so that it would be challenging to fit more than a couple of models.
+Given this issue, we choose the training/testing sets to hold 52/18 weeks: this is the greater train-to-test ratio we deem usable ($\approx 75\%/25\%$), while maintaining computational feasibility.
+The assessing of the lockdown influence will be discussed in Section~\ref{sec:methodology}.
 
@@
 
-In \figurename~\ref{fig:direct-multi-seasonality} we see both approaches applied to both stations. Those plots show an interesting property of our seasonalities:
+In \figurename~\ref{fig:direct-multi-seasonality} we see both approaches applied to both stations.
+Those plots show interesting properties of our seasonalities, probably given by the method (local mean) and the fact that the weekly period is an exact multiple of the daily period:
 \begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
@@
 Still, if we notice that the main timeseries lies between values of 0 and 100, we can find out that around half of the value will be composed by the trend, a 30\% by the joined seasonality components, and the remaining value by the residuals.
-We will need to better handle the spikes that start from the middle of that dataset.
+We will need to apply some transformation to flatten the spikes that start from the middle of that dataset.
 
-\section{Methodology}
-Having two different datasets, we need to fit two separate models that represent the two stations.
-During this estimation, our aim is to obtain a model that generalises our training dataset in such a way that forecasting becomes effective, also assessing errors analytically.
+\section{Methodology}\label{sec:methodology}
+Having two different datasets, we will operate on them separately and then we will compare the model interpretation.
+
+For each station, we fit a model that generalises our training dataset in such a way that forecasting becomes effective, also assessing errors analytically.
 Obtaining such a reliable model will be the first step to verify any significant difference forecasting over the lockdown period.
 
-The first model we are going to estimate is the ARIMA model, including the seasonality components.
+The model we are going to estimate is the ARIMA model, including the seasonality components.
 To identify the order of the different components described by ARIMA, we will analyse the autocorrelation and partial autocorrelation of our datasets.
@@
 
-Since our dataset shows a relevant trend component, resulting in non-constant mean, we consider the use of the \textit{diff}\footnote{The \textit{diff} operator applies the \textit{d}-order differences, by default on lag 1.
-All \textit{diff} operators we will apply use $d=1$.} operator to smooth the mean.
+Since our dataset shows a relevant trend component, resulting in non-constant mean, we consider the use of the \textit{diff}\footnote{The \textit{diff} operator applies the \textit{d}-order differences.
+All \textit{diff} operators we apply use $d=1$.} operator to smooth the mean.
 After applying the \textit{diff} operator with lag 168, in conjunction with the lag 1 \textit{diff} or not, we come up with two possible models by checking the ACF and PACF plots:
@@
 \end{itemize}
-We also employ the \textit{auto.arima} stepwise methodology\footnote{The \textit{auto.arima} stepwise methodology fits different ARIMA models within a range of orders, and based on each result's information criteria it will try to minimize it by changing each time one of the \textit{p}, \textit{d}, \textit{q}, \textit{P}, \textit{D}, or \textit{Q} values, depending on the configured limits.}, locking the \textit{diff} values to 0 or 1, to get some hints on the possible models.
+We also employ the \textit{auto.arima} stepwise methodology\footnote{The \textit{auto.arima} stepwise methodology fits different ARIMA models within a range of orders.
+Based on each result's information criterion (Akaike), it changes each time one of the \textit{p}, \textit{d}, \textit{q}, \textit{P}, \textit{D}, or \textit{Q} values, depending on the configured limits.
+These changes minimizes the information criteria of the final model.}, locking the \textit{diff} values to 0 or 1, to get some hints on the possible models.
 
 Following both the analytical ways and the \textit{auto.arima} hints, and refining the models through further inspection of ACF/PACF plots of the residuals, we end up with three characteristic models.
-We selected them through the Akaike Information Criterion (AIC), being the only three out of the fitted models with $\mathrm{AIC} < 120.000$: $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$, $\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$, and $\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$.
+We selected them through the Akaike Information Criterion (AIC)\footnote{Actually, we could have used the corrected AIC as it is usually more precise, but the values of AIC and AICc are equal to the first decimal for all our models ($\lesssim 10^{-7}\%$ difference).}, being the only three out of the fitted models with $\mathrm{AIC} < 120.000$: $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$, $\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$, and $\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$.
 
@@
 
-In this case, the employment of \textit{auto.arima} stepwise methodology does not lead to any useful model, since it obtains $\mathrm{AICc}=\inf$ for the two manually-chosen models, that have an AICc value around -84.500, while the best model found by that methodology has an AICc around -81.000.
+In this case, the employment of \textit{auto.arima} stepwise methodology\footnote{The \textit{auto.arima} stepwise methodology fits different ARIMA models within a range of orders.
+Based on each result's information criterion (Akaike), it changes each time one of the \textit{p}, \textit{d}, \textit{q}, \textit{P}, \textit{D}, or \textit{Q} values, depending on the configured limits.
+These changes minimizes the information criteria of the final model.} does not lead to any useful model, since it obtains $\mathrm{AIC}=\inf$ for the two manually-chosen models, that have an AIC value around -84.500, while the best model found by that methodology has an AIC around -81.000.
 This can be caused by approximation errors\footnote{The \textit{auto.arima} methodology was only used with initial approximation, due to high computing time.}, as we supposed earlier.

### 2020-06-09 project.Rmd
more fixes on the draft and project code

@@
 blocks$interval <- seconds(blocks$end - blocks$start + 1)
-blocks$start <- seis.first.second + seconds(blocks$start - 1)
-blocks$end <- seis.first.second + seconds(blocks$end - 1)
+blocks$start <- seis.first.second + seconds(blocks$start + 1)
+blocks$end <- seis.first.second + seconds(blocks$end + 1)
 plot(blocks$start, blocks$interval,
@@
      main="Consecutive seconds of missing values",
-     ylab="Seconds", xlab="Interval start",
+     ylab="Seconds", xlab="Interval start (mm-'yy)",
      xaxt="n", xlim=c(min(seis$utc_second), max(seis$utc_second))
 )
-axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-%y", cex=0.8)
+axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-'%y", cex=0.8)
 abline(h=3600, col="green", lty="dotted")
@@
 blocks$interval <- hours(blocks$end - blocks$start + 1)
-blocks$start <- seis.first.second + seconds(blocks$start - 1)
-blocks$end <- seis.first.second + seconds(blocks$end - 1)
+blocks$start <- seis.first.second + hours(blocks$start + 1)
+blocks$end <- seis.first.second + hours(blocks$end + 1)
 plot(c(rbind(blocks$start, blocks$end, NA)), rep(hour(blocks$interval), each=3),
@@
   scale_x_continuous(breaks=hours.from.start/168, labels=format(month.starts, "%m-%y")) +
-  ylab("mean hourly movement (nm)") + xlab(NULL) + ggtitle("Training set") + labs(caption = "Data in original scale")
+  ylab("mean hourly movement (nm)") + xlab(NULL) + labs(title = "Training set", caption = "Data in original scale")
+week.starts <- seq.POSIXt(from=floor_date(seis.h.start, 'week'), to=ceiling_date(seis.h.end, 'week'), by = "1 week")
+hours.from.start <- as.numeric(week.starts-seis.h.start)*24+168
+autoplot(seis.h.ts, type="l") +
+  scale_x_continuous(breaks=hours.from.start/168, labels=format(week.starts, "%d-%m-%y"), limits = c(11,15)-3/24) +
+  scale_y_continuous(limits = c(min(seis.h.ts[(10*168):(16*168)]), max(seis.h.ts[(10*168):(16*168)]))) +
+  ylab("mean hourly movement (nm)") + xlab(NULL) + labs(title = "Training set", subtitle = "4-weeks zoom", caption = "Data in original scale")
 dec.addi <- decompose(seasonality.aggregate(seis.h.ts))

### 2020-06-09 project.after.Rmd
more fixes on the draft and project code

@@
   autoplot(ts(seasonbase.before.lockdown.24.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
-    geom_line(aes(y=season.before.lockdown.24.df$figure, col="Multi daily")) +
+    geom_line(aes(y=season.before.lockdown.24$figure, col="Multi daily")) +
     scale_colour_manual("", values=c("purple", "yellow"), breaks=c("Direct daily", "Multi daily")),
   autoplot(ts(seasonbase.before.lockdown.168.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
-    geom_line(aes(y=season.before.lockdown.168.df$figure + rep(season.before.lockdown.24.df$figure, 7), col="Multi w+d")) +
+    geom_line(aes(y=season.before.lockdown.168.df$figure + rep(season.before.lockdown.24$figure, 7), col="Multi w+d")) +
     #geom_line(aes(y=rep(seasonbase.before.lockdown.24$figure, 7), col="Direct daily"), size=1.5) +

### 2020-06-09 project.R

@@ -343,13 +343,19 @@ show_overlap_and_diff <- function(base, compare, base.desc, compare.desc, xlab, 
 
 # We plot the dataset for a first visual inspection
 month.starts <- seq.POSIXt(from=floor_date(seis.h.start, 'month'), to=ceiling_date(seis.h.end, 'month'), by = "1 month")
-hours.from.start <- as.numeric(month.starts-seis.h.start)*24+seasonality.period
+hours.from.start <- as.numeric(month.starts-seis.h.start)*24+168
 autoplot(seis.h.ts, type="l") +
-  scale_x_continuous(breaks=hours.from.start/seasonality.period, labels=format(month.starts, "%m-%y")) +
-  ylab("mean hourly movement (nm)") + xlab(NULL) + ggtitle("Training set") + labs(caption = "Data in original scale")
+  scale_x_continuous(breaks=hours.from.start/168, labels=format(month.starts, "%m-%y")) +
+  ylab("mean hourly movement (nm)") + xlab(NULL) + labs(title = "Training set", caption = "Data in original scale")
+week.starts <- seq.POSIXt(from=floor_date(seis.h.start, 'week'), to=ceiling_date(seis.h.end, 'week'), by = "1 week")
+hours.from.start <- as.numeric(week.starts-seis.h.start)*24+168
+autoplot(seis.h.ts, type="l") +
+  scale_x_continuous(breaks=hours.from.start/168, labels=format(week.starts, "%d-%m-%y"), limits = c(11,15)-3/24) +
+  scale_y_continuous(limits = c(min(seis.h.ts[(10*168):(16*168)]), max(seis.h.ts[(10*168):(16*168)]))) +
+  ylab("mean hourly movement (nm)") + xlab(NULL) + labs(title = "Training set", subtitle = "4-weeks zoom", caption = "Data in original scale")
 # We see that the fluctuations doesn't seem to increase with the trend, so we use an additive decomposition
 dec.addi <- decompose(seasonality.aggregate(seis.h.ts))
-autoplot(dec.addi) + labs(title="Decomposition of additive time series", subtitle = paste0("Seasonality ", seasonality.period)) + xlab(seasonality.xlab)
+autoplot(dec.addi) + labs(title="Decomposition of additive time series", subtitle = paste0("Seasonality ", seasonality.period)) + xlab("Weeks")
 
 
 

### 2020-06-09 project.Rmd
plot periodgram on original scale

@@
 ggplot(pgram.data) +
-  geom_line(aes(freq, log(spec))) +
-  xlab("frequency") + ylab("log(spectrum)") +
+  geom_line(aes(freq, spec)) +
+  xlab("frequency") + ylab("spectrum") +
   labs(title = "Periodgram", caption = paste0("bandwidth = ", pgram$bandwidth, "    df = ", pgram$df)) ->
@@
   scale_x_continuous(sec.axis = sec_axis(~1/./3600, breaks=c(1, 3, 2*c(1,2,3,4,6,12), 24*7), labels = as.integer, name="hours")) +
-  scale_y_continuous(sec.axis = sec_axis(~exp(.), breaks=mean(pgram$spec) + sd(pgram$spec) * c(1,3), labels=c("1", "2", "3")[c(1,3)], name="spectrum/sd ratio")) +
-  geom_hline(yintercept=log(mean(pgram$spec) + sd(pgram$spec) * c(1:3)), linetype="dotted", color = ratio.color(mean(pgram$spec) + sd(pgram$spec) * c(1:3))) +
-  geom_point(data=data.frame(freq=top95freq, spec=top95pc), aes(freq, log(spec), color=ratio.color(spec)), shape="o", size=4, show.legend = F) +
-  geom_label_repel(data=data.frame(freq=top95freq, spec=top95pc)[1:shown_labels,], aes(freq, log(spec), label=round(1/freq/3600), color=ratio.color(spec)), segment.color="grey", show.legend=F) +
+  scale_y_continuous(sec.axis = sec_axis(~., breaks=mean(pgram$spec) + sd(pgram$spec) * c(1,3), labels=c("1", "2", "3")[c(1,3)], name="spectrum/sd ratio")) +
+  geom_hline(yintercept=(mean(pgram$spec) + sd(pgram$spec) * c(1:3)), linetype="dotted", color = ratio.color(mean(pgram$spec) + sd(pgram$spec) * c(1:3))) +
+  geom_point(data=data.frame(freq=top95freq, spec=top95pc), aes(freq, (spec), color=ratio.color(spec)), shape="o", size=4, show.legend = F) +
+  geom_label_repel(data=data.frame(freq=top95freq, spec=top95pc)[1:shown_labels,], aes(freq, (spec), label=round(1/freq/3600), color=ratio.color(spec)), segment.color="grey", show.legend=F) +
   scale_color_manual(values = c("darkgreen", "blue", "red")[start_ratio:3])

### 2020-06-09 project.after.Rmd
use original model for fitting lockdown data

@@
 cat("SARIMA(", paste0(chosen.model[1:3], collapse = ","), ")(", paste0(chosen.model[4:6], collapse = ","), ")[", frequency(final.training), "]", sep = "")
-if (!exists("final.model")) {
-  cat(" fitting...")
-  final.model <- Arima(final.training, chosen.model[1:3], chosen.model[4:6], lambda=chosen.lambda)
-  cat("DONE")
-}
-if (!exists("final.forecast")) {
+if (!exists("full.forecast")) {
   cat(" forecasting...")
-  final.forecast <- forecast(final.model, h=168*8, lambda=chosen.lambda)
+  full.forecast <- forecast(models[[fit.name]], h=168*26, lambda = chosen.lambda)
   cat("DONE")
@@
 cat("\n")
-plotting.model <- final.model
-plotting.forecast <- final.forecast
+plotting.model <- models[[fit.name]]
+plotting.forecast <- full.forecast
 print(plotting.model)
@@
 ))
-print(plot_forecast(plotting.forecast) + autolayer(final.testing, series="Original") + ylab("Mean hourly movement (nm)") + xlab("Weeks") + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2))
+print(plot_forecast(plotting.forecast) + autolayer(ts_c(`Before March '20`=seis.h.test, `From March '20`=final.testing), series="Original") + ylab("Mean hourly movement (nm)") + xlab("Weeks") + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2))
 plots <- show_overlap_and_diff(
-  final.training, plotting.forecast$fitted, xlab=seasonality.xlab,
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
   base.desc = "training set",
@@
 cat("Test set inside the CI\n")
-cat("80% CI \t ", print_dec(mean(plotting.forecast$lower[,1] < final.testing & final.testing < plotting.forecast$upper[,1], na.rm = T)*100), "%\n")
-cat("95% CI \t ", print_dec(mean(plotting.forecast$lower[,2] < final.testing & final.testing < plotting.forecast$upper[,2], na.rm = T)*100), "%\n")
+cat("CI   ", "Overall ", "Before  ", "After  \n")
+cat("80 % ",
+    print_dec(mean(plotting.forecast$lower[,1] < c(seis.h.test, final.testing) & c(seis.h.test, final.testing) < plotting.forecast$upper[,1], na.rm = T)*100),
+    "% ",
+    print_dec(mean(plotting.forecast$lower[1:length(seis.h.test),1] < seis.h.test & seis.h.test < plotting.forecast$upper[1:length(seis.h.test),1], na.rm = T)*100),
+    "% ",
+    print_dec(mean(plotting.forecast$lower[(length(seis.h.test)+1):dim(plotting.forecast$lower)[1],1] < final.testing & final.testing < plotting.forecast$upper[(length(seis.h.test)+1):dim(plotting.forecast$lower)[1],1], na.rm = T)*100),
+    "%\n")
+cat("95 % ",
+    print_dec(mean(plotting.forecast$lower[,2] < c(seis.h.test, final.testing) & c(seis.h.test, final.testing) < plotting.forecast$upper[,2], na.rm = T)*100),
+    "% ",
+    print_dec(mean(plotting.forecast$lower[1:length(seis.h.test),2] < seis.h.test & seis.h.test < plotting.forecast$upper[1:length(seis.h.test),2], na.rm = T)*100),
+    "% ",
+    print_dec(mean(plotting.forecast$lower[(length(seis.h.test)+1):dim(plotting.forecast$lower)[1],2] < final.testing & final.testing < plotting.forecast$upper[(length(seis.h.test)+1):dim(plotting.forecast$lower)[1],2], na.rm = T)*100),
+    "%\n")
 
 plots <- show_overlap_and_diff(
-  final.testing, plotting.forecast$mean, xlab=seasonality.xlab,
+  c(seis.h.test, final.testing), plotting.forecast$mean, xlab=seasonality.xlab,
   base.desc = "test set",
@@
 
-print(accuracy(plotting.forecast, final.testing))
+plots <- show_overlap_and_diff(
+  seis.h.test, window(plotting.forecast$mean, start=c(53,1), end=c(70,168)), xlab=seasonality.xlab,
+  base.desc = "test set before lockdown",
+  compare.desc = paste0("forecast with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="%",
+  diff.ylab.sec.accuracy = .01
+)
+print(plot_grid(plots[[1]],
+                plots[[2]],
+                nrow = 2
+                ))
+
+plots <- show_overlap_and_diff(
+  final.testing, window(plotting.forecast$mean, start=c(71,1), end=c(78,168)), xlab=seasonality.xlab,
+  base.desc = "test set during lockdown",
+  compare.desc = paste0("forecast with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="%",
+  diff.ylab.sec.accuracy = .01
+)
+print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
+                plots[[2]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
+                nrow = 2
+                ))
+
+print(accuracy(plotting.forecast, c(seis.h.test, final.testing)))
 # qqPlot(plotting.model$residuals, main = paste0("Q-Q plot for ", plotting.forecast$method), xlab = "residuals")

### 2020-06-09 project.Rmd
fixes and cleanups

@@
 seis.missing.ids <- which(is.na(seis.complete$min))
-
+#seis.complete$utc_second[seis.missing.ids]
 seis.present <- length(seis$utc_second)
@@
 text(x=blocks$start[which(blocks$interval > 3600)], y=blocks$interval[which(blocks$interval > 3600)], labels = paste0(blocks$start[which(blocks$interval > 3600)], " - ", blocks$end[which(blocks$interval > 3600)]), cex=0.9, pos=4)
-remove(seis.complete, seis.missing.ids)
+remove(seis.complete, seis.missing.ids, seis.present, seis.total, seis.missing, block_start, block_end, blocks)
 ```
@@
 seis.missing.ids <- which(is.na(seis.complete$mean_movement_s))
-
+# seis.complete$utc_hour[seis.missing.ids]
 seis.present <- length(seis.h$utc_hour)
@@
 text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=1)
-remove(seis.complete, seis.missing.ids)
-
+blocks$start
 blocks$interval
+remove(seis.complete, seis.missing.ids, seis.present, seis.total, seis.missing, block_start, block_end, blocks)
 
@@
 
+remove(pgram, pgram.data, filter_freq, top95pc, top95freq, top95seas.s, top95seas.h, ratio.color, periodgram)
 gc()
@@
 legend("topleft", legend=weekdays(seis.train.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
+remove(seis.h.24)
 
@@
   nrow = 2)
-
 ```
@@
 trend.s <- stats::filter(seasonality.aggregate(seis.h.train), side=2, filter=weights.s)
-stats::filter(seasonality.aggregate(seis.h.train), sides=2, filter = weights.s) %>% autoplot()
-stats::filter(seasonality.aggregate(seis.h.train) %>% diff(), sides=2, filter = weights.s) %>% autoplot()
-stats::filter(seasonality.aggregate(seis.h.train) %>% diff(lag=168), sides=2, filter = weights.s) %>% autoplot()
-stats::filter(seasonality.aggregate(seis.h.train) %>% diff() %>% diff(lag=168), sides=2, filter = weights.s) %>% autoplot()
-mean((seasonality.aggregate(seis.h.train) %>% diff(lag=168))[(1+(168*4)):(1+(168*6))])
 
@@
 
+```{r cleanup seasonalities}
+remove(matrix, weights, weights.s, ts)
+gc()
+```
+
 ### Multi trend extraction
@@
 
+remove(aggregated)
+
 ```

### 2020-06-09 project.after.Rmd
fixes and cleanups

--- 2020-06-09 project.after.Rmd
@@
 final.testing <- build_ts(seis.h, periods = c(seasonality.period), start = seis.train.start, end = seis.test.final.end, force_ms = T)
-final.testing <- window(final.testing, start = c(end(final.training)[1] + 1, 1))
+final.testing <- window(final.testing, start = c(end(seis.h.test)[1] + 1, 1))
 if (!is.null(chosen.lambda)) {
@@
 ))
-print(plot_forecast(plotting.forecast) + autolayer(final.testing, series="Original") + ylab("Mean hourly movement (nm)") + xlab("Weeks") + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2))
+lockdown.x <- as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1
+print(
+  plot_forecast(plotting.forecast)
+      + autolayer(ts_c(`Before March '20`=seis.h.test, `From March '20`=final.testing))
+      + ylab("Mean hourly movement (nm)")
+      + xlab("Weeks")
+      + geom_vline(aes(xintercept=lockdown.x), linetype=2)
+)
 plots <- show_overlap_and_diff(
-  final.training, plotting.forecast$fitted, xlab=seasonality.xlab,
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
   base.desc = "training set",
@@
 print(plot_grid(plots[[1]], plots[[2]], nrow = 2))
-
+seis.h.test.len <- length(seis.h.test)
+final.testing.len <- length(final.testing)
+full.forecast.len <- seis.h.test.len + final.testing.len
+full.forecast.before.ids <- 1:seis.h.test.len
+full.forecast.after.ids <- (seis.h.test.len+1):full.forecast.len
 cat("Test set inside the CI\n")
-cat("80% CI \t ", print_dec(mean(plotting.forecast$lower[,1] < final.testing & final.testing < plotting.forecast$upper[,1], na.rm = T)*100), "%\n")
-cat("95% CI \t ", print_dec(mean(plotting.forecast$lower[,2] < final.testing & final.testing < plotting.forecast$upper[,2], na.rm = T)*100), "%\n")
+cat("CI   ", "Overall ", "Before  ", "After  \n")
+cat("80 % ",
+    print_dec(mean(plotting.forecast$lower[,1] < c(seis.h.test, final.testing) & c(seis.h.test, final.testing) < plotting.forecast$upper[,1], na.rm = T)*100),
+    "% ",
+    print_dec(mean(plotting.forecast$lower[full.forecast.before.ids,1] < seis.h.test & seis.h.test < plotting.forecast$upper[full.forecast.before.ids,1], na.rm = T)*100),
+    "% ",
+    print_dec(mean(plotting.forecast$lower[full.forecast.after.ids,1] < final.testing & final.testing < plotting.forecast$upper[full.forecast.after.ids,1], na.rm = T)*100),
+    "%\n")
+cat("95 % ",
+    print_dec(mean(plotting.forecast$lower[,2] < c(seis.h.test, final.testing) & c(seis.h.test, final.testing) < plotting.forecast$upper[,2], na.rm = T)*100),
+    "% ",
+    print_dec(mean(plotting.forecast$lower[full.forecast.before.ids,2] < seis.h.test & seis.h.test < plotting.forecast$upper[full.forecast.before.ids,2], na.rm = T)*100),
+    "% ",
+    print_dec(mean(plotting.forecast$lower[full.forecast.after.ids,2] < final.testing & final.testing < plotting.forecast$upper[full.forecast.after.ids,2], na.rm = T)*100),
+    "%\n")
 
 plots <- show_overlap_and_diff(
-  final.testing, plotting.forecast$mean, xlab=seasonality.xlab,
+  c(seis.h.test, final.testing), plotting.forecast$mean, xlab=seasonality.xlab,
   base.desc = "test set",
@@
 )
-print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
-                plots[[2]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
-                nrow = 2
-                ))
+print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), plots[[2]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), nrow = 2))
 
-print(accuracy(plotting.forecast, final.testing))
+plots <- show_overlap_and_diff(
+  seis.h.test, window(plotting.forecast$mean, start=c(53,1), end=c(70,168)), xlab=seasonality.xlab,
+  base.desc = "test set before lockdown",
+  compare.desc = paste0("forecast with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="%",
+  diff.ylab.sec.accuracy = .01
+)
+print(plot_grid(plots[[1]], plots[[2]], nrow = 2))
+
+plots <- show_overlap_and_diff(
+  final.testing, window(plotting.forecast$mean, start=c(71,1), end=c(78,168)), xlab=seasonality.xlab,
+  base.desc = "test set during lockdown",
+  compare.desc = paste0("forecast with ", plotting.forecast$method),
+  overlap.ylab = "Mean hourly movement (nm)",
+  diff.title = "Difference between ts",
+  diff.ylab = "nm", diff.ylab.sec="%",
+  diff.ylab.sec.accuracy = .01
+)
+print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), plots[[2]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), nrow = 2))
+
+print(accuracy(plotting.forecast, c(seis.h.test, final.testing)))
 # qqPlot(plotting.model$residuals, main = paste0("Q-Q plot for ", plotting.forecast$method), xlab = "residuals")

### 2020-06-09 project.R

@@ -789,39 +789,49 @@ chosen.model <- c(2,0,0,0,1,1)
 # We make our final train/test split date on 2nd of March
 seis.test.final.start <- seis.test.end + seconds(1)
 seis.test.final.end <- seis.test.final.start + days(7*8) - seconds(1)
-final.training <- build_ts(seis.h, periods = c(seasonality.period), start = seis.train.start, end = seis.test.end, force_ms = T)
 final.testing <- build_ts(seis.h, periods = c(seasonality.period), start = seis.train.start, end = seis.test.final.end, force_ms = T)
 final.testing <- window(final.testing, start = c(end(final.training)[1] + 1, 1))
-if (!is.null(chosen.lambda)) {
-  cat(chosen.lambda, BoxCox.lambda(final.training), chosen.lambda == BoxCox.lambda(final.training), "\n")
-}
-# Then, we fit the chosen model on the new training set
+# Then, we forecast the chosen model until the end of April
 fit.name <- paste0("fit", paste0(chosen.model, collapse = ""))
-forecast.name <- paste0("forecast", paste0(chosen.model, collapse = ""))
-cat("SARIMA(", paste0(chosen.model[1:3], collapse = ","), ")(", paste0(chosen.model[4:6], collapse = ","), ")[", frequency(final.training), "]", sep = "")
-if (!exists("final.model")) {
-  cat(" fitting...")
-  final.model <- Arima(final.training, chosen.model[1:3], chosen.model[4:6], lambda=chosen.lambda)
-  cat("DONE")
-}
-if (!exists("final.forecast")) {
+plotting.model <- models[[fit.name]]
+cat("SARIMA(", paste0(chosen.model[1:3], collapse = ","), ")(", paste0(chosen.model[4:6], collapse = ","), ")[", seasonality.period, "]", sep = "")
+if (!exists("full.forecast")) {
   cat(" forecasting...")
-  final.forecast <- forecast(final.model, h=168*8, lambda=chosen.lambda)
+  full.forecast <- forecast(plotting.model, h=168*26, lambda = chosen.lambda)
   cat("DONE")
 }
 cat("\n")
+plotting.forecast <- full.forecast
 # We can directly assess how the forecasts behave
-plotting.model <- final.model
-plotting.forecast <- final.forecast
 # plot forecasts
[…]

### 2020-06-09 draft.tex
improvements on ARIMA report and some plots

@@
 To deal with them when R methods need complete timeseries, we define different completion strategies, depending on the seasonality, using the R function \texttt{na.aggregate}.
-
+%
 \begin{figure}[t]
@@
 \end{figure}
-
+%
 \subsubsection{Seasonality assumptions}
@@
 Still, we recognize that the already performed transformations impose an human-based time alteration to the data itself (Section \ref{sec:cleanup-transform}).
-
+%
 \begin{figure}[h]
@@
 \end{figure}
-
+%
 The method we are going to apply is based on the Fourier Transform (FT), and generates a \textit{periodgram} highlighting the most interesting frequencies that might correspond to a seasonality component.
@@
 We can analyse the resulting periodgrams for Uccle and Membach stations in \figurename{s} \ref{fig:periodgram:uccs} and \ref{fig:periodgram:mems}.
-
+%
 \begin{figure}[h]
@@
 \end{figure}
-
+%
 Both highlight multiple 24-hour periodicities.
@@
 Since they do not produce particularly different seasonalities, we choose to keep the results obtained by applying the MA smoothing filter, for its significance in terms of seasonally adjusted data.
-
+%
 \begin{figure}[h]
@@
 	\end{subfigure}
-	\begin{center}
-		\caption{Timeseries decomposition}{\small Only the second method results are shown for clarity}
+	\caption{Timeseries decomposition}{\centering\small Only the second method results are shown for clarity. Week 10 includes January $1^{st}$.\\}
 	\label{fig:decomposed}
-	\end{center}
 \end{figure}
-
+%
 Since the periodgrams suggested the presence of two significative periodicities, we also tested the extraction of both corresponding seasonalities:
@@
 \end{enumerate}
-
+%
 \begin{figure}[h]
@@
 	\end{subfigure}
-	\begin{center}
-		\caption{Seasonality figures}{\small\textit{Direct} refers to single seasonality extraction\\\textit{Multi} refers to both seasonalities extracted at the same time}
+	\caption{Seasonality figures}{\centering\small\textit{Direct} refers to single seasonality extraction\\\centering\textit{Multi} refers to both seasonalities extracted at the same time\\}
 	\label{fig:direct-multi-seasonality}
-	\end{center}
 \end{figure}
-
+%
 In \figurename~\ref{fig:direct-multi-seasonality} we see both approaches applied to both stations.
@@
 
+From tha same decomposition results, we notice that data around January $1^{st}$ show a significant drop in values.
+That is highlighted by both trends, where the overall minimum falls in that period, and by the residuals for Uccle data, showing a clear overestimation.
+
 \section{Methodology}\label{sec:methodology}
@@
 
-For each station, we fit a model that generalises our training dataset in such a way that forecasting becomes effective, also assessing errors analytically.
+For each station, we fit a model that generalises our training dataset in such a way that forecasting until March $1^{st}$ becomes effective, also assessing errors analytically.
 Obtaining such a reliable model will be the first step to verify any significant difference forecasting over the lockdown period.
 
-The model we are going to estimate is the ARIMA model, including the seasonality components.
+The model we fit is the ARIMA model, including the seasonality components.
 To identify the order of the different components described by ARIMA, we will analyse the autocorrelation and partial autocorrelation of our datasets.
@@
 	\item non-seasonal MA process of order $q$ is described by an exponentially or sigmoidally decaying PACF, and by a cut-off of significant ACF values at lag $q$
-	\item seasonal AR process of order $P$ is described by an exponentially or sigmoidally decay at seasonal lags on ACF values, and by a cut-off of significant PACF values at each seasonal lag until the $P^{th}$
-	\item seasonal MA process of order $Q$ is described by an exponentially or sigmoidally decay at seasonal lags on PACF values, and by a cut-off of significant ACF values at each seasonal lag until the $Q^{th}$
+	\item seasonal AR process of order $P$ is described by an exponential or sigmoidal decay at seasonal lags on ACF values, and by a cut-off of significant PACF values at each seasonal lag until the $P^{th}$
+	\item seasonal MA process of order $Q$ is described by an exponential or sigmoidal decay at seasonal lags on PACF values, and by a cut-off of significant ACF values at each seasonal lag until the $Q^{th}$
 	\item given how significance is assessed for the ACF/PACF spikes, one over 20 spikes (5\%) could be \textit{coincidentally significant} or \textit{coincidentally \textbf{non}-significant}
 \end{itemize}
+
+The same model will be then used to forecast other 8 weeks from March $2^{nd}$, so that we can assess the forecast power over the lockdown period.
+Given the fact that both our series are quite stable over time (\figurename{s} \ref{fig:decomposed:uccs}, \ref{fig:mems:box-cox:data}), we can expect that the predicted variability for further forecasts will not significantly increase.
+
+Also, we will decompose the seasonalities from data before and after the lockdown, in order to assess any significant change regarding that aspect.
+
+Per our hypotesis, we expect the Uccle station's data to show significant changes after the lockdown, for its position near Brussels, while we expect Membach's forecasts not to show any significant change, for its position is further from busy cities.
+
+We will not directly compare the two models, since data from the two stations have different orders of magnitude.
 
@@
 \end{itemize}
-We also employ the \textit{auto.arima} stepwise methodology\footnote{The \textit{auto.arima} stepwise methodology fits different ARIMA models within a range of orders.
+We also employ the \textit{auto.arima} stepwise methodology\footnote{\label{footnote:auto.arima}The \textit{auto.arima} stepwise methodology fits different ARIMA models within a range of orders.
 Based on each result's information criterion (Akaike), it changes each time one of the \textit{p}, \textit{d}, \textit{q}, \textit{P}, \textit{D}, or \textit{Q} values, depending on the configured limits.
@@
 Following both the analytical ways and the \textit{auto.arima} hints, and refining the models through further inspection of ACF/PACF plots of the residuals, we end up with three characteristic models.
-We selected them through the Akaike Information Criterion (AIC)\footnote{Actually, we could have used the corrected AIC as it is usually more precise, but the values of AIC and AICc are equal to the first decimal for all our models ($\lesssim 10^{-7}\%$ difference).}, being the only three out of the fitted models with $\mathrm{AIC} < 120.000$: $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$, $\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$, and $\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$.
+We selected them through the Akaike Information Criterion (AIC)\footnote{Actually, we could have used the corrected AIC (AICc) as it is usually more precise, but the values of AIC and AICc are equal to the first decimal for all our models ($\lesssim 10^{-7}\%$ difference).}, being the only three out of the fitted models with $\mathrm{AIC} < 120.000$: $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$, $\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$, and $\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$.
 
@@
 Nevertheless, MAPE is not a single valid criterion for selection, since a low error predicting the training set itself does not necessarily mean a good forecast power.
-
+%
 \begin{figure}[h]
@@
 	\end{subfigure}
-	\begin{center}
 	\caption{Plots for $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$}
 	\label{fig:arima}
-	\end{center}
 \end{figure}
-
+%
 The ACF and PACF of the residuals for all three models are very similar (\figurename~\ref{fig:arima-26}), even if the models themselves have great differences in meaning and in number of parameters: they highlight a clear seasonal ARMA process with periodicity 24.
 This shows how our previous conclusion that the 168 hours seasonality could include the 24 hours one is flawed.
-Unfortunately, the parameters estimator we are using does not permit to estimate multiple seasonalities\footnote{\textit{Arima} method from the \textit{forecast} R package, only specifying the seasonal and non-seasonal orders.}.
+Unfortunately, the R function we are using does not permit to estimate multiple seasonalities\footnote{\textit{Arima} method from the \textit{forecast} R package, only specifying the seasonal and non-seasonal orders.}.
 
@@
 This is expected, since normally the model is more fitted to the training data than to unknown future data.
-Also, it must be noted that a January $1^{st}$ period lies in our test dataset, and without being able to include a yearly seasonality, that leads to greater errors just from those couple of weeks (\figurename~\ref{fig:arima-forecast:200011}).
 
+Also, it must be noted that a January $1^{st}$ period lies in our test dataset, that is already noticed as low-valued period (\figurename~\ref{fig:decomposed}).
+We can notice a clear overestimation during those couple of weeks (around week 62 on \figurename~\ref{fig:arima-forecast}).
+%
 \begin{figure}[h]
@@
 	\end{subfigure}
-	\begin{center}
 	\caption{ARIMA forecasts}{\footnotesize$\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$ behaves like $\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$ with faster increasing variablity}
 	\label{fig:arima-forecast}
-	\end{center}
 \end{figure}
-
+%
 We can see in \figurename~\ref{fig:arima-forecast} the forecast behaviour for our models.
@@
 
-Also, ARIMA parameters roots have values within the unit circle for the first two models, while for $\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$ two of the roots fall outside the unit circle, with values of $\approx1.49$ and $\approx1.80$.
+%Also, ARIMA parameters roots have values within the unit circle for the first two models, while for $\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$ two of the roots fall outside the unit circle, with values of $\approx1.49$ and $\approx-1.80$, making the models invalid.
 
@@
 This will probably cause some approximation errors that could rig the final results.
-
+%
 \begin{figure}[h]
@@
 	\end{subfigure}
-	\begin{center}
 	\caption{Results of Box-Cox transformation}
 	\label{fig:mems:box-cox}
-	\end{center}
 \end{figure}
+%
+The transformed data still shows no constant mean nor variance, even if they show less variability around the peaks with respect to the original data, confirmed by the Q-Q plot (\figurename~\ref{fig:mems:box-cox:qq}).
+Thus, we consider the usage of the \textit{diff} operator, at lag 1 and 168.
+Interestingly, the analysis of ACF and PACF plots hints to similar models to the untransformed Uccle data:
+$\mathrm{ARIMA}(0,1,0)(0,1,1)_{168}$ and $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$. All ACF and PACF plots for Membach model estimation are not shown for brevity.
 
-The transformed data still shows no constant mean, even if it has a way better shape than the original data, confirmed by the Q-Q plot (\figurename~\ref{fig:mems:box-cox:qq}).
-Thus, we consider the usage of the \textit{diff} operator, at lag 1 and 168.
-Interestingly, the analysis of ACF and PACF plots hints to similar models to the Uccle station:
-$\mathrm{ARIMA}(0,1,0)(0,1,1)_{168}$ and $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$.
+In this case, the employment of \textit{auto.arima} stepwise methodology$^{\ref{footnote:auto.arima}}$ highlights the first model improved with one non-seasonal MA parameter more.
+This is confirmed by the ACF/PACF of its residuals.
+The analysis of ACF/PACF plots of the residuals of the second model do not lead to any improvement.
+The ACF and PACF plots of both models are similar to the ones from the chosen model for the Uccle station (\figurename~\ref{fig:arima-24}), and we can notice the same 24 hours remaining periodicity.
 
-In this case, the employment of \textit{auto.arima} stepwise methodology\footnote{The \textit{auto.arima} stepwise methodology fits different ARIMA models within a range of orders.
-Based on each result's information criterion (Akaike), it changes each time one of the \textit{p}, \textit{d}, \textit{q}, \textit{P}, \textit{D}, or \textit{Q} values, depending on the configured limits.
-These changes minimizes the information criteria of the final model.} does not lead to any useful model, since it obtains $\mathrm{AIC}=\inf$ for the two manually-chosen models, that have an AIC value around -84.500, while the best model found by that methodology has an AIC around -81.000.
-This can be caused by approximation errors\footnote{The \textit{auto.arima} methodology was only used with initial approximation, due to high computing time.}, as we supposed earlier.
-
-The analysis of ACF/PACF plots of the residuals from both models do not lead to any new model but, as with the Uccle station, we can notice a clear 24 hours periodicity.
-
-Both models have valid p-value results for the Box-Pierce test, similar AICc values, and similar MAPE on the training set.
+The Box-Pierce test statistics for both models accept the independence hypothesis between residuals.
+The two models have similar AICc values, and similar MAPE on the training set ($\approx6\%$).
 Thus, we can assess the forecasts for both models in \figurename~\ref{fig:mems:arima}.
-
+%
 \begin{figure}[h]
@@
 	\end{subfigure}
-	\begin{center}
 	\caption{ARIMA forecasts}
 	\label{fig:mems:arima}
-	\end{center}
 \end{figure}
-
-The first model catches a rapidly-increasing trend, while the latter predicts more stable values moving the most part of the increasing trend into increasing variability.
+%
+The first model catches a rapidly-increasing trend, while the latter predicts more stable values moving the most part of the increasing trend into slightly increasing variability.
 Visually, both could be acceptable, even if the latter seems to better behave on \textit{normal} days losing precision on high-valued ones, while the first tries to better accomodate the high-valued days losing precision on the \textit{normal} ones.
@@
 Also considering the MAPE on the test set, the latter model fits the data better with a $16\%$ error, instead of the $18\%$ for the first model.
-
-Parameters roots for both models have values within the unit circle.
+%
+%Parameters roots for both models have values within the unit circle.
 
@@
 
+TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
+
+As a side note, we followed the same procedure without applying the Box-Cox transformation, but the best model that could be chosen had more parameters and was less precise than the chosen model on the transformed data.
+
+TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO 
 %\subsubsection{MSTL}
@@
 This interval estimation lies on the assumption that values for each single hour in the hourly/weekly seasonality is Normally distributed, a slightly different assumption that is not necessarily refuted by the already seen Q-Q plots.
-
+%
 \begin{figure}[h]
@@
 	\end{subfigure}
-	\begin{center}
-		\caption{Seasonalities computed before and after the lockdown}{\small Seasonalities extracted together, confidence bands $\pm\mathrm{sd}$}
+	\caption{Seasonalities computed before and after the lockdown}{\centering\small Seasonalities extracted together, confidence bands $\pm\mathrm{sd}$\\}
 	\label{fig:season.lockdown}
-	\end{center}
 \end{figure}
-
+%
 As we can see in \figurename~\ref{fig:season.lockdown}, the seasonalities' shapes seem different, in particular comparing the weekly seasonality.
@@
 	\end{subfigure}
-	\begin{center}
 	\caption{Seasonal plots for Membach station}
 	\label{fig:season.lockdown:ms:mems}
-	\end{center}
 \end{figure}
-
+%
 The differences in shape are more highlighted checking the seasonal plots \cite{hyndman2018forecasting} in \figurename~\ref{fig:season.lockdown:ms:mems}.
@@
 \subsubsection{Forecast divergence}
+%
 \begin{figure}[h]
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-7"}
+		\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-9"}
 		\caption{Uccle station}
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-7"}
+		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-9"}
 		\caption{Membach station}
@@
 	\end{subfigure}
-	\begin{center}
-		\caption{Forecasts from March $2^{nd}$.}{\small In red the actual values, in black the forecasts. March $14^{th}$ at midnight is highlighted.}
+	\caption{Forecasts from March $2^{nd}$.}{\centering\small In red the actual values, in black the forecasts. March $14^{th}$ at midnight is highlighted.\\}
 	\label{fig:forecast.lockdown}
-	\end{center}
 \end{figure}
-
+%
 By fitting the model\footnote{Data is different, because we are fitting the model on the original train+test dataset, but if we follow similar steps for both stations we can see that the final model we shall choose is the same.

### 2020-06-09 project.Rmd
improvements on ARIMA report and some plots

@@
   autoplot(aggregated-season.24$seasonal-season.168$seasonal-season.168$trend) + labs(title="Residuals", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(t=-6, b=-6, unit="pt")),
-  autoplot(season.168$figure + rep(season.24$figure, 7)) + labs(title="Weekly seasonality (sum)", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(r=10, b=-6, unit="pt")),
+  autoplot(season.168$figure + rep(season.24$figure, 7)) + labs(title="Composed seasonality", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(r=10, b=-6, unit="pt")),
   nrow = 3,

### 2020-06-09 project.mems.Rmd
improvements on ARIMA report and some plots

@@
   ))
-  print(plot_forecast(plotting.forecast) + autolayer(seis.h.test, series="Original") + ylab("Mean hourly movement (nm)"))
+  print(plot_forecast(plotting.forecast) + autolayer(seis.h.test, series="Original") + ylab("Mean hourly movement (nm)") + xlab(seasonality.xlab))
   plots <- show_overlap_and_diff(

### 2020-06-09 project.R

@@ -748,7 +748,7 @@ for (model in to_test) {
     nrow = 2
   ))
   # plot forecasts
-  print(plot_forecast(plotting.forecast) + autolayer(seis.h.test, series="Original") + ylab("Mean hourly movement (nm)"))
+  print(plot_forecast(plotting.forecast) + autolayer(seis.h.test, series="Original") + ylab("Mean hourly movement (nm)") + xlab(seasonality.xlab))
   # print stats about data falling inside 80% and 90% CIs
   cat("Test set inside the CI\n")
   cat("80% CI \t ", print_dec(mean(plotting.forecast$lower[,1] < seis.h.test & seis.h.test < plotting.forecast$upper[,1], na.rm = T)*100), "%\n")

### 2020-06-09 project.Rmd
cleanup some labels

@@
 ```{r test_train split ts, include=FALSE}
+data.lab <- "Hourly mean (nm)"
+
 seis.h.start <- as_datetime('2018-10-29T00:00:00')
@@
   scale_x_continuous(breaks=hours.from.start/168, labels=format(month.starts, "%m-%y")) +
-  ylab("mean hourly movement (nm)") + xlab(NULL) + labs(title = "Training set", caption = "Data in original scale")
+  ylab(data.lab) + xlab(NULL) + labs(title = "Training set", caption = "Data in original scale")
 week.starts <- seq.POSIXt(from=floor_date(seis.h.start, 'week'), to=ceiling_date(seis.h.end, 'week'), by = "1 week")
@@
   scale_y_continuous(limits = c(min(seis.h.ts[(10*168):(16*168)]), max(seis.h.ts[(10*168):(16*168)]))) +
-  ylab("mean hourly movement (nm)") + xlab(NULL) + labs(title = "Training set", subtitle = "4-weeks zoom", caption = "Data in original scale")
+  ylab(data.lab) + xlab(NULL) + labs(title = "Training set", subtitle = "4-weeks zoom", caption = "Data in original scale")
 dec.addi <- decompose(seasonality.aggregate(seis.h.ts))
@@
   scale_x_continuous(breaks=hours.from.train.start/168, labels=format(month.starts, "%m-%y")) +
-  ylab("mean hourly movement (nm)") + xlab(NULL) + ggtitle("Training set") + labs(caption = "Data in original scale")
-gghistogram(seis.h.train, add.normal = T) + xlab("mean hourly movement (nm)") + labs(caption = "Data in original scale")
+  ylab(data.lab) + xlab(NULL) + ggtitle("Training set") + labs(caption = "Data in original scale")
+gghistogram(seis.h.train, add.normal = T) + xlab(data.lab) + labs(caption = "Data in original scale")
 gg_qqplot(seis.h.train) + 
@@
   scale_x_continuous(breaks=hours.from.train.start/168, labels=format(month.starts, "%m-%y")) +
-  ylab("mean hourly movement (nm)") + xlab(NULL) + ggtitle("Training set") + labs(caption = "Data in log-scale")
-gghistogram(log(seis.h.train), add.normal = T) + xlab("mean hourly movement (nm)") + labs(caption = "Data in log-scale")
+  ylab(data.lab) + xlab(NULL) + ggtitle("Training set") + labs(caption = "Data in log-scale")
+gghistogram(log(seis.h.train), add.normal = T) + xlab(data.lab) + labs(caption = "Data in log-scale")
 gg_qqplot(log(seis.h.train)) + 
@@
   scale_x_continuous(breaks=hours.from.train.start/168, labels=format(month.starts, "%m-%y")) +
-  ylab("mean hourly movement (nm)") + xlab(NULL) + ggtitle("Training set") + labs(caption = paste0("Data Box-Cox transformed with lambda = ", train.bc.lambda))
-gghistogram(train.bc, add.normal = T) + xlab("mean hourly movement (nm)") + labs(caption = paste0("Data Box-Cox transformed with lambda = ", train.bc.lambda))
+  ylab(data.lab) + xlab(NULL) + ggtitle("Training set") + labs(caption = paste0("Data Box-Cox transformed with lambda = ", train.bc.lambda))
+gghistogram(train.bc, add.normal = T) + xlab(data.lab) + labs(caption = paste0("Data Box-Cox transformed with lambda = ", train.bc.lambda))
 gg_qqplot(train.bc) + 
@@
       geom_line(aes(y=base, col="red"), show.legend = F) +
-      labs(title=paste0(base.desc, " vs ", compare.desc))
+      labs(title=NULL, subtitle=paste0(base.desc, " vs ", compare.desc))
   diff <- compare-base
@@
   compare.desc = "reconstructed serie (manual)",
-  overlap.ylab = "Mean hourly movement (nm)",
+  overlap.ylab = data.lab,
   diff.title = "Difference between ts",
@@
   compare.desc = "reconstructed serie (decompose additive)",
-  overlap.ylab = "Mean hourly movement (nm)",
+  overlap.ylab = data.lab,
   diff.title = "Difference between ts",
@@
   compare.desc = "168+24 from ds",
-  overlap.ylab = "Mean hourly movement (nm)",
+  overlap.ylab = data.lab,
   diff.title = "Difference between 168",

### 2020-06-09 project.after.Rmd
cleanup some labels

--- 2020-06-09 project.after.Rmd
@@
 ))
-print(plot_forecast(plotting.forecast) + autolayer(final.testing, series="Original") + ylab("Mean hourly movement (nm)") + xlab("Weeks") + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2))
+lockdown.x <- as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1
+print(
+  plot_forecast(plotting.forecast)
+      + autolayer(ts_c(`Before March '20`=seis.h.test, `From March '20`=final.testing))
+      + ylab(data.lab)
+      + xlab("Weeks")
+      + geom_vline(aes(xintercept=lockdown.x), linetype=2)
+)
 plots <- show_overlap_and_diff(
-  final.training, plotting.forecast$fitted, xlab=seasonality.xlab,
+  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
   base.desc = "training set",
   compare.desc = paste0("refit with ", plotting.forecast$method),
-  overlap.ylab = "Mean hourly movement (nm)",
+  overlap.ylab = data.lab,
   diff.title = "Difference between ts",
@@
 print(plot_grid(plots[[1]], plots[[2]], nrow = 2))
-
+seis.h.test.len <- length(seis.h.test)
+final.testing.len <- length(final.testing)
+full.forecast.len <- seis.h.test.len + final.testing.len
+full.forecast.before.ids <- 1:seis.h.test.len
+full.forecast.after.ids <- (seis.h.test.len+1):full.forecast.len
 cat("Test set inside the CI\n")
-cat("80% CI \t ", print_dec(mean(plotting.forecast$lower[,1] < final.testing & final.testing < plotting.forecast$upper[,1], na.rm = T)*100), "%\n")
-cat("95% CI \t ", print_dec(mean(plotting.forecast$lower[,2] < final.testing & final.testing < plotting.forecast$upper[,2], na.rm = T)*100), "%\n")
+cat("CI   ", "Overall ", "Before  ", "After  \n")
+cat("80 % ",
+    print_dec(mean(plotting.forecast$lower[,1] < c(seis.h.test, final.testing) & c(seis.h.test, final.testing) < plotting.forecast$upper[,1], na.rm = T)*100),
+    "% ",
+    print_dec(mean(plotting.forecast$lower[full.forecast.before.ids,1] < seis.h.test & seis.h.test < plotting.forecast$upper[full.forecast.before.ids,1], na.rm = T)*100),
+    "% ",
+    print_dec(mean(plotting.forecast$lower[full.forecast.after.ids,1] < final.testing & final.testing < plotting.forecast$upper[full.forecast.after.ids,1], na.rm = T)*100),
+    "%\n")
+cat("95 % ",
+    print_dec(mean(plotting.forecast$lower[,2] < c(seis.h.test, final.testing) & c(seis.h.test, final.testing) < plotting.forecast$upper[,2], na.rm = T)*100),
+    "% ",
+    print_dec(mean(plotting.forecast$lower[full.forecast.before.ids,2] < seis.h.test & seis.h.test < plotting.forecast$upper[full.forecast.before.ids,2], na.rm = T)*100),
+    "% ",
+    print_dec(mean(plotting.forecast$lower[full.forecast.after.ids,2] < final.testing & final.testing < plotting.forecast$upper[full.forecast.after.ids,2], na.rm = T)*100),
+    "%\n")
 
 plots <- show_overlap_and_diff(
-  final.testing, plotting.forecast$mean, xlab=seasonality.xlab,
+  c(seis.h.test, final.testing), plotting.forecast$mean, xlab=seasonality.xlab,
   base.desc = "test set",
   compare.desc = paste0("forecast with ", plotting.forecast$method),
-  overlap.ylab = "Mean hourly movement (nm)",
-  diff.title = "Difference between ts",
+  overlap.ylab = data.lab,
+  diff.title = NULL,
   diff.ylab = "nm", diff.ylab.sec="%",
@@
 )
-print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
-                plots[[2]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
-                nrow = 2
-                ))
+print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), plots[[2]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), nrow = 2))
 
-print(accuracy(plotting.forecast, final.testing))
+plots <- show_overlap_and_diff(
+  seis.h.test, window(plotting.forecast$mean, start=c(53,1), end=c(70,168)), xlab=seasonality.xlab,
+  base.desc = "test set before lockdown",
+  compare.desc = paste0("forecast with ", plotting.forecast$method),
+  overlap.ylab = data.lab,
+  diff.title = NULL,
+  diff.ylab = "nm", diff.ylab.sec="%",
+  diff.ylab.sec.accuracy = .01
+)
+print(plot_grid(plots[[1]], plots[[2]], nrow = 2))
+
+plots <- show_overlap_and_diff(
+  final.testing, window(plotting.forecast$mean, start=c(71,1), end=c(78,168)), xlab=seasonality.xlab,
+  base.desc = "test set during lockdown",
+  compare.desc = paste0("forecast with ", plotting.forecast$method),
+  overlap.ylab = data.lab,
+  diff.title = NULL,
+  diff.ylab = "nm", diff.ylab.sec="%",
+  diff.ylab.sec.accuracy = .01
+)
+print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), plots[[2]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), nrow = 2))
+
+print(accuracy(plotting.forecast, c(seis.h.test, final.testing)))
 # qqPlot(plotting.model$residuals, main = paste0("Q-Q plot for ", plotting.forecast$method), xlab = "residuals")
@@
   geom_line(aes(Hour, figure, col=Weekday)) +
-  ylab("Mean hourly movement (nm)") +
+  ylab(data.lab) +
   labs(title="Seasonal plot: before lockdown") +
@@
   geom_line(aes(Hour, figure, col=Weekday)) +
-  ylab("Mean hourly movement (nm)") +
+  ylab(data.lab) +
   labs(title="Seasonal plot: after lockdown") +

### 2020-06-09 project.mems.Rmd
cleanup some labels

--- 2020-06-06 project.mems.Rmd
@@
   ))
-  print(plot_forecast(plotting.forecast) + autolayer(seis.h.test, series="Original") + ylab("Mean hourly movement (nm)"))
+  print(plot_forecast(plotting.forecast) + autolayer(seis.h.test, series="Original") + ylab(data.lab) + xlab(seasonality.xlab))
   plots <- show_overlap_and_diff(
@@
     compare.desc = paste0("refit with ", plotting.forecast$method),
-    overlap.ylab = "Mean hourly movement (nm)",
-    diff.title = "Difference between ts",
+    overlap.ylab = data.lab,
+    diff.title = NULL,
     diff.ylab = "nm", diff.ylab.sec="%",
@@
     compare.desc = paste0("forecast with ", plotting.forecast$method),
-    overlap.ylab = "Mean hourly movement (nm)",
-    diff.title = "Difference between ts",
+    overlap.ylab = data.lab,
+    diff.title = NULL,
     diff.ylab = "nm", diff.ylab.sec="%",
@@
   compare.desc = paste0("refit with ", plotting.forecast$method),
-  overlap.ylab = "Mean hourly movement (nm)",
-  diff.title = "Difference between ts",
+  overlap.ylab = data.lab,
+  diff.title = NULL,
   diff.ylab = "nm", diff.ylab.sec="% over ts span",
@@
   compare.desc = paste0("refit with ", plotting.forecast$method),
-  overlap.ylab = "Mean hourly movement (nm)",
+  overlap.ylab = data.lab,
   diff.title = "Difference between ts",

### 2020-06-09 project.R

@@ -280,6 +280,7 @@ gc()
 
 
 
+data.lab <- "Hourly mean (nm)"
 # We now proceed selecting the dataset timespan for train and test subsets.
 # We will use the first 52 weeks starting from 2018-10-29 as training, and the following 18 weeks until 2020-03-02 as test
 seis.h.start <- as_datetime('2018-10-29T00:00:00')
@@ -316,7 +317,7 @@ show_overlap_and_diff <- function(base, compare, base.desc, compare.desc, xlab, 
   overlap_plot <-
     autoplot(compare, xlab=xlab, ylab=overlap.ylab) +
     geom_line(aes(y=base, col="red"), show.legend = F) +
-    labs(title=paste0(base.desc, " vs ", compare.desc))
+    labs(title=NULL, subtitle=paste0(base.desc, " vs ", compare.desc))
   # The we compare the difference
   diff <- compare-base
   # Here we compute and print some error statistics
@@ -346,13 +347,13 @@ month.starts <- seq.POSIXt(from=floor_date(seis.h.start, 'month'), to=ceiling_da
 hours.from.start <- as.numeric(month.starts-seis.h.start)*24+168
 autoplot(seis.h.ts, type="l") +
   scale_x_continuous(breaks=hours.from.start/168, labels=format(month.starts, "%m-%y")) +
-  ylab("mean hourly movement (nm)") + xlab(NULL) + labs(title = "Training set", caption = "Data in original scale")
+  ylab(data.lab) + xlab(NULL) + labs(title = "Training set", caption = "Data in original scale")
 week.starts <- seq.POSIXt(from=floor_date(seis.h.start, 'week'), to=ceiling_date(seis.h.end, 'week'), by = "1 week")
 hours.from.start <- as.numeric(week.starts-seis.h.start)*24+168
 autoplot(seis.h.ts, type="l") +
   scale_x_continuous(breaks=hours.from.start/168, labels=format(week.starts, "%d-%m-%y"), limits = c(11,15)-3/24) +
   scale_y_continuous(limits = c(min(seis.h.ts[(10*168):(16*168)]), max(seis.h.ts[(10*168):(16*168)]))) +
-  ylab("mean hourly movement (nm)") + xlab(NULL) + labs(title = "Training set", subtitle = "4-weeks zoom", caption = "Data in original scale")
[…]

### 2020-06-10 draft.tex
cleanup some labels
various tweaks to the report

@@
 	\item In days with missing values there are placeholder values instead a missing row, that would make the following rows misinterpreted for measurements on the wrong second.
-The placeholder values are one of these two min-max pairs: $(0,0)\, (-1,1)$
+The placeholder values are one of these two min-max pairs: $(0,0),\, (-1,1)$.
 \end{enumerate}
@@
 Data from each station have different domains, as distinct equipment can register a specific amplitude of the seismic movements.
-In fact, we found some values\footnote{Values out of domain for both stations are less than 10 out of $\approx50$ million.} outside the domains, and we considered them as missing values.
+In fact, we found some values\footnote{Values out of domain for each station are less than 10 out of $\approx50$ million.} outside the domains, and we consider them as missing values.
 
-In particular, we find that out of $\approx50$ million seconds, from 27-10-2018 to 28-05-2020, only 0.67\% (Uccle) and 0.37\% (Membach) have missing values.
+In particular, we find that out of $\approx50$ million seconds, from 27-10-2018 to 28-05-2020, only 0.67\% (Uccle) and 0.37\% (Membach) are missing values.
 We will see later how they are distributed for each station.
@@
 To deal with them when R methods need complete timeseries, we define different completion strategies, depending on the seasonality, using the R function \texttt{na.aggregate}.
-%
+
 \begin{figure}[t]
@@
 Still, we recognize that the already performed transformations impose an human-based time alteration to the data itself (Section \ref{sec:cleanup-transform}).
-%
+
 \begin{figure}[h]
@@
 %
-The method we are going to apply is based on the Fourier Transform (FT), and generates a \textit{periodgram} highlighting the most interesting frequencies that might correspond to a seasonality component.
+The method we are going to apply is based on the Fourier Transform (FT), and generates a \textit{periodgram} highlighting the frequencies that might correspond to a seasonality component \cite[chapter 11]{brockwell1991time}.
 Applying the FT for any frequency, we transform our 1-dimensional data into a 2-dimensional data mapped around a circle.
@@
 
-A common operation to increase the difference between significant and non-significant periodicities is to apply a MA smoothing filter with a smaller window than the interesting ones.
+A common operation to increase the difference between significant and non-significant periodicities is to apply a Moving Average (MA) smoothing filter with a smaller window than the interesting ones.
 We will use such a filter with $p=2$.
 
-On this note we must specify that, when a significant periodicity is found, it influences smaller periodicities given by $f*\{2,3,\dots\}$ the significant frequency.
+On this note we must specify that, when a significant periodicity is found, it influences smaller periodicities given by $f*\{2,3,\dots\}$ the significant frequency \cite{brockwell1991time}.
 Also, if any periodicity is found that is longer than the original dataset, it must be considered as simple chance, rather than serendipity.
@@
 We can analyse the resulting periodgrams for Uccle and Membach stations in \figurename{s} \ref{fig:periodgram:uccs} and \ref{fig:periodgram:mems}.
-%
+An higher power value for a frequency means that the corresponding periodicity is meaningful for describing the dataset.
+
 \begin{figure}[h]
@@
 
-In particular from the Uccle station, we can see a great number of spikes, that will probably undermine our models.
-
 \subsubsection{Dataset division}
@@
 Given this issue, we choose the training/testing sets to hold 52/18 weeks: this is the greater train-to-test ratio we deem usable ($\approx 75\%/25\%$), while maintaining computational feasibility.
-The assessing of the lockdown influence will be discussed in Section~\ref{sec:methodology}.
+The assessment of the lockdown influence will be discussed in Section~\ref{sec:methodology}.
 
@@
 Also, please note that we are going to have a timeseries object indexed by weeks, so that it contains the 168 hours between each integer index.
-Since we know that the longest missing values series is 90 hours long, we can apply a simple completion strategy for serving those commands a filled timeseries: we will substitute missing values with the mean of the week they are in; in R terms \texttt{na.aggregate(x, floor)}.
+Since we know that the longest missing values series is 90 hours long, we can apply a simple completion strategy for serving those commands a filled timeseries: we will substitute missing values with the mean of the week they are in; in R, \texttt{na.aggregate(x, floor)}.
 
@@
 Since they do not produce particularly different seasonalities, we choose to keep the results obtained by applying the MA smoothing filter, for its significance in terms of seasonally adjusted data.
-%
+
 \begin{figure}[h]
@@
 \end{enumerate}
-%
+
 \begin{figure}[h]
@@
 
-Since our dataset shows a relevant trend component, resulting in non-constant mean, we consider the use of the \textit{diff}\footnote{The \textit{diff} operator applies the \textit{d}-order differences.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.7\linewidth]{"project.uccs_files/figure-latex/diffs/d168-acf-pacf-1"}
+	\caption{ACF and PACF plots after \textit{diff} at lag 168}
+	\label{fig:uccs:p-acf}
+\end{figure}
+%
+Since our dataset shows a relevant trend component, resulting in non-constant mean, we consider the usage of the \textit{diff}\footnote{The \textit{diff} operator applies the \textit{d}-order differences.
 All \textit{diff} operators we apply use $d=1$.} operator to smooth the mean.
@@
 \begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
-	\item from using both \textit{diff}s, the ACF and PACF on the long run hint to $\mathrm{ARIMA}(0,1,0)(0,1,1)_{168}$
-	\item from using just the lag 168 \textit{diff}, the PACF on the first lags and the ACF/PACF combination on the long run hint to $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$
+	\item from using both \textit{diff}s, the ACF and PACF on the long run hint to $\mathrm{ARIMA}(0,1,0)(0,1,1)_{168}$ (plots omitted for brevity)
+	\item from using just the lag 168 \textit{diff}, the PACF on the first lags and the ACF/PACF combination on the long run hint to $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$ (\figurename~\ref{fig:uccs:p-acf})
 \end{itemize}
@@
 Nevertheless, MAPE is not a single valid criterion for selection, since a low error predicting the training set itself does not necessarily mean a good forecast power.
-%
+
 \begin{figure}[h]
@@
 We can notice a clear overestimation during those couple of weeks (around week 62 on \figurename~\ref{fig:arima-forecast}).
-%
+
 \begin{figure}[h]
@@
 This will probably cause some approximation errors that could rig the final results.
-%
+
 \begin{figure}[h]
@@
 Thus, we consider the usage of the \textit{diff} operator, at lag 1 and 168.
-Interestingly, the analysis of ACF and PACF plots hints to similar models to the untransformed Uccle data:
-$\mathrm{ARIMA}(0,1,0)(0,1,1)_{168}$ and $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$. All ACF and PACF plots for Membach model estimation are not shown for brevity.
+The analysis of ACF and PACF plots hints to two models: $\mathrm{ARIMA}(0,1,0)(0,1,1)_{168}$ and $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$.
+All ACF and PACF plots for Membach model estimation are not shown for brevity.
 
@@
 Thus, we can assess the forecasts for both models in \figurename~\ref{fig:mems:arima}.
-%
+
 \begin{figure}[h]
@@
 
+Also, it is important to notice that while the first model includes in the confidence intervals negative values, non existent in our domain, the latter directs all the variability to higher values, while maintaining a consistent lower bound.
+This appears to be a good feature to correctly model our data.
+
 Analytically, we have quite a difference between the two models in terms of adherence of the confidence intervals:
@@
 
-TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
+Since checking the ACF and PACF plots for the original data could lead to a model with the same order, we fitted it for comparison.
+The MAPEs are higher but comparable, $7\%$ on the test set and $19\%$ on the train set, the Box-Pierce test refutes the independence between residuals, and $88\%/94\%$ of test data fits inside the $80\%$ and $95\%$ CI.
+Finally, it shows a variability that tends to admit lower values than the training data shows, including negative ones, while having a lower acceptance for higher values.
+We omit plots related to this model, since such results confirm the usage of the already selected model on the transformed data.
 
-As a side note, we followed the same procedure without applying the Box-Cox transformation, but the best model that could be chosen had more parameters and was less precise than the chosen model on the transformed data.
-
-TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO 
+It is important to highlight that no model we assessed found any pattern for predicting the high-valued spikes.
 %\subsubsection{MSTL}
@@
 
+Still, we can assess how the lockdown period is forecasted for both stations, and then compare their efficency.
+Also, we can extract the seasonalities, using the same MA filters we already used, separately from the periods before and during the lockdown, and compare also these results.
+
 \subsubsection{Seasonalities divergence}
-We can still compare how the seasonalities differ with respect with each original dataset and between each other, by extracting the seasonalities for each station from before and after the lockdown (March $14^{th}$).
-
 We will highlight the confidence bands for an estimated error of 67\%.
 This interval estimation lies on the assumption that values for each single hour in the hourly/weekly seasonality is Normally distributed, a slightly different assumption that is not necessarily refuted by the already seen Q-Q plots.
-%
+
 \begin{figure}[h]
@@
 	\end{subfigure}
-	\caption{Seasonalities computed before and after the lockdown}{\centering\small Seasonalities extracted together, confidence bands $\pm\mathrm{sd}$\\}
+	\caption{Seasonalities computed before and after the lockdown}{\centering\small Seasonalities extracted together with MA filters, confidence bands $\pm\mathrm{sd}$\\}
 	\label{fig:season.lockdown}
@@
 It is possible to notice a couple of changes: the Monday to Friday shapes are heavily flattened almost on the weekend values, and all the values are flattened towards 0.
-Both flattenings can be noticed checking the Uccle station, not showed for brevity.
+Both flattenings can be noticed also checking the Uccle station, not showed for brevity.
 
 \subsubsection{Forecast divergence}
-%
+
 \begin{figure}[h]
@@
 \section{Conclusions}
+TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
+
+Interestingly, the analysis of ACF and PACF plots for Membach hints to similar models to the untransformed Uccle data:
+$\mathrm{ARIMA}(0,1,0)(0,1,1)_{168}$ and $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$.
+
+TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
+
 \subsubsection{Lockdown influence}
 \subsection{Limitations and future works}
+forecast residuals of model with 24h seasonality, and
+
+seasonalities divergence -> check normality assumption for each hour to validate the gaussian-empirical confidence bands
+
 The main problems of this projects that we identified are multiple seasonalities and the Normality assumption.

### 2020-06-10 draft.tex
complete all but conclusions

@@
 
-We will not directly compare the two models, since data from the two stations have different orders of magnitude.
+We will not directly compare the two models, since data from the two stations have different orders of magnitude, and that difference should be firstly assessed in order to meaningfully compare the models.
 
@@
 In this case, the employment of \textit{auto.arima} stepwise methodology$^{\ref{footnote:auto.arima}}$ highlights the first model improved with one non-seasonal MA parameter more.
-This is confirmed by the ACF/PACF of its residuals.
-The analysis of ACF/PACF plots of the residuals of the second model do not lead to any improvement.
+This is confirmed by the ACF/PACF plots of its residuals.
+Their analysis for the second model do not lead to any improvement.
 The ACF and PACF plots of both models are similar to the ones from the chosen model for the Uccle station (\figurename~\ref{fig:arima-24}), and we can notice the same 24 hours remaining periodicity.
@@
 The Box-Pierce test statistics for both models accept the independence hypothesis between residuals.
-The two models have similar AICc values, and similar MAPE on the training set ($\approx6\%$).
+The two models have similar AIC values, and similar MAPE on the training set ($\approx6\%$).
 Thus, we can assess the forecasts for both models in \figurename~\ref{fig:mems:arima}.
@@
 Finally, it shows a variability that tends to admit lower values than the training data shows, including negative ones, while having a lower acceptance for higher values.
-We omit plots related to this model, since such results confirm the usage of the already selected model on the transformed data.
+We omit plots related to this model, since such results confirm the choice of the model on the transformed data.
 
@@
 Also, we can extract the seasonalities, using the same MA filters we already used, separately from the periods before and during the lockdown, and compare also these results.
-
-\subsubsection{Seasonalities divergence}
-We will highlight the confidence bands for an estimated error of 67\%.
-This interval estimation lies on the assumption that values for each single hour in the hourly/weekly seasonality is Normally distributed, a slightly different assumption that is not necessarily refuted by the already seen Q-Q plots.
 
@@
 %
+\subsubsection{Seasonalities divergence}
+We highlight the confidence bands for an estimated error of 67\%.
+This interval estimation lies on the assumption that values for each single hour in the hourly/weekly seasonality is Normally distributed, a slightly different assumption that is not necessarily refuted by the already seen Q-Q plots.
+
 As we can see in \figurename~\ref{fig:season.lockdown}, the seasonalities' shapes seem different, in particular comparing the weekly seasonality.
@@
 \subsubsection{Forecast divergence}
+Since both our models forecast stable values with almost stable variance (\figurename{s}~\ref{fig:arima-forecast:200011}, \ref{fig:mems:arima:200011}), we can forecast another 8 weeks ahead, until April $26^{th}$, and then compare the efficency of the results with the original test set.
+These 8 weeks comprise 6 weeks after March $14^{th}$, first day of lockdown in Belgium.
+
+As we can see in \figurename~\ref{fig:forecast.lockdown}, at both stations the models overestimate starting after the lockdown.
+Uccle data in particular show a greater overestimation, that can also be noticed by looking at the difference between the forecasted and measured values.
 
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-9"}
+		\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-5"}
 		\caption{Uccle station}
@@
 	\begin{subfigure}{.5\linewidth}
-		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-9"}
+		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-5"}
 		\caption{Membach station}
@@
 %
-By fitting the model\footnote{Data is different, because we are fitting the model on the original train+test dataset, but if we follow similar steps for both stations we can see that the final model we shall choose is the same.
-Also, the $\lambda$ parameter for the Box-Cox transformation of Membach data is estimated equally.} on data until March $1^{st}$ we can compute forecasts for 8 weeks ahead, until April $26^{th}$.
-As we can see in \figurename~\ref{fig:forecast.lockdown}, both stations overestimate starting after the lockdown.
-Uccle data in particular show a greater overestimation that can be noticed also by looking at the difference between the forecasted and measured values.
+We assess analytical values obtained from the new 8 weeks forecast:
+\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
+	\item the MAPE for Uccle triples with respect to the original test set to $\approx46\%$, and only 39\% and 63\% of the test data falls inside the confidence intevals (80\% and 95\%).
+	\item regarding the Membach station forecasts, the MAPE is reduced to $14\%$, while 82\% and 96\% of data falls inside the confidence intervals.
+\end{itemize}
 
-Analytically, the MAPE for Uccle doubles with respect to the previous test set ($\approx34\%$), and only 58\% and 78\% of the test data falls inside the confidence intevals (80\% and 95\%).
-Analysing the Membach station, we find those measures have comparable values to the previous test set: the MAPE remains at $16.6\%$, while 74\% and 90\% of data falls inside the confidence intervals.
+Thus, we notice a visible degradation in performance for Uccle station data, and a slight improvement for Membach station's.
+The slight improvement is most certainly driven by the single peak, with lower values than previous peaks, and by the increased variability for data further from the end of the training set.
 
@@
 \section{Conclusions}
+The models we choose for both stations show interesting predictive efficiency,
+
 TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
@@
 The main problems of this projects that we identified are multiple seasonalities and the Normality assumption.
+
+% comment by prof on some arima plots
+% If you're trying to compare what happens before and after lock-down, you should probably assess the forecast for the period without lock-down and then compare with how it performs for forecasting lock-down values.
+A large difference (forecast performing worse during lock-down) would indicate that the model doesn't work during lock-down because something has changed in the behaviour of the data, thus confirming your hypothesis
 

### 2020-06-11 draft.tex
Completed draft, removed notice

@@
 
-\title{Seismic Time Series Analysis\\Statistics for Spatio-Temporal Data\\\textit{-- project draft --}}
+\title{Seismic Time Series Analysis\\Statistics for Spatio-Temporal Data}
 \author{Marco Ciotola, 848222}
@@
 	\end{subfigure}
-	\caption{Forecasts from March $2^{nd}$.}{\centering\small In red the actual values, in black the forecasts. March $14^{th}$ at midnight is highlighted.\\}
+	\caption{Forecasts from March $2^{nd}$.}{\centering\small The forecasted values are on the background. March $14^{th}$ at midnight is highlighted.\\}
 	\label{fig:forecast.lockdown}
@@
 Thus, we notice a visible degradation in performance for Uccle station data, and a slight improvement for Membach station's.
-The slight improvement is most certainly driven by the single peak, with lower values than previous peaks, and by the increased variability for data further from the end of the training set.
-
-%MEMS
-%Test set inside the CI
-%80% CI 	  74.18 %
-%95% CI 	  89.58 %
-%Errors:  forecast with ARIMA(2,0,0)(0,1,1)[168] 
-%MSE= 117.9258   MAE= 7.709243   MAPE= 16.60103   RMSE= 10.85937   R2= 0.4777765
-
-%UCCS
-%Test set inside the CI
-%80% CI 	  57.51 %
-%95% CI 	  78.42 %
-%Errors:  forecast with ARIMA(2,0,0)(0,1,1)[168] 
-%MSE= 670897.8   MAE= 664.5903   MAPE= 33.94793   RMSE= 819.0835   R2= 0.5188538
-
-%Uccle: remodel on 52+18 (until 1st March 2020), forecast 2 months (8 weeks) => diff in MAPE
-%Membach: remodel on 52+18 (until 1st March 2020), forecast 2 months (8 weeks) => diff in MAPE
+The slight improvement is most certainly driven by the single peak with lower values than the multiple previous peaks, and by the increased variability for data further from the end of the training set.
 
 \section{Conclusions}
-The models we choose for both stations show interesting predictive efficiency,
+The final models for both stations show interesting predictive efficiency, even if analysing the ACF and PACF plots of their residuals we notice a remaining daily seasonality that is not explained by the models (\figurename~\ref{fig:arima-24}).
 
-TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
+The initial assumptions on the seasonalities are confirmed by our study, given the periodgrams (\figurename~\ref{fig:periodgram}) and the efficiency and structure of the models (seasonal terms).
+Though, the origin of those assumptions (human activity) cannot be solely confirmed by our methodology.
 
-Interestingly, the analysis of ACF and PACF plots for Membach hints to similar models to the untransformed Uccle data:
-$\mathrm{ARIMA}(0,1,0)(0,1,1)_{168}$ and $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$.
+Interestingly, both final models have the same order, $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$: still, they cannot be actually compared, given the Box-Cox transformation applied to the Membach data.
+% Still, a similar interpreation could be done
 
-TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
+\paragraph{Lockdown influence} The assessment of the lockdown influence on seismic data show relevant information for further studies.
 
-\subsubsection{Lockdown influence}
+Extracting both daily and weekly seasonalities with the local mean methodology (\figurename~\ref{fig:season.lockdown}), it is not possible to appreciate any statistically significant difference from before and during the lockdown on both datasets.
+
+Checking the seasonal plots of the complete weekly seasonality (\figurename~\ref{fig:season.lockdown:ms:mems}), it is possible to notice a flattening of the values towards 0, in particular regarding Monday to Friday values.
+
+The Uccle station, located near the Belgium capital city, show a relevant difference in forecasts after the lockdown, that are heavily out of the confidence bands, reaching over 45\% of MAPE.
+This is the opposite for data from Membach station, located in a small town near a natural park, that show no significant difference in forecasts performance.
+
+The forecast results hint to a significant correlation between the location of the seismic station and the influence of the lockdown.
+
 \subsection{Limitations and future works}
-forecast residuals of model with 24h seasonality, and
+The main problems identified during this analysis are the length and number of the datasets, the multiple seasonalities and the Normality assumption.
 
-seasonalities divergence -> check normality assumption for each hour to validate the gaussian-empirical confidence bands
+Even if after a first decomposition we chose to only consider the weekly seasonality, our models show a remaining daily seasonality.
+This could be addressed with various approach.
+One initial test could be to forecast a different model on the residuals with a daily seasonality, so that the composition of the two models could include both seasonalities.
 
-The main problems of this projects that we identified are multiple seasonalities and the Normality assumption.
+A second mean to address the multi-seasonalities is through a harmonic regression approach.
+Using that we could model multiple seasonalities at once through Fourier terms.
+One difficulty we could encounter is that our weekly seasonality appear to be flat for most of the week (\figurename~\ref{fig:season.lockdown}), and such a behaviour is not easy to model with a few Fourier terms per seasonality, leading to computationally heavy modelling and forecasting.
 
-% comment by prof on some arima plots
-% If you're trying to compare what happens before and after lock-down, you should probably assess the forecast for the period without lock-down and then compare with how it performs for forecasting lock-down values.
-A large difference (forecast performing worse during lock-down) would indicate that the model doesn't work during lock-down because something has changed in the behaviour of the data, thus confirming your hypothesis
+Different models, such as STL (Seasonal and Trend decomposition using Loess) or TBATS (Trigonometric seasonality, Box-Cox transformation, ARMA errors, Trend and Seasonal components), are suggested in the literature \cite[chapter 12]{hyndman2018forecasting} to address multiple and complex seasonalities, also when they change in time.
+These could also better tackle the missing Normality assumption of our data.
 
-The tested models do not effectively include both daily and weekly seasonalities, even if initially we thought so.
-Also they cannot include the apparent yearly seasonality, mainly because of the short timeseries.
-To overcome this limitation, we could try different models, also ARIMA-based, that can estimate multiple seasonalities.
-The first that we tested models seasonalites through Fourier terms.
-The main problem that lead us not to go further with that approach is that our seasonalities cannot be expressed by a combination of few sine and cosine curves.
-This requires a high number of curves to estimate, that requires an heavy computational power.
+A future study should be done when more data is available, in order to assess the presence of higher order seasonalities, such as a yearly one.
 
-To compensate those difficulties, in the literature it is suggested to test TBATS models, which combines Fourier terms, Box-Cox transformation, and exponential smoothing.
-This seems a suitable approach given the analyses we performed until now, that can possibly address the non-Normal distribution.
+The confidence bands used to assess any significant change in seasonalities lie on another Normality assumption, that data for each hour of the day/week follow a Normal distribution.
+This should be assessed in a future study integration.
 
-The final suggestion for future work would be to better analyse the differences between stations, in such a way to better correlate different influences of the lockdown on the areas near each station.
+On a more broad perspective, other analyses should be done on the lockdown impact, and more generally on isolating different factors that influence our datasets.
+Repeating the study on the other 7 stations will give interesting insights.
 
-\newpage
 \bibliography{project}

### 2020-06-11 project.after.Rmd
Completed draft, removed notice

@@
       + geom_vline(aes(xintercept=lockdown.x), linetype=2)
+      + scale_color_viridis_d()
 )
@@
 plot_grid(
-  autoplot(ts(season.after.lockdown.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
+  autoplot(ts(season.after.lockdown.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + labs(subtitle = "Daily seasonality", title=NULL) +
     geom_ribbon(aes(ymin = season.after.lockdown.24$figure  - season.after.lockdown.24$sd,  ymax = season.after.lockdown.24$figure  + season.after.lockdown.24$sd),  alpha=0.5, fill = "mediumpurple1") +
@@
     scale_colour_manual("", values=c("purple", "red"), breaks=c("After lockdown", "Before lockdown")),
-  autoplot(ts(season.after.lockdown.168.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
+  autoplot(ts(season.after.lockdown.168.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + labs(subtitle = "Weekly seasonality", title=NULL) +
     geom_ribbon(aes(ymin = season.after.lockdown.168.df$figure  - season.after.lockdown.168.df$sd,  ymax = season.after.lockdown.168.df$figure  + season.after.lockdown.168.df$sd),  alpha=0.5, fill = "mediumpurple1") +
@@
 
-autoplot(ts(seasonbase.after.lockdown.168.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
+autoplot(ts(seasonbase.after.lockdown.168.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + labs(subtitle = "Complete weekly seasonality", title=NULL) +
   geom_ribbon(aes(ymin = (seasonbase.after.lockdown.168.df$figure - seasonbase.after.lockdown.168.df$sd), ymax = (seasonbase.after.lockdown.168.df$figure + seasonbase.after.lockdown.168.df$sd)), fill = "mediumpurple1", alpha=0.5) +

### 2020-06-11 project.mems.Rmd
Completed draft, removed notice

@@
 for (model in to_test) {
-  # model <- c(4,1,2,0,1,0)
+  # model <- c(2,0,0,0,1,1)
   fit.name <- paste0("fit", paste0(model, collapse = ""))

### 2020-06-11 project.R

@@ -812,6 +812,7 @@ print(
     + ylab(data.lab)
     + xlab("Weeks")
     + geom_vline(aes(xintercept=lockdown.x), linetype=2)
+    + scale_color_viridis_d()
 )
 # print stats about data falling inside 80% and 90% CIs
 cat("Test set inside the CI\n")

### 2020-06-12 draft.tex
Last fixes before delivery

@@
 The correct estimation of a background noise, correlated or not with human activity, could help institutions to better isolate less prominent earthquakes and similar events from the seismic data.
-This could result in better understanding of the frequency of earthquakes and by studying their more-precise evolution over time.
+This could result in better understanding of the frequency of earthquakes by studying their more precise evolution over time.
 
 \subsubsection{Data source and description}
-The Royal Observatory of Belgium (ROB) gathers and publishes \cite{RoyalDataPolicy,Data20200402} data about vertical ground displacement from 9 stations through the country, linking them to seismic events whenever possible \cite{DataEvents20200402}.
+The Royal Observatory of Belgium (ROB) gathers and publishes \cite{RoyalDataPolicy,Data20200402} data about vertical ground displacement from 9 stations through the Country, linking them to seismic events whenever possible \cite{DataEvents20200402}.
 They also offer a data visualization tool for the same data \cite{DataVisualization20200402}.
@@
 	\item \textbf{Uccle}, \textbf{Sart-Tilman}, and \textbf{Ében-Émael} are near a city (Brussels, Liege, and Maastricht).
-	\item \textbf{Ostenda} This station is near the city of Ostenda, that borders with the sea.
-	\item \textbf{Membach} and \textbf{Dourbes} are near a natural park.
+	\item \textbf{Ostenda} is near the city of Ostenda, that borders with the sea.
+	\item \textbf{Membach} and \textbf{Dourbes} are in towns near a natural park.
 \end{itemize}
@@
 	\item For merging days, we need to add a column with the second-precise datetime
-	\item In days with missing values there are placeholder values instead a missing row, that would make the following rows misinterpreted for measurements on the wrong second.
+	\item In days with missing values there are placeholder values instead of a missing row, that would make the following rows misinterpreted for measurements at the wrong second.
 The placeholder values are one of these two min-max pairs: $(0,0),\, (-1,1)$.
@@
 \subsubsection{Data aggregation and initial transformation}
-Since we are not interested in a second-precise analysis, and it would be computationally expensive to deal with the whole dataset, we aggregate data by hour.
+Since we are not interested in a second-precise analysis, and it would be computationally expensive to deal with the whole dataset, we aggregate the data by hour.
 The aggregation results in around 14 thousand points, that can still be captured by interpretable models.
 
-While performing this operation, we should consider the correct time-zone: any seasonal analysis dealing with UTC dates would be out of phase for the half-year of legal time, with respect to the half-year of solar time.
-With the help of the \texttt{lubridate} R package we converted times to local time-zones, and then we kept them as absolute values, because all the R functions we use work on the UTC conversion of datetimes by default.
+While performing this operation, we should consider the correct time-zone: as we anticipated, any seasonal analysis dealing with UTC dates would be out of phase for the half-year of legal time, with respect to the half-year of solar time.
+Employing the \texttt{lubridate} R package we converted times to local time-zones, and then we kept them as absolute instants, because all the R functions we use work on the UTC conversion of datetimes by default.
 This transformation made it possible to have each hour in the timeseries to represent the local time in Belgium (\textit{Europe/Brussels}), without being influenced by the legal time.
 
-The aggregation was performed considering the mean total ground displacement, for the following reasons:
+The aggregation was performed considering the mean total ground displacement at each hour, for the following reasons:
 \begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
@@
 
-Since our main focus is on the seasonalities themselves, we aim to prevent assumptions based on this objective.
-In other words, we want to avoid searching for correlation with human activity while imposing human-based seasonality periods.
+Since one of the main focuses is on the seasonalities themselves, we aim to avoid searching for correlation with human activity while imposing human-based seasonality periods.
 Thus, we use an analytical method to extract seasonalities that the data itself suggests to us.
 Still, we recognize that the already performed transformations impose an human-based time alteration to the data itself (Section \ref{sec:cleanup-transform}).
+
+The method we are going to apply is based on the Fourier Transform (FT), and generates a \textit{periodgram} highlighting the frequencies that might correspond to a seasonality component \cite[chapter 11]{brockwell1991time}.
+Applying the FT for any frequency, we transform our 1-dimensional data into 2-dimensional data mapped around a circle.
+Then, we can sum all the resulting vectors (that are our new data points) to obtain a vector for each frequency, whose absolute value represents the power level of the same frequency.
+An high power level represents a significant periodicity on that frequency in our original data, since it means that the corresponding 2-dimensional data behaves in a highly similar way during each cycle.
 
@@
 %
-The method we are going to apply is based on the Fourier Transform (FT), and generates a \textit{periodgram} highlighting the frequencies that might correspond to a seasonality component \cite[chapter 11]{brockwell1991time}.
-Applying the FT for any frequency, we transform our 1-dimensional data into a 2-dimensional data mapped around a circle.
-Then, we can sum all the resulting vectors (that are our new data points) to obtain a vector for each frequency, whose absolute value represents the power level of the same frequency.
-An high power level represents a significant periodicity on that frequency in our original data.
-
 A common operation to increase the difference between significant and non-significant periodicities is to apply a Moving Average (MA) smoothing filter with a smaller window than the interesting ones.
@@
 
-On this note we must specify that, when a significant periodicity is found, it influences smaller periodicities given by $f*\{2,3,\dots\}$ the significant frequency \cite{brockwell1991time}.
+On this note we must specify that, when a significant periodicity is found, it influences smaller periodicities given by $f*\{2,3,\dots\}$ the significant frequency \cite[chapter 11]{brockwell1991time}.
 Also, if any periodicity is found that is longer than the original dataset, it must be considered as simple chance, rather than serendipity.
@@
 	\end{subfigure}
-	\caption{Periodgram results}
+	\caption{Periodgram plots}
 	\label{fig:periodgram}
@@
 %
-Both highlight multiple 24-hour periodicities.
+Both plots highlight multiple 24-hour periodicities.
 This is given by approximation factors, since we round the divisions result to the nearest natural number.
-It also highlights that using a single 24-hour seasonality could leave some information on the remaining components.
+It also highlights that using a single 24-hour seasonality could leave some information on the residuals.
 Still, in both stations it is the most significant periodicity.
@@
 \subsubsection{Dataset division}
-As best practices suggest, we are going to divide the dataset into training and testing datasets with a ratio around $85\%/15\% - 75\%/25\%$.
-Since one of our objectives is to assess any significant difference from before and after the Belgium lockdown (on March $14^{th}$ the \textit{soft} lockdown, while on March $18^{th}$ the complete lockdown), we selected data from the start of our timeseries until 20-01-2020 (64 weeks $\approx 82\%$) as training, while the test follows and ends on 27-04-2020 (14 weeks $\approx 18\%$).
+As best practices suggest, we divide the dataset into training and testing datasets with a ratio around $85\%/15\% - 75\%/25\%$.
+Since one of our objectives is to assess any significant difference from before and after the Belgium lockdown (on March $14^{th}$ the \textit{soft} lockdown, while on March $18^{th}$ the complete lockdown), we selected data from the start of our timeseries until 20-01-2020 (64 weeks $\approx 82\%$) as training, while the test follows ending on 27-04-2020 (14 weeks $\approx 18\%$).
 
@@
 Given this issue, we choose the training/testing sets to hold 52/18 weeks: this is the greater train-to-test ratio we deem usable ($\approx 75\%/25\%$), while maintaining computational feasibility.
-The assessment of the lockdown influence will be discussed in Section~\ref{sec:methodology}.
+The assessment of the lockdown influence will be discussed in Section~\ref{sec:methodology:lockdown}.
 
-Some R commands cannot handle timeseries with missing values, but as we have seen both stations datasets include missing values.
+Some R functions cannot handle timeseries with missing values, but as we have seen both stations' datasets include missing values.
 Also, please note that we are going to have a timeseries object indexed by weeks, so that it contains the 168 hours between each integer index.
@@
 \end{enumerate}
-While doing so, we tested different smoothing filters: simple filter (p=seasonality), Spencer's 15-point filter and MA smoothing filter (f=seasonality).
+While doing so, we tested different smoothing filters: simple filter (p=seasonality), Spencer's 15-point filter and MA smoothing filter (p=seasonality).
 Since they do not produce particularly different seasonalities, we choose to keep the results obtained by applying the MA smoothing filter, for its significance in terms of seasonally adjusted data.
@@
 	\end{subfigure}
-	\caption{Timeseries decomposition}{\centering\small Only the second method results are shown for clarity. Week 10 includes January $1^{st}$.\\}
+	\caption{Timeseries decomposition}{\centering\small Only the second method results are shown for clarity.
+Week 10 includes January $1^{st}$.\\\centering x-axes on left columns represent weeks, on right columns represent hours.\\}
 	\label{fig:decomposed}
@@
 %
-Since the periodgrams suggested the presence of two significative periodicities, we also tested the extraction of both corresponding seasonalities:
+Since the periodgrams suggested the presence of two significative periodicities, we also tested the extraction of both corresponding seasonalities (\figurename~\ref{fig:decomposed}):
 \begin{enumerate}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
-	\item Apply the MA smoothing filter (f=24) on the dataset to extract the trend component
+	\item Apply the MA smoothing filter (p=24) on the dataset to extract the trend component
 	\item Remove the trend from the dataset and extract the 24-hour seasonal figure using the local mean method
 	\item Remove the 24-hour seasonal components to obtain a deseasoned dataset
-	\item Apply the MA smoothing filter (f=168) on the deseasoned dataset to extract the updated trend component
+	\item Apply the MA smoothing filter (p=168) on the deseasoned dataset to extract the updated trend component
 	\item Remove the trend from the already deseasoned dataset and extract the 168-hour seasonal figure using the local mean method
@@
 
-From tha same decomposition results, we notice that data around January $1^{st}$ show a significant drop in values.
+From the same decomposition results, we notice that data around January $1^{st}$ show a significant drop in values.
 That is highlighted by both trends, where the overall minimum falls in that period, and by the residuals for Uccle data, showing a clear overestimation.
@@
 \section{Methodology}\label{sec:methodology}
-Having two different datasets, we will operate on them separately and then we will compare the model interpretation.
+Having two different datasets, we will operate on them separately and then we will compare the model interpretations.
 
@@
 
-The model we fit is the ARIMA model, including the seasonality components.
-To identify the order of the different components described by ARIMA, we will analyse the autocorrelation and partial autocorrelation of our datasets.
+The model we fit is the ARIMA\footnote{AutoRegressive Integrated Moving Average} model, including the seasonality components.
+To identify the order of the different components described by ARIMA, we will analyse the autocorrelation (ACF) and partial autocorrelation (PACF) functions of our datasets.
 As a reminder:
@@
 The same model will be then used to forecast other 8 weeks from March $2^{nd}$, so that we can assess the forecast power over the lockdown period.
-Given the fact that both our series are quite stable over time (\figurename{s} \ref{fig:decomposed:uccs}, \ref{fig:mems:box-cox:data}), we can expect that the predicted variability for further forecasts will not significantly increase.
+Given the fact that both our series are quite stable over time (\figurename{s} \ref{fig:decomposed:uccs}, \ref{fig:mems:box-cox:data}), we can expect that the predicted variability for further forecasts will not significantly increase, compromising the long-term forecasts.
 
@@
 This shows how our previous conclusion that the 168 hours seasonality could include the 24 hours one is flawed.
-Unfortunately, the R function we are using does not permit to estimate multiple seasonalities\footnote{\textit{Arima} method from the \textit{forecast} R package, only specifying the seasonal and non-seasonal orders.}.
+Unfortunately, how we are using the R function do not permit us to estimate multiple seasonalities\footnote{\textit{Arima} method from the \textit{forecast} R package, only specifying the seasonal and non-seasonal orders.}.
 
@@
 
-Also, it must be noted that a January $1^{st}$ period lies in our test dataset, that is already noticed as low-valued period (\figurename~\ref{fig:decomposed}).
+Also, it must be noted that a January $1^{st}$ period lies in our test dataset, that was already noticed as low-valued period (\figurename~\ref{fig:decomposed}).
 We can notice a clear overestimation during those couple of weeks (around week 62 on \figurename~\ref{fig:arima-forecast}).
@@
 	\end{subfigure}
-	\caption{ARIMA forecasts}{\footnotesize$\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$ behaves like $\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$ with faster increasing variablity}
+	\caption{ARIMA forecasts}{\centering\footnotesize$\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$ behaves like $\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$ with faster increasing variablity\\}
 	\label{fig:arima-forecast}
@@
 \subsubsection{ARIMA}
-Similarly to data from Uccle, data from Membach doesn't respect the Normality assumption.
+Similarly to data from Uccle, data from Membach does not respect the Normality assumption.
 We tested if the Q-Q plot showed significantly better approximation of the Normal distribution after applying the \textit{log} or \textit{Box-Cox} transformations, and the latter makes the data approach Normality with $\lambda\approx-1$ (\figurename~\ref{fig:mems:box-cox}).
@@
 
-In this case, the employment of \textit{auto.arima} stepwise methodology$^{\ref{footnote:auto.arima}}$ highlights the first model improved with one non-seasonal MA parameter more.
-This is confirmed by the ACF/PACF plots of its residuals.
-Their analysis for the second model do not lead to any improvement.
+The employment of \textit{auto.arima} stepwise methodology$^{\ref{footnote:auto.arima}}$ highlights the first model improved with one more non-seasonal MA parameter.
+This is confirmed by the ACF/PACF plots of its residuals, while their analysis for the second model do not lead to any improvement.
 The ACF and PACF plots of both models are similar to the ones from the chosen model for the Uccle station (\figurename~\ref{fig:arima-24}), and we can notice the same 24 hours remaining periodicity.
@@
 As we see, the first model over-generalises the variability, while the latter seems more precise, even if losing some information.
-Also considering the MAPE on the test set, the latter model fits the data better with a $16\%$ error, instead of the $18\%$ for the first model.
+Considering the MAPE on the test set, the latter model fits the data better with a $16\%$ error, instead of the $18\%$ for the first model.
 %
@@
 
-\subsection{Lockdown influence}
+\subsection{Lockdown influence}\label{sec:methodology:lockdown}
 Even if the two models have the same order, they cannot be directly compared using the Information Criteria, for the Membach model works after a transformation pass, and they are fitted from different data.
@@
 
-As we can see in \figurename~\ref{fig:forecast.lockdown}, at both stations the models overestimate starting after the lockdown.
-Uccle data in particular show a greater overestimation, that can also be noticed by looking at the difference between the forecasted and measured values.
+As we can see in \figurename~\ref{fig:forecast.lockdown}, at both stations the models overestimate starting after the lockdown. Uccle data in particular show a greater overestimation.
 
@@
 \begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
-	\item the MAPE for Uccle triples with respect to the original test set to $\approx46\%$, and only 39\% and 63\% of the test data falls inside the confidence intevals (80\% and 95\%).
+	\item the MAPE for Uccle triples with respect to the original test set to $\approx46\%$, and only 39\% and 63\% of the lockdown data falls inside the confidence intevals (80\% and 95\%).
 	\item regarding the Membach station forecasts, the MAPE is reduced to $14\%$, while 82\% and 96\% of data falls inside the confidence intervals.
@@
 Checking the seasonal plots of the complete weekly seasonality (\figurename~\ref{fig:season.lockdown:ms:mems}), it is possible to notice a flattening of the values towards 0, in particular regarding Monday to Friday values.
+This seem to confirm the initial hypotesis of lockdown influence.
 
@@
 A second mean to address the multi-seasonalities is through a harmonic regression approach.
-Using that we could model multiple seasonalities at once through Fourier terms.
+Using that, we could model multiple seasonalities at once through Fourier terms.
 One difficulty we could encounter is that our weekly seasonality appear to be flat for most of the week (\figurename~\ref{fig:season.lockdown}), and such a behaviour is not easy to model with a few Fourier terms per seasonality, leading to computationally heavy modelling and forecasting.
@@
 Different models, such as STL (Seasonal and Trend decomposition using Loess) or TBATS (Trigonometric seasonality, Box-Cox transformation, ARMA errors, Trend and Seasonal components), are suggested in the literature \cite[chapter 12]{hyndman2018forecasting} to address multiple and complex seasonalities, also when they change in time.
-These could also better tackle the missing Normality assumption of our data.
+These could better tackle the missing Normality assumption of our data, along with relaxing the initial transformation between timezones.
 
@@
 On a more broad perspective, other analyses should be done on the lockdown impact, and more generally on isolating different factors that influence our datasets.
-Repeating the study on the other 7 stations will give interesting insights.
+Repeating the study on the other 7 stations will surely give interesting insights.
 

### 2020-06-12 project.Rmd
Last fixes before delivery

@@
 axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-'%y", cex=0.8)
-text(x=blocks$start, y=hour(blocks$interval), labels = paste0(format(blocks$start, "%y-%m-%d %H:%M")), cex=0.9, pos=2)
-text(x=blocks$end, y=hour(blocks$interval), labels = paste0(format(blocks$end, "%y-%m-%d %H:%M")), cex=0.9, pos=4)
-text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=3)
-text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=1)
+text(x=blocks$start, y=hour(blocks$interval) - .2, labels = paste0(format(blocks$start, "%y-%m-%d %H:%M")), cex=0.9, pos=2)
+# text(x=blocks$end, y=hour(blocks$interval), labels = paste0(format(blocks$end, "%y-%m-%d %H:%M")), cex=0.9, pos=4)
+text(x=blocks$end, y=hour(blocks$interval) - .2, labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=4)
+#text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=3)
+#text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=1)
 blocks$start
@@
   geom_line(aes(freq, spec)) +
-  xlab("frequency") + ylab("spectrum") +
-  labs(title = "Periodgram", caption = paste0("bandwidth = ", pgram$bandwidth, "    df = ", pgram$df)) ->
+  xlab("frequency") + ylab("power level") +
+  labs(title = paste("Periodgram -", description), caption = paste0("bandwidth = ", pgram$bandwidth, "    df = ", pgram$df)) ->
   periodgram
 
-start_ratio <- 2
-filter_freq <- (pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*start_ratio)) 
+start_ratio <- 3
+filter_freq <- (pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*start_ratio)) & 1/pgram$freq/3600 < 24*365
 top95pc <- pgram$spec[filter_freq]
@@
   scale_x_continuous(sec.axis = sec_axis(~1/./3600, breaks=c(1, 3, 2*c(1,2,3,4,6,12), 24*7), labels = as.integer, name="hours")) +
-  scale_y_continuous(sec.axis = sec_axis(~., breaks=mean(pgram$spec) + sd(pgram$spec) * c(1,3), labels=c("1", "2", "3")[c(1,3)], name="spectrum/sd ratio")) +
-  geom_hline(yintercept=(mean(pgram$spec) + sd(pgram$spec) * c(1:3)), linetype="dotted", color = ratio.color(mean(pgram$spec) + sd(pgram$spec) * c(1:3))) +
+  scale_y_continuous(n.breaks=6) +
+  #scale_y_continuous(sec.axis = sec_axis(~., breaks=mean(pgram$spec) + sd(pgram$spec) * c(start_ratio:3), labels=c("1", "2", "3")[c(start_ratio:3)])) +
+  geom_hline(yintercept=(mean(pgram$spec) + sd(pgram$spec) * c(start_ratio:3)), linetype="dotted", color = ratio.color(mean(pgram$spec) + sd(pgram$spec) * c(start_ratio:3))) +
   geom_point(data=data.frame(freq=top95freq, spec=top95pc), aes(freq, (spec), color=ratio.color(spec)), shape="o", size=4, show.legend = F) +

### 2020-06-12 project.R

@@ -20,7 +20,7 @@ print_dec <- function(x, decimals=2) trimws(format(round(x, decimals), nsmall=de
 # - proj.data.RData
 # - proj.mems.RData (optional)
 # - proj.uccs.RData (optional)
-# setwd("/media/data/Documents/Uni/spatiotemp/villalobos/project/")
+# setwd("/media/data/Documents/Uni/spatiotemp/project/")
 # Load the datasets. proj.data.RData contains seis.uccs and seis.mems data.frame objects
 load('proj.data.RData')
 
@@ -197,10 +197,8 @@ plot(c(rbind(blocks$start, blocks$end, NA)), rep(hour(blocks$interval), each=3),
      xaxt="n", xlim=c(min(seis$utc_second), max(seis$utc_second))
 )
 axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-'%y", cex=0.8)
-text(x=blocks$start, y=hour(blocks$interval), labels = paste0(format(blocks$start, "%y-%m-%d %H:%M")), cex=0.9, pos=2)
-text(x=blocks$end, y=hour(blocks$interval), labels = paste0(format(blocks$end, "%y-%m-%d %H:%M")), cex=0.9, pos=4)
-text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=3)
-text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=1)
+text(x=blocks$start, y=hour(blocks$interval) - .2, labels = paste0(format(blocks$start, "%y-%m-%d %H:%M")), cex=0.9, pos=2)
+text(x=blocks$end, y=hour(blocks$interval) - .2, labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=4)
 # remove from RAM copy of dataset and cleanup
 remove(seis.complete, seis.missing.ids)
 gc()
@@ -241,20 +239,19 @@ build_ts <- function(dset, periods=NULL, start=NULL, end=NULL, force_ms=T) {
 
 # To analytically select the period, we follow these steps, using the periodgram
[…]

### 2020-06-14 project.Rmd
slides v1

@@
   scale_y_continuous(limits = c(min(seis.h.ts[(10*168):(16*168)]), max(seis.h.ts[(10*168):(16*168)]))) +
-  ylab(data.lab) + xlab(NULL) + labs(title = "Training set", subtitle = "4-weeks zoom", caption = "Data in original scale")
+  ylab(data.lab) + xlab(NULL) + labs(title = paste("4-weeks zoom -", description), caption = "Data in original scale")
 dec.addi <- decompose(seasonality.aggregate(seis.h.ts))
@@
 plot_grid(
-  autoplot(seis.h.train, main = "Hourly seismic movement", xlab = seasonality.xlab, ylab = "mean movement (nm)"),
-  gghistogram(seis.h.train, add.normal = T) + xlab("mean movement (nm)"),
+  autoplot(seis.h.train, main = "Hourly seismic movement", xlab = seasonality.xlab, ylab = data.lab),
+  gghistogram(seis.h.train, add.normal = T) + xlab(data.lab),
   nrow = 2)
@@
 plot_grid(plots[[1]], plots[[2]], nrow = 2)
+autoplot(seis.h.train) +
+  labs(title=paste("MA filtered trend -", description)) +
+  scale_x_continuous(n.breaks=6, labels=function(x) format(seis.h.start + weeks(x-1), "%d-%m-'%y")) +
+  geom_line(aes(y=trend.s, col="red"), show.legend=F) +
+  ylab(data.lab)
 

### 2020-06-14 project.uccs.Rmd
slides v1

@@
 # UCCS
-domain <- c(-350000,350000); pattern <- "UCCS-.+\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
+domain <- c(-350000,350000); pattern <- "UCCS-.+\\.csv"; name <- "UCCS"; description <- "Uccle, Brussels"; proj_file <- 'proj.uccs.RData'
 ```

### 2020-06-14 slides.tex
slides v1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND THEMES
%----------------------------------------------------------------------------------------

\documentclass{beamer}
\usepackage{subcaption}

\mode<presentation> {

	% The Beamer class comes with a number of default slide themes
	% which change the colors and layouts of slides. Below this is a list
	% of all the themes, uncomment each in turn to see what they look like.

	%\usetheme{default}
	%\usetheme{AnnArbor}
	%\usetheme{Antibes} %three-row header
	%\usetheme{Bergen} %side
	%\usetheme{Berkeley} %side
	%\usetheme{Berlin} %dots by sub + subsec header
	%\usetheme[secheader]{Boadilla} % maybe
	%\usetheme{Boadilla} % maybe
	%\usetheme{CambridgeUS}
	%\usetheme{Copenhagen} %heavy sec+subsec header
	%\usetheme{Darmstadt} % dots by sec + subsec header
	%\usetheme{Dresden} % maybe
	\usetheme{Frankfurt} % maybe % dots by sec 
	%\usetheme{Goettingen}
	%\usetheme{Hannover}
	%\usetheme{Ilmenau}
	%\usetheme{JuanLesPins}
	%\usetheme{Luebeck}
	%\usetheme{Madrid} % maybe
	%\usetheme{Malmoe}
	%\usetheme{Marburg}
	%\usetheme{Montpellier}
	%\usetheme{PaloAlto}
	%\usetheme{Pittsburgh}
	%\usetheme{Rochester}
	%\usetheme{Singapore} % maybe
	%\usetheme{Szeged}
	%\usetheme{Warsaw}

	% As well as themes, the Beamer class has a number of color themes
	% for any slide theme. Uncomment each of these in turn to see how it
	% changes the colors of your current slide theme.

	%\usecolortheme{albatross}
	%\usecolortheme{beaver} % rosso
	%\usecolortheme{beetle}
	%\usecolortheme{crane}
	%\usecolortheme{dolphin} % viola
	%\usecolortheme{dove}
	%\usecolortheme{fly}
	%\usecolortheme{lily} % viola, se fosse rosso...
	%\usecolortheme{orchid}
	%\usecolortheme{rose}
	%\usecolortheme{seagull}
	%\usecolortheme{seahorse}
	\usecolortheme{whale}
    %\usecolortheme{wolverine}

	%\setbeamertemplate{footline} % To remove the footer line in all slides uncomment this line
	%\setbeamertemplate{footline}[page number] % To replace the footer line in all slides with a simple slide count uncomment this line

	\setbeamertemplate{navigation symbols}{} % To remove the navigation symbols from the bottom of all slides uncomment this line

}

\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
%\usepackage{enumitem}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Seismic Time Series Analysis]{Seismic Time Series Analysis} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Marco Ciotola} % Your name
\institute[Ca' Foscari] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
	Ca' Foscari University\\ % Your institution for the title page
	\medskip
	\textit{848222@stud.unive.it} % Your email address
}
\date{June 16, 2020} % Date, can be changed to a custom date

\begin{document}

\begin{frame}[plain,noframenumbering]
	\titlepage % Print the title page as the first slide
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

%------------------------------------------------
\section{Introduction} % Sections can be created in order to organize your presentation into discrete blocks, all sections and subsections are automatically printed in the table of contents as an overview of the talk
%------------------------------------------------
\subsection{Goal}
\begin{frame}{Lockdown influence on seismic surveys}
	\begin{columns}[c]
		\column{0.666\textwidth}
		\centering\large COVID-19 outbreak\\
		\vspace{.5em}\centering\large$\Downarrow$\vspace{.5em}\\
		\centering\large Lockdown\\\small(March $14^{th}$, 2020) \\
		\vspace{.5em}\centering\large$\Downarrow$\vspace{.5em}\\
		\centering\large \textbf{Influence on background noise}\\
		\vspace{.5em}\centering\large$\Downarrow$\vspace{.5em}\\
		\centering\large \textit{Human activity correlation}\note{Our study cannot fulfill this point} \\
	\end{columns}
\end{frame}

\subsection{Dataset}

\begin{frame}{Data source}
	\begin{columns}[c]
		\column{0.666\textwidth}
		\begin{columns}[c]
			\column{0.13\textwidth}
			\column{0.2\textwidth}
			\begin{figure}
				\includegraphics[width=\linewidth,clip=true]{slides_files/ROB-logo}
			\end{figure}
			\column{0.57\textwidth}\vspace{.4em}\\
			\small Royal Observatory of Belgium\\\scriptsize\textit{Seismology-Gravimetry}
			\column{0.1\textwidth}
		\end{columns}
		\begin{figure}
			\includegraphics[width=.8\linewidth]{slides_files/stations-map-smol-2}
		\end{figure}
		\column{0.3\textwidth}
		\centering 9 stations
		\\\vspace{.3em}$\downarrow$
		\\\vspace{.3em}2 selected
		\\~\\~
		\\\vspace{.25em}\textbf{Uccle}\\in Brussels\\~
		\\\vspace{.25em}\textbf{Membach}\\mountain area\\near a national park\\~
		\column{0.033\textwidth}
	\end{columns}
\end{frame}

\begin{frame}{Data description}
	\begin{columns}[T]
		\column{.5\textwidth}
		(min, max) displacements\hfill$\rightarrow$\\~\\
		second precision\hfill$\rightarrow$\\~\\
		UTC times\hfill$\rightarrow$\\~\\
		missing values\hfill$\rightarrow$
		\column{.5\textwidth}
		total displacement (difference)\\~\\
		hourly mean\\~\\
		legal-time conversion\\~\\
		fill with weekly mean\\(when needed)
	\end{columns}
\end{frame}

\begin{frame}{Dataset division}\centering
	%Data from October $29^{th}$, 2018 to April $27^{th}$ 2020\\~\\
	We want to be able to verify our models on pre-lockdown data
	\vspace{0.1\textheight}
	\begin{columns}
		\column{0.90\textwidth}
		\begin{columns}
			\column{0.54\textwidth}\hspace{-.005\textwidth}29-10-2018\\\hspace{-.005\textwidth}\tiny$\vert$
			\column{0.255\textwidth}\centering
			28-10-2019\\\tiny$\vert$
			\column{0.205\textwidth}\centering
			02-03-2020\\\tiny$\vert$
		\end{columns}
		\begin{columns}
			\column{0.6667\textwidth}$\underbrace{\rule[.5em]{\textwidth}{0.15em}}$
			\column{0.2307\textwidth}$\underbrace{\rule[.5em]{\textwidth}{0.15em}}$
			\column{0.1026\textwidth}$\underbrace{\rule[.5em]{\textwidth}{0.15em}}$
		\end{columns}
		\begin{columns}
			\column{0.6667\textwidth}\centering
			Training set - \textbf{67\%}
			\column{0.2307\textwidth}\centering
			\hspace{-.3em}Test set - \textbf{23\%}\hspace{-.3em}
			\column{0.1026\textwidth}\centering
			\textbf{10\%}
		\end{columns}
	\end{columns}
	\vspace{0.1\textheight}
	Fit on training $\rightarrow$ validate on test $\rightarrow$ assess lockdown influence
\end{frame}

%------------------------------------------------
\section{Initial analysis}
%------------------------------------------------

\subsection{Seasonality}
\begin{frame}{\subsecname~- Visual inspection}
	\centering Plot review\\
	$\Downarrow$\\ %hints
	Daily and weekly seasonalities\\~
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/whole data plot-2"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/whole data plot-2"}
		\end{subfigure}
		\caption{Zoom from January $6^{th}$ to February $3^{rd}$ 2019}
	\end{figure}
\end{frame}

\begin{frame}{\subsecname~- Periodgram}
	\centering Fourier Transform with different frequencies\\
	$\Downarrow$\\ %reveals
	Daily and weekly seasonalities\\~
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/periodgram-1"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/periodgram-1"}
		\end{subfigure}
	\end{figure}
\end{frame}

\subsection{Decomposition}
\begin{frame}{\subsecname~- Trend extraction}\centering
	Tested simple, Spencer's and Moving Average filters
	\begin{figure}
		\includegraphics[width=.55\linewidth]{"project.uccs_files/figure-latex/trend MA-2"}
	\end{figure}
	\large$\Downarrow$\\\normalsize\textbf{Moving Average}\\\small better seasonal features
\end{frame}

\begin{frame}{\subsecname~- Seasonality extraction}
	\centering MA filter for detrending\\+\\local mean method%\\
%	single \& multiple seasonalities extraction
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/extract both 24 and 168-2"}
			\caption{Uccle, Brussels}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/extract both 24 and 168-2"}
			\caption{Membach}
		\end{subfigure}
	\end{figure}
	\large$\Downarrow$\\
	\normalsize\textbf{Only weekly seasonality}\\\small fitting models
\end{frame}

\subsection{Lockdown impact}
\begin{frame}{\subsecname~- Multiple seasonalities}\centering
	Seasonalities extracted together
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/seasonalities/before-after-compare-1"}
			\caption{Uccle, Brussels}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/before-after-compare-1"}
			\caption{Membach}
		\end{subfigure}
	\end{figure}
	\large$\Downarrow$\\
	\normalsize No significant difference
\end{frame}

\begin{frame}{\subsecname~- Weekly seasonality}\centering
	Only weekly seasonality extracted
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/seasonplot-mean-sd-1"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/seasonplot-mean-sd-2"}
		\end{subfigure}
	\end{figure}
	\large$\Downarrow$\\
	\vspace{.3em}\normalsize Monday to Friday lower values\\
	\vspace{.3em}Overall flattening towards 0
\end{frame}

%------------------------------------------------
\section{Parameters estimation}
%------------------------------------------------
\subsection{Uccle}

\begin{frame}{\subsecname~- ARMA orders}
	\begin{figure}[h]
		\centering
		\caption{ACF and PACF plots after \textit{diff} at lag 168}
		\includegraphics[width=0.7\linewidth]{"project.uccs_files/figure-latex/diffs/d168-acf-pacf-1"}
	\end{figure}
	\begin{columns}
		\column{.22\linewidth}
		\column{.28\linewidth}\centering
		$\Downarrow$\\
		non-seasonal AR\\
		order 2
		\column{.38\linewidth}\centering
		$\Downarrow$\\
		seasonal MA\\
		order 1
		\column{.12\linewidth}
	\end{columns}~\\~
%	\centering	This hints to $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$
\end{frame}

\begin{frame}{\subsecname~- Remaining seasonality}
	\centering
	ACF and PACF plots highlight 24-hours seasonality on residuals\\~
	\begin{figure}[h]
		\centering
		\includegraphics[width=0.7\linewidth]{"project.uccs_files/figure-latex/arima-35"}
	\end{figure}
\end{frame}

\begin{frame}{\subsecname~- Models comparison}
	\begin{table}
		\begin{tabular}{c | c c c}
%			\toprule
			               &     & (refit) & Box-Pierce \\
			\textbf{ARIMA} & AIC & MAPE    & p-value \\
			\midrule
			$(2,0,0)(0,1,1)_{168}$ & 119700.1 & 5.89\% & 0.333 \\
			$(3,1,1)(0,1,1)_{168}$ & 119513.4 & 6.10\% & 0.854 \\
			$(4,1,2)(0,1,1)_{168}$ & 119396.9 & 6.13\% & 0.109\\
%			\bottomrule
		\end{tabular}
	\end{table}
	\vspace{0.1\textheight}\centering
	Other models have higher AIC and MAPE, while p-value $< 0.05$
\end{frame}

\begin{frame}{\subsecname~- Models validation}
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/arima-37"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/arima-69"}
		\end{subfigure}
	\end{figure}
	\begin{columns}[T]
		\column{.28\linewidth}\raggedleft~\\~\\\vspace{1em}accepted model $\Rightarrow$\hspace{-1.3em}
		\column{.72\linewidth}
		\begin{table}
			\begin{tabular}{c | c c c}
				%			\toprule
				\textbf{ARIMA} & MAPE   & 80\% CI & 95\% CI \\
				\midrule
				$\mathbf{(2,0,0)(0,1,1)_{168}}$ & 17.48\% & $\approx81\%$ & $\approx91\%$ \\
				$(3,1,1)(0,1,1)_{168}$ & 15.18\% & $\approx96\%$ & $\approx98\%$ \\
				$(4,1,2)(0,1,1)_{168}$ & 14.68\% & $\approx98\%$ & $\approx99\%$ \\
				%			\bottomrule
			\end{tabular}
		\end{table}
		\column{.05\linewidth}
	\end{columns}
\end{frame}

\subsection{Membach}

\begin{frame}{\subsecname~- Data transformation}\centering
	Spikes in data make it heavily non-Normal\\$\Downarrow$\\
	Transform using Box-Cox with $\lambda\approx-1$\\ %\quad\Rightarrow\quad y'=(y-1)/y$\\
	\vspace{.5em}\small Safe since all our data is positive\\\vspace{1em}\large$\Downarrow$\vspace{.5em}\\
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/training data transformation plots-7"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/training data transformation plots-9"}
		\end{subfigure}
	\end{figure}
\end{frame}

\begin{frame}{\subsecname~- Models validation}
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/arima-21"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/arima-13"}
		\end{subfigure}
	\end{figure}
	\begin{columns}[T]
		\column{.28\linewidth}\raggedleft~\\~\\~\\~\\\vspace{1em}accepted model $\Rightarrow$ \hspace{-1.5em}
		\column{.72\linewidth}
		\begin{table}
			\begin{tabular}{c | c c c}
				%			\toprule
				\textbf{ARIMA} & MAPE   & 80\% CI & 95\% CI \\
				\midrule
				$(0,1,0)(0,1,1)_{168}$ & 17.25\% & $\approx99\%$ & $100\%$ \\
				$(0,1,1)(0,1,1)_{168}$ & 17.98\% & $\approx98\%$ & $\approx99\%$ \\
				$\mathbf{(2,0,0)(0,1,1)_{168}}$ & 16.14\% & $\approx73\%$ & $\approx89\%$ \\
				%			\bottomrule
			\end{tabular}
		\end{table}
		\column{.05\linewidth}
	\end{columns}
\end{frame}

\subsection{Lockdown impact}

\begin{frame}{\subsecname~- Forecasts assessment}
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-5"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-5"}
		\end{subfigure}
	\end{figure}
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-9"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-9"}
		\end{subfigure}
	\end{figure}
\end{frame}

\begin{frame}{\subsecname~- Models performance}
	\centering Different stations behaviour\\before lockdown $\rightarrow$ during lockdown\\~\\
	\begin{columns}[T]
		\column{.12\linewidth}
		\column{.3\linewidth}\centering
		\textbf{Uccle, Brussels}~~~
		\column{.15\linewidth}\centering
		\column{.35\linewidth}\centering
		\textbf{Membach}
		\column{.08\linewidth}
	\end{columns}
	\begin{columns}[T]
		\column{.12\linewidth}
		\column{.15\linewidth}\raggedleft
		${16\%} \rightarrow$\\
		${73\%} \rightarrow$\\
		${89\%} \rightarrow$\\
		\column{.12\linewidth}\raggedright
		$46\%$\\
		$39\%$\\
		$63\%$\\
		\column{.21\linewidth}\centering
		\textbf{$<$ MAPE $>$}\\\textbf{$<$ 80\% CI $>$}\\\textbf{$<$ 95\% CI $>$}\\
		\column{.17\linewidth}\raggedleft
		$17\% \rightarrow$\\
		$81\% \rightarrow$\\
		$91\% \rightarrow$\\
		\column{.15\linewidth}\raggedright
		${14\%}$\\
		${82\%}$\\
		${96\%}$\\
		\column{.08\linewidth}
	\end{columns}~\\
	\begin{columns}[T]
		\column{.5\linewidth}\centering
		$\Downarrow$\\
		Performance degradation
		\column{.5\linewidth}\centering
		$\Downarrow$\\
		Slight performance improvement\\
		\textit{lower peaks}
	\end{columns}
\end{frame}

%------------------------------------------------
\section{Conclusion}
%------------------------------------------------
\subsection{Models}

\begin{frame}{Models performance}
	- Periodgrams and models confirm \textbf{weekly} and \textbf{daily seasonality}

	~

	- Both chosen models have \textbf{interesting forecasting power}

	~ ~ \textit{Missed daily seasonality}

	~

	- Models have \textbf{same number of parameters} for each component

	~ ~ \textit{Membach data is transformed}

	~

	- Stable \textbf{seasonality explains most of the variability}

	~ ~ \textit{Membach peaks not explained}
\end{frame}

\begin{frame}{Lockdown influence}
	Mixed findings:

	~

	~ - Influence results \textbf{non-significant on seasonality}

	~

	~ - Influence on forecasts \textbf{depends on the station}

	~ ~ ~ \textit{Correlation with station location}

\end{frame}

\subsection{Future Works}
\begin{frame}{Future Works}
	- Models with \textbf{multiple seasonality} components

	~ ~ \textit{MSTL, TBATS, ARIMA with Fourier Terms, ...}

	~

	- Analysis on the other \textbf{7 stations}

	~

	- Analysis when \textbf{$\mathbf{> 2}$ years} of data available

	~

	- Different factors modify \textbf{lockdown impact}

	~ ~ \textit{Location, dimension of nearest city, ...}
\end{frame}

\title{Thank you for your attention}
\date{}
\author{}
\institute{}
\begin{frame}[plain,noframenumbering]
	\titlepage
\end{frame}

\end{document}

### 2020-06-15 project.Rmd
separate and better Lockdown impact

@@
 ```{r show_overlap_and_diff, include=FALSE}
-show_overlap_and_diff <- function(base, compare, base.desc, compare.desc, xlab, overlap.ylab, diff.title, diff.ylab, diff.ylab.sec, diff.ylab.sec.accuracy=.1) {
+show_overlap_and_diff <- function(base, compare, base.desc, compare.desc, xlab, overlap.ylab, diff.title, diff.ylab, diff.ylab.sec, diff.ylab.sec.accuracy=.1, base.lab="red", compare.lab="black", overlap.legend=F) {
   overlap_plot <-
-    autoplot(compare, xlab=xlab, ylab=overlap.ylab) +
-      geom_line(aes(y=base, col="red"), show.legend = F) +
+    autoplot(compare, xlab=xlab, ylab=overlap.ylab, series=compare.lab, show.legend = overlap.legend) +
+      geom_line(aes(y=base, col=base.lab), show.legend = overlap.legend) +
       labs(title=NULL, subtitle=paste0(base.desc, " vs ", compare.desc))

### 2020-06-15 project.after.Rmd
separate and better Lockdown impact

@@
   final.testing, window(plotting.forecast$mean, start=c(71,1), end=c(78,168)), xlab=seasonality.xlab,
-  base.desc = "test set during lockdown",
+  base.desc = "test set before lockdown",
+  base.lab = "Actual",
   compare.desc = paste0("forecast with ", plotting.forecast$method),
+  compare.lab = "Forecast",
+  overlap.legend = T,
   overlap.ylab = data.lab,
@@
 print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), plots[[2]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), nrow = 2))
+print(
+  plots[[1]]
+    + geom_vline(aes(xintercept=lockdown.x), linetype=2)
+    + labs(title=paste0("Forecasts during lockdown with ", plotting.forecast$method), subtitle=NULL)
+)
 

### 2020-06-15 slides.tex
separate and better Lockdown impact

@@
 	\begin{figure}
-		\begin{subfigure}{.45\linewidth}
+		\begin{subfigure}{.49\linewidth}
 			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/whole data plot-2"}
 		\end{subfigure}
-		\begin{subfigure}{.45\linewidth}
+		\begin{subfigure}{.49\linewidth}
 			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/whole data plot-2"}
@@
 	\begin{figure}
-		\begin{subfigure}{.45\linewidth}
+		\begin{subfigure}{.49\linewidth}
 			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/periodgram-1"}
 		\end{subfigure}
-		\begin{subfigure}{.45\linewidth}
+		\begin{subfigure}{.49\linewidth}
 			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/periodgram-1"}
@@
 	\begin{figure}
-		\includegraphics[width=.55\linewidth]{"project.uccs_files/figure-latex/trend MA-2"}
+		\includegraphics[width=.6\linewidth]{"project.uccs_files/figure-latex/trend MA-2"}
 	\end{figure}
@@
 	\begin{figure}
-		\begin{subfigure}{.45\linewidth}
+		\begin{subfigure}{.49\linewidth}
 			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/extract both 24 and 168-2"}
@@
 		\end{subfigure}
-		\begin{subfigure}{.45\linewidth}
+		\begin{subfigure}{.49\linewidth}
 			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/extract both 24 and 168-2"}
@@
 	\normalsize\textbf{Only weekly seasonality}\\\small fitting models
-\end{frame}
-
-\subsection{Lockdown impact}
-\begin{frame}{\subsecname~- Multiple seasonalities}\centering
-	Seasonalities extracted together
-	\begin{figure}
-		\begin{subfigure}{.45\linewidth}
-			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/seasonalities/before-after-compare-1"}
-			\caption{Uccle, Brussels}
-		\end{subfigure}
-		\begin{subfigure}{.45\linewidth}
-			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/before-after-compare-1"}
-			\caption{Membach}
-		\end{subfigure}
-	\end{figure}
-	\large$\Downarrow$\\
-	\normalsize No significant difference
-\end{frame}
-
-\begin{frame}{\subsecname~- Weekly seasonality}\centering
-	Only weekly seasonality extracted
-	\begin{figure}
-		\begin{subfigure}{.45\linewidth}
-			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/seasonplot-mean-sd-1"}
-		\end{subfigure}
-		\begin{subfigure}{.45\linewidth}
-			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/seasonplot-mean-sd-2"}
-		\end{subfigure}
-	\end{figure}
-	\large$\Downarrow$\\
-	\vspace{.3em}\normalsize Monday to Friday lower values\\
-	\vspace{.3em}Overall flattening towards 0
 \end{frame}
@@
 	\begin{figure}
-		\begin{subfigure}{.45\linewidth}
+		\begin{subfigure}{.49\linewidth}
 			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/arima-37"}
 		\end{subfigure}
-		\begin{subfigure}{.45\linewidth}
+		\begin{subfigure}{.49\linewidth}
 			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/arima-69"}
@@
 	\begin{figure}
-		\begin{subfigure}{.45\linewidth}
+		\begin{subfigure}{.49\linewidth}
 			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/training data transformation plots-7"}
 		\end{subfigure}
-		\begin{subfigure}{.45\linewidth}
+		\begin{subfigure}{.49\linewidth}
 			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/training data transformation plots-9"}
@@
 	\begin{figure}
-		\begin{subfigure}{.45\linewidth}
+		\begin{subfigure}{.49\linewidth}
 			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/arima-21"}
 		\end{subfigure}
-		\begin{subfigure}{.45\linewidth}
+		\begin{subfigure}{.49\linewidth}
 			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/arima-13"}
@@
 
-\subsection{Lockdown impact}
+%------------------------------------------------
+\section{Lockdown impact}
+%------------------------------------------------
 
-\begin{frame}{\subsecname~- Forecasts assessment}
+\subsection{Seasonalities}
+
+\begin{frame}{Multiple seasonalities}\centering
+	Seasonalities extracted together
 	\begin{figure}
-		\begin{subfigure}{.45\linewidth}
-			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-5"}
+		\begin{subfigure}{.49\linewidth}
+			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/seasonalities/before-after-compare-1"}
+			\caption{Uccle, Brussels}
 		\end{subfigure}
-		\begin{subfigure}{.45\linewidth}
-			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-5"}
+		\begin{subfigure}{.49\linewidth}
+			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/before-after-compare-1"}
+			\caption{Membach}
 		\end{subfigure}
 	\end{figure}
+	\large$\Downarrow$\\
+	\normalsize No significant difference
+\end{frame}
+
+\begin{frame}{Weekly seasonality}\centering
+	Only weekly seasonality extracted
 	\begin{figure}
-		\begin{subfigure}{.45\linewidth}
-			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-9"}
+		\begin{subfigure}{.49\linewidth}
+			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/seasonplot-mean-sd-1"}
 		\end{subfigure}
-		\begin{subfigure}{.45\linewidth}
-			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-9"}
+		\begin{subfigure}{.49\linewidth}
+			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/seasonplot-mean-sd-2"}
 		\end{subfigure}
 	\end{figure}
+	\large$\Downarrow$\\
+	\vspace{.3em}\normalsize Monday to Friday lower values\\
+	\vspace{.3em}Overall flattening towards 0
 \end{frame}
 
-\begin{frame}{\subsecname~- Models performance}
+\subsection{Forecasts}
+
+\begin{frame}{Forecasts assessment}
+	\begin{columns}[c]
+		\column{.5\linewidth}
+		\begin{figure}
+			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-5"}
+		\end{figure}
+		\begin{figure}
+			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-10"}
+		\end{figure}
+		\column{.5\linewidth}\centering
+		\textbf{Uccle, Brussels}\\
+		lockdown $\rightarrow$ during lockdown\\~\\
+		\begin{columns}[T]
+			\column{.2\linewidth}
+			\column{.33\linewidth}\raggedleft
+			\textbf{MAPE}\\\textbf{80\% CI}\\\textbf{95\% CI}\\
+			\column{.2\linewidth}\raggedleft
+			${16\%} \rightarrow$\\
+			${73\%} \rightarrow$\\
+			${89\%} \rightarrow$\\
+			\column{.1\linewidth}\raggedright
+			$46\%$\\
+			$39\%$\\
+			$63\%$\\
+			\column{.27\linewidth}
+		\end{columns}~\\
+		$\Downarrow$\\
+		Performance degradation
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Forecasts assessment}
+	\begin{columns}[c]
+		\column{.5\linewidth}
+		\begin{figure}
+			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-5"}
+		\end{figure}
+		\begin{figure}
+			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-10"}
+		\end{figure}
+		\column{.5\linewidth}\centering
+		\textbf{Membach}\\
+		lockdown $\rightarrow$ during lockdown\\~\\
+		\begin{columns}[T]
+		\column{.2\linewidth}
+			\column{.33\linewidth}\raggedleft
+			\textbf{MAPE}\\\textbf{80\% CI}\\\textbf{95\% CI}\\
+			\column{.2\linewidth}\raggedleft
+			$17\% \rightarrow$\\
+			$81\% \rightarrow$\\
+			$91\% \rightarrow$\\
+			\column{.1\linewidth}\raggedright
+			${14\%}$\\
+			${82\%}$\\
+			${96\%}$\\
+			\column{.27\linewidth}
+		\end{columns}~\\
+		$\Downarrow$\\
+		Slight performance improvement\\
+		\textit{lower peaks}
+	\end{columns}
+\end{frame}
+
+\begin{frame}<0>[noframenumbering]{Models performance}
 	\centering Different stations behaviour\\before lockdown $\rightarrow$ during lockdown\\~\\
